### 2025-04-25 manuscript_results.Rmd
Add files via upload

---
title: "Final manuscript"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
---

This document summarizes the results of the 12 models that were developed: AUTO ARIMA, AUTO ADJACENT, AUTO EPIWEEK, ES27 TEMPERATURE, ES27 ARIMA, ES27 ADJACENT, ES27 EPIWEEK, ES64 TEMPERATURE, ES64 ARIMA, ES64 ADJACENT, ES64 EPIWEEK, ES64 TEMPERATURE. It has a series of histograms and maps that compare the results of these models based on their WIS. You can read more about the specific covariates and ensemble techniques of these models in their own Rmarkdown documents.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries.

```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
library("ggplot2")
library("sf")
library("forcats")
library("ggplot2")
library("sf")
library("dplyr")
library("scales")  # For `breaks_extended`
library("ggplot2")
library("broom")
library("fields")

```

Loading the results of each model and the shapefiles of the maps.

```{r}
load("NHSH_hospitalizations_seed_nolog_correct/ES_ARIMA/ARIMA_MODELS_influenza_hospitalization_nolog.Rdata")
load("NHSH_hospitalizations_seed_nolog_correct/ES_ADJACENT/ADJACENT_MODELS_influenza_hospitalization_nolog.Rdata")
load("NHSH_hospitalizations_seed_nolog_correct/ES_EPIWEEK/EPIWEEK_MODELS_influenza_hospitalization_nolog.Rdata")
load("NHSH_hospitalizations_seed_nolog_correct/ES_TEMPERATURE/TEMPERATURE_MODELS_influenza_hospitalization_nolog.Rdata")
load("NHSH_hospitalizations_seed_nolog_correct/ES_AVERAGE/AVERAGE_MODELS_influenza_hospitalization_nolog.Rdata")

states <- read_sf("NHSH_hospitalizations_seed_nolog_correct/shapefiles/cb_2018_us_state_500k.shp")

states <- states %>%
  rename(STATE = NAME)

```

------------
1 Week ahead
------------
```{r}

# Creating new columns with Epidemiological weeks based on target_end_week
AUTO_ARIMA_WEEK1$epiweek <- MMWRweek(AUTO_ARIMA_WEEK1$target_end_date)$MMWRweek
AUTO_ADJACENT_WEEK1$epiweek <- MMWRweek(AUTO_ADJACENT_WEEK1$target_end_date)$MMWRweek
AUTO_EPIWEEK_WEEK1$epiweek <- MMWRweek(AUTO_EPIWEEK_WEEK1$target_end_date)$MMWRweek
AUTO_TEMPERATURE_WEEK1$epiweek <- MMWRweek(AUTO_TEMPERATURE_WEEK1$target_end_date)$MMWRweek
AUTO_AVERAGE_WEEK1$epiweek <- MMWRweek(AUTO_AVERAGE_WEEK1$target_end_date)$MMWRweek

ES27_ARIMA_WEEK1$epiweek <- MMWRweek(ES27_ARIMA_WEEK1$target_end_date)$MMWRweek
ES27_ADJACENT_WEEK1$epiweek <- MMWRweek(ES27_ADJACENT_WEEK1$target_end_date)$MMWRweek
ES27_EPIWEEK_WEEK1$epiweek <- MMWRweek(ES27_EPIWEEK_WEEK1$target_end_date)$MMWRweek
ES27_TEMPERATURE_WEEK1$epiweek <- MMWRweek(ES27_TEMPERATURE_WEEK1$target_end_date)$MMWRweek
ES27_AVERAGE_WEEK1$epiweek <- MMWRweek(ES27_AVERAGE_WEEK1$target_end_date)$MMWRweek

ES64_ARIMA_WEEK1$epiweek <- MMWRweek(ES64_ARIMA_WEEK1$target_end_date)$MMWRweek
ES64_ADJACENT_WEEK1$epiweek <- MMWRweek(ES64_ADJACENT_WEEK1$target_end_date)$MMWRweek
ES64_EPIWEEK_WEEK1$epiweek <- MMWRweek(ES64_EPIWEEK_WEEK1$target_end_date)$MMWRweek
ES64_TEMPERATURE_WEEK1$epiweek <- MMWRweek(ES64_TEMPERATURE_WEEK1$target_end_date)$MMWRweek
ES64_AVERAGE_WEEK1$epiweek <- MMWRweek(ES64_AVERAGE_WEEK1$target_end_date)$MMWRweek

# Dataframe that will be analysed for 1 Week Ahead
df_W1_NoLg <- data.frame(
  STATE = AUTO_ARIMA_WEEK1$State,
  Julian_date = AUTO_ARIMA_WEEK1$target_end_date,
  epiweek = AUTO_ARIMA_WEEK1$epiweek,
  AUTO_AR_NoLg=AUTO_ARIMA_WEEK1$WIS, 
  AUTO_ADJ_NoLg=AUTO_ADJACENT_WEEK1$WIS, 
  AUTO_EPI_NoLg=AUTO_EPIWEEK_WEEK1$WIS,
  AUTO_TEMP_NoLg=AUTO_TEMPERATURE_WEEK1$WIS,
  AUTO_AVG_NoLg=AUTO_AVERAGE_WEEK1$WIS,
  ES27_AR_NoLg=ES27_ARIMA_WEEK1$WIS, 
  ES27_ADJ_NoLg=ES27_ADJACENT_WEEK1$WIS, 
  ES27_EPI_NoLg=ES27_EPIWEEK_WEEK1$WIS,
  ES27_TEMP_NoLg=ES27_TEMPERATURE_WEEK1$WIS,
  ES27_AVG_NoLg=ES27_AVERAGE_WEEK1$WIS,
  ES64_AR_NoLg=ES64_ARIMA_WEEK1$WIS, 
  ES64_ADJ_NoLg=ES64_ADJACENT_WEEK1$WIS, 
  ES64_EPI_NoLg=ES64_EPIWEEK_WEEK1$WIS,
  ES64_TEMP_NoLg=ES64_TEMPERATURE_WEEK1$WIS,
  ES64_AVG_NoLg=ES64_AVERAGE_WEEK1$WIS
)

head(df_W1_NoLg)
```

------------
2 Weeks ahead
-------------
```{r}
# Creating new columns with Epidemiological weeks based on target_end_week
AUTO_ARIMA_WEEK2$epiweek <- MMWRweek(AUTO_ARIMA_WEEK2$target_end_date)$MMWRweek
AUTO_ADJACENT_WEEK2$epiweek <- MMWRweek(AUTO_ADJACENT_WEEK2$target_end_date)$MMWRweek
AUTO_EPIWEEK_WEEK2$epiweek <- MMWRweek(AUTO_EPIWEEK_WEEK2$target_end_date)$MMWRweek
AUTO_TEMPERATURE_WEEK2$epiweek <- MMWRweek(AUTO_TEMPERATURE_WEEK2$target_end_date)$MMWRweek
AUTO_AVERAGE_WEEK2$epiweek <- MMWRweek(AUTO_AVERAGE_WEEK2$target_end_date)$MMWRweek

ES27_ARIMA_WEEK2$epiweek <- MMWRweek(ES27_ARIMA_WEEK2$target_end_date)$MMWRweek
ES27_ADJACENT_WEEK2$epiweek <- MMWRweek(ES27_ADJACENT_WEEK2$target_end_date)$MMWRweek
ES27_EPIWEEK_WEEK2$epiweek <- MMWRweek(ES27_EPIWEEK_WEEK2$target_end_date)$MMWRweek
ES27_TEMPERATURE_WEEK2$epiweek <- MMWRweek(ES27_TEMPERATURE_WEEK2$target_end_date)$MMWRweek
ES27_AVERAGE_WEEK2$epiweek <- MMWRweek(ES27_AVERAGE_WEEK2$target_end_date)$MMWRweek

ES64_ARIMA_WEEK2$epiweek <- MMWRweek(ES64_ARIMA_WEEK2$target_end_date)$MMWRweek
ES64_ADJACENT_WEEK2$epiweek <- MMWRweek(ES64_ADJACENT_WEEK2$target_end_date)$MMWRweek
ES64_EPIWEEK_WEEK2$epiweek <- MMWRweek(ES64_EPIWEEK_WEEK2$target_end_date)$MMWRweek
ES64_TEMPERATURE_WEEK2$epiweek <- MMWRweek(ES64_TEMPERATURE_WEEK2$target_end_date)$MMWRweek
ES64_AVERAGE_WEEK2$epiweek <- MMWRweek(ES64_AVERAGE_WEEK2$target_end_date)$MMWRweek

# Dataframe that will be analysed for 2 Weeks Ahead
df_W2_NoLg <- data.frame(
  STATE = AUTO_ARIMA_WEEK2$State,
  Julian_date = AUTO_ARIMA_WEEK2$target_end_date,
  epiweek = AUTO_ARIMA_WEEK2$epiweek,
  AUTO_AR_NoLg=AUTO_ARIMA_WEEK2$WIS, 
  AUTO_ADJ_NoLg=AUTO_ADJACENT_WEEK2$WIS, 
  AUTO_EPI_NoLg=AUTO_EPIWEEK_WEEK2$WIS,
  AUTO_TEMP_NoLg=AUTO_TEMPERATURE_WEEK2$WIS,
  AUTO_AVG_NoLg=AUTO_AVERAGE_WEEK2$WIS,

  ES27_AR_NoLg=ES27_ARIMA_WEEK2$WIS, 
  ES27_ADJ_NoLg=ES27_ADJACENT_WEEK2$WIS, 
  ES27_EPI_NoLg=ES27_EPIWEEK_WEEK2$WIS,
  ES27_TEMP_NoLg=ES27_TEMPERATURE_WEEK2$WIS,
  ES27_AVG_NoLg=ES27_AVERAGE_WEEK2$WIS,

  ES64_AR_NoLg=ES64_ARIMA_WEEK2$WIS, 
  ES64_ADJ_NoLg=ES64_ADJACENT_WEEK2$WIS, 
  ES64_EPI_NoLg=ES64_EPIWEEK_WEEK2$WIS,
  ES64_TEMP_NoLg=ES64_TEMPERATURE_WEEK2$WIS,
  ES64_AVG_NoLg=ES64_AVERAGE_WEEK2$WIS

)

head(df_W2_NoLg)
```

-------------
3 Weeks ahead
-------------
```{r}
# Creating new columns with Epidemiological weeks based on target_end_week
AUTO_ARIMA_WEEK3$epiweek <- MMWRweek(AUTO_ARIMA_WEEK3$target_end_date)$MMWRweek
AUTO_ADJACENT_WEEK3$epiweek <- MMWRweek(AUTO_ADJACENT_WEEK3$target_end_date)$MMWRweek
AUTO_EPIWEEK_WEEK3$epiweek <- MMWRweek(AUTO_EPIWEEK_WEEK3$target_end_date)$MMWRweek
AUTO_TEMPERATURE_WEEK3$epiweek <- MMWRweek(AUTO_TEMPERATURE_WEEK3$target_end_date)$MMWRweek
AUTO_AVERAGE_WEEK3$epiweek <- MMWRweek(AUTO_AVERAGE_WEEK3$target_end_date)$MMWRweek

ES27_ARIMA_WEEK3$epiweek <- MMWRweek(ES27_ARIMA_WEEK3$target_end_date)$MMWRweek
ES27_ADJACENT_WEEK3$epiweek <- MMWRweek(ES27_ADJACENT_WEEK3$target_end_date)$MMWRweek
ES27_EPIWEEK_WEEK3$epiweek <- MMWRweek(ES27_EPIWEEK_WEEK3$target_end_date)$MMWRweek
ES27_TEMPERATURE_WEEK3$epiweek <- MMWRweek(ES27_TEMPERATURE_WEEK3$target_end_date)$MMWRweek
ES27_AVERAGE_WEEK3$epiweek <- MMWRweek(ES27_AVERAGE_WEEK3$target_end_date)$MMWRweek

ES64_ARIMA_WEEK3$epiweek <- MMWRweek(ES64_ARIMA_WEEK3$target_end_date)$MMWRweek
ES64_ADJACENT_WEEK3$epiweek <- MMWRweek(ES64_ADJACENT_WEEK3$target_end_date)$MMWRweek
ES64_EPIWEEK_WEEK3$epiweek <- MMWRweek(ES64_EPIWEEK_WEEK3$target_end_date)$MMWRweek
ES64_TEMPERATURE_WEEK3$epiweek <- MMWRweek(ES64_TEMPERATURE_WEEK3$target_end_date)$MMWRweek
ES64_AVERAGE_WEEK3$epiweek <- MMWRweek(ES64_AVERAGE_WEEK3$target_end_date)$MMWRweek

# Dataframe that will be analysed for 3 Weeks Ahead
df_W3_NoLg <- data.frame(
  STATE = AUTO_ARIMA_WEEK3$State,
  Julian_date = AUTO_ARIMA_WEEK3$target_end_date,
  epiweek = AUTO_ARIMA_WEEK3$epiweek,
  AUTO_AR_NoLg=AUTO_ARIMA_WEEK3$WIS, 
  AUTO_ADJ_NoLg=AUTO_ADJACENT_WEEK3$WIS, 
  AUTO_EPI_NoLg=AUTO_EPIWEEK_WEEK3$WIS,
  AUTO_TEMP_NoLg=AUTO_TEMPERATURE_WEEK3$WIS,
  AUTO_AVG_NoLg=AUTO_AVERAGE_WEEK3$WIS,

  ES27_AR_NoLg=ES27_ARIMA_WEEK3$WIS, 
  ES27_ADJ_NoLg=ES27_ADJACENT_WEEK3$WIS, 
  ES27_EPI_NoLg=ES27_EPIWEEK_WEEK3$WIS,
  ES27_TEMP_NoLg=ES27_TEMPERATURE_WEEK3$WIS,
  ES27_AVG_NoLg=ES27_AVERAGE_WEEK3$WIS,

  ES64_AR_NoLg=ES64_ARIMA_WEEK3$WIS, 
  ES64_ADJ_NoLg=ES64_ADJACENT_WEEK3$WIS, 
  ES64_EPI_NoLg=ES64_EPIWEEK_WEEK3$WIS,
  ES64_TEMP_NoLg=ES64_TEMPERATURE_WEEK3$WIS,
  ES64_AVG_NoLg=ES64_AVERAGE_WEEK3$WIS
)

head(df_W3_NoLg)
```

-------------
4 Weeks ahead
-------------
```{r}
# Creating new columns with Epidemiological weeks based on target_end_week
AUTO_ARIMA_WEEK4$epiweek <- MMWRweek(AUTO_ARIMA_WEEK4$target_end_date)$MMWRweek
AUTO_ADJACENT_WEEK4$epiweek <- MMWRweek(AUTO_ADJACENT_WEEK4$target_end_date)$MMWRweek
AUTO_EPIWEEK_WEEK4$epiweek <- MMWRweek(AUTO_EPIWEEK_WEEK4$target_end_date)$MMWRweek
AUTO_TEMPERATURE_WEEK4$epiweek <- MMWRweek(AUTO_TEMPERATURE_WEEK4$target_end_date)$MMWRweek
AUTO_AVERAGE_WEEK4$epiweek <- MMWRweek(AUTO_AVERAGE_WEEK4$target_end_date)$MMWRweek

ES27_ARIMA_WEEK4$epiweek <- MMWRweek(ES27_ARIMA_WEEK4$target_end_date)$MMWRweek
ES27_ADJACENT_WEEK4$epiweek <- MMWRweek(ES27_ADJACENT_WEEK4$target_end_date)$MMWRweek
ES27_EPIWEEK_WEEK4$epiweek <- MMWRweek(ES27_EPIWEEK_WEEK4$target_end_date)$MMWRweek
ES27_TEMPERATURE_WEEK4$epiweek <- MMWRweek(ES27_TEMPERATURE_WEEK4$target_end_date)$MMWRweek
ES27_AVERAGE_WEEK4$epiweek <- MMWRweek(ES27_AVERAGE_WEEK4$target_end_date)$MMWRweek

ES64_ARIMA_WEEK4$epiweek <- MMWRweek(ES64_ARIMA_WEEK4$target_end_date)$MMWRweek
ES64_ADJACENT_WEEK4$epiweek <- MMWRweek(ES64_ADJACENT_WEEK4$target_end_date)$MMWRweek
ES64_EPIWEEK_WEEK4$epiweek <- MMWRweek(ES64_EPIWEEK_WEEK4$target_end_date)$MMWRweek
ES64_TEMPERATURE_WEEK4$epiweek <- MMWRweek(ES64_TEMPERATURE_WEEK4$target_end_date)$MMWRweek
ES64_AVERAGE_WEEK4$epiweek <- MMWRweek(ES64_AVERAGE_WEEK4$target_end_date)$MMWRweek

# Dataframe that will be analysed for 4 Weeks Ahead
df_W4_NoLg <- data.frame(
  STATE = AUTO_ARIMA_WEEK4$State,
  Julian_date = AUTO_ARIMA_WEEK4$target_end_date,
  epiweek = AUTO_ARIMA_WEEK4$epiweek,
  AUTO_AR_NoLg=AUTO_ARIMA_WEEK4$WIS, 
  AUTO_ADJ_NoLg=AUTO_ADJACENT_WEEK4$WIS, 
  AUTO_EPI_NoLg=AUTO_EPIWEEK_WEEK4$WIS,
  AUTO_TEMP_NoLg=AUTO_TEMPERATURE_WEEK4$WIS,
  AUTO_AVG_NoLg=AUTO_AVERAGE_WEEK4$WIS,

  ES27_AR_NoLg=ES27_ARIMA_WEEK4$WIS, 
  ES27_ADJ_NoLg=ES27_ADJACENT_WEEK4$WIS, 
  ES27_EPI_NoLg=ES27_EPIWEEK_WEEK4$WIS,
  ES27_TEMP_NoLg=ES27_TEMPERATURE_WEEK4$WIS,
  ES27_AVG_NoLg=ES27_AVERAGE_WEEK4$WIS,

  ES64_AR_NoLg=ES64_ARIMA_WEEK4$WIS, 
  ES64_ADJ_NoLg=ES64_ADJACENT_WEEK4$WIS, 
  ES64_EPI_NoLg=ES64_EPIWEEK_WEEK4$WIS,
  ES64_TEMP_NoLg=ES64_TEMPERATURE_WEEK4$WIS,
  ES64_AVG_NoLg=ES64_AVERAGE_WEEK4$WIS  
)

head(df_W4_NoLg)
```

Filter only the flu season

```{r}
# Filter the dataframe for epiweek >= 40 or epiweek <= 20
filtered_df_W1_NoLg <- df_W1_NoLg %>%
  filter(epiweek >= 40 | epiweek <= 20)

# Display the filtered dataset
head(filtered_df_W1_NoLg)

# Filter the dataframe for epiweek >= 40 or epiweek <= 20
filtered_df_W2_NoLg <- df_W2_NoLg %>%
  filter(epiweek >= 40 | epiweek <= 20)

# Display the filtered dataset
head(filtered_df_W2_NoLg)

# Filter the dataframe for epiweek >= 40 or epiweek <= 20
filtered_df_W3_NoLg <- df_W3_NoLg %>%
  filter(epiweek >= 40 | epiweek <= 20)

# Display the filtered dataset
head(filtered_df_W3_NoLg)

# Filter the dataframe for epiweek >= 40 or epiweek <= 20
filtered_df_W4_NoLg <- df_W4_NoLg %>%
  filter(epiweek >= 40 | epiweek <= 20)

# Display the filtered dataset
head(filtered_df_W4_NoLg)
```

```{r}
library(dplyr)
library(ggplot2)

# Combine all datasets and add a column to identify them
df_combined_ <- bind_rows(
  df_W1_NoLg %>% mutate(Dataset = "W1"),
  df_W2_NoLg %>% mutate(Dataset = "W2"),
  df_W3_NoLg %>% mutate(Dataset = "W3"),
  df_W4_NoLg %>% mutate(Dataset = "W4")
) %>%
  select(STATE, epiweek, AUTO_AR_NoLg, Dataset, Julian_date) %>%
  rename(WIS = AUTO_AR_NoLg)  # Rename for consistency

# Define the epiweek values to highlight
highlight_weeks <- c(21, 41)

# Get corresponding Julian dates for these epiweeks
highlight_dates <- df_combined_ %>%
  filter(epiweek %in% highlight_weeks) %>%
  select(epiweek, Julian_date) %>%
  distinct()

# Plot
plt0 <- ggplot(df_combined_, aes(x = Julian_date, y = WIS, group = STATE)) + 
  geom_vline(data = highlight_dates, aes(xintercept = Julian_date), color = "red", linetype = "dashed", size = 1) +  
  geom_line(color = "black", size = 0.2) +  
  facet_wrap(~ Dataset, ncol = 2, scales = "free_y") +  
  theme_minimal() +  
  labs(title = "Baseline Model WIS for 48 states on the contiguous U.S.",#,
#       subtitle = "Dashed red lines indicates epidemiological weeks outsise of the influenza season",
       x = "Month", 
       y = "WIS") +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position = "none") + 
  scale_x_date(labels = scales::date_format("%b"), breaks = "2 month")  # Format x-axis to show months and set monthly breaks

# Print the plot
plt0

ggsave(plt0, height = 5, width = 6, filename="fig0.jpg")  
```

Calculate mean weighted interval score for all forecast weeks on each model.

```{r}

# Define the function
calculate_mean_wis <- function(data) {
  data %>%
    group_by(STATE) %>%
    summarize(
      AUTO_AR_NoLg = mean(AUTO_AR_NoLg, na.rm = TRUE),
      AUTO_ADJ_NoLg = mean(AUTO_ADJ_NoLg, na.rm = TRUE),
      AUTO_EPI_NoLg = mean(AUTO_EPI_NoLg, na.rm = TRUE),
      AUTO_TMP_NoLg = mean(AUTO_TEMP_NoLg, na.rm = TRUE),
      AUTO_AVG_NoLg = mean(AUTO_AVG_NoLg, na.rm = TRUE),
      ES27_AR_NoLg = mean(ES27_AR_NoLg, na.rm = TRUE),
      ES27_ADJ_NoLg = mean(ES27_ADJ_NoLg, na.rm = TRUE),
      ES27_EPI_NoLg = mean(ES27_EPI_NoLg, na.rm = TRUE),
      ES27_TMP_NoLg = mean(ES27_TEMP_NoLg, na.rm = TRUE),
      ES27_AVG_NoLg = mean(ES27_AVG_NoLg, na.rm = TRUE),
      ES64_AR_NoLg = mean(ES64_AR_NoLg, na.rm = TRUE),
      ES64_ADJ_NoLg = mean(ES64_ADJ_NoLg, na.rm = TRUE),
      ES64_EPI_NoLg = mean(ES64_EPI_NoLg, na.rm = TRUE),
      ES64_TMP_NoLg = mean(ES64_TEMP_NoLg, na.rm = TRUE),
      ES64_AVG_NoLg = mean(ES64_AVG_NoLg, na.rm = TRUE)

    )
}

# Now you can use the function with any dataframe
W1_NoLg <- calculate_mean_wis(filtered_df_W1_NoLg)
W2_NoLg <- calculate_mean_wis(filtered_df_W2_NoLg)
W3_NoLg <- calculate_mean_wis(filtered_df_W3_NoLg)
W4_NoLg <- calculate_mean_wis(filtered_df_W4_NoLg)

# Display the resulting dataframe
head(W1_NoLg)
head(W2_NoLg)
head(W3_NoLg)
head(W4_NoLg)

```

Here we have boxplots of the mean(WIS) by each state.

```{r}
# Combine the data into one data frame
combined_data <- data.frame(
  week = rep(c("W1", "W2", "W3", "W4"), each = 48 * 15),

  Models = rep(c("AUTO_AR", "ES27_AR", "ES64_AR", "AUTO_ADJ", "ES27_ADJ", "ES64_ADJ", "AUTO_TMP", "ES27_TMP", "ES64_TMP", "AUTO_EPI", "ES27_EPI", "ES64_EPI","AUTO_AVG", "ES27_AVG", "ES64_AVG"), each=48, times = 4),

  value = c(W1_NoLg$AUTO_AR_NoLg, W1_NoLg$ES27_AR_NoLg, W1_NoLg$ES64_AR_NoLg, 
            W1_NoLg$AUTO_ADJ_NoLg, W1_NoLg$ES27_ADJ_NoLg, W1_NoLg$ES64_ADJ_NoLg, 
            W1_NoLg$AUTO_TMP_NoLg, W1_NoLg$ES27_TMP_NoLg, W1_NoLg$ES64_TMP_NoLg, 
            W1_NoLg$AUTO_EPI_NoLg, W1_NoLg$ES27_EPI_NoLg, W1_NoLg$ES64_EPI_NoLg,
            W1_NoLg$AUTO_AVG_NoLg, W1_NoLg$ES27_AVG_NoLg, W1_NoLg$ES64_AVG_NoLg,

            W2_NoLg$AUTO_AR_NoLg, W2_NoLg$ES27_AR_NoLg, W2_NoLg$ES64_AR_NoLg, 
            W2_NoLg$AUTO_ADJ_NoLg, W2_NoLg$ES27_ADJ_NoLg, W2_NoLg$ES64_ADJ_NoLg, 
            W2_NoLg$AUTO_TMP_NoLg, W2_NoLg$ES27_TMP_NoLg, W2_NoLg$ES64_TMP_NoLg, 
            W2_NoLg$AUTO_EPI_NoLg, W2_NoLg$ES27_EPI_NoLg, W2_NoLg$ES64_EPI_NoLg,
            W2_NoLg$AUTO_AVG_NoLg, W2_NoLg$ES27_AVG_NoLg, W2_NoLg$ES64_AVG_NoLg,

            W3_NoLg$AUTO_AR_NoLg, W3_NoLg$ES27_AR_NoLg, W3_NoLg$ES64_AR_NoLg,
            W3_NoLg$AUTO_ADJ_NoLg, W3_NoLg$ES27_ADJ_NoLg, W3_NoLg$ES64_ADJ_NoLg, 
            W3_NoLg$AUTO_TMP_NoLg, W3_NoLg$ES27_TMP_NoLg, W3_NoLg$ES64_TMP_NoLg, 
            W3_NoLg$AUTO_EPI_NoLg, W3_NoLg$ES27_EPI_NoLg, W3_NoLg$ES64_EPI_NoLg,
            W3_NoLg$AUTO_AVG_NoLg, W3_NoLg$ES27_AVG_NoLg, W3_NoLg$ES64_AVG_NoLg,

            W4_NoLg$AUTO_AR_NoLg, W4_NoLg$ES27_AR_NoLg, W4_NoLg$ES64_AR_NoLg,
            W4_NoLg$AUTO_ADJ_NoLg, W4_NoLg$ES27_ADJ_NoLg, W4_NoLg$ES64_ADJ_NoLg, 
            W4_NoLg$AUTO_TMP_NoLg, W4_NoLg$ES27_TMP_NoLg, W4_NoLg$ES64_TMP_NoLg, 
            W4_NoLg$AUTO_EPI_NoLg, W4_NoLg$ES27_EPI_NoLg, W4_NoLg$ES64_EPI_NoLg,
            W4_NoLg$AUTO_AVG_NoLg, W4_NoLg$ES27_AVG_NoLg, W4_NoLg$ES64_AVG_NoLg
            )
)

# Create a new column for categories
combined_data <- combined_data %>%
  mutate(category = case_when(
    grepl("AR", Models) ~ "AR",
    grepl("EPI", Models) ~ "EPI",
    grepl("TMP", Models) ~ "TMP",
    grepl("ADJ", Models) ~ "ADJ",
    grepl("AVG", Models) ~ "AVG",
  ))

# Define color ramps for each category
colors_ar <- c("#a6cee3","#226e83", "#08306b")
colors_adj <- c("#fb9a99","red3", "#a11c3e")
colors_epi <- c("#cab2d6", "purple2", "#5e2b7b")
colors_tmp <- c("lightgreen","green3","#319045")

# Create a custom color mapping for each variable
custom_colors <- c(
  "AUTO_AR" = colors_ar[1], "ES27_AR" = colors_ar[2], "ES64_AR" = colors_ar[3],
  "AUTO_EPI" = colors_epi[1], "ES27_EPI" = colors_epi[2], "ES64_EPI" = colors_epi[3],
  "AUTO_TMP" = colors_tmp[1], "ES27_TMP" = colors_tmp[2], "ES64_TMP" = colors_tmp[3],
  "AUTO_ADJ" = colors_adj[1], "ES27_ADJ" = colors_adj[2], "ES64_ADJ" = colors_adj[3],
  "AUTO_AVG" = colors_adj[1], "ES27_AVG" = colors_adj[2], "ES64_AVG" = colors_adj[3]
)

  # Create the box plot with additional facet for categories
plt1<-  ggplot(combined_data, aes(x = week, y = value, fill = Models)) +
    geom_boxplot(color = "black") +
    scale_fill_manual(values = custom_colors) +
    facet_wrap(category ~ Models , nrow = 2) +
    labs(
      title = "A) Models without log-back transformations",
      x = "",
      y = "Mean Weighted Interval Score (WIS)"
    ) +
  scale_y_continuous(limits = c(0, 650)) +  # Set y-axis range
    theme_minimal()
plt1
ggsave(plt1, height = 8, width = 10, filename="fig1A.jpg")  
```

Loading the results of each model and the shapefiles of the maps.

```{r}
load("NHSH_hospitalizations_seed/ES_ARIMA/ARIMA_MODELS_influenza_hospitalization.Rdata")
load("NHSH_hospitalizations_seed/ES_ADJACENT/ADJACENT_MODELS_influenza_hospitalization.Rdata")
load("NHSH_hospitalizations_seed/ES_EPIWEEK/EPIWEEK_MODELS_influenza_hospitalization.Rdata")
load("NHSH_hospitalizations_seed/ES_TEMPERATURE/TEMPERATURE_MODELS_influenza_hospitalization.Rdata")
load("NHSH_hospitalizations_seed/ES_AVERAGE/AVERAGE_MODELS_influenza_hospitalization.Rdata")

states <- read_sf("NHSH_hospitalizations_seed/shapefiles/cb_2018_us_state_500k.shp")

states <- states %>%
  rename(STATE = NAME)

```

------------
1 Week ahead
------------
```{r}
# Creating new columns with Epidemiological weeks based on target_end_week
AUTO_ARIMA_WEEK1$epiweek <- MMWRweek(AUTO_ARIMA_WEEK1$target_end_date)$MMWRweek
AUTO_ADJACENT_WEEK1$epiweek <- MMWRweek(AUTO_ADJACENT_WEEK1$target_end_date)$MMWRweek
AUTO_EPIWEEK_WEEK1$epiweek <- MMWRweek(AUTO_EPIWEEK_WEEK1$target_end_date)$MMWRweek
AUTO_TEMPERATURE_WEEK1$epiweek <- MMWRweek(AUTO_TEMPERATURE_WEEK1$target_end_date)$MMWRweek
AUTO_AVERAGE_WEEK1$epiweek <- MMWRweek(AUTO_AVERAGE_WEEK1$target_end_date)$MMWRweek

ES27_ARIMA_WEEK1$epiweek <- MMWRweek(ES27_ARIMA_WEEK1$target_end_date)$MMWRweek
ES27_ADJACENT_WEEK1$epiweek <- MMWRweek(ES27_ADJACENT_WEEK1$target_end_date)$MMWRweek
ES27_EPIWEEK_WEEK1$epiweek <- MMWRweek(ES27_EPIWEEK_WEEK1$target_end_date)$MMWRweek
ES27_TEMPERATURE_WEEK1$epiweek <- MMWRweek(ES27_TEMPERATURE_WEEK1$target_end_date)$MMWRweek
ES27_AVERAGE_WEEK1$epiweek <- MMWRweek(ES27_AVERAGE_WEEK1$target_end_date)$MMWRweek

ES64_ARIMA_WEEK1$epiweek <- MMWRweek(ES64_ARIMA_WEEK1$target_end_date)$MMWRweek
ES64_ADJACENT_WEEK1$epiweek <- MMWRweek(ES64_ADJACENT_WEEK1$target_end_date)$MMWRweek
ES64_EPIWEEK_WEEK1$epiweek <- MMWRweek(ES64_EPIWEEK_WEEK1$target_end_date)$MMWRweek
ES64_TEMPERATURE_WEEK1$epiweek <- MMWRweek(ES64_TEMPERATURE_WEEK1$target_end_date)$MMWRweek
ES64_AVERAGE_WEEK1$epiweek <- MMWRweek(ES64_AVERAGE_WEEK1$target_end_date)$MMWRweek

# Dataframe that will be analysed for 1 Week Ahead
df_W1 <- data.frame(
  STATE = AUTO_ARIMA_WEEK1$State,
  Julian_date = AUTO_ARIMA_WEEK1$target_end_date,
  epiweek = AUTO_ARIMA_WEEK1$epiweek,
  AUTO_AR=AUTO_ARIMA_WEEK1$WIS, 
  AUTO_ADJ=AUTO_ADJACENT_WEEK1$WIS, 
  AUTO_EPI=AUTO_EPIWEEK_WEEK1$WIS,
  AUTO_TEMP=AUTO_TEMPERATURE_WEEK1$WIS,
  AUTO_AVG=AUTO_AVERAGE_WEEK1$WIS,

  ES27_AR=ES27_ARIMA_WEEK1$WIS, 
  ES27_ADJ=ES27_ADJACENT_WEEK1$WIS, 
  ES27_EPI=ES27_EPIWEEK_WEEK1$WIS,
  ES27_TEMP=ES27_TEMPERATURE_WEEK1$WIS,
  ES27_AVG=ES27_AVERAGE_WEEK1$WIS,

  ES64_AR=ES64_ARIMA_WEEK1$WIS, 
  ES64_ADJ=ES64_ADJACENT_WEEK1$WIS, 
  ES64_EPI=ES64_EPIWEEK_WEEK1$WIS,
  ES64_TEMP=ES64_TEMPERATURE_WEEK1$WIS,
  ES64_AVG=ES64_AVERAGE_WEEK1$WIS  
)

head(df_W1)
```

------------
2 Weeks ahead
-------------
```{r}
# Creating new columns with Epidemiological weeks based on target_end_week
AUTO_ARIMA_WEEK2$epiweek <- MMWRweek(AUTO_ARIMA_WEEK2$target_end_date)$MMWRweek
AUTO_ADJACENT_WEEK2$epiweek <- MMWRweek(AUTO_ADJACENT_WEEK2$target_end_date)$MMWRweek
AUTO_EPIWEEK_WEEK2$epiweek <- MMWRweek(AUTO_EPIWEEK_WEEK2$target_end_date)$MMWRweek
AUTO_TEMPERATURE_WEEK2$epiweek <- MMWRweek(AUTO_TEMPERATURE_WEEK2$target_end_date)$MMWRweek
AUTO_AVERAGE_WEEK2$epiweek <- MMWRweek(AUTO_AVERAGE_WEEK2$target_end_date)$MMWRweek

ES27_ARIMA_WEEK2$epiweek <- MMWRweek(ES27_ARIMA_WEEK2$target_end_date)$MMWRweek
ES27_ADJACENT_WEEK2$epiweek <- MMWRweek(ES27_ADJACENT_WEEK2$target_end_date)$MMWRweek
ES27_EPIWEEK_WEEK2$epiweek <- MMWRweek(ES27_EPIWEEK_WEEK2$target_end_date)$MMWRweek
ES27_TEMPERATURE_WEEK2$epiweek <- MMWRweek(ES27_TEMPERATURE_WEEK2$target_end_date)$MMWRweek
ES27_AVERAGE_WEEK2$epiweek <- MMWRweek(ES27_AVERAGE_WEEK2$target_end_date)$MMWRweek

ES64_ARIMA_WEEK2$epiweek <- MMWRweek(ES64_ARIMA_WEEK2$target_end_date)$MMWRweek
ES64_ADJACENT_WEEK2$epiweek <- MMWRweek(ES64_ADJACENT_WEEK2$target_end_date)$MMWRweek
ES64_EPIWEEK_WEEK2$epiweek <- MMWRweek(ES64_EPIWEEK_WEEK2$target_end_date)$MMWRweek
ES64_TEMPERATURE_WEEK2$epiweek <- MMWRweek(ES64_TEMPERATURE_WEEK2$target_end_date)$MMWRweek
ES64_AVERAGE_WEEK2$epiweek <- MMWRweek(ES64_AVERAGE_WEEK2$target_end_date)$MMWRweek

# Dataframe that will be analysed for 2 Week Ahead
df_W2 <- data.frame(
  STATE = AUTO_ARIMA_WEEK2$State,
  Julian_date = AUTO_ARIMA_WEEK2$target_end_date,
  epiweek = AUTO_ARIMA_WEEK2$epiweek,
  AUTO_AR=AUTO_ARIMA_WEEK2$WIS, 
  AUTO_ADJ=AUTO_ADJACENT_WEEK2$WIS, 
  AUTO_EPI=AUTO_EPIWEEK_WEEK2$WIS,
  AUTO_TEMP=AUTO_TEMPERATURE_WEEK2$WIS,
  AUTO_AVG=AUTO_AVERAGE_WEEK2$WIS,

  ES27_AR=ES27_ARIMA_WEEK2$WIS, 
  ES27_ADJ=ES27_ADJACENT_WEEK2$WIS, 
  ES27_EPI=ES27_EPIWEEK_WEEK2$WIS,
  ES27_TEMP=ES27_TEMPERATURE_WEEK2$WIS,
  ES27_AVG=ES27_AVERAGE_WEEK2$WIS,

  ES64_AR=ES64_ARIMA_WEEK2$WIS, 
  ES64_ADJ=ES64_ADJACENT_WEEK2$WIS, 
  ES64_EPI=ES64_EPIWEEK_WEEK2$WIS,
  ES64_TEMP=ES64_TEMPERATURE_WEEK2$WIS,
  ES64_AVG=ES64_AVERAGE_WEEK2$WIS  
)

head(df_W2)
```

-------------
3 Weeks ahead
-------------
```{r}
# Creating new columns with Epidemiological weeks based on target_end_week
AUTO_ARIMA_WEEK3$epiweek <- MMWRweek(AUTO_ARIMA_WEEK3$target_end_date)$MMWRweek
AUTO_ADJACENT_WEEK3$epiweek <- MMWRweek(AUTO_ADJACENT_WEEK3$target_end_date)$MMWRweek
AUTO_EPIWEEK_WEEK3$epiweek <- MMWRweek(AUTO_EPIWEEK_WEEK3$target_end_date)$MMWRweek
AUTO_TEMPERATURE_WEEK3$epiweek <- MMWRweek(AUTO_TEMPERATURE_WEEK3$target_end_date)$MMWRweek
AUTO_AVERAGE_WEEK3$epiweek <- MMWRweek(AUTO_AVERAGE_WEEK3$target_end_date)$MMWRweek

ES27_ARIMA_WEEK3$epiweek <- MMWRweek(ES27_ARIMA_WEEK3$target_end_date)$MMWRweek
ES27_ADJACENT_WEEK3$epiweek <- MMWRweek(ES27_ADJACENT_WEEK3$target_end_date)$MMWRweek
ES27_EPIWEEK_WEEK3$epiweek <- MMWRweek(ES27_EPIWEEK_WEEK3$target_end_date)$MMWRweek
ES27_TEMPERATURE_WEEK3$epiweek <- MMWRweek(ES27_TEMPERATURE_WEEK3$target_end_date)$MMWRweek
ES27_AVERAGE_WEEK3$epiweek <- MMWRweek(ES27_AVERAGE_WEEK3$target_end_date)$MMWRweek

ES64_ARIMA_WEEK3$epiweek <- MMWRweek(ES64_ARIMA_WEEK3$target_end_date)$MMWRweek
ES64_ADJACENT_WEEK3$epiweek <- MMWRweek(ES64_ADJACENT_WEEK3$target_end_date)$MMWRweek
ES64_EPIWEEK_WEEK3$epiweek <- MMWRweek(ES64_EPIWEEK_WEEK3$target_end_date)$MMWRweek
ES64_TEMPERATURE_WEEK3$epiweek <- MMWRweek(ES64_TEMPERATURE_WEEK3$target_end_date)$MMWRweek
ES64_AVERAGE_WEEK3$epiweek <- MMWRweek(ES64_AVERAGE_WEEK3$target_end_date)$MMWRweek

# Dataframe that will be analysed for 2 Week Ahead
df_W3 <- data.frame(
  STATE = AUTO_ARIMA_WEEK3$State,
  Julian_date = AUTO_ARIMA_WEEK3$target_end_date,
  epiweek = AUTO_ARIMA_WEEK3$epiweek,
  AUTO_AR=AUTO_ARIMA_WEEK3$WIS, 
  AUTO_ADJ=AUTO_ADJACENT_WEEK3$WIS, 
  AUTO_EPI=AUTO_EPIWEEK_WEEK3$WIS,
  AUTO_TEMP=AUTO_TEMPERATURE_WEEK3$WIS,
  AUTO_AVG=AUTO_AVERAGE_WEEK3$WIS,

  ES27_AR=ES27_ARIMA_WEEK3$WIS, 
  ES27_ADJ=ES27_ADJACENT_WEEK3$WIS, 
  ES27_EPI=ES27_EPIWEEK_WEEK3$WIS,
  ES27_TEMP=ES27_TEMPERATURE_WEEK3$WIS,
  ES27_AVG=ES27_AVERAGE_WEEK3$WIS,

  ES64_AR=ES64_ARIMA_WEEK3$WIS, 
  ES64_ADJ=ES64_ADJACENT_WEEK3$WIS, 
  ES64_EPI=ES64_EPIWEEK_WEEK3$WIS,
  ES64_TEMP=ES64_TEMPERATURE_WEEK3$WIS,
  ES64_AVG=ES64_AVERAGE_WEEK3$WIS  
)

head(df_W3)

```

-------------
4 Weeks ahead
-------------
```{r}
# Creating new columns with Epidemiological weeks based on target_end_week
AUTO_ARIMA_WEEK4$epiweek <- MMWRweek(AUTO_ARIMA_WEEK4$target_end_date)$MMWRweek
AUTO_ADJACENT_WEEK4$epiweek <- MMWRweek(AUTO_ADJACENT_WEEK4$target_end_date)$MMWRweek
AUTO_EPIWEEK_WEEK4$epiweek <- MMWRweek(AUTO_EPIWEEK_WEEK4$target_end_date)$MMWRweek
AUTO_TEMPERATURE_WEEK4$epiweek <- MMWRweek(AUTO_TEMPERATURE_WEEK4$target_end_date)$MMWRweek
AUTO_AVERAGE_WEEK4$epiweek <- MMWRweek(AUTO_AVERAGE_WEEK4$target_end_date)$MMWRweek

ES27_ARIMA_WEEK4$epiweek <- MMWRweek(ES27_ARIMA_WEEK4$target_end_date)$MMWRweek
ES27_ADJACENT_WEEK4$epiweek <- MMWRweek(ES27_ADJACENT_WEEK4$target_end_date)$MMWRweek
ES27_EPIWEEK_WEEK4$epiweek <- MMWRweek(ES27_EPIWEEK_WEEK4$target_end_date)$MMWRweek
ES27_TEMPERATURE_WEEK4$epiweek <- MMWRweek(ES27_TEMPERATURE_WEEK4$target_end_date)$MMWRweek
ES27_AVERAGE_WEEK4$epiweek <- MMWRweek(ES27_AVERAGE_WEEK4$target_end_date)$MMWRweek

ES64_ARIMA_WEEK4$epiweek <- MMWRweek(ES64_ARIMA_WEEK4$target_end_date)$MMWRweek
ES64_ADJACENT_WEEK4$epiweek <- MMWRweek(ES64_ADJACENT_WEEK4$target_end_date)$MMWRweek
ES64_EPIWEEK_WEEK4$epiweek <- MMWRweek(ES64_EPIWEEK_WEEK4$target_end_date)$MMWRweek
ES64_TEMPERATURE_WEEK4$epiweek <- MMWRweek(ES64_TEMPERATURE_WEEK4$target_end_date)$MMWRweek
ES64_AVERAGE_WEEK4$epiweek <- MMWRweek(ES64_AVERAGE_WEEK4$target_end_date)$MMWRweek

# Dataframe that will be analysed for 2 Week Ahead
df_W4 <- data.frame(
  STATE = AUTO_ARIMA_WEEK4$State,
  Julian_date = AUTO_ARIMA_WEEK4$target_end_date,
  epiweek = AUTO_ARIMA_WEEK4$epiweek,
  AUTO_AR=AUTO_ARIMA_WEEK4$WIS, 
  AUTO_ADJ=AUTO_ADJACENT_WEEK4$WIS, 
  AUTO_EPI=AUTO_EPIWEEK_WEEK4$WIS,
  AUTO_TEMP=AUTO_TEMPERATURE_WEEK4$WIS,
  AUTO_AVG=AUTO_AVERAGE_WEEK4$WIS,

  ES27_AR=ES27_ARIMA_WEEK4$WIS, 
  ES27_ADJ=ES27_ADJACENT_WEEK4$WIS, 
  ES27_EPI=ES27_EPIWEEK_WEEK4$WIS,
  ES27_TEMP=ES27_TEMPERATURE_WEEK4$WIS,
  ES27_AVG=ES27_AVERAGE_WEEK4$WIS,

  ES64_AR=ES64_ARIMA_WEEK4$WIS, 
  ES64_ADJ=ES64_ADJACENT_WEEK4$WIS, 
  ES64_EPI=ES64_EPIWEEK_WEEK4$WIS,
  ES64_TEMP=ES64_TEMPERATURE_WEEK4$WIS,
  ES64_AVG=ES64_AVERAGE_WEEK4$WIS  
)

head(df_W4)
```

Filter only the flu season

```{r}
# Filter the dataframe for epiweek >= 40 or epiweek <= 20
filtered_df_W1 <- df_W1 %>%
  filter(epiweek >= 40 | epiweek <= 20)

# Display the filtered dataset
head(filtered_df_W1)

# Filter the dataframe for epiweek >= 40 or epiweek <= 20
filtered_df_W2 <- df_W2 %>%
  filter(epiweek >= 40 | epiweek <= 20)

# Display the filtered dataset
head(filtered_df_W2)

# Filter the dataframe for epiweek >= 40 or epiweek <= 20
filtered_df_W3 <- df_W3 %>%
  filter(epiweek >= 40 | epiweek <= 20)

# Display the filtered dataset
head(filtered_df_W3)

# Filter the dataframe for epiweek >= 40 or epiweek <= 20
filtered_df_W4 <- df_W4 %>%
  filter(epiweek >= 40 | epiweek <= 20)

# Display the filtered dataset
head(filtered_df_W4)
```

Calculate mean weighted interval score for all forecast weeks on each model.

```{r}

# Define the function
calculate_mean_wis <- function(data) {
  data %>%
    group_by(STATE) %>%
    summarize(
      AUTO_AR_Lg = mean(AUTO_AR, na.rm = TRUE),
      AUTO_ADJ_Lg = mean(AUTO_ADJ, na.rm = TRUE),
      AUTO_EPI_Lg = mean(AUTO_EPI, na.rm = TRUE),
      AUTO_TMP_Lg = mean(AUTO_TEMP, na.rm = TRUE),
      ES27_AR_Lg = mean(ES27_AR, na.rm = TRUE),
      ES27_ADJ_Lg = mean(ES27_ADJ, na.rm = TRUE),
      ES27_EPI_Lg = mean(ES27_EPI, na.rm = TRUE),
      ES27_TMP_Lg = mean(ES27_TEMP, na.rm = TRUE),
      ES64_AR_Lg = mean(ES64_AR, na.rm = TRUE),
      ES64_ADJ_Lg = mean(ES64_ADJ, na.rm = TRUE),
      ES64_EPI_Lg = mean(ES64_EPI, na.rm = TRUE),
      ES64_TMP_Lg = mean(ES64_TEMP, na.rm = TRUE)
    )
}

# Now you can use the function with any dataframe
W1_Lg <- calculate_mean_wis(filtered_df_W1)
W2_Lg <- calculate_mean_wis(filtered_df_W2)
W3_Lg <- calculate_mean_wis(filtered_df_W3)
W4_Lg <- calculate_mean_wis(filtered_df_W4)

# Display the resulting dataframe
head(W1_Lg)
head(W2_Lg)
head(W3_Lg)
head(W4_Lg)
```

Here we have boxplots of the mean(WIS) by each state.

```{r}
# Combine the data into one data frame
combined_data <- data.frame(
  week = rep(c("W1", "W2", "W3", "W4"), each = 48 * 15),

  Models = rep(c("AUTO_AR_Lg", "ES27_AR_Lg", "ES64_AR_Lg", "AUTO_ADJ_Lg", "ES27_ADJ_Lg", "ES64_ADJ_Lg", "AUTO_TMP_Lg", "ES27_TMP_Lg", "ES64_TMP_Lg", "AUTO_EPI_Lg", "ES27_EPI_Lg", "ES64_EPI_Lg"), each=48, times = 4),

  value = c(W1_Lg$AUTO_AR_Lg, W1_Lg$ES27_AR_Lg, W1_Lg$ES64_AR_Lg, W1_Lg$AUTO_ADJ_Lg, W1_Lg$ES27_ADJ_Lg, W1_Lg$ES64_ADJ_Lg, W1_Lg$AUTO_TMP_Lg, W1_Lg$ES27_TMP_Lg, W1_Lg$ES64_TMP_Lg, W1_Lg$AUTO_EPI_Lg, W1_Lg$ES27_EPI_Lg, W1_Lg$ES64_EPI_Lg,

            W2_Lg$AUTO_AR_Lg, W2_Lg$ES27_AR_Lg, W2_Lg$ES64_AR_Lg, W2_Lg$AUTO_ADJ_Lg, W2_Lg$ES27_ADJ_Lg, W2_Lg$ES64_ADJ_Lg, W2_Lg$AUTO_TMP_Lg, W2_Lg$ES27_TMP_Lg, W2_Lg$ES64_TMP_Lg, W2_Lg$AUTO_EPI_Lg, W2_Lg$ES27_EPI_Lg, W2_Lg$ES64_EPI_Lg,

            W3_Lg$AUTO_AR_Lg, W3_Lg$ES27_AR_Lg, W3_Lg$ES64_AR_Lg, W3_Lg$AUTO_ADJ_Lg, W3_Lg$ES27_ADJ_Lg, W3_Lg$ES64_ADJ_Lg, W3_Lg$AUTO_TMP_Lg, W3_Lg$ES27_TMP_Lg, W3_Lg$ES64_TMP_Lg, W3_Lg$AUTO_EPI_Lg, W3_Lg$ES27_EPI_Lg, W3_Lg$ES64_EPI_Lg,

            W4_Lg$AUTO_AR_Lg, W4_Lg$ES27_AR_Lg, W4_Lg$ES64_AR_Lg, W4_Lg$AUTO_ADJ_Lg, W4_Lg$ES27_ADJ_Lg, W4_Lg$ES64_ADJ_Lg, W4_Lg$AUTO_TMP_Lg, W4_Lg$ES27_TMP_Lg, W4_Lg$ES64_TMP_Lg, W4_Lg$AUTO_EPI_Lg, W4_Lg$ES27_EPI_Lg, W4_Lg$ES64_EPI_Lg)
)

# Create a new column for categories
combined_data <- combined_data %>%
  mutate(category = case_when(
    grepl("AR", Models) ~ "AR",
    grepl("EPI", Models) ~ "EPI",
    grepl("TMP", Models) ~ "TMP",
    grepl("ADJ", Models) ~ "ADJ"
  ))

# Define color ramps for each category
colors_ar <- c("#a6cee3","#226e83", "#08306b")
colors_adj <- c("#fb9a99","red3", "#a11c3e")
colors_epi <- c("#cab2d6", "purple2", "#5e2b7b")
colors_tmp <- c("lightgreen","green3","#319045")

# Create a custom color mapping for each variable
custom_colors <- c(
  "AUTO_AR_Lg" = colors_ar[1], "ES27_AR_Lg" = colors_ar[2], "ES64_AR_Lg" = colors_ar[3],
  "AUTO_EPI_Lg" = colors_epi[1], "ES27_EPI_Lg" = colors_epi[2], "ES64_EPI_Lg" = colors_epi[3],
  "AUTO_TMP_Lg" = colors_tmp[1], "ES27_TMP_Lg" = colors_tmp[2], "ES64_TMP_Lg" = colors_tmp[3],
  "AUTO_ADJ_Lg" = colors_adj[1], "ES27_ADJ_Lg" = colors_adj[2], "ES64_ADJ_Lg" = colors_adj[3]
)

  # Create the box plot with additional facet for categories
plt2<-  ggplot(combined_data, aes(x = week, y = value, fill = Models)) +
    geom_boxplot(color = "black") +
    scale_fill_manual(values = custom_colors) +
    facet_wrap(category ~ Models , nrow = 2) +
    labs(
      title = "B)Models with log-back transformations",
      x = "",
      y = "Weighted Interval Score (WIS)"
    ) +
  scale_y_continuous(limits = c(0, 650)) +  # Set y-axis range
    theme_minimal()
plt2
ggsave(plt2, height = 8, width = 10, filename="Fig1B.jpg")  
```

COMBINING THE RESULTS WITH LOG AND NO LOG TRASNFORMATION

```{r}
W1<-merge(W1_Lg,W1_NoLg, by = "STATE")
W2<-merge(W2_Lg,W2_NoLg, by = "STATE")
W3<-merge(W3_Lg,W3_NoLg, by = "STATE")
W4<-merge(W4_Lg,W4_NoLg, by = "STATE")

```

Here we have boxplots of the mean(WIS) by each state.

Here I include a column with the best model result based on the lowest mean(WIS) of each state.

```{r}
# BEST RESULT

# Extract the columns of interest 
cols <- colnames(W1)[-1]
# Initialize a vector to store results
W1$Best_Result <- character(nrow(W1))

# Give me the model with lower WIS value in the best result column
for (i in 1:nrow(W1)) {
  # Find the column name with the minimum value for each row
  W1$Best_Result[i] <- cols[which.min(W1[i, cols])]
}

# REORDER BY FREQUENCY
W1$Best_Result <- fct_infreq(W1$Best_Result)
# Create a new column to classify the model type
W1$Model_Type <- ifelse(grepl("_NoLg$", W1$Best_Result), "Without log-back transformation", "With log-back transformation")

# Print the first rows
head(W1)

################################################

# Extract the columns of interest 
cols <- colnames(W2)[-1]
# Initialize a vector to store results
W2$Best_Result <- character(nrow(W2))
# Give me the model with lower WIS value in the best result column
for (i in 1:nrow(W2)) {
  # Find the column name with the minimum value for each row
  W2$Best_Result[i] <- cols[which.min(W2[i, cols])]
}

# REORDER BY FREQUENCY
W2$Best_Result <- fct_infreq(W2$Best_Result)
# Create a new column to classify the model type
W2$Model_Type <- ifelse(grepl("_NoLg$", W2$Best_Result), "Without log-back transformation", "With log-back transformation")

# Print merged results
head(W2)

######################################
# BEST RESULT
# Extract the columns of interest 
cols <- colnames(W3)[-1]
# Initialize a vector to store results
W3$Best_Result <- character(nrow(W3))

# Give me the model with lower WIS value in the best result column
for (i in 1:nrow(W3)) {
  # Find the column name with the minimum value for each row
  W3$Best_Result[i] <- cols[which.min(W3[i, cols])]
}

# REORDER BY FREQUENCY
W3$Best_Result <- fct_infreq(W3$Best_Result)
# Create a new column to classify the model type
W3$Model_Type <- ifelse(grepl("_NoLg$", W3$Best_Result), "Without log-back transformation", "With log-back transformation")

# Print merged results
head(W3)

###################################
# BEST RESULT
# Extract the columns of interest
cols <- colnames(W4)[-1]
# Initialize a vector to store results
W4$Best_Result <- character(nrow(W4))
# Give me the model with lower WIS value in the best result column
for (i in 1:nrow(W4)) {
  # Find the column name with the minimum value for each row
  W4$Best_Result[i] <- cols[which.min(W4[i, cols])]
}
W4$Best_Result <- fct_infreq(W4$Best_Result)
# Create a new column to classify the model type
W4$Model_Type <- ifelse(grepl("_NoLg$", W4$Best_Result), "Without log-back transformation", "With log-back transformation")
# Print merged results
head(W4)

```

Now, let's plot the best models by each state for each forecast horizon (1-4 weeks ahead).

```{r}

# Create a custom color mapping for each variable
custom_colors <- c(
  "AUTO_AR_Lg" = colors_ar[1], "ES27_AR_Lg" = colors_ar[2], "ES64_AR_Lg" = colors_ar[3],
  "AUTO_EPI_Lg" = colors_epi[1], "ES27_EPI_Lg" = colors_epi[2], "ES64_EPI_Lg" = colors_epi[3],
  "AUTO_TMP_Lg" = colors_tmp[1], "ES27_TMP_Lg" = colors_tmp[2], "ES64_TMP_Lg" = colors_tmp[3],
  "AUTO_ADJ_Lg" = colors_adj[1], "ES27_ADJ_Lg" = colors_adj[2], "ES64_ADJ_Lg" = colors_adj[3],

  "AUTO_AR_NoLg" = colors_ar[1], "ES27_AR_NoLg" = colors_ar[2], "ES64_AR_NoLg" = colors_ar[3],
  "AUTO_EPI_NoLg" = colors_epi[1], "ES27_EPI_NoLg" = colors_epi[2], "ES64_EPI_NoLg" = colors_epi[3],
  "AUTO_TMP_NoLg" = colors_tmp[1], "ES27_TMP_NoLg" = colors_tmp[2], "ES64_TMP_NoLg" = colors_tmp[3],
  "AUTO_ADJ_NoLg" = colors_adj[1], "ES27_ADJ_NoLg" = colors_adj[2], "ES64_ADJ_NoLg" = colors_adj[3]
)

# --------- WEEK1 MODELS ------------- #
plot1<- ggplot(W1, aes(x = Best_Result, fill = Best_Result)) +  
  geom_bar() +  
  scale_fill_manual(values = custom_colors) +  
  labs(title = "Best model on the 48 U.S. states (W1)",       
       x = "", y = "Number of states") +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  facet_wrap(~ Model_Type, scales = "free_x")

plot1

#ggsave("week1_best_models.jpg", plot1)

# --------- WEEK2 MODELS ------------- #
plot2<-ggplot(W2, aes(x = Best_Result, fill = Best_Result))  +  
  geom_bar() +  
  scale_fill_manual(values = custom_colors) +  
  labs(title = "Best model on the 48 U.S. states (W2)",       
       x = "", y = "Number of states") +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  facet_wrap(~ Model_Type, scales = "free_x")

#ggsave("week2_best_models.jpg", plot2)

plot2
# --------- WEEK3 MODELS ------------- #
plot3<-ggplot(W3, aes(x = Best_Result, fill = Best_Result))  +  
  geom_bar() +  
  scale_fill_manual(values = custom_colors) +  
  labs(title = "Best model on the 48 U.S. states (W3)",       
       x = "", y = "Number of states") +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  facet_wrap(~ Model_Type, scales = "free_x")
plot3

#ggsave("week3_best_models.jpg", plot3)

# --------- WEEK4 MODELS ------------- #
plot4<-ggplot(W4, aes(x = Best_Result, fill = Best_Result)) +  
  geom_bar() +  
  scale_fill_manual(values = custom_colors) +  
  labs(title = "Best model on the 48 U.S. states (W4)",       
       x = "", y = "Number of states") +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
  facet_wrap(~ Model_Type, scales = "free_x")
plot4
#ggsave("week4_best_models.jpg", plot4)

W1
```

Let's plot all results

```{r}
library(ggplot2)
library(dplyr)

# Add a column to indicate the week ahead
W1$Week_Ahead <- "W1"
W2$Week_Ahead <- "W2"
W3$Week_Ahead <- "W3"
W4$Week_Ahead <- "W4"

# Combine all weeks into one dataframe
all_weeks <- bind_rows(W1, W2, W3, W4)

# Create the plot
combined_plot <- ggplot(all_weeks, aes(x = Best_Result, fill = Best_Result)) +
  geom_bar() +
  scale_fill_manual(values = custom_colors) +
  labs(title = "Best model based on the lowest mean WIS on 48 U.S. states",
       x = "", y = "Number of states") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(Week_Ahead ~ Model_Type, scales = "free_x")

# Save and display the plot
ggsave("Fig2.jpg", combined_plot, width = 10, height = 8)
print(combined_plot)

```

 Wilcox test with Holm p-value ajustment

1 WEEK AHEAD

```{r}
library(lme4)
library(ez)

# Reshape the data into long format
long_data <- W1 %>%
  pivot_longer(cols = c(-STATE,-Best_Result,-Model_Type,-Week_Ahead) , names_to = "Type", values_to = "Value")#%>%
#  mutate(Type = recode(Type, "AUTO_AR" = "AUTO_AAR"))

wide_data <- long_data %>%
  pivot_wider(names_from = Type, values_from = Value)

# Get all model types except "AUTO_AAR"
model_types <- setdiff(names(wide_data), c("STATE", "AUTO_AR_NoLg","Best_Result","Model_Type","Week_Ahead"))

# Perform Wilcoxon test for each model type against AUTO_AAR
wilcox_results <- map_df(model_types, function(model) {
  test <- wilcox.test(wide_data[[model]], wide_data$AUTO_AR_NoLg, paired = TRUE, p.adjust.method ="holm")
  data.frame(Model = model, p_value = test$p.value, W = test$statistic)
})

# View results
p_values<-wilcox_results$p_value

p_values<-p.adjust(p_values, method = "holm")

p_values<-data.frame(p_values)

#########################################################################
p_df1 <- data.frame(Model = model_types, PValue = p_values, WeekAhead=1)
p_df1 <-drop_na(p_df1)
p_df1

```

2 WEEKS AHEAD Post Hoc Wilcox test

```{r}
# Reshape the data into long format
long_data <- W2 %>%
  pivot_longer(cols = c(-STATE,-Best_Result,-Model_Type,-Week_Ahead) , names_to = "Type", values_to = "Value")#%>%

################################
wide_data <- long_data %>%
  pivot_wider(names_from = Type, values_from = Value)

# Get all model types except "AUTO_AAR"
model_types <- setdiff(names(wide_data), c("STATE", "AUTO_AR_NoLg","Best_Result","Model_Type","Week_Ahead"))

# Perform Wilcoxon test for each model type against AUTO_AAR
wilcox_results <- map_df(model_types, function(model) {
  test <- wilcox.test(wide_data[[model]], wide_data$AUTO_AR_NoLg, paired = TRUE, p.adjust.method ="holm")
  data.frame(Model = model, p_value = test$p.value, W = test$statistic)
})

# View results
print(wilcox_results)

p_values<-wilcox_results$p_value 
p_values<-p.adjust(p_values, method = "holm")
p_values<-data.frame(p_values)

#########################################################################
p_df2 <- data.frame(Model = model_types, PValue = p_values, WeekAhead=2)
p_df2 <-drop_na(p_df2)
p_df2

```

3 WEEKS AHEAD Post Hoc Wilcox test

```{r}
# Reshape the data into long format
long_data <- W3 %>%
  pivot_longer(cols = c(-STATE,-Best_Result,-Model_Type,-Week_Ahead) , names_to = "Type", values_to = "Value")#%>%

######################
wide_data <- long_data %>%
  pivot_wider(names_from = Type, values_from = Value)

# Get all model types except "AUTO_AAR"
model_types <- setdiff(names(wide_data), c("STATE", "AUTO_AR_NoLg","Best_Result","Model_Type","Week_Ahead"))

# Perform Wilcoxon test for each model type against AUTO_AAR
wilcox_results <- map_df(model_types, function(model) {
  test <- wilcox.test(wide_data[[model]], wide_data$AUTO_AR_NoLg, paired = TRUE, p.adjust.method ="holm")
  data.frame(Model = model, p_value = test$p.value, W = test$statistic)
})

# View results
print(wilcox_results)

p_values<-wilcox_results$p_value 
p_values<-p.adjust(p_values, method = "holm")
p_values<-data.frame(p_values)

#########################################################################
p_df3 <- data.frame(Model = model_types, PValue = p_values, WeekAhead=3)
p_df3 <-drop_na(p_df3)
p_df3
```

4 WEEKS AHEAD Post Hoc Wilcox test

```{r}
# Reshape the data into long format
long_data <- W4 %>%
  pivot_longer(cols = c(-STATE,-Best_Result,-Model_Type,-Week_Ahead) , names_to = "Type", values_to = "Value")#%>%
#  mutate(Type = recode(Type, "AUTO_AR" = "AUTO_AR"))

######################
wide_data <- long_data %>%
  pivot_wider(names_from = Type, values_from = Value)

# Get all model types except "AUTO_AAR"
model_types <- setdiff(names(wide_data), c("STATE", "AUTO_AR_NoLg","Best_Result","Model_Type","Week_Ahead"))

# Perform Wilcoxon test for each model type against AUTO_AAR
wilcox_results <- map_df(model_types, function(model) {
  test <- wilcox.test(wide_data[[model]], wide_data$AUTO_AR_NoLg, paired = TRUE, p.adjust.method ="holm")
  data.frame(Model = model, p_value = test$p.value, W = test$statistic)
})

# View results
print(wilcox_results)

p_values<-wilcox_results$p_value

p_values<-p.adjust(p_values, method = "holm")

p_values<-data.frame(p_values)

#########################################################################
p_df4 <- data.frame(Model = model_types, PValue = p_values, WeekAhead=4)
p_df4 <-drop_na(p_df4)
p_df4

```

Let's calculate the mean(WIS) improvement relative to the AUTO ARIMA model. We will compare each model with the AUTO ARIMA model for the same states. Later we sum the results of these comparisons. Positive results indicate that there was a general improvement in the mean(WIS) for a given model type among all states. Note that larger values on specific states may override a general trend and make the results less informative. So I do not strongly advice this analysis.

```{r}
calculate_percentage_of_improvement <- function(data) {
  return(data.frame(
    AUTO_AR_NoLg = (data$AUTO_AR_NoLg - data$AUTO_AR_NoLg),
    ES27_AR_NoLg = ((data$AUTO_AR_NoLg - data$ES27_AR_NoLg) / data$AUTO_AR_NoLg) * 100,
    ES64_AR_NoLg = ((data$AUTO_AR_NoLg - data$ES64_AR_NoLg) / data$AUTO_AR_NoLg) * 100,
    AUTO_ADJ_NoLg = ((data$AUTO_AR_NoLg - data$AUTO_ADJ_NoLg) / data$AUTO_AR_NoLg) * 100,
    ES27_ADJ_NoLg = ((data$AUTO_AR_NoLg - data$ES27_ADJ_NoLg) / data$AUTO_AR_NoLg) * 100,
    ES64_ADJ_NoLg = ((data$AUTO_AR_NoLg - data$ES64_ADJ_NoLg) / data$AUTO_AR_NoLg) * 100,
    AUTO_TMP_NoLg = ((data$AUTO_AR_NoLg - data$AUTO_TMP_NoLg) / data$AUTO_AR_NoLg) * 100,
    ES27_TMP_NoLg = ((data$AUTO_AR_NoLg - data$ES27_TMP_NoLg) / data$AUTO_AR_NoLg) * 100,
    ES64_TMP_NoLg = ((data$AUTO_AR_NoLg - data$ES64_TMP_NoLg) / data$AUTO_AR_NoLg) * 100,
    AUTO_EPI_NoLg = ((data$AUTO_AR_NoLg - data$AUTO_EPI_NoLg) / data$AUTO_AR_NoLg) * 100,
    ES27_EPI_NoLg = ((data$AUTO_AR_NoLg - data$ES27_EPI_NoLg) / data$AUTO_AR_NoLg) * 100,
    ES64_EPI_NoLg = ((data$AUTO_AR_NoLg - data$ES64_EPI_NoLg) / data$AUTO_AR_NoLg) * 100,

    AUTO_AR_Lg = ((data$AUTO_AR_NoLg - data$AUTO_AR_Lg) / data$AUTO_AR_NoLg) * 100,
    ES27_AR_Lg = ((data$AUTO_AR_NoLg - data$ES27_AR_Lg) / data$AUTO_AR_NoLg) * 100,
    ES64_AR_Lg = ((data$AUTO_AR_NoLg - data$ES64_AR_Lg) / data$AUTO_AR_NoLg) * 100,
    AUTO_ADJ_Lg = ((data$AUTO_AR_NoLg - data$AUTO_ADJ_Lg) / data$AUTO_AR_NoLg) * 100,
    ES27_ADJ_Lg = ((data$AUTO_AR_NoLg - data$ES27_ADJ_Lg) / data$AUTO_AR_NoLg) * 100,
    ES64_ADJ_Lg = ((data$AUTO_AR_NoLg - data$ES64_ADJ_Lg) / data$AUTO_AR_NoLg) * 100,
    AUTO_TMP_Lg = ((data$AUTO_AR_NoLg - data$AUTO_TMP_Lg) / data$AUTO_AR_NoLg) * 100,
    ES27_TMP_Lg = ((data$AUTO_AR_NoLg - data$ES27_TMP_Lg) / data$AUTO_AR_NoLg) * 100,
    ES64_TMP_Lg = ((data$AUTO_AR_NoLg - data$ES64_TMP_Lg) / data$AUTO_AR_NoLg) * 100,
    AUTO_EPI_Lg = ((data$AUTO_AR_NoLg - data$AUTO_EPI_Lg) / data$AUTO_AR_NoLg) * 100,
    ES27_EPI_Lg = ((data$AUTO_AR_NoLg - data$ES27_EPI_Lg) / data$AUTO_AR_NoLg) * 100,
    ES64_EPI_Lg = ((data$AUTO_AR_NoLg - data$ES64_EPI_Lg) / data$AUTO_AR_NoLg) * 100
  ))
}

# Calculate percentage of improvemtn 
W1_percentage_of_improvement <- calculate_percentage_of_improvement(W1)
W2_percentage_of_improvement <- calculate_percentage_of_improvement(W2)
W3_percentage_of_improvement <- calculate_percentage_of_improvement(W3)
W4_percentage_of_improvement <- calculate_percentage_of_improvement(W4)

```

Now let's plot a map of percentage of WIS improvement for each state.

Mean and sd percentage of improvement compared to auto_arima

```{r}
# List of column names
#variables <- c("ES27_AR", "ES64_AR", "AUTO_TMP", "ES27_TMP", "ES64_TMP", 
#               "AUTO_EPI", "ES27_EPI", "ES64_EPI", "AUTO_ADJ", "ES27_ADJ", "ES64_ADJ")
variables <-c(names(W1_percentage_of_improvement))

# Create an empty dataframe to store results
summary_impr <- data.frame(WeekAhead = character(), Model = character(), m = numeric(), sd = numeric(), stringsAsFactors = FALSE)

# List of datasets
datasets <- list("1" = W1_percentage_of_improvement, 
                 "2" = W2_percentage_of_improvement, 
                 "3" = W3_percentage_of_improvement, 
                 "4" = W4_percentage_of_improvement)

# Loop through datasets and variables
for (dataset_name in names(datasets)) {
  dataset <- datasets[[dataset_name]]  # Get the dataset

  for (var in variables) {
    data <- dataset[[var]]
    mean_val <- mean(data, na.rm = TRUE)
    sd_val <- sd(data, na.rm = TRUE)

    # Append results to dataframe
    summary_impr <- rbind(summary_impr, data.frame(WeekAhead = dataset_name, Model = var, m = mean_val, sd = sd_val))

    # Generate histogram
    p <- ggplot(dataset, aes(x = .data[[var]])) +
      geom_histogram(color = "black", fill = "lightblue", bins = 30, alpha = 0.7) +
      geom_vline(aes(xintercept = mean_val), color = "red", linetype = "dashed", linewidth = 1) +
      geom_vline(aes(xintercept = mean_val - sd_val), color = "blue", linetype = "dotted", linewidth = 1) +
      geom_vline(aes(xintercept = mean_val + sd_val), color = "blue", linetype = "dotted", linewidth = 1) +
      labs(title = paste("Histogram of", var, "in", dataset_name),
           x = var, 
           y = "Frequency") +
      theme_minimal()

    print(p)  # Display the plot
  }
}

summary_impr$WeekAhead<-as.numeric(summary_impr$WeekAhead)

summary_impr
```

Summary of Post Hoc Wilcox test with  p.adjust.method Benjamini-Yekutieli (BY)

```{r}

all_p_values<-rbind(p_df1,p_df2,p_df3,p_df4)

#names(all_p_values)[names(all_p_values) == "Models"] <- "Model"

p_values_and_impr <- merge(summary_impr, all_p_values, by = c("WeekAhead", "Model"))

# Create a new column "Model_Type" based on whether "NoLg" is in the Model name
p_values_and_impr$Model_Type <- ifelse(grepl("NoLg", p_values_and_impr$Model), "NoLog", "Log")

# View the updated dataframe
head(p_values_and_impr)
p_values_and_impr
```

```{r}

library(ggplot2)
library(dplyr)
library(forcats)

# Define significance threshold  
alpha <- 0.05

# Get model order based on WeekAhead == 1  
model_order <- p_values_and_impr %>%
  filter(WeekAhead == 1) %>%
  arrange(desc(m)) %>%
  pull(Model)  # Extract ordered model names

# Prepare data with significance flag and ordering  
p_df <- p_values_and_impr %>%
  mutate(Significance = ifelse(p_values < alpha, "Significant", "Not Significant"),
         Model_Type = ifelse(grepl("NoLg", Model), "Without log-back transformation", "With log-back transformation"),
         Model = factor(Model, levels = model_order))  # Apply ordered factor

# Heatmap plot with models ordered by WeekAhead == 1  
plot_m_heatmap <- ggplot(p_df, aes(x = factor(WeekAhead), y = Model, fill = m)) +
  geom_tile(aes(fill = ifelse(Significance == "Significant", m, NA)), color = "black") +  # Fill only significant ones
  geom_tile(data = p_df %>% filter(Significance == "Not Significant"), 
            aes(x = factor(WeekAhead), y = Model), fill = "gray80", color = "black") +  # Gray for non-significant
  geom_text(aes(label = paste0("m:", round(m, 2),", sd:", round(sd, 2))), 
            color = "black", size = 4) +  # Add text labels
  scale_fill_gradient(low = "red", high = "cyan", na.value = "gray80", 
                      name = "Mean Improvement", limits = c(-50, 50)) +  # Fixed color scale from -30 to 30
  labs(title = "Model improvement (%) comparing each model' WIS with the baseline across 48 states",
       subtitle = "Gray boxes indicate models not significantly different from the baseline model based on Wilcoxon test with Holm's adjusted p-values",
       x = "Week Ahead",
       y = "") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Model_Type, scales = "free_y")  # Separate NoLg and other models

# Print plot
print(plot_m_heatmap)
ggsave("Fig3.jpg", plot_m_heatmap, width = 15, height = 7 )

```

Friedman test for ADJACENT MODELS

```{r}

library(ggplot2)
library(reshape2)
library(dplyr)
library(tidyr)

# List of datasets
datasets <- list(W1 = W1, W2 = W2, W3 = W3, W4 = W4)

# Function to compute Wilcoxon test and prepare the data for plotting
run_pairwise_tests <- function(data, week) {
  long_data <- data %>%
    pivot_longer(cols = c(AUTO_ADJ_Lg, ES27_ADJ_Lg, ES64_ADJ_Lg),
                 names_to = "Type",
                 values_to = "Value")

  pairwise_results <- pairwise.wilcox.test(long_data$Value, long_data$Type,
                                           p.adjust.method = "holm", paired = TRUE)

  # Convert p-value matrix to long format
  p_df <- as.data.frame(as.table(pairwise_results$p.value)) %>%
    rename(Comparison1 = Var1, Comparison2 = Var2, P_value = Freq) %>%
    mutate(Significance = ifelse(P_value < 0.05, "Significant", "Not Significant"),
           WeekAhead = week)  # Add a column for dataset identifier

  return(p_df)
}

# Apply the function to all datasets and combine results
p_values_df <- bind_rows(lapply(names(datasets), function(w) run_pairwise_tests(datasets[[w]], w)))

adjacent_variance_test<-ggplot(p_values_df, aes(x = Comparison1, y = Comparison2, fill = Significance)) +
  geom_tile(color = "black") +  # Add black gridlines
  geom_text(aes(label = round(P_value, 3)), color = "black", size = 4) +  # Add p-value labels
  scale_fill_manual(values = c("Significant" = "cyan", "Not Significant" = "gray80"), 
                    name = "Significance", na.value = "white", drop = TRUE) +  # Remove NA from legend
  labs(title = "ADJACENT model family pairwise Wilcoxon test",
       subtitle = "Gray indicates not significant differences (Holm-adjusted p-values)",
       x = "",
       y = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ WeekAhead)  # Separate plots for each dataset

adjacent_variance_test

ggsave(adjacent_variance_test,filename="Fig4.jpg", width = 6, height = 5)

```

Now let's plot a map of percentage of WIS improvement for each state.

```{r}
# MAP 1 week ahead

W1_map <- data.frame(
  STATE = W1$STATE,
  percentage_improvement = W1_percentage_of_improvement$AUTO_ADJ_Lg,
  AUTO_ADJ=W1$AUTO_ADJ_Lg
)

# Merge the best model data with the states map
W1_map <- states %>%
  left_join(W1_map, by = c("STATE")) %>%
  drop_na()

################################### MAP 2 weeka ahead

W2_map <- data.frame(
  STATE = W2$STATE,
  percentage_improvement = W2_percentage_of_improvement$AUTO_ADJ_Lg,
  AUTO_ADJ=W2$AUTO_ADJ_Lg
)

# Merge the best model data with the states map
W2_map <- states %>%
  left_join(W2_map, by = c("STATE"))%>%
  drop_na()

##################################### MAP 3 weeka ahead

W3_map <- data.frame(
  STATE = W3$STATE,
  percentage_improvement = W3_percentage_of_improvement$AUTO_ADJ_Lg,
  AUTO_ADJ=W3$AUTO_ADJ_Lg
)

# Merge the best model data with the states map
W3_map <- states %>%
  left_join(W3_map, by = c("STATE"))%>%
  drop_na()

############################### MAP 4 weeka ahead

W4_map <- data.frame(
  STATE = W4$STATE,
  percentage_improvement = W4_percentage_of_improvement$AUTO_ADJ_Lg,
  AUTO_ADJ=W4$AUTO_ADJ_Lg
)

# Merge the best model data with the states map
W4_map <- states %>%
  left_join(W4_map, by = c("STATE"))%>%
  drop_na()
```

Let's plot some maps with the best models in each state and the percentage of improvement compared to the AUTO ARIMA models for the same state.
Here I define some colors categories that I will use on the next maps.

1 week ahead percentage of improvement

```{r}

# Assuming merged_data already includes the necessary columns: model and percentage_improveme
# Create the map plot for 1 Week Ahead with best models and percentage improvement
ES_1WEEK <- ggplot(W1_map) +
  geom_sf(aes(fill = AUTO_ADJ)) +  # Fill based on the best model
  scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 300)) +  
  ggtitle("Mean WIS on each state (W1)") +
  theme_light() +
  theme(legend.position = "right") +
  geom_sf_text(data = W1_map, aes(label = round(AUTO_ADJ,1)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE) +  # Display percentage improvement in each state
  labs(
    fill = "Value",  # Label for the legend
    subtitle = "Based on 2023 and 2024 influenza seasons"
    , 
       x = "",
       y = ""
  )  # Adding a subtitle

x_limits <- c(-125, -67)  # Set the desired longitude range
y_limits <- c(25, 50)    # Set the desired latitude range

ES_1WEEK<- ES_1WEEK + coord_sf(xlim = x_limits, ylim = y_limits)

ggsave(ES_1WEEK, height = 4.5, width = 6.5,  filename="Fig4A.jpg")  
```
2 weeks ahead

```{r}

# Assuming merged_data already includes the necessary columns: model and percentage_improveme
# Create the map plot for 1 Week Ahead with best models and percentage improvement
ES_2WEEK <- ggplot(W2_map) +
  geom_sf(aes(fill = AUTO_ADJ)) +  # Fill based on the best model
  scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 300)) +  
  ggtitle("Mean WIS on each state (W2)") +
  theme_light() +
  theme(legend.position = "right") +
  geom_sf_text(data = W2_map, aes(label = round(AUTO_ADJ,1)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE) +  # Display percentage improvement in each state
  labs(
    fill = "Value",  # Label for the legend
    subtitle = "Based on 2023 and 2024 influenza seasons"
        , 
       x = "",
       y = ""
  )  # Adding a subtitle

x_limits <- c(-125, -67)  # Set the desired longitude range
y_limits <- c(25, 50)    # Set the desired latitude range

ES_2WEEK <-ES_2WEEK + coord_sf(xlim = x_limits, ylim = y_limits)

ggsave(ES_2WEEK, height = 4.5, width = 6.5,  filename="Fig4B.jpg")  
```
3 weeks ahead

```{r}

# Assuming merged_data already includes the necessary columns: model and percentage_improveme
# Create the map plot for 1 Week Ahead with best models and percentage improvement
# Create the map plot for 1 Week Ahead with best models and percentage improvement
ES_3WEEK <- ggplot(W3_map) +
  geom_sf(aes(fill = AUTO_ADJ)) +  # Fill based on the best model
  scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 300)) +  
  ggtitle("Mean WIS on each state (W3)") +
  theme_light() +
  theme(legend.position = "right") +
  geom_sf_text(data = W3_map, aes(label = round(AUTO_ADJ,1)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE) +  # Display percentage improvement in each state
  labs(
    fill = "Value",  # Label for the legend
    subtitle = "Based on 2023 and 2024 influenza seasons"
        , 
       x = "",
       y = ""
  )  # Adding a subtitle

x_limits <- c(-125, -67)  # Set the desired longitude range
y_limits <- c(25, 50)    # Set the desired latitude range

ES_3WEEK<- ES_3WEEK + coord_sf(xlim = x_limits, ylim = y_limits)

ggsave(ES_3WEEK, height = 4.5, width = 6.5,  filename="Fig4C.jpg")  
```
4 weeks ahead

```{r}

ES_4WEEK <- ggplot(W4_map) +
  geom_sf(aes(fill = AUTO_ADJ)) +
  scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", 
                       midpoint = 100, limits = c(0, 300)) +  
  ggtitle("Mean WIS on each state (W4)") +
  theme_light() +
  theme(legend.position = "right") +
  geom_sf_text(data = W4_map, aes(label = round(AUTO_ADJ,1)),  
               size = 3,
               color = "black",
               check_overlap = TRUE) +
  labs(
    fill = "Value",  
    subtitle = "Based on 2023 and 2024 influenza seasons"
        , 
       x = "",
       y = ""
  )

x_limits <- c(-125, -67)  
y_limits <- c(25, 50)

ES_4WEEK<-ES_4WEEK + coord_sf(xlim = x_limits, ylim = y_limits)

ggsave(ES_4WEEK,height = 4.5, width = 6.5,  filename="Fig4D.jpg")  
```

Loading datasets for regression analysis

```{r}
pop_data <- read.csv("NHSH_hospitalizations_seed/regression_features/population_data.csv") # resident population and population density
sovi_data <- read.csv("NHSH_hospitalizations_seed/regression_features/sovi_2010.2014.csv") # SOVI index
bric_data<-read.csv("NHSH_hospitalizations_seed/regression_features/bric2015.csv") # BRIC index
humidity_data<-read.csv("NHSH_hospitalizations_seed/regression_features/humidity_climatology_1990_2020.csv") # ERA5 Specific Humidity
temperature_data<-read.csv("NHSH_hospitalizations_seed/regression_features/temperature_climatology_1990_2020.csv") # ERA5 Specific Temperature

```

REGRESSION MODELS

We will run the regression for evaluating if the best model log(WIS) which represent its performance is related to given independent variables.

```{r}

# List of data frames
WIS_dataframes <- list(W1 = W1_map, W2 = W2_map, W3 = W3_map, W4=W4_map)
regression_models<-data.frame()

```

Percentage of AUTO_ADJ mean WIS

AUTO_ADJ WIS x RESIDENT POPULATION 2020

```{r}

for(i in c(1,2,3,4)){

  # Combining data for the same states
  WIS_pop_data <- inner_join(WIS_dataframes[[i]], pop_data, by = "STATE")
  # Fitting the regression model
  model <- lm((WIS_pop_data$AUTO_ADJ) ~ (WIS_pop_data$Resident_population_2020))
  # View the model summary
  model_summary <- summary(model)
  # Getting the R and p values for the plot
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # Saving the results in a dataframe
  regression_models2 <- data.frame(
    independent_variable = "Resident Population (2020)", 
    Week_Ahead = i, 
    r_squared = r_squared, 
    p_value = p_value
  )
  # Append to the main results dataframe
  regression_models <- rbind(regression_models, regression_models2)
}

regression_models
```

AUTO_ADJ WIS x POPULATION DENSITY

```{r}

for(i in c(1,2,3,4)){

  WIS_pop_data <- inner_join(WIS_dataframes[[i]], pop_data, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_pop_data$AUTO_ADJ) ~ (WIS_pop_data$Population_density_2020))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="Population Density (2020)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # Append to the main results dataframe
  regression_models<-rbind(regression_models,regression_models2)
}

```

Here we will weight the Social Vulnerability Index (SoVI) and Baseline Resilience Indicators for Communities (BRIC) county indexes which for each states based on the population size in each county.

```{r}
#############################################################
# SOVI BY STATE
# weighted by population size in each county
sovi_by_state <- sovi_data %>%
  filter(!is.nan(sovi)) %>%  # Exclude rows where 'sovi' is NaN
  group_by(state.name) %>%
  summarize(weighted_mean = weighted.mean(sovi, w = population.2020, na.rm = TRUE))

colnames(sovi_by_state)[1] <- "STATE"

#############################################################
# BRIC BY STATE
# weighted by population size in each county

bric_by_state <- bric_data %>%
  group_by(state.name) %>%
  summarize(across(16:22, ~ weighted.mean(.x, w = population.2020, na.rm = TRUE), .names = "weighted_mean_{col}"))
colnames(bric_by_state)[1] <- "STATE"

```

AUTO_ADJ WIS x SOVI

```{r}

for(i in c(1,2,3,4)){

  WIS_sovi<-NULL
  WIS_sovi <- inner_join(WIS_dataframes[[i]], sovi_by_state, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_sovi$AUTO_ADJ) ~ (WIS_sovi$weighted_mean))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="Social Vulnerability Index (SoVI)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending the results
  regression_models<-rbind(regression_models,regression_models2)
}

```

AUTO_ADJ WIS x BRIC SOCIAL

```{r}

for(i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model  
  model <- lm((WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.social))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Social (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending the results
  regression_models<-rbind(regression_models,regression_models2)
}

```

AUTO_ADJ WIS x BRIC ECONOMIC

```{r}
for(i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.economic))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Economic (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  regression_models<-rbind(regression_models,regression_models2)
}

```

AUTO_ADJ WIS x BRIC Infrastructure

```{r}

for (i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.infrastructure))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Infrastructure (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending results
  regression_models<-rbind(regression_models,regression_models2)
}

```
AUTO_ADJ WIS x BRIC institutional

```{r}

for (i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.institutional))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Institutional (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending results
  regression_models<-rbind(regression_models,regression_models2)
}

```

AUTO_ADJ WIS x BRIC community

```{r}

for(i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[1]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.community))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Community (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  regression_models<-rbind(regression_models,regression_models2)
}

```

AUTO_ADJ WIS x BRIC environment

```{r}

for(i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.environment))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Environment (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  regression_models<-rbind(regression_models,regression_models2)
}

```
AUTO_ADJ WIS x BRIC total

```{r}

for(i in 1:4){
  # BRIC INDEX
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.total))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC total (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  regression_models<-rbind(regression_models,regression_models2)
}
```

TEMPERATUE - ERA5

Now we will look at regression models that uses mean temperature and specific humidity.

-------------------

AUTO_ADJ WIS x Temperature ERA5 Data

```{r}
for(i in 1:4){
  # TEMPERATURE DATA from ERA5
  WIS_temp <- inner_join(WIS_dataframes[[i]], temperature_data, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_temp$AUTO_ADJ) ~ (WIS_temp$mean))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="Mean Temperature (1990-2020)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending results
  regression_models<-rbind(regression_models,regression_models2)
}
```

AUTO_ADJ WIS x Specific Humidity ERA5 Data

```{r}
library(ggpubr)

# regression results 
print(regression_models)

for(i in 1:4){
  # HUMIDITY DATA from ERA5
  WIS_humidity <- inner_join(WIS_dataframes[[i]], humidity_data, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_humidity$AUTO_ADJ) ~ (WIS_humidity$mean))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="Mean Specific Humidty (1990-2020)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending results
  regression_models<-rbind(regression_models,regression_models2)
}

# Plot the table
regression_models
```

SUMMARY OF RESULTS

```{r}
# Define a significance threshold (p < 0.05)
regression_models <- regression_models %>%
  mutate(significance = ifelse(p_value < 0.05, "Significant", "Not Significant"))

# Plot
reg2<-ggplot(regression_models, aes(x = Week_Ahead, y = independent_variable, fill = significance)) +  
  geom_tile(color = "black") +  # Add black borders to squares
  geom_text(aes(label = round(r_squared, 3)), color = "black", size = 6) +  # Text inside boxes
  scale_fill_manual(values = c("Significant" = "cyan", "Not Significant" = "gray")) +  
  labs(title = "Regression models",
       subtitle = "Mean WIS as dependent variable",
       x = "Week Ahead",
       y = "Independent Variables",
       fill = "Significance") +  
  theme_minimal() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        axis.text.y = element_text(size = 10))

reg2

ggsave(reg2, height = 8, width = 10, filename="Fig4E.jpg")  
#ggsave(reg2, filename="reg2.jpg",height = 8, width = 10)
```

Let's plot some maps with the with the mean (%) of difference between the AUTO_ADJ and the AUTO ARIMA models for the same state.

1 week ahead percentage of improvement

```{r}
# Assuming merged_data already includes the necessary columns: model and percentage_improveme
# Create the map plot for 1 Week Ahead with best models and percentage improvement
ES_1WEEK <- ggplot(W1_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(high = "darkred", mid = "lightyellow", low = "blue", midpoint = 0, limits = c(-120, 10)) +  
  ggtitle("Mean WIS improvement in percentage (W1)") +
  theme_light() +
  theme(legend.position = "right") +
  geom_sf_text(data = W1_map, aes(label = round(percentage_improvement,1)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE) +  # Display percentage improvement in each state
  labs(
    fill = "%",  # Label for the legend
    subtitle = "(Baseline - AUTO_ADJ_LB) / (Baseline*100)"
            , 
       x = "",
       y = ""
  )  # Adding a subtitle

x_limits <- c(-125, -67)  # Set the desired longitude range
y_limits <- c(25, 50)    # Set the desired latitude range

ES_1WEEK<-ES_1WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
ES_1WEEK
ggsave(ES_1WEEK,height = 4.5, width = 6.5,  filename="Fig5A.jpg")  
```

2 weeks ahead

```{r}

# Assuming merged_data already includes the necessary columns: model and percentage_improveme
# Create the map plot for 1 Week Ahead with best models and percentage improvement
ES_2WEEK <- ggplot(W2_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(high = "darkred", mid = "lightyellow", low = "blue", midpoint = 0, limits = c(-120, 10)) +  
  ggtitle("Mean WIS improvement in percentage (W2)") +
  theme_light() +
  theme(legend.position = "right") +
  geom_sf_text(data = W2_map, aes(label = round(percentage_improvement,1)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE) +  # Display percentage improvement in each state
  labs(
    fill = "%",  # Label for the legend
    subtitle = "(Baseline - AUTO_ADJ_LB) / (Baseline*100)"
            , 
       x = "",
       y = ""
  )  # Adding a subtitle

x_limits <- c(-125, -67)  # Set the desired longitude range
y_limits <- c(25, 50)    # Set the desired latitude range

ES_2WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
ES_2WEEK
ggsave(ES_2WEEK, height = 4.5, width = 6.5, filename="Fig5B.jpg")  
```

3 weeks ahead

```{r}

# Assuming merged_data already includes the necessary columns: model and percentage_improveme
# Create the map plot for 1 Week Ahead with best models and percentage improvement
# Create the map plot for 1 Week Ahead with best models and percentage improvement
ES_3WEEK <- ggplot(W3_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(high = "darkred", mid = "lightyellow", low = "blue", midpoint = 0, limits = c(-120, 10)) +  
  ggtitle("Mean WIS improvement in percentage (W3)") +
  theme_light() +
  theme(legend.position = "right") +
  geom_sf_text(data = W3_map, aes(label = round(percentage_improvement,1)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE) +  # Display percentage improvement in each state
  labs(
    fill = "%",  # Label for the legend
    subtitle = "(Baseline - AUTO_ADJ_LB) / (Baseline*100)"
            , 
       x = "",
       y = ""
  )  # Adding a subtitle

x_limits <- c(-125, -67)  # Set the desired longitude range
y_limits <- c(25, 50)    # Set the desired latitude range

ES_3WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
ES_3WEEK
ggsave(ES_3WEEK, height = 4.5, width = 6.5, filename="Fig5C.jpg")  
```
4 weeks ahead

```{r}

# Assuming merged_data already includes the necessary columns: model and percentage_improveme
# Create the map plot for 1 Week Ahead with best models and percentage improvement
ES_4WEEK <- ggplot(W4_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(high = "darkred", mid = "lightyellow", low = "blue", midpoint = 0, limits = c(-120, 10)) +  
  ggtitle("Mean WIS improvement in percentage (W4)") +
  theme_light() +
  theme(legend.position = "right") +
  geom_sf_text(data = W4_map, aes(label = round(percentage_improvement,1)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE) +  # Display percentage improvement in each state
  labs(
    fill = "%",  # Label for the legend
    subtitle = "(Baseline - AUTO_ADJ_LB) / (Baseline*100)"
            , 
       x = "",
       y = ""
  )  # Adding a subtitle

x_limits <- c(-125, -67)  # Set the desired longitude range
y_limits <- c(25, 50)    # Set the desired latitude range

ES_4WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
ES_4WEEK
ggsave(ES_4WEEK, height = 4.5, width = 6.5,  filename="Fig5D.jpg")  
```
REGRESSION MODELS

We will run the regression for evaluating if the best model log(WIS) which represent its performance is related to given independent variables.

```{r}

# List of data frames
WIS_dataframes <- list(W1 = W1_map, W2 = W2_map, W3 = W3_map, W4=W4_map)
regression_models<-data.frame()

```

Percentage of improvement regression analysis

AUTO_ADJ WIS x RESIDENT POPULATION 2020

```{r}

for(i in c(1,2,3,4)){

  # Combining data for the same states
  WIS_pop_data <- inner_join(WIS_dataframes[[i]], pop_data, by = "STATE")
  # Fitting the regression model
  model <- lm((WIS_pop_data$percentage_improvement) ~ (WIS_pop_data$Resident_population_2020))
  # View the model summary
  model_summary <- summary(model)
  # Getting the R and p values for the plot
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # Saving the results in a dataframe
  regression_models2 <- data.frame(
    independent_variable = "Resident Population (2020)", 
    Week_Ahead = i, 
    r_squared = r_squared, 
    p_value = p_value
  )
  # Append to the main results dataframe
  regression_models <- rbind(regression_models, regression_models2)
}

regression_models
```

AUTO_ADJ WIS x POPULATION DENSITY

```{r}

for(i in c(1,2,3,4)){

  WIS_pop_data <- inner_join(WIS_dataframes[[i]], pop_data, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_pop_data$percentage_improvement) ~ (WIS_pop_data$Population_density_2020))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="Population Density (2020)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # Append to the main results dataframe
  regression_models<-rbind(regression_models,regression_models2)
}

```

Here we will weight the Social Vulnerability Index (SoVI) and Baseline Resilience Indicators for Communities (BRIC) county indexes which for each states based on the population size in each county.

```{r}
#############################################################
# SOVI BY STATE
# weighted by population size in each county
sovi_by_state <- sovi_data %>%
  filter(!is.nan(sovi)) %>%  # Exclude rows where 'sovi' is NaN
  group_by(state.name) %>%
  summarize(weighted_mean = weighted.mean(sovi, w = population.2020, na.rm = TRUE))

colnames(sovi_by_state)[1] <- "STATE"

#############################################################
# BRIC BY STATE
# weighted by population size in each county

bric_by_state <- bric_data %>%
  group_by(state.name) %>%
  summarize(across(16:22, ~ weighted.mean(.x, w = population.2020, na.rm = TRUE), .names = "weighted_mean_{col}"))
colnames(bric_by_state)[1] <- "STATE"

```

AUTO_ADJ WIS x SOVI

```{r}

for(i in c(1,2,3,4)){

  WIS_sovi<-NULL
  WIS_sovi <- inner_join(WIS_dataframes[[i]], sovi_by_state, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_sovi$percentage_improvement) ~ (WIS_sovi$weighted_mean))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="Social Vulnerability Index (SoVI)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending the results
  regression_models<-rbind(regression_models,regression_models2)
}

```

AUTO_ADJ WIS x BRIC SOCIAL

```{r}

for(i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model  
  model <- lm((WIS_bric$percentage_improvement) ~ (WIS_bric$weighted_mean_z_bric.social))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Social (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending the results
  regression_models<-rbind(regression_models,regression_models2)
}

```

AUTO_ADJ WIS x BRIC ECONOMIC

```{r}
for(i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_bric$percentage_improvement) ~ (WIS_bric$weighted_mean_z_bric.economic))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Economic (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  regression_models<-rbind(regression_models,regression_models2)
}

```

AUTO_ADJ WIS x BRIC Infrastructure

```{r}

for (i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_bric$percentage_improvement) ~ (WIS_bric$weighted_mean_z_bric.infrastructure))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Infrastructure (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending results
  regression_models<-rbind(regression_models,regression_models2)
}

```

AUTO_ADJ WIS x BRIC institutional

```{r}

for (i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_bric$percentage_improvement) ~ (WIS_bric$weighted_mean_z_bric.institutional))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Institutional (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending results
  regression_models<-rbind(regression_models,regression_models2)
}

```

AUTO_ADJ WIS x BRIC community

```{r}

for(i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[1]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_bric$percentage_improvement) ~ (WIS_bric$weighted_mean_z_bric.community))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Community (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  regression_models<-rbind(regression_models,regression_models2)
}

```

AUTO_ADJ WIS x BRIC environment

```{r}

for(i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_bric$percentage_improvement) ~ (WIS_bric$weighted_mean_z_bric.environment))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Environment (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  regression_models<-rbind(regression_models,regression_models2)
}

```
AUTO_ADJ WIS x BRIC total

```{r}

for(i in 1:4){
  # BRIC INDEX
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_bric$percentage_improvement) ~ (WIS_bric$weighted_mean_z_bric.total))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC total (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  regression_models<-rbind(regression_models,regression_models2)
}
```

TEMPERATUE - ERA5

Now we will look at regression models that uses mean temperature and specific humidity.

-------------------

AUTO_ADJ WIS x Temperature ERA5 Data

```{r}

for(i in 1:4){
  # TEMPERATURE DATA from ERA5
  WIS_temp <- inner_join(WIS_dataframes[[i]], temperature_data, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_temp$percentage_improvement) ~ (WIS_temp$mean))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="Mean Temperature (1990-2020)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending results
  regression_models<-rbind(regression_models,regression_models2)
}
```

AUTO_ADJ WIS x Specific Humidity ERA5 Data

```{r}
library(ggpubr)

# regression results 
print(regression_models)

for(i in 1:4){
  # HUMIDITY DATA from ERA5
  WIS_humidity <- inner_join(WIS_dataframes[[i]], humidity_data, by = "STATE")
  # Fit the regression model
  model <- lm((WIS_humidity$percentage_improvement) ~ (WIS_humidity$mean))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="Mean Specific Humidty (1990-2020)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending results
  regression_models<-rbind(regression_models,regression_models2)
}

# Plot the table
regression_models
```

SUMMARY OF RESULTS

```{r}
# Define a significance threshold (p < 0.05)
regression_models <- regression_models %>%
  mutate(significance = ifelse(p_value < 0.05, "Significant", "Not Significant"))

# Plot
reg2<-ggplot(regression_models, aes(x = Week_Ahead, y = independent_variable, fill = significance)) +  
  geom_tile(color = "black") +  # Add black borders to squares
  geom_text(aes(label = round(r_squared, 3)), color = "black", size = 6) +  # Text inside boxes
  scale_fill_manual(values = c("Significant" = "cyan", "Not Significant" = "gray")) +  
  labs(title = "Regression models",
       subtitle = "Percentage of WIS improvement as dependent variable",
       x = "Week Ahead",
       y = "Independent Variables",
       fill = "Significance") +    
  theme_minimal() +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        axis.text.y = element_text(size = 10))

reg2
ggsave(reg2, filename="Fig5E.jpg",height = 8, width = 10)
```

Now let's evaluate which was the mean WIS improvement by epidemiological week

```{r}

# WEEK 1
W1_AUTO_AR_NoLg_WIS_BY_EPI <- filtered_df_W1_NoLg %>%
  group_by(epiweek) %>%
  summarize(AUTO_AR_NoLg = mean(AUTO_AR_NoLg, na.rm = TRUE))

W1_AUTO_ADJ_WIS_BY_EPI <- filtered_df_W1 %>%
  group_by(epiweek) %>%
  summarize(AUTO_ADJ = mean(AUTO_ADJ, na.rm = TRUE))
# WEEK 2
W2_AUTO_AR_NoLg_WIS_BY_EPI <- filtered_df_W2_NoLg %>%
  group_by(epiweek) %>%
  summarize(AUTO_AR_NoLg = mean(AUTO_AR_NoLg, na.rm = TRUE))

W2_AUTO_ADJ_WIS_BY_EPI <- filtered_df_W2 %>%
  group_by(epiweek) %>%
  summarize(AUTO_ADJ = mean(AUTO_ADJ, na.rm = TRUE))

# WEEK 3
W3_AUTO_AR_NoLg_WIS_BY_EPI <- filtered_df_W3_NoLg %>%
  group_by(epiweek) %>%
  summarize(AUTO_AR_NoLg = mean(AUTO_AR_NoLg, na.rm = TRUE))

W3_AUTO_ADJ_WIS_BY_EPI <- filtered_df_W3 %>%
  group_by(epiweek) %>%
  summarize(AUTO_ADJ = mean(AUTO_ADJ, na.rm = TRUE))

# WEEK 4
W4_AUTO_AR_NoLg_WIS_BY_EPI <- filtered_df_W4_NoLg %>%
  group_by(epiweek) %>%
  summarize(AUTO_AR_NoLg = mean(AUTO_AR_NoLg, na.rm = TRUE))

W4_AUTO_ADJ_WIS_BY_EPI <- filtered_df_W4 %>%
  group_by(epiweek) %>%
  summarize(AUTO_ADJ = mean(AUTO_ADJ, na.rm = TRUE))

filtered_df_W1
filtered_df_W2

```

Absolute WIS difference by epidemiological week

```{r}

# Define pairs of data frames
auto_df <- list(W1_AUTO_AR_NoLg_WIS_BY_EPI, W2_AUTO_AR_NoLg_WIS_BY_EPI, W3_AUTO_AR_NoLg_WIS_BY_EPI, W4_AUTO_AR_NoLg_WIS_BY_EPI)
adj_df <- list(W1_AUTO_ADJ_WIS_BY_EPI, W2_AUTO_ADJ_WIS_BY_EPI, W3_AUTO_ADJ_WIS_BY_EPI, W4_AUTO_ADJ_WIS_BY_EPI)
week_labels <- c("W1", "W2", "W3", "W4")  # Labels for plots

# Function to process each pair
for (i in 1:4) {
  merged_df <- merge(auto_df[[i]], adj_df[[i]], by = "epiweek") %>%
    mutate(percentage_diff = AUTO_AR_NoLg - AUTO_ADJ) %>%
    filter((epiweek >= 40 & epiweek <= 53) | (epiweek >= 1 & epiweek <= 20)) %>%
    mutate(epiweek_order = factor(epiweek, levels = c(40:53, 1:20)))

  mean_value <- mean(merged_df$percentage_diff, na.rm = TRUE)

  plots <- ggplot(merged_df, aes(x = epiweek_order, y = percentage_diff, fill = percentage_diff > 0)) +
    geom_bar(stat = "identity") +
    scale_fill_manual(values = c("TRUE" = "red", "FALSE" = "blue"), guide = "none") +  # Blue for positive, Red for negative
    geom_hline(yintercept = mean_value, linetype = "dashed", color = "black") +
    labs(x = "Epiweek", y = "Absolute WIS difference",
         title = paste("Mean absolute WIS improvement by epidemiological week across 48 states (", week_labels[i],")"),
         subtitle = "Baseline - AUTO_ADJ_LB") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    annotate("text", x = 1, y = mean_value, label = paste("Mean:", round(mean_value, 2)), color = "black", vjust = -1)

  print(plots)  # Print each plot inside the loop

    # Save the plot with a unique filename
  ggsave(filename = paste0("Fig6_", week_labels[i], ".png"),
         plot = plots, width = 8, height = 6, dpi = 300)
}

```

Absolute WIS difference by epidemiological week

```{r}
library(dplyr)
library(ggplot2)

# Define lists of data frames
auto_df <- list(filtered_df_W1_NoLg, filtered_df_W2_NoLg, filtered_df_W3_NoLg, filtered_df_W4_NoLg)
adj_df <- list(filtered_df_W1, filtered_df_W2, filtered_df_W3, filtered_df_W4)

data__<-merge(filtered_df_W1_NoLg,filtered_df_W1)

library(dplyr)

# Summing values by Julian date across all states
summed_data <- data__ %>%
  group_by(Julian_date) %>%  # Group only by Julian date
  summarize(
    AUTO_AR_NoLg = sum(AUTO_AR_NoLg, na.rm = TRUE),  # Sum AUTO_AR_NoLg across all states
    AUTO_ADJ = sum(AUTO_ADJ, na.rm = TRUE)           # Sum ES27_EPI across all states
  )

```

```{r}
library(dplyr)
library(ggplot2)

# Create a list of data frames for the four weeks
weeks <- 1:4
summed_data_list <- list()

for (week in weeks) {
  # Merge the data frames for the current week
  data__ <- merge(get(paste0("filtered_df_W", week, "_NoLg")),
                  get(paste0("filtered_df_W", week)))

  # Sum values by Julian date across all states
  summed_data <- data__ %>%
    group_by(Julian_date) %>%
    summarize(
      AUTO_AR_NoLg = sum(AUTO_AR_NoLg, na.rm = TRUE),  # Sum across all states
      AUTO_ADJ = sum(AUTO_ADJ, na.rm = TRUE)           # Sum across all states
    ) %>%
    # Calculate the difference ratio
    mutate(Difference = 1 - (AUTO_AR_NoLg / AUTO_ADJ),
           Week = paste0("Week ", week))

  # Store the summarized data in the list
  summed_data_list[[week]] <- summed_data
}

# Combine all weeks into a single data frame
summed_data_all <- bind_rows(summed_data_list)

# Plot the difference using ggplot2
ggplot(summed_data_all, aes(x = Julian_date, y = Difference, color = Week)) +
  geom_line(size = 1) +        # Line plot
  geom_point(size = 2) +       # Points on the line
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", size = 1) +  # Reference line at 0
  theme_minimal() +
  labs(
    title = "Difference between AUTO_AR_NoLg and AUTO_ADJ across States for 4 Weeks",
    x = "Julian Date",
    y = "Difference (1 - AUTO_AR_NoLg / AUTO_ADJ)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### 2025-04-25 NHSH_hospitalizations_seed/ES_TEMPERATURE/ES_TEMPERATURE.R
update

@@ -1,0 +1,400 @@
+#' ES_TEMPERATURE
+#' This function utilizes ensembles and single automatic ARIMAX models which have mean temperature as exogenous variables.
+#' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
+#' It return some metrics that evaluate the performance of the models:
+#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
+#' The user can choose  single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).
+#' The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window.
+#' 
+#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. states.
+#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
+#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
+#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
+#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
+#' @param week_lag An integer specifying the week lag between the exogenous variables and the ILI cases in the state under analysis. Default is \code{1}.
+#' @param temperature_data A data frame containing weekly mean temperature data for ALL U.S. states for the same weeks as the ILI data.
+#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
+#'
+#' @return A list containing the forecast results and performance metrics.
+
+ES_TEMPERATURE<-function(current_state_tb, week_lag=1, temperature_data=NULL, auto=FALSE, n_weeks_ahead=1, ES27=TRUE, ES64=FALSE, window=104){
+  
+  set.seed(1)
+  # The model run the ES27 or the ES64.
+  ES27=!ES64 
+  # Empty list that will contain forecasts, predictive quantiles and number of models.
+  results<-listenv()   
+  if(ES27){
+    pdq=c(0,1,2) # Possible ARIMA pdq's.
[…]

### 2025-04-25 NHSH_hospitalizations_seed/ES_ARIMA/ES_ARIMA.Rmd
update

---
title: "ARIMA ENSEMBLE MODELS - Influenza hospitalization data"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

This function utilizes ensembles and single automatic ARIMA models. The function fits a rolling window of 104 weeks to generate forecasts. It returns some metrics that evaluate the performance of the models: target_end_date, abs_error, cases, forecast, number of models, weighted interval score (WIS) and predictive quantiles (%). The user defines if it will use an AUTO ARIMA (auto=TRUE), or ensembles of 27 (ES27=TRUE) or ensembles of 64 models (ES64=TRUE). It can also choose the number of weeks ahead for each forecast, and the size of the rolling window which is set as 104 (2 years).

```{r}
knitr::opts_chunk$set(echo = TRUE)
```
!!!!!!!!!!!!!!!!!!!!
LOADING THE PACKAGES
!!!!!!!!!!!!!!!!!!!!
```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
```
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
LOADING DATASET AND FUNCTIONS 
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

```{r}
###########################################
#       LOADING AND CLEANING THE DATASET  #        
###########################################

# Loads the ADJACENT states models
source("ES_ARIMA.R", local = TRUE, chdir = TRUE)

# Loads the ILI dataset
my_data = read.csv("treated_influenza_hosp_dataframe_v2.csv")
my_data$target_end_date<-as.Date(my_data$target_end_date) # set the dates as dates

list_of_states <- split(my_data, my_data$state_name)

```

AUTO ARIMA WEEK1

```{r}
start_time <- Sys.time()

AUTO_ARIMA_WEEK1_list <- mclapply(list_of_states, ES_ARIMA, auto=TRUE, n_weeks_ahead=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
AUTO_ARIMA_WEEK1 <- bind_rows(AUTO_ARIMA_WEEK1_list, .id = "State")
```

AUTO ARIMA WEEK2

```{r}
start_time <- Sys.time()

AUTO_ARIMA_WEEK2_list <- mclapply(list_of_states , ES_ARIMA, auto=TRUE, n_weeks_ahead=2, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
AUTO_ARIMA_WEEK2 <- bind_rows(AUTO_ARIMA_WEEK2_list, .id = "State")

```

AUTO ARIMA WEEK3

```{r}
start_time <- Sys.time()

AUTO_ARIMA_WEEK3_list <- mclapply(list_of_states , ES_ARIMA, auto=TRUE, n_weeks_ahead=3, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
AUTO_ARIMA_WEEK3 <- bind_rows(AUTO_ARIMA_WEEK3_list, .id = "State")

```

AUTO ARIMA WEEK4

```{r}
start_time <- Sys.time()

AUTO_ARIMA_WEEK4_list <- mclapply(list_of_states, ES_ARIMA, auto=TRUE, n_weeks_ahead=4 ,mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
AUTO_ARIMA_WEEK4 <- bind_rows(AUTO_ARIMA_WEEK4_list, .id = "State")

```

```{r}
save.image("ARIMA_MODELS_influenza_hospitalization.Rdata")
```

ES27 ARIMA WEEK1

```{r}
start_time <- Sys.time()

ES27_ARIMA_WEEK1_list <- mclapply(list_of_states, ES_ARIMA, ES27=TRUE, n_weeks_ahead=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_ARIMA_WEEK1 <- bind_rows(ES27_ARIMA_WEEK1_list, .id = "State")

```

ES27 ARIMA WEEK2

```{r}
start_time <- Sys.time()

ES27_ARIMA_WEEK2_list <- mclapply(list_of_states, ES_ARIMA, ES27=TRUE, n_weeks_ahead=2, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_ARIMA_WEEK2 <- bind_rows(ES27_ARIMA_WEEK2_list, .id = "State")

```

ES27 ARIMA WEEK3

```{r}
start_time <- Sys.time()

ES27_ARIMA_WEEK3_list <- mclapply(list_of_states, ES_ARIMA, ES27=TRUE, n_weeks_ahead=3, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_ARIMA_WEEK3 <- bind_rows(ES27_ARIMA_WEEK3_list, .id = "State")

```

ES27 ARIMA WEEK4

```{r}
start_time <- Sys.time()

ES27_ARIMA_WEEK4_list <- mclapply(list_of_states, ES_ARIMA, ES27=TRUE, n_weeks_ahead=4 ,mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_ARIMA_WEEK4 <- bind_rows(ES27_ARIMA_WEEK4_list, .id = "State")

```

```{r}
save.image("ARIMA_MODELS_influenza_hospitalization.Rdata")
```

ES64 ARIMA WEEK1

```{r}
start_time <- Sys.time()

ES64_ARIMA_WEEK1_list <- mclapply(list_of_states, ES_ARIMA, ES64=TRUE, n_weeks_ahead=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_ARIMA_WEEK1 <- bind_rows(ES64_ARIMA_WEEK1_list, .id = "State")

```

ES64 ARIMA WEEK2

```{r}
start_time <- Sys.time()

ES64_ARIMA_WEEK2_list <- mclapply(list_of_states, ES_ARIMA, ES64=TRUE, n_weeks_ahead=2, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_ARIMA_WEEK2 <- bind_rows(ES64_ARIMA_WEEK2_list, .id = "State")

```

ES64 ARIMA WEEK3

```{r}
start_time <- Sys.time()

ES64_ARIMA_WEEK3_list <- mclapply(list_of_states, ES_ARIMA, ES64=TRUE, n_weeks_ahead=3, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_ARIMA_WEEK3 <- bind_rows(ES64_ARIMA_WEEK3_list, .id = "State")

```

ES64 ARIMA WEEK4

```{r}
start_time <- Sys.time()

ES64_ARIMA_WEEK4_list <- mclapply(list_of_states, ES_ARIMA, ES64=TRUE, n_weeks_ahead=4 ,mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_ARIMA_WEEK4 <- bind_rows(ES64_ARIMA_WEEK4_list, .id = "State")

```

```{r}
save.image("ARIMA_MODELS_influenza_hospitalization.Rdata")
```

### 2025-04-25 NHSH_hospitalizations_seed/hospitalization_results.Rmd
update

---
title: "INFLUENZA HOSPITALIZATIONS RESULTS"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
---

This document summarizes the results of the 12 models that were developed: AUTO ARIMA, AUTO ADJACENT, AUTO EPIWEEK, ES27 TEMPERATURE, ES27 ARIMA, ES27 ADJACENT, ES27 EPIWEEK, ES64 TEMPERATURE, ES64 ARIMA, ES64 ADJACENT, ES64 EPIWEEK, ES64 TEMPERATURE. It has a series of histograms and maps that compare the results of these models based on their WIS. You can read more about the specific covariates and ensemble techniques of these models in their own Rmarkdown documents.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries.

```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
library("ggplot2")
library("sf")
library("forcats")
library("ggplot2")
library("sf")
library("dplyr")
library("scales")  # For `breaks_extended`
library("ggplot2")
library("broom")
library("fields")

```

Loading the results of each model and the shapefiles of the maps.

```{r}
load("ES_ARIMA/ARIMA_MODELS_influenza_hospitalization.Rdata")
load("ES_ADJACENT/ADJACENT_MODELS_influenza_hospitalization.Rdata")
load("ES_EPIWEEK/EPIWEEK_MODELS_influenza_hospitalization.Rdata")
load("ES_TEMPERATURE/TEMPERATURE_MODELS_influenza_hospitalization.Rdata")

states <- read_sf("shapefiles/cb_2018_us_state_500k.shp")

states <- states %>%
  rename(STATE = NAME)

```

The results come as lists of dataframes with all states (e.g. AUTO_ARIMA_WEEK1_list) or as a single dataframe which combines all states (e.g. AUTO_ARIMA_WEEK1). Each model was run for forecasting 1-4 weeks ahead (e.g. AUTO_ARIMA_WEEK1, AUTO_ARIMA_WEEK2, AUTO_ARIMA_WEEK3, AUTO_ARIMA_WEEK4).

For easier visualization create some variables and plot some of the results for a specific model:

```{r}
# Here we extract the results of the ES27_ARIMA model which was utilized to forecast 1 WEEK ahead for the state of Alabama.

my_wis<- AUTO_ARIMA_WEEK1_list$Alabama[['WIS']] # get the actual cases
my_cases<- AUTO_ARIMA_WEEK1_list$Alabama[['cases']] # get the abs_error
my_forecasts<- AUTO_ARIMA_WEEK1_list$Alabama[['forecasts']] # get the number of models
my_abs_error<- AUTO_ARIMA_WEEK1_list$Alabama[['abs_error']] # get the abs_error
my_number_of_models<- AUTO_ARIMA_WEEK1_list$Alabama[['Number_of_models']] # get the number of models
my_dates<- AUTO_ARIMA_WEEK1_list$Alabama[['target_end_date']] # get the dates

# Here we will take a look at the WIS for 3 different states
Alabama<- AUTO_ARIMA_WEEK1_list$Alabama[['WIS']] # get the actual cases
Colorado<- AUTO_ARIMA_WEEK1_list$Colorado[['WIS']] # get the forecasts
Montana<- AUTO_ARIMA_WEEK1_list$Montana[['WIS']] # get the forecasts

# Assuming your data is in a data frame with columns 'my_dates', 'my_cases', and 'my_wis'
wis_data <- data.frame(
  my_dates = my_dates, # Ensure dates are in Date format
  Alabama = Alabama,
  Colorado = Colorado,
  Montana= Montana# Take the absolute value of WIS
)

# Calculate the mean WIS
mean_Alabama <- round(mean(wis_data$Alabama), 6)
mean_Colorado <- round(mean(wis_data$Colorado), 6)
mean_Montana <- round(mean(wis_data$Montana), 6)

# Let's look at WIS data for 3 different states
ggplot(wis_data) +
  geom_line(aes(x = my_dates, y = Alabama, color = "Alabama")) + # Plot cases and set legend label
  geom_line(aes(x = my_dates, y = Colorado, color = "Colorado")) + # Plot cases and set legend label
  geom_line(aes(x = my_dates, y = Montana, color = "Montana")) + # Plot WIS and set legend label
  scale_color_manual(values = c("Alabama" = "red", "Colorado" = "blue","Montana" = "green")) + # Define colors
  labs(
    title = "AUTO ARIMA weighted interval scores (WIS) for each target week (1 Wk ahead)",
    x = "",
    y = "WIS",
    color = "States", # Title for the legend
    subtitle = paste("MEAN WIS at Alabama:", mean_Alabama, "Colorado:", mean_Colorado, "Montana:", mean_Montana)
  ) +
  theme_minimal()

```

------------
1 Week ahead
------------
```{r}
# Creating new columns with Epidemiological weeks based on target_end_week
AUTO_ARIMA_WEEK1$epiweek <- MMWRweek(AUTO_ARIMA_WEEK1$target_end_date)$MMWRweek
AUTO_ADJACENT_WEEK1$epiweek <- MMWRweek(AUTO_ADJACENT_WEEK1$target_end_date)$MMWRweek
AUTO_EPIWEEK_WEEK1$epiweek <- MMWRweek(AUTO_EPIWEEK_WEEK1$target_end_date)$MMWRweek
AUTO_TEMPERATURE_WEEK1$epiweek <- MMWRweek(AUTO_TEMPERATURE_WEEK1$target_end_date)$MMWRweek
ES27_ARIMA_WEEK1$epiweek <- MMWRweek(ES27_ARIMA_WEEK1$target_end_date)$MMWRweek
ES27_ADJACENT_WEEK1$epiweek <- MMWRweek(ES27_ADJACENT_WEEK1$target_end_date)$MMWRweek
ES27_EPIWEEK_WEEK1$epiweek <- MMWRweek(ES27_EPIWEEK_WEEK1$target_end_date)$MMWRweek
ES27_TEMPERATURE_WEEK1$epiweek <- MMWRweek(ES27_TEMPERATURE_WEEK1$target_end_date)$MMWRweek
ES64_ARIMA_WEEK1$epiweek <- MMWRweek(ES64_ARIMA_WEEK1$target_end_date)$MMWRweek
ES64_ADJACENT_WEEK1$epiweek <- MMWRweek(ES64_ADJACENT_WEEK1$target_end_date)$MMWRweek
ES64_EPIWEEK_WEEK1$epiweek <- MMWRweek(ES64_EPIWEEK_WEEK1$target_end_date)$MMWRweek
ES64_TEMPERATURE_WEEK1$epiweek <- MMWRweek(ES64_TEMPERATURE_WEEK1$target_end_date)$MMWRweek

# Dataframe that will be analysed for 1 Week Ahead
df_W1 <- data.frame(
  STATE = AUTO_ARIMA_WEEK1$State,
  Julian_date = AUTO_ARIMA_WEEK1$target_end_date,
  epiweek = AUTO_ARIMA_WEEK1$epiweek,
  AUTO_AR=AUTO_ARIMA_WEEK1$WIS, 
  AUTO_ADJ=AUTO_ADJACENT_WEEK1$WIS, 
  AUTO_EPI=AUTO_EPIWEEK_WEEK1$WIS,
  AUTO_TEMP=AUTO_TEMPERATURE_WEEK1$WIS,
  ES27_AR=ES27_ARIMA_WEEK1$WIS, 
  ES27_ADJ=ES27_ADJACENT_WEEK1$WIS, 
  ES27_EPI=ES27_EPIWEEK_WEEK1$WIS,
  ES27_TEMP=ES27_TEMPERATURE_WEEK1$WIS,
  ES64_AR=ES64_ARIMA_WEEK1$WIS, 
  ES64_ADJ=ES64_ADJACENT_WEEK1$WIS, 
  ES64_EPI=ES64_EPIWEEK_WEEK1$WIS,
  ES64_TEMP=ES64_TEMPERATURE_WEEK1$WIS
)

head(df_W1)
```

------------
2 Weeks ahead
-------------
```{r}
# Creating new columns with Epidemiological weeks based on target_end_week
AUTO_ARIMA_WEEK2$epiweek <- MMWRweek(AUTO_ARIMA_WEEK2$target_end_date)$MMWRweek
AUTO_ADJACENT_WEEK2$epiweek <- MMWRweek(AUTO_ADJACENT_WEEK2$target_end_date)$MMWRweek
AUTO_EPIWEEK_WEEK2$epiweek <- MMWRweek(AUTO_EPIWEEK_WEEK2$target_end_date)$MMWRweek
AUTO_TEMPERATURE_WEEK2$epiweek <- MMWRweek(AUTO_TEMPERATURE_WEEK2$target_end_date)$MMWRweek
ES27_ARIMA_WEEK2$epiweek <- MMWRweek(ES27_ARIMA_WEEK2$target_end_date)$MMWRweek
ES27_ADJACENT_WEEK2$epiweek <- MMWRweek(ES27_ADJACENT_WEEK2$target_end_date)$MMWRweek
ES27_EPIWEEK_WEEK2$epiweek <- MMWRweek(ES27_EPIWEEK_WEEK2$target_end_date)$MMWRweek
ES27_TEMPERATURE_WEEK2$epiweek <- MMWRweek(ES27_TEMPERATURE_WEEK2$target_end_date)$MMWRweek
ES64_ARIMA_WEEK2$epiweek <- MMWRweek(ES64_ARIMA_WEEK2$target_end_date)$MMWRweek
ES64_ADJACENT_WEEK2$epiweek <- MMWRweek(ES64_ADJACENT_WEEK2$target_end_date)$MMWRweek
ES64_EPIWEEK_WEEK2$epiweek <- MMWRweek(ES64_EPIWEEK_WEEK2$target_end_date)$MMWRweek
ES64_TEMPERATURE_WEEK2$epiweek <- MMWRweek(ES64_TEMPERATURE_WEEK2$target_end_date)$MMWRweek

# Dataframe that will be analysed for 2 Weeks Ahead
df_W2 <- data.frame(
  STATE = AUTO_ARIMA_WEEK2$State,
  Julian_date = AUTO_ARIMA_WEEK2$target_end_date,
  epiweek = AUTO_ARIMA_WEEK2$epiweek,
  AUTO_AR=AUTO_ARIMA_WEEK2$WIS, 
  AUTO_ADJ=AUTO_ADJACENT_WEEK2$WIS, 
  AUTO_EPI=AUTO_EPIWEEK_WEEK2$WIS,
  AUTO_TEMP=AUTO_TEMPERATURE_WEEK2$WIS,
  ES27_AR=ES27_ARIMA_WEEK2$WIS, 
  ES27_ADJ=ES27_ADJACENT_WEEK2$WIS, 
  ES27_EPI=ES27_EPIWEEK_WEEK2$WIS,
  ES27_TEMP=ES27_TEMPERATURE_WEEK2$WIS,
  ES64_AR=ES64_ARIMA_WEEK2$WIS, 
  ES64_ADJ=ES64_ADJACENT_WEEK2$WIS, 
  ES64_EPI=ES64_EPIWEEK_WEEK2$WIS,
  ES64_TEMP=ES64_TEMPERATURE_WEEK2$WIS
)

head(df_W2)
```

-------------
3 Weeks ahead
-------------
```{r}
# Creating new columns with Epidemiological weeks based on target_end_week
AUTO_ARIMA_WEEK3$epiweek <- MMWRweek(AUTO_ARIMA_WEEK3$target_end_date)$MMWRweek
AUTO_ADJACENT_WEEK3$epiweek <- MMWRweek(AUTO_ADJACENT_WEEK3$target_end_date)$MMWRweek
AUTO_EPIWEEK_WEEK3$epiweek <- MMWRweek(AUTO_EPIWEEK_WEEK3$target_end_date)$MMWRweek
AUTO_TEMPERATURE_WEEK3$epiweek <- MMWRweek(AUTO_TEMPERATURE_WEEK3$target_end_date)$MMWRweek
ES27_ARIMA_WEEK3$epiweek <- MMWRweek(ES27_ARIMA_WEEK3$target_end_date)$MMWRweek
ES27_ADJACENT_WEEK3$epiweek <- MMWRweek(ES27_ADJACENT_WEEK3$target_end_date)$MMWRweek
ES27_EPIWEEK_WEEK3$epiweek <- MMWRweek(ES27_EPIWEEK_WEEK3$target_end_date)$MMWRweek
ES27_TEMPERATURE_WEEK3$epiweek <- MMWRweek(ES27_TEMPERATURE_WEEK3$target_end_date)$MMWRweek
ES64_ARIMA_WEEK3$epiweek <- MMWRweek(ES64_ARIMA_WEEK3$target_end_date)$MMWRweek
ES64_ADJACENT_WEEK3$epiweek <- MMWRweek(ES64_ADJACENT_WEEK3$target_end_date)$MMWRweek
ES64_EPIWEEK_WEEK3$epiweek <- MMWRweek(ES64_EPIWEEK_WEEK3$target_end_date)$MMWRweek
ES64_TEMPERATURE_WEEK3$epiweek <- MMWRweek(ES64_TEMPERATURE_WEEK3$target_end_date)$MMWRweek

# Dataframe that will be analysed for 3 Weeks Ahead
df_W3 <- data.frame(
  STATE = AUTO_ARIMA_WEEK3$State,
  Julian_date = AUTO_ARIMA_WEEK3$target_end_date,
  epiweek = AUTO_ARIMA_WEEK3$epiweek,
  AUTO_AR=AUTO_ARIMA_WEEK3$WIS, 
  AUTO_ADJ=AUTO_ADJACENT_WEEK3$WIS, 
  AUTO_EPI=AUTO_EPIWEEK_WEEK3$WIS,
  AUTO_TEMP=AUTO_TEMPERATURE_WEEK3$WIS,
  ES27_AR=ES27_ARIMA_WEEK3$WIS, 
  ES27_ADJ=ES27_ADJACENT_WEEK3$WIS, 
  ES27_EPI=ES27_EPIWEEK_WEEK3$WIS,
  ES27_TEMP=ES27_TEMPERATURE_WEEK3$WIS,
  ES64_AR=ES64_ARIMA_WEEK3$WIS, 
  ES64_ADJ=ES64_ADJACENT_WEEK3$WIS, 
  ES64_EPI=ES64_EPIWEEK_WEEK3$WIS,
  ES64_TEMP=ES64_TEMPERATURE_WEEK3$WIS
)

head(df_W3)
```

-------------
4 Weeks ahead
-------------
```{r}
# Creating new columns with Epidemiological weeks based on target_end_week
AUTO_ARIMA_WEEK4$epiweek <- MMWRweek(AUTO_ARIMA_WEEK4$target_end_date)$MMWRweek
AUTO_ADJACENT_WEEK4$epiweek <- MMWRweek(AUTO_ADJACENT_WEEK4$target_end_date)$MMWRweek
AUTO_EPIWEEK_WEEK4$epiweek <- MMWRweek(AUTO_EPIWEEK_WEEK4$target_end_date)$MMWRweek
AUTO_TEMPERATURE_WEEK4$epiweek <- MMWRweek(AUTO_TEMPERATURE_WEEK4$target_end_date)$MMWRweek
ES27_ARIMA_WEEK4$epiweek <- MMWRweek(ES27_ARIMA_WEEK4$target_end_date)$MMWRweek
ES27_ADJACENT_WEEK4$epiweek <- MMWRweek(ES27_ADJACENT_WEEK4$target_end_date)$MMWRweek
ES27_EPIWEEK_WEEK4$epiweek <- MMWRweek(ES27_EPIWEEK_WEEK4$target_end_date)$MMWRweek
ES27_TEMPERATURE_WEEK4$epiweek <- MMWRweek(ES27_TEMPERATURE_WEEK4$target_end_date)$MMWRweek
ES64_ARIMA_WEEK4$epiweek <- MMWRweek(ES64_ARIMA_WEEK4$target_end_date)$MMWRweek
ES64_ADJACENT_WEEK4$epiweek <- MMWRweek(ES64_ADJACENT_WEEK4$target_end_date)$MMWRweek
ES64_EPIWEEK_WEEK4$epiweek <- MMWRweek(ES64_EPIWEEK_WEEK4$target_end_date)$MMWRweek
ES64_TEMPERATURE_WEEK4$epiweek <- MMWRweek(ES64_TEMPERATURE_WEEK4$target_end_date)$MMWRweek

# Dataframe that will be analysed for 4 Weeks Ahead
df_W4 <- data.frame(
  STATE = AUTO_ARIMA_WEEK4$State,
  Julian_date = AUTO_ARIMA_WEEK4$target_end_date,
  epiweek = AUTO_ARIMA_WEEK4$epiweek,
  AUTO_AR=AUTO_ARIMA_WEEK4$WIS, 
  AUTO_ADJ=AUTO_ADJACENT_WEEK4$WIS, 
  AUTO_EPI=AUTO_EPIWEEK_WEEK4$WIS,
  AUTO_TEMP=AUTO_TEMPERATURE_WEEK4$WIS,
  ES27_AR=ES27_ARIMA_WEEK4$WIS, 
  ES27_ADJ=ES27_ADJACENT_WEEK4$WIS, 
  ES27_EPI=ES27_EPIWEEK_WEEK4$WIS,
  ES27_TEMP=ES27_TEMPERATURE_WEEK4$WIS,
  ES64_AR=ES64_ARIMA_WEEK4$WIS, 
  ES64_ADJ=ES64_ADJACENT_WEEK4$WIS, 
  ES64_EPI=ES64_EPIWEEK_WEEK4$WIS,
  ES64_TEMP=ES64_TEMPERATURE_WEEK4$WIS
)

head(df_W4)
```

Filter only the flu season

```{r}
# Filter the dataframe for epiweek >= 40 or epiweek <= 20
filtered_df_W1 <- df_W1 %>%
  filter(epiweek >= 40 | epiweek <= 20)

# Display the filtered dataset
head(filtered_df_W1)

# Filter the dataframe for epiweek >= 40 or epiweek <= 20
filtered_df_W2 <- df_W2 %>%
  filter(epiweek >= 40 | epiweek <= 20)

# Display the filtered dataset
head(filtered_df_W2)

# Filter the dataframe for epiweek >= 40 or epiweek <= 20
filtered_df_W3 <- df_W3 %>%
  filter(epiweek >= 40 | epiweek <= 20)

# Display the filtered dataset
head(filtered_df_W3)

# Filter the dataframe for epiweek >= 40 or epiweek <= 20
filtered_df_W4 <- df_W4 %>%
  filter(epiweek >= 40 | epiweek <= 20)

# Display the filtered dataset
head(filtered_df_W1)
```

Calculate mean weighted interval score for all forecast weeks on each model.

```{r}

# Define the function
calculate_mean_wis <- function(data) {
  data %>%
    group_by(STATE) %>%
    summarize(
      AUTO_AR = mean(AUTO_AR, na.rm = TRUE),
      AUTO_ADJ = mean(AUTO_ADJ, na.rm = TRUE),
      AUTO_EPI = mean(AUTO_EPI, na.rm = TRUE),
      AUTO_TMP = mean(AUTO_TEMP, na.rm = TRUE),
      ES27_AR = mean(ES27_AR, na.rm = TRUE),
      ES27_ADJ = mean(ES27_ADJ, na.rm = TRUE),
      ES27_EPI = mean(ES27_EPI, na.rm = TRUE),
      ES27_TMP = mean(ES27_TEMP, na.rm = TRUE),
      ES64_AR = mean(ES64_AR, na.rm = TRUE),
      ES64_ADJ = mean(ES64_ADJ, na.rm = TRUE),
      ES64_EPI = mean(ES64_EPI, na.rm = TRUE),
      ES64_TMP = mean(ES64_TEMP, na.rm = TRUE)
    )
}

# Now you can use the function with any dataframe
W1 <- calculate_mean_wis(filtered_df_W1)
W2 <- calculate_mean_wis(filtered_df_W2)
W3 <- calculate_mean_wis(filtered_df_W3)
W4 <- calculate_mean_wis(filtered_df_W4)

# Display the resulting dataframe
head(W1)
head(W2)
head(W3)
head(W4)
```

Here we have boxplots of the mean(WIS) by each state.

```{r}
# Combine the data into one data frame
combined_data <- data.frame(
  week = rep(c("W1", "W2", "W3", "W4"), each = 47 * 12),

  variable = rep(c("AUTO_AR", "ES27_AR", "ES64_AR", "AUTO_ADJ", "ES27_ADJ", "ES64_ADJ", "AUTO_TMP", "ES27_TMP", "ES64_TMP", "AUTO_EPI", "ES27_EPI", "ES64_EPI"), each=47, times = 4),

  value = c(W1$AUTO_AR, W1$ES27_AR, W1$ES64_AR, W1$AUTO_ADJ, W1$ES27_ADJ, W1$ES64_ADJ, W1$AUTO_TMP, W1$ES27_TMP, W1$ES64_TMP, W1$AUTO_EPI, W1$ES27_EPI, W1$ES64_EPI,
            W2$AUTO_AR, W2$ES27_AR, W2$ES64_AR, W2$AUTO_ADJ, W2$ES27_ADJ, W2$ES64_ADJ, W2$AUTO_TMP, W2$ES27_TMP, W2$ES64_TMP, W2$AUTO_EPI, W2$ES27_EPI, W2$ES64_EPI,
            W3$AUTO_AR, W3$ES27_AR, W3$ES64_AR, W3$AUTO_ADJ, W3$ES27_ADJ, W3$ES64_ADJ, W3$AUTO_TMP, W3$ES27_TMP, W3$ES64_TMP, W3$AUTO_EPI, W3$ES27_EPI, W3$ES64_EPI,
            W4$AUTO_AR, W4$ES27_AR, W4$ES64_AR, W4$AUTO_ADJ, W4$ES27_ADJ, W4$ES64_ADJ, W4$AUTO_TMP, W4$ES27_TMP, W4$ES64_TMP, W4$AUTO_EPI, W4$ES27_EPI, W4$ES64_EPI)
)

# Create a new column for categories
combined_data <- combined_data %>%
  mutate(category = case_when(
    grepl("AR", variable) ~ "AR",
    grepl("EPI", variable) ~ "EPI",
    grepl("TMP", variable) ~ "TMP",
    grepl("ADJ", variable) ~ "ADJ"
  ))

# Define color ramps for each category
colors_ar <- c("#a6cee3","#226e83", "#08306b")
colors_adj <- c("#fb9a99","red3", "#a11c3e")
colors_epi <- c("#cab2d6", "purple2", "#5e2b7b")
colors_tmp <- c("lightgreen","green3","#319045")

# Create a custom color mapping for each variable
custom_colors <- c(
  "AUTO_AR" = colors_ar[1], "ES27_AR" = colors_ar[2], "ES64_AR" = colors_ar[3],
  "AUTO_EPI" = colors_epi[1], "ES27_EPI" = colors_epi[2], "ES64_EPI" = colors_epi[3],
  "AUTO_TMP" = colors_tmp[1], "ES27_TMP" = colors_tmp[2], "ES64_TMP" = colors_tmp[3],
  "AUTO_ADJ" = colors_adj[1], "ES27_ADJ" = colors_adj[2], "ES64_ADJ" = colors_adj[3]
)

  # Create the box plot with additional facet for categories
  ggplot(combined_data, aes(x = week, y = value, fill = variable)) +
    geom_boxplot(color = "black") +
    scale_fill_manual(values = custom_colors) +
    facet_wrap(category ~ variable , nrow = 2) +
    labs(
      title = "Models' mean(WIS) for 47 U.S. states by weeks ahead",
      x = "",
      y = "Weighted Interval Score (WIS)"
    ) +
    theme_minimal()

  # Create the box plot with additional facet for categories
  ggplot(combined_data, aes(x = week, y = log(value), fill = variable)) +
    geom_boxplot(color = "black") +
    scale_fill_manual(values = custom_colors) +
    facet_wrap(category ~ variable , nrow = 2) +
    labs(
      title = "Models' mean(WIS) for 47 U.S. states by weeks ahead on a log-scale",
      x = "",
      y = "log(WIS)"
    ) +
    theme_minimal()

```

Here I include a column with the best model result based on the lowest mean(WIS) of each state.

```{r}
# BEST RESULT

# Extract the columns of interest 
cols <- colnames(W1)[-1]
# Initialize a vector to store results
W1$Best_Result <- character(nrow(W1))

# Give me the model with lower WIS value in the best result column
for (i in 1:nrow(W1)) {
  # Find the column name with the minimum value for each row
  W1$Best_Result[i] <- cols[which.min(W1[i, cols])]
}

# REORDER BY FREQUENCY
W1$Best_Result <- fct_infreq(W1$Best_Result)

# Print merged results
head(W1)

################################################

# Extract the columns of interest 
cols <- colnames(W2)[-1]
# Initialize a vector to store results
W2$Best_Result <- character(nrow(W2))
# Give me the model with lower WIS value in the best result column
for (i in 1:nrow(W2)) {
  # Find the column name with the minimum value for each row
  W2$Best_Result[i] <- cols[which.min(W2[i, cols])]
}

# REORDER BY FREQUENCY
W2$Best_Result <- fct_infreq(W2$Best_Result)

# Print merged results
head(W2)

######################################
# BEST RESULT
# Extract the columns of interest 
cols <- colnames(W3)[-1]
# Initialize a vector to store results
W3$Best_Result <- character(nrow(W3))

# Give me the model with lower WIS value in the best result column
for (i in 1:nrow(W3)) {
  # Find the column name with the minimum value for each row
  W3$Best_Result[i] <- cols[which.min(W3[i, cols])]
}

# REORDER BY FREQUENCY
W3$Best_Result <- fct_infreq(W3$Best_Result)

# Print merged results
head(W3)

###################################
# BEST RESULT
# Extract the columns of interest
cols <- colnames(W4)[-1]
# Initialize a vector to store results
W4$Best_Result <- character(nrow(W4))
# Give me the model with lower WIS value in the best result column
for (i in 1:nrow(W4)) {
  # Find the column name with the minimum value for each row
  W4$Best_Result[i] <- cols[which.min(W4[i, cols])]
}
W4$Best_Result <- fct_infreq(W4$Best_Result)

# Print merged results
head(W4)

```

Now, let's plot the best models by each state for each forecast horizon (1-4 weeks ahead).

```{r}
# --------- WEEK1 MODELS ------------- #
ggplot(W1, aes(x = Best_Result, fill = Best_Result)) +
  geom_bar() +
  scale_fill_manual(values = custom_colors) +
  labs(title = "Best Model based on mean(WIS) for 47 U.S. states (1 Week Ahead)",
       x = "", y = "Number of states") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# --------- WEEK2 MODELS ------------- #
ggplot(W2, aes(x = Best_Result, fill = Best_Result)) +
  geom_bar() +
  scale_fill_manual(values = custom_colors) +
  labs(title = "Best Model based on mean(WIS) for 47 U.S. states (2 Weeks Ahead)",
       x = "", y = "Number of states") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# --------- WEEK3 MODELS ------------- #
ggplot(W3, aes(x = Best_Result, fill = Best_Result)) +
  geom_bar() +
  scale_fill_manual(values = custom_colors) +
  labs(title = "Best Model based on mean(WIS) for 47 U.S. states (3 Weeks Ahead)",
       x = "", y = "Number of states") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# --------- WEEK4 MODELS ------------- #
ggplot(W4, aes(x = Best_Result, fill = Best_Result)) +
  geom_bar() +
  scale_fill_manual(values = custom_colors) +
  labs(title = "Best Model based on mean(WIS) for 47 U.S. states (4 Weeks Ahead)",
       x = "", y = "Number of states") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

This plot combines the 4 forecast horizons and the counts of best models in a single plot.

```{r}
models <- readr::read_csv("models.csv") %>% 
  mutate(ex.var = factor(ex.var,
                         levels = c("none",
                                    "mean of ILI in adjacent states in previous week",
                                    "temperature",
                                    "mean of ILI in epiweek - 52 and ILI in epiweek - 104")
                         )
         )

thegrid <- expand_grid(model = models$model, target=paste0("W",1:4))

BestResults <- W1 %>% select(STATE, W1 = Best_Result) %>% 
  full_join(W2 %>% select(STATE, W2 = Best_Result)) %>%  
  full_join(W3 %>% select(STATE, W3 = Best_Result)) %>%  
  full_join(W4 %>% select(STATE, W4 = Best_Result)) %>% 
  pivot_longer(-STATE, names_to = "target", values_to = "Best_Result") %>% 
  group_by(target, Best_Result) %>% 
  summarise(n_best = n()) %>% 
  rename(model = Best_Result) %>% 
  ungroup() %>% 
  # complete(target, model) %>% 
  full_join(thegrid, by = c("target", "model")) %>% 
  full_join(models) %>% 
  replace_na(list(n_best = 0))

g <- BestResults %>% 
  arrange(ex.var) %>% 
  ggplot(aes(x=factor(model,
                      levels = models %>% arrange(ensemble.size, ex.var) %>% pull(model)), 
             y=n_best, 
             group=target, 
             fill=ex.var,
             alpha=target)) + 
  geom_bar(
           position="dodge",
           stat='identity') +

  labs(title = "Best Model based on WIS for 47 U.S. states (1 to 4 Weeks Ahead)",
       x = "", y="Number of states for which model is best") +
  theme_bw() +
  scale_alpha_manual(values=c(.4,.6,.8,1), na.translate = FALSE) +
  scale_fill_manual(values=c("#08306b","red3","#319045","#5e2b7b"),
                    na.translate = FALSE) +
  theme(panel.grid.major.x = element_blank(),
        axis.ticks.x = element_blank(), 
        axis.text.x = element_text(angle = 45, size = 7, vjust = 0.5, hjust=1),
        strip.placement = "outside", 
        strip.background = element_rect(fill=NA,colour="grey50"),
        panel.spacing=unit(0,"cm"),
        legend.position="none")  # This line removes the legend
g

#g %>% plotly::ggplotly()
```

Now let's plot maps with the best models we found for each state based on the best mean(WIS).

```{r}
#################################################
# ES27 ARIMAX by ADJACENT STATES - 1 WEEK AHEAD #
#################################################

category_colors <- c(
  "AUTO_AR" = "#a6cee3", "ES27_AR" = "#226e83", "ES64_AR" = "#08306b",
  "AUTO_ADJ" = "#fb9a99", "ES27_ADJ" = "red3", "ES64_ADJ" = "#a11c3e",
  "AUTO_TMP" = "lightgreen", "ES27_TMP" = "green3", "ES64_TMP" = "#319045",
  "AUTO_EPI" = "#cab2d6", "ES27_EPI" = "purple2", "ES64_EPI" = "#5e2b7b"
)

map_week1<-left_join(states, W1, by=join_by("STATE"))%>%
  drop_na()

ES_1WEEK<- ggplot(map_week1, fill ="lightgrey") +  theme_light()  + geom_sf(aes(fill=Best_Result)) +  scale_fill_manual(values = category_colors) +  ggtitle("Best model based on mean(WIS) (1 Week Ahead)")

x_limits <- c(-130, -65)  # Set the desired longitude range
y_limits <- c(20, 55)    # Set the desired latitude range

ES_1WEEK + coord_sf(xlim = x_limits, ylim = y_limits)

###################################################
# ES27 ARIMAX by ADJACENT STATES - 2 WEEKS AHEAD #
#################################################

map_week2<-left_join(states, W2, by=join_by("STATE"))%>%
  drop_na()

MAP_WEEK2<- ggplot(map_week2, fill ="lightgrey") +  theme_light()  + geom_sf(aes(fill=Best_Result)) +  scale_fill_manual(values = category_colors) +  ggtitle("Best model based on mean(WIS) (2 Weeks Aheads)")

x_limits <- c(-130, -65)  # Set the desired longitude range
y_limits <- c(20, 55)    # Set the desired latitude range

MAP_WEEK2 + coord_sf(xlim = x_limits, ylim = y_limits)

###################################################
# ES27 ARIMAX by ADJACENT STATES - 3 WEEKS AHEAD #
#################################################

map_week3<-left_join(states, W3, by=join_by("STATE"))%>%
  drop_na()

MAP_WEEK3<- ggplot(map_week3, fill ="lightgrey") +  theme_light()  + geom_sf(aes(fill=Best_Result)) +  scale_fill_manual(values = category_colors) +  ggtitle("Best model based on mean(WIS) (3 Weeks Aheads)")

x_limits <- c(-130, -65)  # Set the desired longitude range
y_limits <- c(20, 55)    # Set the desired latitude range

MAP_WEEK3 + coord_sf(xlim = x_limits, ylim = y_limits)

###################################################
# ES27 ARIMAX by ADJACENT STATES - 4 WEEKS AHEAD #
#################################################

map_week4<-left_join(states, W4, by=join_by("STATE"))%>%
  drop_na()

MAP_WEEK4<- ggplot(map_week4, fill ="lightgrey") +  theme_light()  + geom_sf(aes(fill=Best_Result)) +  scale_fill_manual(values = category_colors) +  ggtitle("Best model based on mean(WIS) (4 Weeks Aheads)")

x_limits <- c(-130, -65)  # Set the desired longitude range
y_limits <- c(20, 55)    # Set the desired latitude range

MAP_WEEK4 + coord_sf(xlim = x_limits, ylim = y_limits) 
```

Now, let's map the mean(WIS) for the AUTO ARIMA model as an example.

```{r}

# Calculate breaks dynamically or use a fixed range
data_range <- c(0,400)# 
breaks <- pretty(data_range, n = 6)  # breaks
labels <- breaks  #

# Define color palette
colors <- colorRampPalette(c("#08306b", "white", "red3"))(100)

MAP_WEEK1 <- ggplot(map_week1) + 
    theme_light() + 
    geom_sf(aes(fill = (AUTO_AR)), color = "black") + 
    scale_fill_gradientn(
        "mean(WIS)",
        colors = colors,
        breaks = breaks,
        labels = labels,
        limits = data_range,
        na.value = "lightgrey"
    ) + 
    ggtitle("MEAN WIS AUTO ARIMA (1 Wk)") + 
    labs(subtitle = "Prevalence: ILI Cases/State Population") +  # Add your subtitle here
    coord_sf(xlim = c(-130, -65), ylim = c(20, 55)) + 
    theme(legend.position = "right")

# Print the map
print(MAP_WEEK1)

# Create the map
MAP_WEEK2 <- ggplot(map_week2) + 
  theme_light() + 
  geom_sf(aes(fill = AUTO_AR), color = "black") + 
  scale_fill_gradientn(
    "mean(WIS)", 
    colors = colors,
    breaks = breaks,
    labels = labels,
    limits = data_range,
    na.value = "lightgrey"  
  ) + 
    ggtitle("MEAN WIS AUTO ARIMA (2 Wk)") + 
    labs(subtitle = "Prevalence: ILI Cases/State Population") +  # Add your subtitle here
  coord_sf(xlim = c(-130, -65), ylim = c(20, 55)) + 
  theme(legend.position = "right")

# Print the map
print(MAP_WEEK2)

################################

MAP_WEEK3 <- ggplot(map_week3) + 
  theme_light() + 
  geom_sf(aes(fill = AUTO_AR), color = "black") + 
  scale_fill_gradientn(
    "mean(WIS)", 
    colors = colors,
    breaks = breaks,
    labels = labels,
    limits = data_range,
    na.value = "lightgrey"  
  ) + 
    ggtitle("MEAN WIS AUTO ARIMA (3 Wk)") + 
  labs(subtitle = "Prevalence: ILI Cases/State Population") +  # Add your subtitle here
  coord_sf(xlim = c(-130, -65), ylim = c(20, 55)) + 
  theme(legend.position = "right")

# Print the map
print(MAP_WEEK3)

#######################
MAP_WEEK4 <- ggplot(map_week4) + 
  theme_light() + 
  geom_sf(aes(fill = AUTO_AR), color = "black") + 
  scale_fill_gradientn(
    "mean(WIS)", 
    colors = colors,
    breaks = breaks,
    labels = labels,
    limits = data_range,
    na.value = "lightgrey"  
  ) + 
  ggtitle("MEAN WIS AUTO ARIMA (4 Wk)") + 
  labs(subtitle = "Prevalence: ILI Cases/State Population") +  # Add your subtitle here
  coord_sf(xlim = c(-130, -65), ylim = c(20, 55)) + 
  theme(legend.position = "right")

# Print the map
print(MAP_WEEK4)

```

Let's plot the same map on a log scale for better visualization.

```{r}

# Calculate breaks dynamically or use a fixed range
data_range <- c(0,7)# 
breaks <- pretty(data_range, n = 6)  # breaks
labels <- breaks  #

# Define color palette
colors <- colorRampPalette(c("#08306b", "white", "red3"))(100)

MAP_WEEK1 <- ggplot(map_week1) + 
    theme_light() + 
    geom_sf(aes(fill =log(AUTO_AR)), color = "black") + 
    scale_fill_gradientn(
        "log(MEAN WIS)",
        colors = colors,
        breaks = breaks,
        labels = labels,
        limits = data_range,
        na.value = "lightgrey"
    ) + 
    ggtitle("log(MEAN WIS) AUTO ARIMA (1 Wk)") + 
    labs(subtitle = "Prevalence: ILI Cases/State Population") +  # Add your subtitle here
    coord_sf(xlim = c(-130, -65), ylim = c(20, 55)) + 
    theme(legend.position = "right")

# Print the map
print(MAP_WEEK1)

# Create the map
MAP_WEEK2 <- ggplot(map_week2) + 
  theme_light() + 
  geom_sf(aes(fill = log(AUTO_AR)), color = "black") + 
  scale_fill_gradientn(
    "log(MEAN WIS)", 
    colors = colors,
    breaks = breaks,
    labels = labels,
    limits = data_range,
    na.value = "lightgrey"  
  ) + 
    ggtitle("log(MEAN WIS) AUTO ARIMA (2 Wk)") + 
    labs(subtitle = "Prevalence: ILI Cases/State Population") +  # Add your subtitle here
  coord_sf(xlim = c(-130, -65), ylim = c(20, 55)) + 
  theme(legend.position = "right")

# Print the map
print(MAP_WEEK2)

################################

MAP_WEEK3 <- ggplot(map_week3) + 
  theme_light() + 
  geom_sf(aes(fill = log(AUTO_AR)), color = "black") + 
  scale_fill_gradientn(
    "log(MEAN WIS)", 
    colors = colors,
    breaks = breaks,
    labels = labels,
    limits = data_range,
    na.value = "lightgrey"  
  ) + 
    ggtitle("log(MEAN WIS) AUTO ARIMA (3 Wk)") + 
  labs(subtitle = "Prevalence: ILI Cases/State Population") +  # Add your subtitle here
  coord_sf(xlim = c(-130, -65), ylim = c(20, 55)) + 
  theme(legend.position = "right")

# Print the map
print(MAP_WEEK3)

#######################
MAP_WEEK4 <- ggplot(map_week4) + 
  theme_light() + 
  geom_sf(aes(fill = log(AUTO_AR)), color = "black") + 
  scale_fill_gradientn(
    "log(MEAN WIS)", 
    colors = colors,
    breaks = breaks,
    labels = labels,
    limits = data_range,
    na.value = "lightgrey"  
  ) + 
  ggtitle("log(MEAN WIS) AUTO ARIMA (4 Wk)") + 
  labs(subtitle = "Prevalence: ILI Cases/State Population") +  # Add your subtitle here
  coord_sf(xlim = c(-130, -65), ylim = c(20, 55)) + 
  theme(legend.position = "right")

# Print the map
print(MAP_WEEK4)

```

Let's calculate the mean(WIS) improvement relative to the AUTO ARIMA model. We will compare each model with the AUTO ARIMA model for the same states. Later we sum the results of these comparisons. Positive results indicate that there was a general improvement in the mean(WIS) for a given model type among all states. Note that larger values on specific states may override a general trend and make the results less informative. So I do not strongly advice this analysis.

```{r}
calculate_percentage_of_improvement <- function(data) {
  return(data.frame(
    AUTO_AR = (data$AUTO_AR - data$AUTO_AR),
    ES27_AR = ((data$AUTO_AR - data$ES27_AR) / data$AUTO_AR) * 100,
    ES64_AR = ((data$AUTO_AR - data$ES64_AR) / data$AUTO_AR) * 100,
    AUTO_ADJ = ((data$AUTO_AR - data$AUTO_ADJ) / data$AUTO_AR) * 100,
    ES27_ADJ = ((data$AUTO_AR - data$ES27_ADJ) / data$AUTO_AR) * 100,
    ES64_ADJ = ((data$AUTO_AR - data$ES64_ADJ) / data$AUTO_AR) * 100,
    AUTO_TMP = ((data$AUTO_AR - data$AUTO_TMP) / data$AUTO_AR) * 100,
    ES27_TMP = ((data$AUTO_AR - data$ES27_TMP) / data$AUTO_AR) * 100,
    ES64_TMP = ((data$AUTO_AR - data$ES64_TMP) / data$AUTO_AR) * 100,
    AUTO_EPI = ((data$AUTO_AR - data$AUTO_EPI) / data$AUTO_AR) * 100,
    ES27_EPI = ((data$AUTO_AR - data$ES27_EPI) / data$AUTO_AR) * 100,
    ES64_EPI = ((data$AUTO_AR - data$ES64_EPI) / data$AUTO_AR) * 100
  ))
}

# Calculate percentage of improvemtn 
W1_percentage_of_improvement <- calculate_percentage_of_improvement(W1)
W2_percentage_of_improvement <- calculate_percentage_of_improvement(W2)
W3_percentage_of_improvement <- calculate_percentage_of_improvement(W3)
W4_percentage_of_improvement <- calculate_percentage_of_improvement(W4)
```

Now let's plot a map of percentage of WIS improvement for each state.

```{r}
# MAP 1 week ahead

W1_map <- data.frame(
  STATE = W1$STATE,
  best_model = W1$Best_Result,
  percentage_improvement = apply(W1_percentage_of_improvement, 1, max)
)

# Merge the best model data with the states map
W1_map <- states %>%
  left_join(W1_map, by = c("STATE")) %>%
  filter(!is.na(percentage_improvement))

################################### MAP 2 weeka ahead

W2_map <- data.frame(
  STATE = W2$STATE,
  best_model = W2$Best_Result,
  percentage_improvement = apply(W2_percentage_of_improvement, 1, max)
)

# Merge the best model data with the states map
W2_map <- states %>%
  left_join(W2_map, by = c("STATE")) %>%
  filter(!is.na(percentage_improvement))

##################################### MAP 3 weeka ahead

W3_map <- data.frame(
  STATE = W3$STATE,
  best_model = W3$Best_Result,
  percentage_improvement = apply(W3_percentage_of_improvement, 1, max)
)

# Merge the best model data with the states map
W3_map <- states %>%
  left_join(W3_map, by = c("STATE")) %>%
  filter(!is.na(percentage_improvement))

############################### MAP 4 weeka ahead

W4_map <- data.frame(
  STATE = W4$STATE,
  best_model = W4$Best_Result,
  percentage_improvement = apply(W4_percentage_of_improvement, 1, max)
)

# Merge the best model data with the states map
W4_map <- states %>%
  left_join(W4_map, by = c("STATE")) %>%
  filter(!is.na(percentage_improvement))
```

Let's plot some maps with the best models in each state and the percentage of improvement compared to the AUTO ARIMA models for the same state.

Here I define some colors categories that I will use on the next maps.

```{r}

# Define colors for the modelss
category_colors <- c(
  "AUTO_AR" = "#a6cee3",
  "ES27_AR" = "#226e83",
  "ES64_AR" = "#08306b",
  "AUTO_ADJ" = "#fb9a99",
  "ES27_ADJ" = "red3",
  "ES64_ADJ" = "#a11c3e",
  "AUTO_TMP" = "lightgreen",
  "ES27_TMP" = "green3",
  "ES64_TMP" = "#319045",
  "AUTO_EPI" = "#cab2d6",
  "ES27_EPI" = "purple2",
  "ES64_EPI" = "#5e2b7b"
)
```

1 week ahead percentage of improvement

```{r}
# Assuming merged_data already includes the necessary columns: model and percentage_improveme
# Create the map plot for 1 Week Ahead with best models and percentage improvement
ES_1WEEK <- ggplot(W1_map) +
  geom_sf(aes(fill = best_model)) +  # Fill based on the best model
  scale_fill_manual(values = category_colors) +
  ggtitle("Best models and % of improvement compared to AUTO ARIMA (Wk1)") +
  theme_light() +
  theme(legend.position = "right") +
  geom_sf_text(data = W1_map, aes(label = round(percentage_improvement,1)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE)+  # Display percentage improvement in each state
  labs(fill = "Model")  # Label for the legend

x_limits <- c(-125, -67)  # Set the desired longitude range
y_limits <- c(25, 50)    # Set the desired latitude range

ES_1WEEK + coord_sf(xlim = x_limits, ylim = y_limits)

```

Histogram of the percentage of improvement for all states - 1 week ahead

```{r}
# Assuming your data is in W4_map$percentage_improvement
# Create a new column to classify positive and negative values
W1_map$category <- ifelse(W1_map$percentage_improvement > 0, "Improved WIS", "Not Improved WIS")

# Create the ggplot histogram
ggplot(W1_map, aes(x = percentage_improvement, fill = category)) +
  geom_histogram(binwidth = 5, color = "white", boundary = 0) +  # Set bin width to 5
  scale_fill_manual(values = c("Improved WIS" = "blue", "Not Improved WIS" = "red")) +  # Color mapping
  scale_x_continuous(breaks = seq(-5, 35, by = 5), limits = c(-5, 35)) +  # Set x-axis breaks
  labs(title = "Mean WIS improvement % by each state compared to AUTO ARIMA (Wk1)",
       x = "Improvement % = (WIS AUTO_ARIMA - WIS BEST MODEL) / WIS AUTO_ARIMA)*100",
       y = "Frequency", fill=NULL) +
  theme_minimal()
```

2 weeks ahead

```{r}

# Assuming merged_data already includes the necessary columns: model and percentage_improveme
# Create the map plot for 1 Week Ahead with best models and percentage improvement
ES_2WEEK <- ggplot(W2_map) +
  geom_sf(aes(fill = best_model)) +  # Fill based on the best model
  scale_fill_manual(values = category_colors) +
  ggtitle("Best models and % of improvement compared to AUTO ARIMA (Wk2)") +
  theme_light() +
  theme(legend.position = "right") +
  geom_sf_text(data = W2_map, aes(label = round(percentage_improvement,1)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE)+  # Display percentage improvement in each state
  labs(fill = "Model")  # Label for the legend

x_limits <- c(-125, -67)  # Set the desired longitude range
y_limits <- c(25, 50)    # Set the desired latitude range

ES_2WEEK + coord_sf(xlim = x_limits, ylim = y_limits)

```

Histogram of the percentage of improvement for all states - 2 weeks ahead

```{r}
# Assuming your data is in W4_map$percentage_improvement
# Create a new column to classify positive and negative values
W2_map$category <- ifelse(W2_map$percentage_improvement > 0, "Improved WIS", "Not Improved WIS")

# Create the ggplot histogram
ggplot(W2_map, aes(x = percentage_improvement, fill = category)) +
  geom_histogram(binwidth = 5, color = "white", boundary = 0) +  # Set bin width to 5
  scale_fill_manual(values = c("Improved WIS" = "blue", "Not Improved WIS" = "red")) +  # Color mapping
  scale_x_continuous(breaks = seq(-5, 35, by = 5), limits = c(-5, 35)) +  # Set x-axis breaks
  labs(title = "Mean WIS improvement % by each state compared to AUTO ARIMA (Wk2)",
       x = "Improvement % = (WIS AUTO_ARIMA - WIS BEST MODEL) / WIS AUTO_ARIMA)*100",
       y = "Frequency", fill=NULL) +
  theme_minimal()
```

3 weeks ahead

```{r}

# Assuming merged_data already includes the necessary columns: model and percentage_improveme
# Create the map plot for 1 Week Ahead with best models and percentage improvement
ES_3WEEK <- ggplot(W3_map) +
  geom_sf(aes(fill = best_model)) +  # Fill based on the best model
  scale_fill_manual(values = category_colors) +
  ggtitle("Best models and % of improvement compared to AUTO ARIMA (Wk3)") +
  theme_light() +
  theme(legend.position = "right") +
  geom_sf_text(data = W3_map, aes(label = round(percentage_improvement,1)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE)+  # Display percentage improvement in each state
  labs(fill = "Model")  # Label for the legend

x_limits <- c(-125, -67)  # Set the desired longitude range
y_limits <- c(25, 50)    # Set the desired latitude range

ES_3WEEK + coord_sf(xlim = x_limits, ylim = y_limits)

```

Histogram of the percentage of improvement for all states - 3 weeks ahead

```{r}
# Assuming your data is in W4_map$percentage_improvement
# Create a new column to classify positive and negative values
W3_map$category <- ifelse(W3_map$percentage_improvement > 0, "Improved WIS", "Not Improved WIS")

# Create the ggplot histogram
ggplot(W3_map, aes(x = percentage_improvement, fill = category)) +
  geom_histogram(binwidth = 5, color = "white", boundary = 0) +  # Set bin width to 5
  scale_fill_manual(values = c("Improved WIS" = "blue", "Not Improved WIS" = "red")) +  # Color mapping
  scale_x_continuous(breaks = seq(-5, 35, by = 5), limits = c(-5, 35)) +  # Set x-axis breaks
  labs(title = "Mean WIS improvement % by each state compared to AUTO ARIMA (Wk3)",
       x = "Improvement % = (WIS AUTO_ARIMA - WIS BEST MODEL) / WIS AUTO_ARIMA)*100",
       y = "Frequency", fill=NULL) +
  theme_minimal()
```

4 weeks ahead

```{r}

# Assuming merged_data already includes the necessary columns: model and percentage_improveme
# Create the map plot for 1 Week Ahead with best models and percentage improvement
ES_4WEEK <- ggplot(W4_map) +
  geom_sf(aes(fill = best_model)) +  # Fill based on the best model
  scale_fill_manual(values = category_colors) +
  ggtitle("Best models and % of improvement compared to AUTO ARIMA (Wk4)") +
  theme_light() +
  theme(legend.position = "right") +
  geom_sf_text(data = W4_map, aes(label = round(percentage_improvement,1)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE)+  # Display percentage improvement in each state
  labs(fill = "Model")  # Label for the legend

x_limits <- c(-125, -67)  # Set the desired longitude range
y_limits <- c(25, 50)    # Set the desired latitude range

ES_4WEEK + coord_sf(xlim = x_limits, ylim = y_limits)

```

Histogram of the percentage of improvement for all states - 4 weeks ahead

```{r}
# Assuming your data is in W4_map$percentage_improvement
# Create a new column to classify positive and negative values
W4_map$category <- ifelse(W4_map$percentage_improvement > 0, "Improved WIS", "Not Improved WIS")

# Create the ggplot histogram
ggplot(W4_map, aes(x = percentage_improvement, fill = category)) +
  geom_histogram(binwidth = 5, color = "white", boundary = 0) +  # Set bin width to 5
  scale_fill_manual(values = c("Improved WIS" = "blue", "Not Improved WIS" = "red")) +  # Color mapping
  scale_x_continuous(breaks = seq(-5, 35, by = 5), limits = c(-5, 35)) +  # Set x-axis breaks
  labs(title = "Mean WIS improvement % by each state compared to AUTO ARIMA (Wk4)",
       x = "Improvement % = (WIS AUTO_ARIMA - WIS BEST MODEL) / WIS AUTO_ARIMA)*100",
       y = "Frequency", fill=NULL) +
  theme_minimal()
```

Now let's evaluate which are the best models by each epidemiological week
Here I define some colors and model order for the next plots.

```{r}
# Define a color palette for models
model_colors <- c(
  "AUTO_AR" = "#a6cee3",  # Blue
  "ES27_AR" = "#226e83",  # Light blue
  "ES64_AR" = "#08306b",  # Green
  "AUTO_ADJ" = "#fb9a99", # Red
  "ES27_ADJ" = "red3", # Light red #fb9a99
  "ES64_ADJ" = "#a11c3e", # Orange
  "AUTO_EPI" = "#cab2d6", # Yellow
  "ES27_EPI" = "purple2", # Lavender
  "ES64_EPI" = "#5e2b7b", # Purple
  "AUTO_TEMP" = "lightgreen",# Light orange
  "ES27_TEMP" = "green3", # Light green
  "ES64_TEMP" = "#319045" # Pale yellow
)

# Define the model order
model_order <- c(
  "AUTO_AR", "ES27_AR", "ES64_AR",
  "AUTO_ADJ", "ES27_ADJ", "ES64_ADJ",
  "AUTO_EPI", "ES27_EPI", "ES64_EPI",
  "AUTO_TEMP", "ES27_TEMP", "ES64_TEMP"
)
```

1 week ahead

```{r}
# Reshape the dataframe to long format for easier processing
df_long <- filtered_df_W1 %>%
  pivot_longer(cols = -c(STATE, Julian_date, epiweek),  # Keep Julian_date, epiweek, and STATE 
               names_to = "model", values_to = "WIS") %>%
  mutate(model = gsub("_WIS", "", model))  # Remove _WIS suffix from model names

# Find the best model by Julian_date and STATE
best_model_per_state_date <- df_long %>%
  group_by(STATE, Julian_date) %>%
  slice(which.min(WIS)) %>%
  ungroup()

# Count occurrences of each best model by epiweeks across all states
counts_per_week <- best_model_per_state_date %>%
  group_by(epiweek, model) %>%
  summarize(count = n(), .groups = 'drop')

# Convert 'model' to a factor with custom levels
counts_per_week$model <- factor(counts_per_week$model, levels = model_order)

# Reorder epiweeks to start from week 39 to 53, then 1 to 38
epiweek_levels <- c(40:52, 1:20)

counts_per_week <- counts_per_week %>%
  mutate(epiweek = factor(epiweek, levels = epiweek_levels))

# Calculate total counts for each epiweek
total_counts_per_week <- counts_per_week %>%
  group_by(epiweek) %>%
  summarize(total_count = sum(count), .groups = 'drop')

# Join total counts back to counts_per_week and calculate percentages
counts_per_week <- counts_per_week %>%
  left_join(total_counts_per_week, by = "epiweek") %>%
  mutate(percentage = (count / total_count) * 100)
# Plot the best model counts as percentages
ggplot() +
  geom_bar(data = counts_per_week, aes(x = epiweek, y = percentage, fill = model), 
           stat = "identity", position = "stack", width = 0.7) +
  scale_x_discrete(labels = as.character(epiweek_levels)) +
  labs(x = "Epidemiological weeks", 
       y = "Percentage of Best Models", 
       fill = "Model", 
       title = "Percentage of best models by epidemiological week (1Wk)",  
       subtitle = "Results for 47 U.S. states") +
  scale_fill_manual(values = model_colors) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1))

```

2 weeks ahead

```{r}
# Reshape the dataframe to long format for easier processing
df_long <- filtered_df_W2 %>%
  pivot_longer(cols = -c(STATE, Julian_date, epiweek),  # Keep Julian_date, epiweek, and STATE 
               names_to = "model", values_to = "WIS") %>%
  mutate(model = gsub("_WIS", "", model))  # Remove _WIS suffix from model names

# Find the best model by Julian_date and STATE
best_model_per_state_date <- df_long %>%
  group_by(STATE, Julian_date) %>%
  slice(which.min(WIS)) %>%
  ungroup()

# Count occurrences of each best model by epiweeks across all states
counts_per_week <- best_model_per_state_date %>%
  group_by(epiweek, model) %>%
  summarize(count = n(), .groups = 'drop')

# Convert 'model' to a factor with custom levels
counts_per_week$model <- factor(counts_per_week$model, levels = model_order)

# Reorder epiweeks to start from week 39 to 53, then 1 to 38
epiweek_levels <- c(40:52, 1:20)

counts_per_week <- counts_per_week %>%
  mutate(epiweek = factor(epiweek, levels = epiweek_levels))

# Calculate total counts for each epiweek
total_counts_per_week <- counts_per_week %>%
  group_by(epiweek) %>%
  summarize(total_count = sum(count), .groups = 'drop')

# Join total counts back to counts_per_week and calculate percentages
counts_per_week <- counts_per_week %>%
  left_join(total_counts_per_week, by = "epiweek") %>%
  mutate(percentage = (count / total_count) * 100)
# Plot the best model counts as percentages
ggplot() +
  geom_bar(data = counts_per_week, aes(x = epiweek, y = percentage, fill = model), 
           stat = "identity", position = "stack", width = 0.7) +
  scale_x_discrete(labels = as.character(epiweek_levels)) +
  labs(x = "Epidemiological weeks", 
       y = "Percentage of Best Models", 
       fill = "Model", 
       title = "Percentage of best models by epidemiological week (2Wk)",  
       subtitle = "Results for 47 U.S. states") +
  scale_fill_manual(values = model_colors) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1))

```

3 weeks ahead

```{r}
# Reshape the dataframe to long format for easier processing
df_long <- filtered_df_W3 %>%
  pivot_longer(cols = -c(STATE, Julian_date, epiweek),  # Keep Julian_date, epiweek, and STATE 
               names_to = "model", values_to = "WIS") %>%
  mutate(model = gsub("_WIS", "", model))  # Remove _WIS suffix from model names

# Find the best model by Julian_date and STATE
best_model_per_state_date <- df_long %>%
  group_by(STATE, Julian_date) %>%
  slice(which.min(WIS)) %>%
  ungroup()

# Count occurrences of each best model by epiweeks across all states
counts_per_week <- best_model_per_state_date %>%
  group_by(epiweek, model) %>%
  summarize(count = n(), .groups = 'drop')

# Convert 'model' to a factor with custom levels
counts_per_week$model <- factor(counts_per_week$model, levels = model_order)

# Reorder epiweeks to start from week 39 to 53, then 1 to 38
epiweek_levels <- c(40:52, 1:20)

counts_per_week <- counts_per_week %>%
  mutate(epiweek = factor(epiweek, levels = epiweek_levels))

# Calculate total counts for each epiweek
total_counts_per_week <- counts_per_week %>%
  group_by(epiweek) %>%
  summarize(total_count = sum(count), .groups = 'drop')

# Join total counts back to counts_per_week and calculate percentages
counts_per_week <- counts_per_week %>%
  left_join(total_counts_per_week, by = "epiweek") %>%
  mutate(percentage = (count / total_count) * 100)
# Plot the best model counts as percentages
ggplot() +
  geom_bar(data = counts_per_week, aes(x = epiweek, y = percentage, fill = model), 
           stat = "identity", position = "stack", width = 0.7) +
  scale_x_discrete(labels = as.character(epiweek_levels)) +
  labs(x = "Epidemiological weeks", 
       y = "Percentage of Best Models", 
       fill = "Model", 
       title = "Percentage of best models by epidemiological week (3Wk)",  
       subtitle = "Results for 47 U.S. states") +
  scale_fill_manual(values = model_colors) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1))

```

4 weeks ahead

```{r}
# Reshape the dataframe to long format for easier processing
df_long <- filtered_df_W4 %>%
  pivot_longer(cols = -c(STATE, Julian_date, epiweek),  # Keep Julian_date, epiweek, and STATE 
               names_to = "model", values_to = "WIS") %>%
  mutate(model = gsub("_WIS", "", model))  # Remove _WIS suffix from model names

# Find the best model by Julian_date and STATE
best_model_per_state_date <- df_long %>%
  group_by(STATE, Julian_date) %>%
  slice(which.min(WIS)) %>%
  ungroup()

# Count occurrences of each best model by epiweeks across all states
counts_per_week <- best_model_per_state_date %>%
  group_by(epiweek, model) %>%
  summarize(count = n(), .groups = 'drop')

# Convert 'model' to a factor with custom levels
counts_per_week$model <- factor(counts_per_week$model, levels = model_order)

# Reorder epiweeks to start from week 39 to 53, then 1 to 38
epiweek_levels <- c(40:52, 1:20)

counts_per_week <- counts_per_week %>%
  mutate(epiweek = factor(epiweek, levels = epiweek_levels))

# Calculate total counts for each epiweek
total_counts_per_week <- counts_per_week %>%
  group_by(epiweek) %>%
  summarize(total_count = sum(count), .groups = 'drop')

# Join total counts back to counts_per_week and calculate percentages
counts_per_week <- counts_per_week %>%
  left_join(total_counts_per_week, by = "epiweek") %>%
  mutate(percentage = (count / total_count) * 100)
# Plot the best model counts as percentages
ggplot() +
  geom_bar(data = counts_per_week, aes(x = epiweek, y = percentage, fill = model), 
           stat = "identity", position = "stack", width = 0.7) +
  scale_x_discrete(labels = as.character(epiweek_levels)) +
  labs(x = "Epidemiological weeks", 
       y = "Percentage of Best Models", 
       fill = "Model", 
       title = "Percentage of best models by epidemiological week (4Wk)",  
       subtitle = "Results for 47 U.S. states") +
  scale_fill_manual(values = model_colors) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1))

```

Loading datasets for regression analysis

```{r}
pop_data <- read.csv("regression_features/population_data.csv") # resident population and population density
sovi_data <- read.csv("regression_features/sovi_2010.2014.csv") # SOVI index
bric_data<-read.csv("regression_features/bric2015.csv") # BRIC index
humidity_data<-read.csv("regression_features/humidity_climatology_1990_2020.csv") # ERA5 Specific Humidity
temperature_data<-read.csv("regression_features/temperature_climatology_1990_2020.csv") # ERA5 Specific Temperature

```

REGRESSION MODELS

We will run the regression for evaluating if the best model log(WIS) which represent its performance is related to given independent variables.

```{r}

# List of data frames
WIS_dataframes <- list(W1 = W1, W2 = W2, W3 = W3, W4=W4)
regression_models<-data.frame()

```

All Regression results

AUTO_ADJ WIS x RESIDENT POPULATION 2020

```{r}

for(i in c(1,2,3,4)){

  # Combining data for the same states
  colnames(WIS_dataframes[[i]])[1] <- "STATE"
  WIS_dataframes[[i]]$STATE <- as.character(WIS_dataframes[[i]]$STATE)
  WIS_pop_data <- inner_join(WIS_dataframes[[i]], pop_data, by = "STATE")
  # Fitting the regression model
  model <- lm(log1p(WIS_pop_data$AUTO_ADJ) ~ log(WIS_pop_data$Resident_population_2020))
  # View the model summary
  model_summary <- summary(model)
  # Getting the R and p values for the plot
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # Saving the results in a dataframe
  regression_models2 <- data.frame(
    independent_variable = "Resident Population (2020)", 
    Week_Ahead = i, 
    r_squared = r_squared, 
    p_value = p_value
  )
  # Append to the main results dataframe
  regression_models <- rbind(regression_models, regression_models2)
}

```

AUTO_ADJ WIS x POPULATION DENSITY

```{r}

for(i in c(1,2,3,4)){

  WIS_pop_data <- inner_join(WIS_dataframes[[i]], pop_data, by = "STATE")
  # Fit the regression model
  model <- lm(log1p(WIS_pop_data$AUTO_ADJ) ~ log(WIS_pop_data$Population_density_2020))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="Population Density (2020)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # Append to the main results dataframe
  regression_models<-rbind(regression_models,regression_models2)
}

```

Here we will weight the Social Vulnerability Index (SoVI) and Baseline Resilience Indicators for Communities (BRIC) county indexes which for each states based on the population size in each county.

```{r}
#############################################################
# SOVI BY STATE
# weighted by population size in each county
sovi_by_state <- sovi_data %>%
  filter(!is.nan(sovi)) %>%  # Exclude rows where 'sovi' is NaN
  group_by(state.name) %>%
  summarize(weighted_mean = weighted.mean(sovi, w = population.2020, na.rm = TRUE))

colnames(sovi_by_state)[1] <- "STATE"

#############################################################
# BRIC BY STATE
# weighted by population size in each county

bric_by_state <- bric_data %>%
  group_by(state.name) %>%
  summarize(across(16:22, ~ weighted.mean(.x, w = population.2020, na.rm = TRUE), .names = "weighted_mean_{col}"))
colnames(bric_by_state)[1] <- "STATE"

```

AUTO_ADJ WIS x SOVI

```{r}

for(i in c(1,2,3,4)){

  WIS_sovi<-NULL
  WIS_sovi <- inner_join(WIS_dataframes[[i]], sovi_by_state, by = "STATE")
  # Fit the regression model
  model <- lm(log(WIS_sovi$AUTO_ADJ) ~ (WIS_sovi$weighted_mean))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="Social Vulnerability Index (SoVI)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending the results
  regression_models<-rbind(regression_models,regression_models2)
}

```

AUTO_ADJ WIS x BRIC SOCIAL

```{r}

for(i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model  
  model <- lm(log(WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.social))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Social (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending the results
  regression_models<-rbind(regression_models,regression_models2)
}

```

AUTO_ADJ WIS x BRIC ECONOMIC

```{r}
for(i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm(log(WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.economic))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Economic (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  regression_models<-rbind(regression_models,regression_models2)
}

```

AUTO_ADJ WIS x BRIC Infrastructure

```{r}

for (i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm(log(WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.infrastructure))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Infrastructure (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending results
  regression_models<-rbind(regression_models,regression_models2)
}

```
AUTO_ADJ WIS x BRIC institutional

```{r}

for (i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm(log(WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.institutional))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Institutional (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending results
  regression_models<-rbind(regression_models,regression_models2)
}

```

AUTO_ADJ WIS x BRIC community

```{r}

for(i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[1]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm(log(WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.community))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Community (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  regression_models<-rbind(regression_models,regression_models2)
}

```

AUTO_ADJ WIS x BRIC environment

```{r}

for(i in 1:4){
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm(log(WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.environment))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC Environment (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  regression_models<-rbind(regression_models,regression_models2)
}

```
AUTO_ADJ WIS x BRIC total

```{r}

for(i in 1:4){
  # BRIC INDEX
  WIS_bric <- inner_join(WIS_dataframes[[i]], bric_by_state, by = "STATE")
  # Fit the regression model
  model <- lm(log(WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.total))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="BRIC total (2015)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  regression_models<-rbind(regression_models,regression_models2)
}
```

TEMPERATUE - ERA5

Now we will look at regression models that uses mean temperature and specific humidity.

-------------------

AUTO_ADJ WIS x Temperature ERA5 Data

```{r}
for(i in 1:4){
  # TEMPERATURE DATA from ERA5
  WIS_temp <- inner_join(WIS_dataframes[[i]], temperature_data, by = "STATE")
  # Fit the regression model
  model <- lm(log(WIS_temp$AUTO_ADJ) ~ (WIS_temp$mean))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="Mean Temperature (1990-2020)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending results
  regression_models<-rbind(regression_models,regression_models2)
}
```

AUTO_ADJ WIS x Specific Humidity ERA5 Data

```{r}
# regression results 
print(regression_models)

for(i in 1:4){
  # HUMIDITY DATA from ERA5
  WIS_humidity <- inner_join(WIS_dataframes[[i]], humidity_data, by = "STATE")
  # Fit the regression model
  model <- lm(log(WIS_humidity$AUTO_ADJ) ~ (WIS_humidity$mean))
  # View the model summary
  model_summary <- summary(model)
  # Extract R-squared and p-value
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)
  # saving the results in a dataframe
  regression_models2<-data.frame("independent_variable"="Mean Specific Humidty (1990-2020)","Week_Ahead"=i, "r_squared"=r_squared,"p_value"= p_value)
  # appending results
  regression_models<-rbind(regression_models,regression_models2)
}
```

SUMMARY OF RESULTS

```{r}
# Add a significance indicator for coloring
data <- regression_models %>%
  mutate(Significant = ifelse(p_value < 0.05, "Significant", "Not Significant"))

# Plot R-squared with conditional coloring and faceting by Week_Ahead
ggplot(data, aes(x = independent_variable, y = r_squared, fill = Significant)) +
  geom_col(color = "black", width = 0.5) +
  scale_fill_manual(values = c("Significant" = "blue", "Not Significant" = "red")) +
  labs(
    title = "Regression Models of the AUTO_ADJ performance (on log(WIS)) for 48 U.S. States (1 to 4 weeks ahead)",
    subtitle = "Blue bars indicate p-values < 0.05",
    x = "Independent Variables",
    y = "R-squared",
    fill = "Significance"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~Week_Ahead, ncol = 1)

# Save the plot with specified size
ggsave("regression_plot.jpg", width = 10, height = 14, dpi = 300)
```

######################
Friedman variance test
######################

1 WEEK AHEAD Post Hoc Wilcox test

```{r}
library(lme4)
library(ez)

# Reshape the data into long format
long_data <- W1 %>%
  pivot_longer(cols = c(-STATE,-Best_Result) , names_to = "Type", values_to = "Value")%>%
  mutate(Type = recode(Type, "AUTO_AR" = "AUTO_AAR"))

######################
# Friedman test since we have related samples (models by each STATE)
# and non-normally distributed data
# therefore, it could be better to use this non-parametric test

week1_wis_test<-friedman.test(Value ~ Type | STATE, data = long_data)
# results
friedman_results_w1<-data.frame("chi_squared"=week1_wis_test$statistic, "p_value"=week1_wis_test$p.value)
#friedman_results_w1

#########################################################################
wilcox_results <- pairwise.wilcox.test(long_data$Value, long_data$Type, 
                                       p.adjust.method ="BY", 
                                       paired = TRUE)
p_values<-wilcox_results$p.value 
p_values<-data.frame(p_values)

#########################################################################
p_df1 <- data.frame(Models = rownames(p_values), PValue = p_values$AUTO_AAR, WeekAhead=1)
p_df1 <-drop_na(p_df1)
#p_df1

```

2 WEEKS AHEAD Post Hoc Wilcox test

```{r}
# Reshape the data into long format
long_data <- W2 %>%
  pivot_longer(cols = c(-STATE,-Best_Result) , names_to = "Type", values_to = "Value")%>%
  mutate(Type = recode(Type, "AUTO_AR" = "AUTO_AAR"))

######################
# Friedman test since we have related samples (models by each STATE)
# and non-normally distributed data
# therefore, it could be better to use this non-parametric test
week2_wis_test<-friedman.test(Value ~ Type | STATE, data = long_data)
# adding results to the dataframe
friedman_results_w2<-data.frame("chi_squared"=week2_wis_test$statistic, "p_value"=week2_wis_test$p.value)

#########################################################################
wilcox_results <- pairwise.wilcox.test(long_data$Value, long_data$Type, 
                                       p.adjust.method ="BY"  , 
                                       paired = TRUE)

p_values<-wilcox_results$p.value 
p_values<-data.frame(p_values)

#########################################################################
p_df2 <- data.frame(Models = rownames(p_values), PValue = p_values$AUTO_AAR, WeekAhead=2)
p_df2 <-drop_na(p_df2)
#p_df2

```

3 WEEKS AHEAD Post Hoc Wilcox test

```{r}
# Reshape the data into long format
long_data <- W3 %>%
  pivot_longer(cols = c(-STATE,-Best_Result) , names_to = "Type", values_to = "Value")%>%
  mutate(Type = recode(Type, "AUTO_AR" = "AUTO_AAR"))

######################
# Friedman test since we have related samples (models by each STATE)
# and non-normally distributed data
# therefore, it could be better to use this non-parametric test
week3_wis_test<-friedman.test(Value ~ Type | STATE, data = long_data)
# adding results to the dataframe
friedman_results_w3<-data.frame("chi_squared"=week3_wis_test$statistic, "p_value"=week3_wis_test$p.value)

#########################################################################
wilcox_results <- pairwise.wilcox.test(long_data$Value, long_data$Type, 
                                       p.adjust.method ="BY"  , 
                                       paired = TRUE)

p_values<-wilcox_results$p.value 
p_values<-data.frame(p_values)

#########################################################################
p_df3 <- data.frame(Models = rownames(p_values), PValue = p_values$AUTO_AAR, WeekAhead=3)
p_df3 <-drop_na(p_df3)
#p_df3
```

4 WEEKS AHEAD Post Hoc Wilcox test

```{r}
# Reshape the data into long format
long_data <- W4 %>%
  pivot_longer(cols = c(-STATE,-Best_Result) , names_to = "Type", values_to = "Value")%>%
  mutate(Type = recode(Type, "AUTO_AR" = "AUTO_AAR"))

######################
# Friedman test since we have related samples (models by each STATE)
# and non-normally distributed data
# therefore, it could be better to use this non-parametric test
week4_wis_test<-friedman.test(Value ~ Type | STATE, data = long_data)
# adding results to the dataframe
friedman_results_w4<-data.frame("chi_squared"=week4_wis_test$statistic, "p_value"=week4_wis_test$p.value)

#########################################################################
wilcox_results <- pairwise.wilcox.test(long_data$Value, long_data$Type, 
                                       p.adjust.method ="BY"  , 
                                       paired = TRUE)

p_values<-wilcox_results$p.value 
p_values<-data.frame(p_values)

#########################################################################
p_df4 <- data.frame(Models = rownames(p_values), PValue = p_values$AUTO_AAR, WeekAhead=4)
p_df4 <-drop_na(p_df4)
#p_df4

```

Friedman test

```{r}
friedman_results<-rbind("1week_ahead"=friedman_results_w1,"2week_ahead"=friedman_results_w2,"3week_ahead"=friedman_results_w3,"4week_ahead"=friedman_results_w4)

weeks_<-data.frame("weeks_ahead"=c(1,2,3,4))

friedman_results<-cbind(friedman_results,weeks_)

# Define significance threshold (e.g., 0.05)
significance_level <- 0.05

# Add a new column to classify significance
friedman_results$significance <- ifelse(friedman_results$p_value < significance_level, "Significant", "Not Significant")

# Create the barplot
ggplot(friedman_results, aes(x = factor(weeks_ahead), y = 1, fill = significance)) +
  geom_tile(color = "black") +  # Add black borders to squares
  scale_fill_manual(values = c("Significant" = "blue", "Not Significant" = "red")) +
  labs(title = "Significance of Friedman tests for models' results according to each state",
       x = "Week Ahead",
       y = "Results",
       fill = "Significance") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_blank(),   # Remove y-axis values
        axis.ticks.y = element_blank())

```

Summary of Post Hoc Wilcox test with  p.adjust.method Benjamini-Yekutieli (BY)

```{r}

all_p_values<-rbind(p_df1,p_df2,p_df3,p_df4)
all_p_values

# Define significance threshold
alpha <- 0.05

# Create significance column
p_df <- all_p_values %>%
  mutate(Significance = ifelse(PValue < alpha, "Significant", "Not Significant"))

# Plot
ggplot(p_df, aes(x = factor(WeekAhead), y = Models, fill = Significance)) +
  geom_tile(color = "black") +  # Add black borders to squares
  scale_fill_manual(values = c("Significant" = "blue", "Not Significant" = "red")) +
  labs(title = "Post Hoc comparisons with AUTO_AR - Wilcox test with BY adjustment",
       subtitle = "Significant for p_value < 0.05",
       x = "Week Ahead",
       y = "Models",
       fill = "") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

This figure is just a example of a probabilistic forecast for 4 weeks ahead.

```{r}
# Load necessary library
library(ggplot2)

# Example data
data <- data.frame(
  horizon = c(1, 2, 3, 4),
  mean = c(10, 20, 30, 40),
  lower = c(5, 9, 15, 20),
  upper = c(15, 28, 40, 60)
)

# Create the plot
ggplot(data, aes(x = horizon, y = mean)) +
  geom_line(color = 'blue', size = 4) +           # Line for the forecast mean
  geom_point(color = 'red', size = 8) +           # Points for the forecast mean
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +  # Confidence interval ribbon
  labs(
    x = 'Forecast Horizon',
    y = 'Forecast Value',
    title = 'Probabilistic Forecast with 99% Confidence Interval'
  ) +
  theme_minimal() +                     # Clean theme
  theme(
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    plot.title = element_text(size = 14, face = 'bold'),
    legend.text = element_text(size = 12),  # Increase legend text size
    legend.title = element_text(size = 12)  # Increase legend title size
  )
```

### 2025-04-25 NHSH_hospitalizations_seed/ES_ADJACENT/ES_ADJACENT_48.R

@@ -1,0 +1,475 @@
+#' ES_ADJACENT 
+#' This function utilizes ensembles and single automatic ARIMAX models which have mean cases by adjacent states as exogenous variables.
+#' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
+#' It return some metrics that evaluate the performance of the models:
+#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
+#' The user can choose single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).
+#' The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window.
+#' 
+#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. states.
+#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
+#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
+#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
+#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
+#' @param week_lag An integer specifying the week lag between the exogenous variables and the ILI cases in the state under analysis. Default is \code{1}.
+#' @param list_of_states A list containing Influenza Like Illness tibbles for ALL U.S. states.
+#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
+#'
+#' @return A list containing the forecast results and performance metrics.
+
+ES_ADJACENT<-function(current_state_tb, auto=FALSE, n_weeks_ahead=1, week_lag=1, ES27=TRUE, ES64=FALSE, window=104, list_of_states=list_of_states){
+  
+  set.seed(1)
+  # The model run the ES27 or the ES64.
+  ES27=!ES64 
[…]

### 2025-04-25 NHSH_hospitalizations_seed_nolog_correct/ES_ADJACENT/ADJACENT.Rmd
update

---
title: "ADJACENT STATES ARIMAX MODEL - Influenza hospitalization data"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

This function utilizes ensembles and single automatic ARIMAX models which have mean cases by adjacent states as exogenous variables. The function fits on a rolling window of previous 104 weeks for the state under analysis and a rolling windows with the same size with 1 week-lag for the exogenous variables to generate forecasts. It return some metrics that evaluate the performance of the mod
els:target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles. The user can choose a  single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE). The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window which is set as 2 years (104 weeks).

```{r}
knitr::opts_chunk$set(echo = TRUE)
```
!!!!!!!!!!!!!!!!!!!
LOADING THE DATASET
!!!!!!!!!!!!!!!!!!!
```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
```
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
LOADING DATASET AND FUNCTIONS 
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

```{r}
###########################################
#       LOADING AND CLEANING THE DATASET  #        
###########################################

# Loads the ADJACENT states models
source("ES_ADJACENT_48_nolog.R", local = TRUE, chdir = TRUE)

# Loads the ILI dataset
my_data = read.csv("treated_influenza_hosp_dataframe_v2.csv")
my_data$target_end_date<-as.Date(my_data$target_end_date) # set the dates as dates

list_of_states <- split(my_data, my_data$state_name)
```

AUTO ADJACENT WEEK1 - ALL STATES

IF YOU WANT A FAST RESULT RUN ONLY THE AUTO_example, the others take longer.

```{r}

start_time <- Sys.time()

AUTO_ADJACENT_WEEK1_list <- mclapply(list_of_states, ES_ADJACENT, auto=TRUE, n_weeks_ahead=1,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# FINAL DATAFRAME
AUTO_ADJACENT_WEEK1 <- bind_rows(AUTO_ADJACENT_WEEK1_list, .id = "State")

```

AUTO ADJACENT WEEK2 - ALL STATES

```{r}
start_time <- Sys.time()

# RUN MODEL 
AUTO_ADJACENT_WEEK2_list <- mclapply(list_of_states, ES_ADJACENT, auto=TRUE, n_weeks_ahead=2,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
AUTO_ADJACENT_WEEK2 <- bind_rows(AUTO_ADJACENT_WEEK2_list, .id = "State")
```

AUTO ADJACENT WEEK3 - ALL STATES

```{r}
start_time <- Sys.time()

AUTO_ADJACENT_WEEK3_list <- mclapply(list_of_states, ES_ADJACENT, auto=TRUE, n_weeks_ahead=3,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# Combine the list of data frames into a single data frame with names as a column
AUTO_ADJACENT_WEEK3 <- bind_rows(AUTO_ADJACENT_WEEK3_list, .id = "State")
```

AUTO ADJACENT WEEK4 - ALL STATES

```{r}
start_time <- Sys.time()

AUTO_ADJACENT_WEEK4_list <- mclapply(list_of_states, ES_ADJACENT, auto=TRUE, n_weeks_ahead=4,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# Combine the list of data frames into a single data frame with names as a column
AUTO_ADJACENT_WEEK4 <- bind_rows(AUTO_ADJACENT_WEEK4_list, .id = "State")
```

```{r}
save.image("ADJACENT_MODELS_influenza_hospitalization_nolog.Rdata")
```

ES27 ADJACENT WEEK1 - ALL STATES

```{r}
start_time <- Sys.time()

ES27_ADJACENT_WEEK1_list <- mclapply(list_of_states, ES_ADJACENT, ES27=TRUE, n_weeks_ahead=1,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
ES27_ADJACENT_WEEK1 <- bind_rows(ES27_ADJACENT_WEEK1_list, .id = "State")
```

ES27 ADJACENT WEEK2 - ALL STATES

```{r}
start_time <- Sys.time()

ES27_ADJACENT_WEEK2_list <- mclapply(list_of_states, ES_ADJACENT, ES27=TRUE, n_weeks_ahead=2,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
ES27_ADJACENT_WEEK2 <- bind_rows(ES27_ADJACENT_WEEK2_list, .id = "State")
```

ES27 ADJACENT WEEK3 - ALL STATES

```{r}
start_time <- Sys.time()

ES27_ADJACENT_WEEK3_list <- mclapply(list_of_states, ES_ADJACENT, ES27=TRUE, n_weeks_ahead=3,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
ES27_ADJACENT_WEEK3 <- bind_rows(ES27_ADJACENT_WEEK3_list, .id = "State")
```

ES27 ADJACENT WEEK4 - ALL STATES

```{r}
start_time <- Sys.time()

ES27_ADJACENT_WEEK4_list <- mclapply(list_of_states, ES_ADJACENT, ES27=TRUE, n_weeks_ahead=4,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
ES27_ADJACENT_WEEK4 <- bind_rows(ES27_ADJACENT_WEEK4_list, .id = "State")
```

```{r}
save.image("ADJACENT_MODELS_influenza_hospitalization_nolog.Rdata")
```

ES64 ADJACENT WEEK1 - ALL STATES

```{r}
start_time <- Sys.time()

ES64_ADJACENT_WEEK1_list <- mclapply(list_of_states, ES_ADJACENT, ES64=TRUE, n_weeks_ahead=1,list_of_states=list_of_states, mc.cores=2,week_lag=1)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
ES64_ADJACENT_WEEK1 <- bind_rows(ES64_ADJACENT_WEEK1_list, .id = "State")
```

ES64 ADJACENT WEEK2 - ALL STATES

```{r}
start_time <- Sys.time()

ES64_ADJACENT_WEEK2_list <- mclapply(list_of_states, ES_ADJACENT, ES64=TRUE, n_weeks_ahead=2,list_of_states=list_of_states, mc.cores=2,week_lag=1)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# Combine the list of data frames into a single data frame with names as a column
ES64_ADJACENT_WEEK2 <- bind_rows(ES64_ADJACENT_WEEK2_list, .id = "State")

```

ES64 ADJACENT WEEK3 - ALL STATES

```{r}
start_time <- Sys.time()

ES64_ADJACENT_WEEK3_list <- mclapply(list_of_states, ES_ADJACENT, ES64=TRUE, n_weeks_ahead=3,list_of_states=list_of_states, mc.cores=2,week_lag=1)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# Combine the list of data frames into a single data frame with names as a column
ES64_ADJACENT_WEEK3 <- bind_rows(ES64_ADJACENT_WEEK3_list, .id = "State")

```

ES64 ADJACENT WEEK4 - ALL STATES

```{r}
start_time <- Sys.time()

ES64_ADJACENT_WEEK4_list <- mclapply(list_of_states, ES_ADJACENT, ES64=TRUE, n_weeks_ahead=4,list_of_states=list_of_states, mc.cores=2,week_lag=1)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# Combine the list of data frames into a single data frame with names as a column
ES64_ADJACENT_WEEK4 <- bind_rows(ES64_ADJACENT_WEEK4_list, .id = "State")

```

```{r}
save.image("ADJACENT_MODELS_influenza_hospitalization_nolog.Rdata")
```

### 2025-04-25 NHSH_hospitalizations_seed_nolog_correct/ES_ARIMA/ES_ARIMA.Rmd
update

--- 2025-04-25 NHSH_hospitalizations_seed/ES_ARIMA/ES_ARIMA.Rmd
@@
 # Loads the ADJACENT states models
-source("ES_ARIMA.R", local = TRUE, chdir = TRUE)
+source("ES_ARIMA_nolog.R", local = TRUE, chdir = TRUE)
 
@@
 ```{r}
-save.image("ARIMA_MODELS_influenza_hospitalization.Rdata")
+save.image("ARIMA_MODELS_influenza_hospitalization_nolog.Rdata")
 ```

### 2025-04-25 NHSH_hospitalizations_seed_nolog_correct/ES_AVERAGE/AVERAGE.Rmd
update

---
title: "AVERAGE STATES ARIMAX MODEL - Influenza hospitalization data - no log-back transformation"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

This function utilizes ensembles and single automatic ARIMAX models which have mean hospitalization on U.S. states as exogenous variables. The function fits on a rolling window of previous 104 weeks for the state under analysis and a rolling windows with the same size with 1 week-lag for the exogenous variables to generate forecasts. It return some metrics that evaluate the performance of the mod
els:target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles. The user can choose a  single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE). The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window which is set as 2 years (104 weeks).

```{r}
knitr::opts_chunk$set(echo = TRUE)
```
!!!!!!!!!!!!!!!!!!!
LOADING THE DATASET
!!!!!!!!!!!!!!!!!!!
```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
```
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
LOADING DATASET AND FUNCTIONS 
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

```{r}
###########################################
#       LOADING AND CLEANING THE DATASET  #        
###########################################

# Loads the AVERAGE states models
source("ES_AVERAGE_48_nolog.R", local = TRUE, chdir = TRUE)

# Loads the ILI dataset
my_data = read.csv("treated_influenza_hosp_dataframe_v2.csv")
my_data$target_end_date<-as.Date(my_data$target_end_date) # set the dates as dates

list_of_states <- split(my_data, my_data$state_name)

```

AUTO AVERAGE WEEK1 - ALL STATES

IF YOU WANT A FAST RESULT RUN ONLY THE AUTO_example, the others take longer.

```{r}

start_time <- Sys.time()

AUTO_AVERAGE_WEEK1_list <- mclapply(list_of_states, ES_AVERAGE, auto=TRUE, n_weeks_ahead=1,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# FINAL DATAFRAME
AUTO_AVERAGE_WEEK1 <- bind_rows(AUTO_AVERAGE_WEEK1_list, .id = "State")

```

AUTO AVERAGE WEEK2 - ALL STATES

```{r}
start_time <- Sys.time()

# RUN MODEL 
AUTO_AVERAGE_WEEK2_list <- mclapply(list_of_states, ES_AVERAGE, auto=TRUE, n_weeks_ahead=2,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
AUTO_AVERAGE_WEEK2 <- bind_rows(AUTO_AVERAGE_WEEK2_list, .id = "State")
```

AUTO AVERAGE WEEK3 - ALL STATES

```{r}
start_time <- Sys.time()

AUTO_AVERAGE_WEEK3_list <- mclapply(list_of_states, ES_AVERAGE, auto=TRUE, n_weeks_ahead=3,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# Combine the list of data frames into a single data frame with names as a column
AUTO_AVERAGE_WEEK3 <- bind_rows(AUTO_AVERAGE_WEEK3_list, .id = "State")
```

AUTO AVERAGE WEEK4 - ALL STATES

```{r}
start_time <- Sys.time()

AUTO_AVERAGE_WEEK4_list <- mclapply(list_of_states, ES_AVERAGE, auto=TRUE, n_weeks_ahead=4,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# Combine the list of data frames into a single data frame with names as a column
AUTO_AVERAGE_WEEK4 <- bind_rows(AUTO_AVERAGE_WEEK4_list, .id = "State")
```

```{r}
save.image("AVERAGE_MODELS_influenza_hospitalization_nolog.Rdata")
```

ES27 AVERAGE WEEK1 - ALL STATES

```{r}
start_time <- Sys.time()

ES27_AVERAGE_WEEK1_list <- mclapply(list_of_states, ES_AVERAGE, ES27=TRUE, n_weeks_ahead=1,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
ES27_AVERAGE_WEEK1 <- bind_rows(ES27_AVERAGE_WEEK1_list, .id = "State")
```

ES27 AVERAGE WEEK2 - ALL STATES

```{r}
start_time <- Sys.time()

ES27_AVERAGE_WEEK2_list <- mclapply(list_of_states, ES_AVERAGE, ES27=TRUE, n_weeks_ahead=2,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
ES27_AVERAGE_WEEK2 <- bind_rows(ES27_AVERAGE_WEEK2_list, .id = "State")
```

ES27 AVERAGE WEEK3 - ALL STATES

```{r}
start_time <- Sys.time()

ES27_AVERAGE_WEEK3_list <- mclapply(list_of_states, ES_AVERAGE, ES27=TRUE, n_weeks_ahead=3,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
ES27_AVERAGE_WEEK3 <- bind_rows(ES27_AVERAGE_WEEK3_list, .id = "State")
```

ES27 AVERAGE WEEK4 - ALL STATES

```{r}
start_time <- Sys.time()

ES27_AVERAGE_WEEK4_list <- mclapply(list_of_states, ES_AVERAGE, ES27=TRUE, n_weeks_ahead=4,list_of_states=list_of_states,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
ES27_AVERAGE_WEEK4 <- bind_rows(ES27_AVERAGE_WEEK4_list, .id = "State")
```

```{r}
save.image("AVERAGE_MODELS_influenza_hospitalization_nolog.Rdata")
```

ES64 AVERAGE WEEK1 - ALL STATES

```{r}
start_time <- Sys.time()

ES64_AVERAGE_WEEK1_list <- mclapply(list_of_states, ES_AVERAGE, ES64=TRUE, n_weeks_ahead=1,list_of_states=list_of_states, mc.cores=2,week_lag=1)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# FINAL DATAFRAME
ES64_AVERAGE_WEEK1 <- bind_rows(ES64_AVERAGE_WEEK1_list, .id = "State")
```

ES64 AVERAGE WEEK2 - ALL STATES

```{r}
start_time <- Sys.time()

ES64_AVERAGE_WEEK2_list <- mclapply(list_of_states, ES_AVERAGE, ES64=TRUE, n_weeks_ahead=2,list_of_states=list_of_states, mc.cores=2,week_lag=1)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# Combine the list of data frames into a single data frame with names as a column
ES64_AVERAGE_WEEK2 <- bind_rows(ES64_AVERAGE_WEEK2_list, .id = "State")

```

ES64 AVERAGE WEEK3 - ALL STATES

```{r}
start_time <- Sys.time()

ES64_AVERAGE_WEEK3_list <- mclapply(list_of_states, ES_AVERAGE, ES64=TRUE, n_weeks_ahead=3,list_of_states=list_of_states, mc.cores=2,week_lag=1)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# Combine the list of data frames into a single data frame with names as a column
ES64_AVERAGE_WEEK3 <- bind_rows(ES64_AVERAGE_WEEK3_list, .id = "State")

```

ES64 AVERAGE WEEK4 - ALL STATES

```{r}
start_time <- Sys.time()

ES64_AVERAGE_WEEK4_list <- mclapply(list_of_states, ES_AVERAGE, ES64=TRUE, n_weeks_ahead=4,list_of_states=list_of_states, mc.cores=2,week_lag=1)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)
# Combine the list of data frames into a single data frame with names as a column
ES64_AVERAGE_WEEK4 <- bind_rows(ES64_AVERAGE_WEEK4_list, .id = "State")

```

```{r}
save.image("AVERAGE_MODELS_influenza_hospitalization_nolog.Rdata")
```

### 2025-04-25 NHSH_hospitalizations_seed_nolog_correct/ES_EPIWEEK/EPIWEEK.Rmd
update

---
title: "EPIWEEK ARIMAX MODELS - Influenza hospitalization data"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

This function utilizes ensembles and single automatic ARIMAX models which have mean cases by epidemiological weeks over the last 2 years as exogenous variables. The function fits on a rolling window of previous 104 weeks for the state under analysis and a rolling windows with the same size with 1 week-lag for the exogenous variables to generate forecasts. It return some metrics that evaluate the performance of the models:target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles. The user can choose a  single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE). The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window which is set as 2 years (104 weeks).

```{r}
knitr::opts_chunk$set(echo = TRUE)
```
!!!!!!!!!!!!!!!!!!!!
LOADING THE PACKAGES
!!!!!!!!!!!!!!!!!!!!
```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
```
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
LOADING DATASET AND FUNCTIONS 
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
```{r}
###########################################
#       LOADING AND CLEANING THE DATASET  #        
###########################################

# Loads the ADJACENT states models
source("ES_EPIWEEK_nolog.R", local = TRUE, chdir = TRUE)

# Loads the ILI dataset
my_data = read.csv("treated_influenza_hosp_dataframe_v2.csv")
my_data$target_end_date<-as.Date(my_data$target_end_date) # set the dates as dates

list_of_states <- split(my_data, my_data$state_name)
```

AUTO EPIWEEK WEEK1

```{r}
start_time <- Sys.time()

# RUN MODEL 
AUTO_EPIWEEK_WEEK1_list <- mclapply(list_of_states, ES_EPIWEEK, auto=TRUE, n_weeks_ahead=1,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# FINAL DATAFRAME
AUTO_EPIWEEK_WEEK1 <- bind_rows(AUTO_EPIWEEK_WEEK1_list, .id = "State")

# SAVE TO CORRECT FOLDER
#write.csv(EPIWEEK/AUTO/AUTO_EPIWEEK_WEEK1, file = "AUTO_EPIWEEK_WEEK1.csv", row.names = TRUE)
```

AUTO EPIWEEK WEEK2

```{r}
start_time <- Sys.time()

# RUN MODEL 
AUTO_EPIWEEK_WEEK2_list <- mclapply(list_of_states, ES_EPIWEEK, auto=TRUE, n_weeks_ahead=2,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# FINAL DATAFRAME
AUTO_EPIWEEK_WEEK2 <- bind_rows(AUTO_EPIWEEK_WEEK2_list, .id = "State")

```

AUTO EPIWEEK WEEK3

```{r}
start_time <- Sys.time()

AUTO_EPIWEEK_WEEK3_list <- mclapply(list_of_states, ES_EPIWEEK, auto=TRUE, n_weeks_ahead=3,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
AUTO_EPIWEEK_WEEK3 <- bind_rows(AUTO_EPIWEEK_WEEK3_list, .id = "State")

```

AUTO EPIWEEK WEEK4

```{r}
start_time <- Sys.time()

AUTO_EPIWEEK_WEEK4_list <- mclapply(list_of_states, ES_EPIWEEK, auto=TRUE, n_weeks_ahead=4,week_lag=1 ,mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
AUTO_EPIWEEK_WEEK4 <- bind_rows(AUTO_EPIWEEK_WEEK4_list, .id = "State")

```

```{r}
save.image("EPIWEEK_MODELS_influenza_hospitalization_nolog.Rdata")

```

ES27 EPIWEEK WEEK1

```{r}
start_time <- Sys.time()

ES27_EPIWEEK_WEEK1_list <- mclapply(list_of_states, ES_EPIWEEK, ES27=TRUE, n_weeks_ahead=1,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_EPIWEEK_WEEK1 <- bind_rows(ES27_EPIWEEK_WEEK1_list, .id = "State")

```

ES27 EPIWEEK WEEK2

```{r}
start_time <- Sys.time()

ES27_EPIWEEK_WEEK2_list <- mclapply(list_of_states, ES_EPIWEEK, ES27=TRUE, n_weeks_ahead=2,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_EPIWEEK_WEEK2 <- bind_rows(ES27_EPIWEEK_WEEK2_list, .id = "State")

```

ES27 EPIWEEK WEEK3

```{r}
start_time <- Sys.time()

ES27_EPIWEEK_WEEK3_list <- mclapply(list_of_states, ES_EPIWEEK, ES27=TRUE, n_weeks_ahead=3,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_EPIWEEK_WEEK3 <- bind_rows(ES27_EPIWEEK_WEEK3_list, .id = "State")

```

ES27 EPIWEEK WEEK4

```{r}
start_time <- Sys.time()

ES27_EPIWEEK_WEEK4_list <- mclapply(list_of_states, ES_EPIWEEK, ES27=TRUE, n_weeks_ahead=4,week_lag=1 ,mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_EPIWEEK_WEEK4 <- bind_rows(ES27_EPIWEEK_WEEK4_list, .id = "State")

```

```{r}
save.image("EPIWEEK_MODELS_influenza_hospitalization_nolog.Rdata")

```

ES64 EPIWEEK WEEK1

```{r}
start_time <- Sys.time()

ES64_EPIWEEK_WEEK1_list <- mclapply(list_of_states, ES_EPIWEEK, ES64=TRUE, n_weeks_ahead=1,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_EPIWEEK_WEEK1 <- bind_rows(ES64_EPIWEEK_WEEK1_list, .id = "State")

```

ES64 EPIWEEK WEEK2

```{r}
start_time <- Sys.time()

ES64_EPIWEEK_WEEK2_list <- mclapply(list_of_states, ES_EPIWEEK, ES64=TRUE, n_weeks_ahead=2,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_EPIWEEK_WEEK2 <- bind_rows(ES64_EPIWEEK_WEEK2_list, .id = "State")

```

ES64 EPIWEEK WEEK3

```{r}
start_time <- Sys.time()

ES64_EPIWEEK_WEEK3_list <- mclapply(list_of_states, ES_EPIWEEK, ES64=TRUE, n_weeks_ahead=3,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_EPIWEEK_WEEK3 <- bind_rows(ES64_EPIWEEK_WEEK3_list, .id = "State")

```

ES64 EPIWEEK WEEK4

```{r}
start_time <- Sys.time()

ES64_EPIWEEK_WEEK4_list <- mclapply(list_of_states, ES_EPIWEEK, ES64=TRUE, n_weeks_ahead=4,week_lag=1,mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_EPIWEEK_WEEK4 <- bind_rows(ES64_EPIWEEK_WEEK4_list, .id = "State")

```

```{r}
save.image("EPIWEEK_MODELS_influenza_hospitalization_nolog.Rdata")

```

### 2025-04-25 NHSH_hospitalizations_seed_nolog_correct/flu_hospitalization_dataset .Rmd
update

---
title: "Influenza Hopistalizations dataset with Florida - NHSN"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading datasets

```{r}
library(purrr)
library(tidyr)
library(dplyr)
library(lubridate)
library(imputeTS)
library(MMWRweek)
```

Influenza hospitalizations

```{r}
influenza_data<-read.csv("Weekly_United_States_Hospitalization_Metrics_by_Jurisdiction__During_Mandatory_Reporting_Period_from_August_1__2020_to_April_30__2024__and_for_Data_Reported_Voluntarily_Beginning_May_1__2024__National_Healthcare_Safety_Network__NHSN__20241001.csv")

state_codes<-read.csv("State_Codes.csv")

```

Choosing U.S. states

```{r}
states=state_codes$abbreviation[1:48] # removing Florida
```

Creating a list with multiple dataframes

```{r}

influenza_data$Week.Ending.Date<-as.Date(influenza_data$Week.Ending.Date)
influenza_data$epiweek<-MMWRweek(influenza_data$Week.Ending.Date)$MMWRweek

# Initialize an empty list to store the results
influenza_hosp_by_state <- list()

# Loop through each state in the 'states' vector
for (i in 1:48) {
  st <- states[i]  # Get the current state
  # Filter and arrange the data for the current state
  influenza_hosp_by_state[[i]] <- influenza_data %>%
    arrange(Week.Ending.Date) %>%  # Order by date
    filter(Geographic.aggregation == st)  # Filter for the current state
}

```

Finding the datasets that have NAs

We need to start from week 12 of all datasets till we remove all initial NAs.
In addition to the initial week 3 states had more NAs on three states, which we replace with zeros.

WV(state code=45): 3 NAs
MN(state code=20): 18 NAs
MA(state code=18): 32 NAs

```{r}
#INITIAL NAs
# Loop through each element in state_data_list
for (i in 1:length(influenza_hosp_by_state)) {

  # Extract the unique state name from the 'state' column
  states_names <- unique(influenza_hosp_by_state[[i]]$Geographic.aggregation)
  # Plot epi_week, setting the title to the state name
  print(sum(is.na(influenza_hosp_by_state[[i]]$Weekly.Total.Influenza.Admissions[12:18])))
}

#MORE NAs
# Loop through each element in state_data_list
for (i in 1:48) {

  state_name <- unique(influenza_hosp_by_state[[i]]$Geographic.aggregation)
  missing_values_count <- sum(is.na(influenza_hosp_by_state[[i]]$Weekly.Total.Influenza.Admissions[12:length(influenza_hosp_by_state[[i]]$Week.Ending.Date)]))

  print(paste("State:", state_name,"State code:", i, "- Missing values:", missing_values_count))
}

#FINAL NAs
# Loop through each element in state_data_list
for (i in 1:length(influenza_hosp_by_state)) {

  # Extract the unique state name from the 'state' column
  states_names <- unique(influenza_hosp_by_state[[i]]$Geographic.aggregation)
  # Plot epi_week, setting the title to the state name
  print(sum(is.na(influenza_hosp_by_state[[i]]$Weekly.Total.Influenza.Admissions[190:198])))
}

```

3 states with NAs that will be replaced with zeros.

```{r}

# Loop through each element in state_data_list
for (i in c(18,20,45)) {

  # Extract the unique state name from the 'state' column
  states_names <- unique(influenza_hosp_by_state[[i]]$Geographic.aggregation)
  # Plot epi_week, setting the title to the state name
  plot(influenza_hosp_by_state[[i]]$Week.Ending.Date[12:198],influenza_hosp_by_state[[i]]$Weekly.Total.Influenza.Admissions[12:198], main = states_names, ylab="Cases", xlab="Dates")
}
```
Noting that the weeks that have NAs are at the end of the last season and in a location with low influenza activity we will replace these values with zero.

```{r}
for (i in c(18, 20, 45)) {
  # Extract the unique state name from the 'Geographic.aggregation' column
  state_name <- unique(influenza_hosp_by_state[[i]]$Geographic.aggregation)

  # Extract the 'Weekly.Total.Influenza.Admissions' column from 12 to 198
  admissions <- influenza_hosp_by_state[[i]]$Weekly.Total.Influenza.Admissions[12:198]

  # Replace only the NA values with zero
  admissions[is.na(admissions)] <- 0

  # Update the original data with the new values
  influenza_hosp_by_state[[i]]$Weekly.Total.Influenza.Admissions[12:198] <- admissions

  # Plot the time series and set the title to the state name
  plot(admissions, 
       type = "l",   # Line plot
       main = state_name, 
       xlab = "Week", 
       ylab = "Influenza Admissions")
}

```
The clean final dataset for all states

```{r}
# Loop through each element in state_data_list
for (i in 1:48) {

  # Extract the unique state name from the 'state' column
  states_names <- unique(influenza_hosp_by_state[[i]]$Geographic.aggregation)
  # Plot epi_week, setting the title to the state name
  plot(influenza_hosp_by_state[[i]]$Week.Ending.Date[12:198],influenza_hosp_by_state[[i]]$Weekly.Total.Influenza.Admissions[12:198], main = states_names)
}

# Combine the list of data frames into a single data frame with names as a column
influenza_hosp_dataframe <- bind_rows(lapply(influenza_hosp_by_state, function(df) df[12:198, ]))
```

Final dataset similar in format that can be read by the ensembles

```{r}
colnames(state_codes)[3] <- "state_name"

final_dataset <- influenza_hosp_dataframe %>%
  select("target_end_date"=Week.Ending.Date, "abbreviation"=Geographic.aggregation,"cases"= Weekly.Total.Influenza.Admissions)

final_dataset <- left_join(final_dataset, state_codes, by = "abbreviation")
final_dataset

# write the dataset to use in the models
write.csv(final_dataset, file = "ES_ADJACENT/treated_influenza_hosp_dataframe_v2.csv")
write.csv(final_dataset, file = "ES_ARIMA/treated_influenza_hosp_dataframe_v2.csv")
write.csv(final_dataset, file = "ES_EPIWEEK/treated_influenza_hosp_dataframe_v2.csv")
write.csv(final_dataset, file = "ES_TEMPERATURE/treated_influenza_hosp_dataframe_v2.csv")

```

### 2025-04-25 NHSH_hospitalizations_seed_nolog_correct/ES_ADJACENT/ES_ADJACENT_48_nolog.R

@@ -1,0 +1,475 @@
+#' ES_ADJACENT 
+#' This function utilizes ensembles and single automatic ARIMAX models which have mean cases by adjacent states as exogenous variables.
+#' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
+#' It return some metrics that evaluate the performance of the models:
+#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
+#' The user can choose single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).
+#' The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window.
+#' 
+#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. states.
+#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
+#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
+#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
+#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
+#' @param week_lag An integer specifying the week lag between the exogenous variables and the ILI cases in the state under analysis. Default is \code{1}.
+#' @param list_of_states A list containing Influenza Like Illness tibbles for ALL U.S. states.
+#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
+#'
+#' @return A list containing the forecast results and performance metrics.
+
+ES_ADJACENT<-function(current_state_tb, auto=FALSE, n_weeks_ahead=1, week_lag=1, ES27=TRUE, ES64=FALSE, window=104, list_of_states=list_of_states){
+  
+  set.seed(1)
+  # The model run the ES27 or the ES64.
+  ES27=!ES64 
[…]

### 2025-04-25 NHSH_hospitalizations_seed_nolog_correct/ES_AVERAGE/ES_AVERAGE_48_nolog.R

@@ -1,0 +1,477 @@
+#' ES_AVERAGE 
+#' This function utilizes ensembles and single automatic ARIMAX models which have mean cases by AVERAGE states as exogenous variables.
+#' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
+#' It return some metrics that evaluate the performance of the models:
+#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
+#' The user can choose single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).
+#' The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window.
+#' 
+#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. states.
+#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
+#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
+#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
+#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
+#' @param week_lag An integer specifying the week lag between the exogenous variables and the ILI cases in the state under analysis. Default is \code{1}.
+#' @param list_of_states A list containing Influenza Like Illness tibbles for ALL U.S. states.
+#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
+#'
+#' @return A list containing the forecast results and performance metrics.
+
+ES_AVERAGE<-function(current_state_tb, auto=FALSE, n_weeks_ahead=1, week_lag=1, ES27=TRUE, ES64=FALSE, window=104, list_of_states=list_of_states){
+  
+  set.seed(1)
+  # The model run the ES27 or the ES64.
+  ES27=!ES64 
+  # Empty list that will contain forecasts, predictive quantiles and number of models.
+  results<-listenv()   
+  if(ES27){
+    pdq=c(0,1,2) # Possible ARIMA pdq's.
[…]

### 2025-04-25 NHSH_hospitalizations_seed_nolog_correct/ES_TEMPERATURE/ES_TEMPERATURE_nolog.R

@@ -1,0 +1,400 @@
+#' ES_TEMPERATURE
+#' This function utilizes ensembles and single automatic ARIMAX models which have mean temperature as exogenous variables.
+#' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
+#' It return some metrics that evaluate the performance of the models:
+#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
+#' The user can choose  single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).
+#' The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window.
+#' 
+#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. states.
+#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
+#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
+#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
+#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
+#' @param week_lag An integer specifying the week lag between the exogenous variables and the ILI cases in the state under analysis. Default is \code{1}.
+#' @param temperature_data A data frame containing weekly mean temperature data for ALL U.S. states for the same weeks as the ILI data.
+#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
+#'
+#' @return A list containing the forecast results and performance metrics.
+
+ES_TEMPERATURE<-function(current_state_tb, week_lag=1, temperature_data=NULL, auto=FALSE, n_weeks_ahead=1, ES27=TRUE, ES64=FALSE, window=104){
+  
+  set.seed(1)
+  # The model run the ES27 or the ES64.
+  ES27=!ES64 
+  # Empty list that will contain forecasts, predictive quantiles and number of models.
+  results<-listenv()   
+  if(ES27){
+    pdq=c(0,1,2) # Possible ARIMA pdq's.
[…]

### 2025-04-25 manuscript_results.Rmd
update

@@
 ---
-title: "Final manuscript"
+title: "Manuscript results: Forecasting influenza using ARIMA and ARIMAX"
 author: "Victor Felix"
@@
 
-This document summarizes the results of the 12 models that were developed: AUTO ARIMA, AUTO ADJACENT, AUTO EPIWEEK, ES27 TEMPERATURE, ES27 ARIMA, ES27 ADJACENT, ES27 EPIWEEK, ES64 TEMPERATURE, ES64 ARIMA, ES64 ADJACENT, ES64 EPIWEEK, ES64 TEMPERATURE.
-It has a series of histograms and maps that compare the results of these models based on their WIS.
-You can read more about the specific covariates and ensemble techniques of these models in their own Rmarkdown documents.
+This document has the results that we present in our manuscript draft.
 
@@
 library("fields")
-
+library("ggpubr")
 ```
@@
 ```{r}
+
 load("NHSH_hospitalizations_seed_nolog_correct/ES_ARIMA/ARIMA_MODELS_influenza_hospitalization_nolog.Rdata")
@@
   AUTO_AVG_NoLg=AUTO_AVERAGE_WEEK1$WIS,
+
   ES27_AR_NoLg=ES27_ARIMA_WEEK1$WIS, 
@@
   ES27_AVG_NoLg=ES27_AVERAGE_WEEK1$WIS,
+
   ES64_AR_NoLg=ES64_ARIMA_WEEK1$WIS, 
@@
 
+Example of influenza hospitalizations forecasts and 95% confidence interval.
+
+```{r}
+
+# Create a data frame with forecasts for plotting
+forecast_df <- data.frame(
+  Week = factor(1:4, labels = paste0("Week ", 1:4)),  # Label weeks
+  Value = c(AUTO_ARIMA_WEEK1_list[[1]]$forecasts[5], 
+            AUTO_ARIMA_WEEK2_list[[1]]$forecasts[5], 
+            AUTO_ARIMA_WEEK3_list[[1]]$forecasts[5], 
+            AUTO_ARIMA_WEEK4_list[[1]]$forecasts[5]),
+  Lower = c(AUTO_ARIMA_WEEK1_list[[1]][["0.025"]][5], 
+            AUTO_ARIMA_WEEK2_list[[1]][["0.025"]][5], 
+            AUTO_ARIMA_WEEK3_list[[1]][["0.025"]][5], 
+            AUTO_ARIMA_WEEK4_list[[1]][["0.025"]][5]),
+  Upper = c(AUTO_ARIMA_WEEK1_list[[1]][["0.975"]][5], 
+            AUTO_ARIMA_WEEK2_list[[1]][["0.975"]][5], 
+            AUTO_ARIMA_WEEK3_list[[1]][["0.975"]][5], 
+            AUTO_ARIMA_WEEK4_list[[1]][["0.975"]][5])
+)
+
+# Plotting forecasts and ribbon
+plt00 <- ggplot(forecast_df, aes(x = Week, y = Value, group = 1)) +
+  geom_line(color = "black", size = 1) +          # Line connecting forecast values
+  geom_point(color = "black", size = 3) +         # Points for each forecast value
+  geom_ribbon(aes(ymin = Lower, ymax = Upper),   # Ribbon for confidence intervals
+              fill = "gray", alpha = 0.25) +
+  labs(title = "Example of influenza hospitalizaions forecasts with 95% confidence intervals",
+       x = "Target Weeks",
+       y = "Forecast Values") +
+  theme_minimal()
+
+plt00
+
+ggsave(plt00, height = 5, width = 6, filename="fig00.jpg") 
+```
+
+Weighted interval scores for each state and each target week.
+
 ```{r}
@@
 df_combined_ <- bind_rows(
-  df_W1_NoLg %>% mutate(Dataset = "W1"),
-  df_W2_NoLg %>% mutate(Dataset = "W2"),
-  df_W3_NoLg %>% mutate(Dataset = "W3"),
-  df_W4_NoLg %>% mutate(Dataset = "W4")
+  df_W1_NoLg %>% mutate(Dataset = "1 week ahead"),
+  df_W2_NoLg %>% mutate(Dataset = "2 weeks ahead"),
+  df_W3_NoLg %>% mutate(Dataset = "3 weeks ahead"),
+  df_W4_NoLg %>% mutate(Dataset = "4 weeks ahead")
 ) %>%
@@
 
-Calculate mean weighted interval score for all forecast weeks on each model.
+Computing mean weighted interval score for all target week for each model.
 
@@
       AUTO_AVG_NoLg = mean(AUTO_AVG_NoLg, na.rm = TRUE),
+
       ES27_AR_NoLg = mean(ES27_AR_NoLg, na.rm = TRUE),
@@
       ES27_AVG_NoLg = mean(ES27_AVG_NoLg, na.rm = TRUE),
+
       ES64_AR_NoLg = mean(ES64_AR_NoLg, na.rm = TRUE),
@@
       ES64_AVG_NoLg = mean(ES64_AVG_NoLg, na.rm = TRUE)
-
     )
@@
 
-Here we have boxplots of the mean(WIS) by each state.
+Boxplots of the mean(WIS) by each state.
 
@@
 colors_tmp <- c("lightgreen","green3","#319045")
+colors_avg <- c("orange","orange2", "orange3")
 
@@
   "AUTO_ADJ" = colors_adj[1], "ES27_ADJ" = colors_adj[2], "ES64_ADJ" = colors_adj[3],
-  "AUTO_AVG" = colors_adj[1], "ES27_AVG" = colors_adj[2], "ES64_AVG" = colors_adj[3]
+  "AUTO_AVG" = colors_avg[1], "ES27_AVG" = colors_avg[2], "ES64_AVG" = colors_avg[3]
 )
@@
     scale_fill_manual(values = custom_colors) +
-    facet_wrap(category ~ Models , nrow = 2) +
+    facet_wrap(category ~ Models , nrow = 3) +
     labs(
       title = "A) Models without log-back transformations",
-      x = "",
+      x = "Target Week",
       y = "Mean Weighted Interval Score (WIS)"
@@
 
-Loading the results of each model and the shapefiles of the maps.
+Loading model with log-back transformations.
 
@@
 load("NHSH_hospitalizations_seed/ES_AVERAGE/AVERAGE_MODELS_influenza_hospitalization.Rdata")
-
-states <- read_sf("NHSH_hospitalizations_seed/shapefiles/cb_2018_us_state_500k.shp")
-
-states <- states %>%
-  rename(STATE = NAME)
-
 ```
@@
       AUTO_TMP_Lg = mean(AUTO_TEMP, na.rm = TRUE),
+      AUTO_AVG_Lg = mean(AUTO_AVG, na.rm = TRUE),
+
       ES27_AR_Lg = mean(ES27_AR, na.rm = TRUE),
@@
       ES27_TMP_Lg = mean(ES27_TEMP, na.rm = TRUE),
+      ES27_AVG_Lg = mean(ES27_AVG, na.rm = TRUE),
+
       ES64_AR_Lg = mean(ES64_AR, na.rm = TRUE),
@@
       ES64_EPI_Lg = mean(ES64_EPI, na.rm = TRUE),
-      ES64_TMP_Lg = mean(ES64_TEMP, na.rm = TRUE)
+      ES64_TMP_Lg = mean(ES64_TEMP, na.rm = TRUE),
+      ES64_AVG_Lg = mean(ES64_AVG, na.rm = TRUE),
+
     )
@@
 
-  Models = rep(c("AUTO_AR_Lg", "ES27_AR_Lg", "ES64_AR_Lg", "AUTO_ADJ_Lg", "ES27_ADJ_Lg", "ES64_ADJ_Lg", "AUTO_TMP_Lg", "ES27_TMP_Lg", "ES64_TMP_Lg", "AUTO_EPI_Lg", "ES27_EPI_Lg", "ES64_EPI_Lg"), each=48, times = 4),
-
-  value = c(W1_Lg$AUTO_AR_Lg, W1_Lg$ES27_AR_Lg, W1_Lg$ES64_AR_Lg, W1_Lg$AUTO_ADJ_Lg, W1_Lg$ES27_ADJ_Lg, W1_Lg$ES64_ADJ_Lg, W1_Lg$AUTO_TMP_Lg, W1_Lg$ES27_TMP_Lg, W1_Lg$ES64_TMP_Lg, W1_Lg$AUTO_EPI_Lg, W1_Lg$ES27_EPI_Lg, W1_Lg$ES64_EPI_Lg,
-
-            W2_Lg$AUTO_AR_Lg, W2_Lg$ES27_AR_Lg, W2_Lg$ES64_AR_Lg, W2_Lg$AUTO_ADJ_Lg, W2_Lg$ES27_ADJ_Lg, W2_Lg$ES64_ADJ_Lg, W2_Lg$AUTO_TMP_Lg, W2_Lg$ES27_TMP_Lg, W2_Lg$ES64_TMP_Lg, W2_Lg$AUTO_EPI_Lg, W2_Lg$ES27_EPI_Lg, W2_Lg$ES64_EPI_Lg,
-
-            W3_Lg$AUTO_AR_Lg, W3_Lg$ES27_AR_Lg, W3_Lg$ES64_AR_Lg, W3_Lg$AUTO_ADJ_Lg, W3_Lg$ES27_ADJ_Lg, W3_Lg$ES64_ADJ_Lg, W3_Lg$AUTO_TMP_Lg, W3_Lg$ES27_TMP_Lg, W3_Lg$ES64_TMP_Lg, W3_Lg$AUTO_EPI_Lg, W3_Lg$ES27_EPI_Lg, W3_Lg$ES64_EPI_Lg,
-
-            W4_Lg$AUTO_AR_Lg, W4_Lg$ES27_AR_Lg, W4_Lg$ES64_AR_Lg, W4_Lg$AUTO_ADJ_Lg, W4_Lg$ES27_ADJ_Lg, W4_Lg$ES64_ADJ_Lg, W4_Lg$AUTO_TMP_Lg, W4_Lg$ES27_TMP_Lg, W4_Lg$ES64_TMP_Lg, W4_Lg$AUTO_EPI_Lg, W4_Lg$ES27_EPI_Lg, W4_Lg$ES64_EPI_Lg)
+  Models = rep(c("AUTO_AR_Lg", "ES27_AR_Lg", "ES64_AR_Lg", "AUTO_ADJ_Lg", "ES27_ADJ_Lg", "ES64_ADJ_Lg", "AUTO_TMP_Lg", "ES27_TMP_Lg", "ES64_TMP_Lg", "AUTO_EPI_Lg", "ES27_EPI_Lg", "ES64_EPI_Lg", "AUTO_AVG_Lg", "ES27_AVG_Lg", "ES64_AVG_Lg"), each=48, times = 4),
+
+  value = c(W1_Lg$AUTO_AR_Lg, W1_Lg$ES27_AR_Lg, W1_Lg$ES64_AR_Lg, 
+            W1_Lg$AUTO_ADJ_Lg, W1_Lg$ES27_ADJ_Lg, W1_Lg$ES64_ADJ_Lg, 
+            W1_Lg$AUTO_TMP_Lg, W1_Lg$ES27_TMP_Lg, W1_Lg$ES64_TMP_Lg, 
+            W1_Lg$AUTO_EPI_Lg, W1_Lg$ES27_EPI_Lg, W1_Lg$ES64_EPI_Lg,
+            W1_Lg$AUTO_AVG_Lg, W1_Lg$ES27_AVG_Lg, W1_Lg$ES64_AVG_Lg,
+
+            W2_Lg$AUTO_AR_Lg, W2_Lg$ES27_AR_Lg, W2_Lg$ES64_AR_Lg, 
+            W2_Lg$AUTO_ADJ_Lg, W2_Lg$ES27_ADJ_Lg, W2_Lg$ES64_ADJ_Lg, 
+            W2_Lg$AUTO_TMP_Lg, W2_Lg$ES27_TMP_Lg, W2_Lg$ES64_TMP_Lg, 
+            W2_Lg$AUTO_EPI_Lg, W2_Lg$ES27_EPI_Lg, W2_Lg$ES64_EPI_Lg,
+            W2_Lg$AUTO_AVG_Lg, W2_Lg$ES27_AVG_Lg, W2_Lg$ES64_AVG_Lg,
+
+            W3_Lg$AUTO_AR_Lg, W3_Lg$ES27_AR_Lg, W3_Lg$ES64_AR_Lg, 
+            W3_Lg$AUTO_ADJ_Lg, W3_Lg$ES27_ADJ_Lg, W3_Lg$ES64_ADJ_Lg, 
+            W3_Lg$AUTO_TMP_Lg, W3_Lg$ES27_TMP_Lg, W3_Lg$ES64_TMP_Lg, 
+            W3_Lg$AUTO_EPI_Lg, W3_Lg$ES27_EPI_Lg, W3_Lg$ES64_EPI_Lg,
+            W3_Lg$AUTO_AVG_Lg, W3_Lg$ES27_AVG_Lg, W3_Lg$ES64_AVG_Lg,
+
+            W4_Lg$AUTO_AR_Lg, W4_Lg$ES27_AR_Lg, W4_Lg$ES64_AR_Lg, 
+            W4_Lg$AUTO_ADJ_Lg, W4_Lg$ES27_ADJ_Lg, W4_Lg$ES64_ADJ_Lg, 
+            W4_Lg$AUTO_TMP_Lg, W4_Lg$ES27_TMP_Lg, W4_Lg$ES64_TMP_Lg, 
+            W4_Lg$AUTO_EPI_Lg, W4_Lg$ES27_EPI_Lg, W4_Lg$ES64_EPI_Lg,
+            W4_Lg$AUTO_AVG_Lg, W4_Lg$ES27_AVG_Lg, W4_Lg$ES64_AVG_Lg)
+
 )
@@
     grepl("TMP", Models) ~ "TMP",
-    grepl("ADJ", Models) ~ "ADJ"
+    grepl("ADJ", Models) ~ "ADJ",
+    grepl("AVG", Models) ~ "AVG",
   ))
@@
   "AUTO_TMP_Lg" = colors_tmp[1], "ES27_TMP_Lg" = colors_tmp[2], "ES64_TMP_Lg" = colors_tmp[3],
-  "AUTO_ADJ_Lg" = colors_adj[1], "ES27_ADJ_Lg" = colors_adj[2], "ES64_ADJ_Lg" = colors_adj[3]
+  "AUTO_ADJ_Lg" = colors_adj[1], "ES27_ADJ_Lg" = colors_adj[2], "ES64_ADJ_Lg" = colors_adj[3],
+  "AUTO_AVG_Lg" = colors_avg[1], "ES27_AVG_Lg" = colors_avg[2], "ES64_AVG_Lg" = colors_avg[3]
 )
@@
     scale_fill_manual(values = custom_colors) +
-    facet_wrap(category ~ Models , nrow = 2) +
+    facet_wrap(category ~ Models , nrow = 3) +
     labs(
-      title = "B)Models with log-back transformations",
+      title = "B) Models with log-back transformations",
       x = "",
@@
 W4<-merge(W4_Lg,W4_NoLg, by = "STATE")
-
-```
-
-Here we have boxplots of the mean(WIS) by each state.
+```
 
@@
   "AUTO_ADJ_Lg" = colors_adj[1], "ES27_ADJ_Lg" = colors_adj[2], "ES64_ADJ_Lg" = colors_adj[3],
+  "AUTO_AVG_Lg" = colors_avg[1], "ES27_AVG_Lg" = colors_avg[2], "ES64_AVG_Lg" = colors_avg[3],
 
@@
   "AUTO_TMP_NoLg" = colors_tmp[1], "ES27_TMP_NoLg" = colors_tmp[2], "ES64_TMP_NoLg" = colors_tmp[3],
-  "AUTO_ADJ_NoLg" = colors_adj[1], "ES27_ADJ_NoLg" = colors_adj[2], "ES64_ADJ_NoLg" = colors_adj[3]
+  "AUTO_ADJ_NoLg" = colors_adj[1], "ES27_ADJ_NoLg" = colors_adj[2], "ES64_ADJ_NoLg" = colors_adj[3],
+  "AUTO_AVG_NoLg" = colors_avg[1], "ES27_AVG_NoLg" = colors_avg[2], "ES64_AVG_NoLg" = colors_avg[3]
 )
@@
 plot2
+
 # --------- WEEK3 MODELS ------------- #
@@
 plot4
+
 #ggsave("week4_best_models.jpg", plot4)
 
-W1
 ```
@@
 
- Wilcox test with Holm p-value ajustment
-
-1 WEEK AHEAD
+ Now we will utilize a Wilcox test with Holm p-value ajustment for repeated comparassion to evaluate the differences in model performances.
+
+1 WEEK AHEAD - Wilcoxon test with Holm p-value adjustment
 
@@
   pivot_longer(cols = c(-STATE,-Best_Result,-Model_Type,-Week_Ahead) , names_to = "Type", values_to = "Value")#%>%
-#  mutate(Type = recode(Type, "AUTO_AR" = "AUTO_AAR"))
 
@@
 wilcox_results <- map_df(model_types, function(model) {
-  test <- wilcox.test(wide_data[[model]], wide_data$AUTO_AR_NoLg, paired = TRUE, p.adjust.method ="holm")
+  test <- wilcox.test(wide_data[[model]], wide_data$AUTO_AR_NoLg, paired = TRUE)
   data.frame(Model = model, p_value = test$p.value, W = test$statistic)
@@
 p_values<-wilcox_results$p_value 
-
 p_values<-p.adjust(p_values, method = "holm")
-
 p_values<-data.frame(p_values)
@@
 #########################################################################
-p_df1 <- data.frame(Model = model_types, PValue = p_values, WeekAhead=1)
-p_df1 <-drop_na(p_df1)
-p_df1
-
-```
-
-2 WEEKS AHEAD Post Hoc Wilcox test
+p_values_wk1 <- data.frame(Model = model_types, PValue = p_values, WeekAhead=1)
+p_values_wk1 <-drop_na(p_values_wk1)
+p_values_wk1
+
+```
+
+2 WEEKS AHEAD - Wilcoxon test with Holm p-value adjustment
 
@@
 #########################################################################
-p_df2 <- data.frame(Model = model_types, PValue = p_values, WeekAhead=2)
-p_df2 <-drop_na(p_df2)
-p_df2
-
-```
-
-3 WEEKS AHEAD Post Hoc Wilcox test
+p_values_wk2 <- data.frame(Model = model_types, PValue = p_values, WeekAhead=2)
+p_values_wk2 <-drop_na(p_values_wk2)
+p_values_wk2
+
+```
+
+3 WEEKS AHEAD - Wilcoxon test with Holm p-value adjustment
 
@@
 #########################################################################
-p_df3 <- data.frame(Model = model_types, PValue = p_values, WeekAhead=3)
-p_df3 <-drop_na(p_df3)
-p_df3
-```
-
-4 WEEKS AHEAD Post Hoc Wilcox test
+p_values_wk3 <- data.frame(Model = model_types, PValue = p_values, WeekAhead=3)
+p_values_wk3 <-drop_na(p_values_wk3)
+p_values_wk3
+```
+
+4 WEEKS AHEAD - Wilcoxon test with Holm p-value adjustment
 
@@
 
-# View results
-print(wilcox_results)
-
+# Perform Wilcoxon test for each model type against AUTO_AAR
 p_values<-wilcox_results$p_value 
-
 p_values<-p.adjust(p_values, method = "holm")
-
 p_values<-data.frame(p_values)
@@
 #########################################################################
-p_df4 <- data.frame(Model = model_types, PValue = p_values, WeekAhead=4)
-p_df4 <-drop_na(p_df4)
-p_df4
+p_values_wk4 <- data.frame(Model = model_types, PValue = p_values, WeekAhead=4)
+p_values_wk4 <-drop_na(p_values_wk4)
+p_values_wk4
 
@@
   return(data.frame(
-    AUTO_AR_NoLg = (data$AUTO_AR_NoLg - data$AUTO_AR_NoLg),
-    ES27_AR_NoLg = ((data$AUTO_AR_NoLg - data$ES27_AR_NoLg) / data$AUTO_AR_NoLg) * 100,
-    ES64_AR_NoLg = ((data$AUTO_AR_NoLg - data$ES64_AR_NoLg) / data$AUTO_AR_NoLg) * 100,
-    AUTO_ADJ_NoLg = ((data$AUTO_AR_NoLg - data$AUTO_ADJ_NoLg) / data$AUTO_AR_NoLg) * 100,
-    ES27_ADJ_NoLg = ((data$AUTO_AR_NoLg - data$ES27_ADJ_NoLg) / data$AUTO_AR_NoLg) * 100,
-    ES64_ADJ_NoLg = ((data$AUTO_AR_NoLg - data$ES64_ADJ_NoLg) / data$AUTO_AR_NoLg) * 100,
-    AUTO_TMP_NoLg = ((data$AUTO_AR_NoLg - data$AUTO_TMP_NoLg) / data$AUTO_AR_NoLg) * 100,
-    ES27_TMP_NoLg = ((data$AUTO_AR_NoLg - data$ES27_TMP_NoLg) / data$AUTO_AR_NoLg) * 100,
-    ES64_TMP_NoLg = ((data$AUTO_AR_NoLg - data$ES64_TMP_NoLg) / data$AUTO_AR_NoLg) * 100,
-    AUTO_EPI_NoLg = ((data$AUTO_AR_NoLg - data$AUTO_EPI_NoLg) / data$AUTO_AR_NoLg) * 100,
-    ES27_EPI_NoLg = ((data$AUTO_AR_NoLg - data$ES27_EPI_NoLg) / data$AUTO_AR_NoLg) * 100,
-    ES64_EPI_NoLg = ((data$AUTO_AR_NoLg - data$ES64_EPI_NoLg) / data$AUTO_AR_NoLg) * 100,
-
-    AUTO_AR_Lg = ((data$AUTO_AR_NoLg - data$AUTO_AR_Lg) / data$AUTO_AR_NoLg) * 100,
-    ES27_AR_Lg = ((data$AUTO_AR_NoLg - data$ES27_AR_Lg) / data$AUTO_AR_NoLg) * 100,
-    ES64_AR_Lg = ((data$AUTO_AR_NoLg - data$ES64_AR_Lg) / data$AUTO_AR_NoLg) * 100,
-    AUTO_ADJ_Lg = ((data$AUTO_AR_NoLg - data$AUTO_ADJ_Lg) / data$AUTO_AR_NoLg) * 100,
-    ES27_ADJ_Lg = ((data$AUTO_AR_NoLg - data$ES27_ADJ_Lg) / data$AUTO_AR_NoLg) * 100,
-    ES64_ADJ_Lg = ((data$AUTO_AR_NoLg - data$ES64_ADJ_Lg) / data$AUTO_AR_NoLg) * 100,
-    AUTO_TMP_Lg = ((data$AUTO_AR_NoLg - data$AUTO_TMP_Lg) / data$AUTO_AR_NoLg) * 100,
-    ES27_TMP_Lg = ((data$AUTO_AR_NoLg - data$ES27_TMP_Lg) / data$AUTO_AR_NoLg) * 100,
-    ES64_TMP_Lg = ((data$AUTO_AR_NoLg - data$ES64_TMP_Lg) / data$AUTO_AR_NoLg) * 100,
-    AUTO_EPI_Lg = ((data$AUTO_AR_NoLg - data$AUTO_EPI_Lg) / data$AUTO_AR_NoLg) * 100,
-    ES27_EPI_Lg = ((data$AUTO_AR_NoLg - data$ES27_EPI_Lg) / data$AUTO_AR_NoLg) * 100,
-    ES64_EPI_Lg = ((data$AUTO_AR_NoLg - data$ES64_EPI_Lg) / data$AUTO_AR_NoLg) * 100
+    AUTO_AR_NoLg = (data$AUTO_AR_NoLg / data$AUTO_AR_NoLg),
+
+    ES27_AR_NoLg = ((1 - (data$ES27_AR_NoLg/data$AUTO_AR_NoLg)) * 100),
+    ES64_AR_NoLg = ((1 - (data$ES64_AR_NoLg/data$AUTO_AR_NoLg )) * 100),
+
+    AUTO_ADJ_NoLg = ((1 - (data$AUTO_ADJ_NoLg/data$AUTO_AR_NoLg)) * 100),
+    ES27_ADJ_NoLg = ((1 - (data$ES27_ADJ_NoLg/data$AUTO_AR_NoLg)) * 100),
+    ES64_ADJ_NoLg = ((1 - (data$ES64_ADJ_NoLg/data$AUTO_AR_NoLg)) * 100),
+
+    AUTO_TMP_NoLg = ((1 - (data$AUTO_TMP_NoLg/data$AUTO_AR_NoLg)) * 100),
+    ES27_TMP_NoLg = ((1 - (data$ES27_TMP_NoLg/data$AUTO_AR_NoLg)) * 100),
+    ES64_TMP_NoLg = ((1 - (data$ES64_TMP_NoLg/data$AUTO_AR_NoLg)) * 100),
+
+    AUTO_EPI_NoLg = ((1 - (data$AUTO_EPI_NoLg/data$AUTO_AR_NoLg)) * 100),
+    ES27_EPI_NoLg = ((1 - (data$ES27_EPI_NoLg/data$AUTO_AR_NoLg)) * 100),
+    ES64_EPI_NoLg = ((1 - (data$ES64_EPI_NoLg/data$AUTO_AR_NoLg)) * 100),
+
+    AUTO_AVG_NoLg = ((1 - (data$AUTO_AVG_NoLg/data$AUTO_AR_NoLg)) * 100),
+    ES27_AVG_NoLg = ((1 - (data$ES27_AVG_NoLg/data$AUTO_AR_NoLg)) * 100),
+    ES64_AVG_NoLg = ((1 - (data$ES64_AVG_NoLg/data$AUTO_AR_NoLg)) * 100),
+
+    AUTO_AR_Lg = ((1 - (data$AUTO_AR_Lg/data$AUTO_AR_NoLg)) * 100),
+    ES27_AR_Lg = ((1 - (data$ES27_AR_Lg/data$AUTO_AR_NoLg)) * 100),
+    ES64_AR_Lg = ((1 - (data$ES64_AR_Lg/data$AUTO_AR_NoLg)) * 100),
+
+    AUTO_ADJ_Lg = ((1 - (data$AUTO_ADJ_Lg/data$AUTO_AR_NoLg)) * 100),
+    ES27_ADJ_Lg = ((1 - (data$ES27_ADJ_Lg/data$AUTO_AR_NoLg)) * 100),
+    ES64_ADJ_Lg = ((1 - (data$ES64_ADJ_Lg/data$AUTO_AR_NoLg)) * 100),
+
+    AUTO_TMP_Lg = ((1 - (data$AUTO_TMP_Lg/data$AUTO_AR_NoLg)) * 100),
+    ES27_TMP_Lg = ((1 - (data$ES27_TMP_Lg/data$AUTO_AR_NoLg)) * 100),
+    ES64_TMP_Lg = ((1 - (data$ES64_TMP_Lg/data$AUTO_AR_NoLg)) * 100),
+
+    AUTO_EPI_Lg = ((1 - (data$AUTO_EPI_Lg/data$AUTO_AR_NoLg)) * 100),
+    ES27_EPI_Lg = ((1 - (data$ES27_EPI_Lg/data$AUTO_AR_NoLg)) * 100),
+    ES64_EPI_Lg = ((1 - (data$ES64_EPI_Lg/data$AUTO_AR_NoLg)) * 100),
+
+    AUTO_AVG_Lg = ((1 - (data$AUTO_AVG_Lg/data$AUTO_AR_NoLg)) * 100),
+    ES27_AVG_Lg = ((1 - (data$ES27_AVG_Lg/data$AUTO_AR_NoLg)) * 100),
+    ES64_AVG_Lg = ((1 - (data$ES64_AVG_Lg/data$AUTO_AR_NoLg)) * 100)    
   ))
@@
 
+W4_percentage_of_improvement
 ```
@@
 # List of column names
-#variables <- c("ES27_AR", "ES64_AR", "AUTO_TMP", "ES27_TMP", "ES64_TMP", 
-#               "AUTO_EPI", "ES27_EPI", "ES64_EPI", "AUTO_ADJ", "ES27_ADJ", "ES64_ADJ")
 variables <-c(names(W1_percentage_of_improvement))
@@
 
-Summary of Post Hoc Wilcox test with  p.adjust.method Benjamini-Yekutieli (BY)
-
-```{r}
-
-all_p_values<-rbind(p_df1,p_df2,p_df3,p_df4)
-
-#names(all_p_values)[names(all_p_values) == "Models"] <- "Model"
+Datasets with differences and Wilcoxon Holm adjusted p-values
+
+```{r}
+#
+
+all_p_values<-rbind(p_values_wk1,p_values_wk2,p_values_wk3,p_values_wk4)
 
@@
 ```
+
+Mean difference and standard deviation
 
@@
   scale_fill_gradient(low = "red", high = "cyan", na.value = "gray80", 
-                      name = "Mean Improvement", limits = c(-50, 50)) +  # Fixed color scale from -30 to 30
-  labs(title = "Model improvement (%) comparing each model' WIS with the baseline across 48 states",
+                      name = "Percentage (%)", limits = c(-35, 35)) +  # Fixed color scale from -30 to 30
+  labs(title = "Mean model improvement and standard deviation comparing each model' WIS with the baseline across 48 states",
        subtitle = "Gray boxes indicate models not significantly different from the baseline model based on Wilcoxon test with Holm's adjusted p-values",
@@
 
-Friedman test for ADJACENT MODELS
-
-```{r}
-
-library(ggplot2)
-library(reshape2)
-library(dplyr)
-library(tidyr)
-
-# List of datasets
-datasets <- list(W1 = W1, W2 = W2, W3 = W3, W4 = W4)
-
-# Function to compute Wilcoxon test and prepare the data for plotting
-run_pairwise_tests <- function(data, week) {
-  long_data <- data %>%
-    pivot_longer(cols = c(AUTO_ADJ_Lg, ES27_ADJ_Lg, ES64_ADJ_Lg),
-                 names_to = "Type",
-                 values_to = "Value")
-
-  pairwise_results <- pairwise.wilcox.test(long_data$Value, long_data$Type,
-                                           p.adjust.method = "holm", paired = TRUE)
-
-  # Convert p-value matrix to long format
-  p_df <- as.data.frame(as.table(pairwise_results$p.value)) %>%
-    rename(Comparison1 = Var1, Comparison2 = Var2, P_value = Freq) %>%
-    mutate(Significance = ifelse(P_value < 0.05, "Significant", "Not Significant"),
-           WeekAhead = week)  # Add a column for dataset identifier
-
-  return(p_df)
-}
-
-# Apply the function to all datasets and combine results
-p_values_df <- bind_rows(lapply(names(datasets), function(w) run_pairwise_tests(datasets[[w]], w)))
-
-adjacent_variance_test<-ggplot(p_values_df, aes(x = Comparison1, y = Comparison2, fill = Significance)) +
-  geom_tile(color = "black") +  # Add black gridlines
-  geom_text(aes(label = round(P_value, 3)), color = "black", size = 4) +  # Add p-value labels
-  scale_fill_manual(values = c("Significant" = "cyan", "Not Significant" = "gray80"), 
-                    name = "Significance", na.value = "white", drop = TRUE) +  # Remove NA from legend
-  labs(title = "ADJACENT model family pairwise Wilcoxon test",
-       subtitle = "Gray indicates not significant differences (Holm-adjusted p-values)",
-       x = "",
-       y = "") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  facet_wrap(~ WeekAhead)  # Separate plots for each dataset
-
-adjacent_variance_test
-
-ggsave(adjacent_variance_test,filename="Fig4.jpg", width = 6, height = 5)
-
-```
-
 Now let's plot a map of percentage of WIS improvement for each state.
@@
   STATE = W1$STATE,
-  percentage_improvement = W1_percentage_of_improvement$AUTO_ADJ_Lg,
-  AUTO_ADJ=W1$AUTO_ADJ_Lg
+  percentage_improvement = W1_percentage_of_improvement$AUTO_AVG_Lg,
+  AUTO_AVG=W1$AUTO_AVG_Lg
 )
@@
   STATE = W2$STATE,
-  percentage_improvement = W2_percentage_of_improvement$AUTO_ADJ_Lg,
-  AUTO_ADJ=W2$AUTO_ADJ_Lg
+  percentage_improvement = W2_percentage_of_improvement$AUTO_AVG_Lg,
+  AUTO_AVG=W2$AUTO_AVG_Lg
 )
@@
   STATE = W3$STATE,
-  percentage_improvement = W3_percentage_of_improvement$AUTO_ADJ_Lg,
-  AUTO_ADJ=W3$AUTO_ADJ_Lg
+  percentage_improvement = W3_percentage_of_improvement$AUTO_AVG_Lg,
+  AUTO_AVG=W3$AUTO_AVG_Lg
 )
@@
   STATE = W4$STATE,
-  percentage_improvement = W4_percentage_of_improvement$AUTO_ADJ_Lg,
-  AUTO_ADJ=W4$AUTO_ADJ_Lg
+  percentage_improvement = W4_percentage_of_improvement$AUTO_AVG_Lg,
+  AUTO_AVG=W4$AUTO_AVG_Lg
 )
@@
 ES_1WEEK <- ggplot(W1_map) +
-  geom_sf(aes(fill = AUTO_ADJ)) +  # Fill based on the best model
-  scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 300)) +  
+  geom_sf(aes(fill = AUTO_AVG)) +  # Fill based on the best model
+  scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
   ggtitle("Mean WIS on each state (W1)") +
@@
   theme(legend.position = "right") +
-  geom_sf_text(data = W1_map, aes(label = round(AUTO_ADJ,1)),  # Round to whole numbers
+  geom_sf_text(data = W1_map, aes(label = round(AUTO_AVG,1)),  # Round to whole numbers
                size = 3,
@@
 
+ES_1WEEK
 ggsave(ES_1WEEK, height = 4.5, width = 6.5,  filename="Fig4A.jpg")  
@@
 ES_2WEEK <- ggplot(W2_map) +
-  geom_sf(aes(fill = AUTO_ADJ)) +  # Fill based on the best model
-  scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 300)) +  
+  geom_sf(aes(fill = AUTO_AVG)) +  # Fill based on the best model
+  scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
   ggtitle("Mean WIS on each state (W2)") +
@@
   theme(legend.position = "right") +
-  geom_sf_text(data = W2_map, aes(label = round(AUTO_ADJ,1)),  # Round to whole numbers
+  geom_sf_text(data = W2_map, aes(label = round(AUTO_AVG,1)),  # Round to whole numbers
                size = 3,
@@
 
+ES_2WEEK
 ggsave(ES_2WEEK, height = 4.5, width = 6.5,  filename="Fig4B.jpg")  
@@
 ES_3WEEK <- ggplot(W3_map) +
-  geom_sf(aes(fill = AUTO_ADJ)) +  # Fill based on the best model
-  scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 300)) +  
+  geom_sf(aes(fill = AUTO_AVG)) +  # Fill based on the best model
+  scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
   ggtitle("Mean WIS on each state (W3)") +
@@
   theme(legend.position = "right") +
-  geom_sf_text(data = W3_map, aes(label = round(AUTO_ADJ,1)),  # Round to whole numbers
+  geom_sf_text(data = W3_map, aes(label = round(AUTO_AVG,1)),  # Round to whole numbers
                size = 3,
@@
 
+ES_3WEEK
+
 ggsave(ES_3WEEK, height = 4.5, width = 6.5,  filename="Fig4C.jpg")  
@@
 ES_4WEEK <- ggplot(W4_map) +
-  geom_sf(aes(fill = AUTO_ADJ)) +
+  geom_sf(aes(fill = AUTO_AVG)) +
   scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", 
-                       midpoint = 100, limits = c(0, 300)) +  
+                       midpoint = 100, limits = c(0, 352)) +  
   ggtitle("Mean WIS on each state (W4)") +
@@
   theme(legend.position = "right") +
-  geom_sf_text(data = W4_map, aes(label = round(AUTO_ADJ,1)),  
+  geom_sf_text(data = W4_map, aes(label = round(AUTO_AVG,1)),  
                size = 3,
@@
 
+ES_4WEEK
 ggsave(ES_4WEEK,height = 4.5, width = 6.5,  filename="Fig4D.jpg")  
 ```
-
 Loading datasets for regression analysis
@@
   # Fitting the regression model
-  model <- lm((WIS_pop_data$AUTO_ADJ) ~ (WIS_pop_data$Resident_population_2020))
+  model <- lm((WIS_pop_data$AUTO_AVG) ~ (WIS_pop_data$Resident_population_2020))
   # View the model summary
@@
   # Fit the regression model
-  model <- lm((WIS_pop_data$AUTO_ADJ) ~ (WIS_pop_data$Population_density_2020))
+  model <- lm((WIS_pop_data$AUTO_AVG) ~ (WIS_pop_data$Population_density_2020))
   # View the model summary
@@
 ```{r}
+
 #############################################################
@@
 
-AUTO_ADJ WIS x SOVI
+AUTO_AVG WIS x SOVI
 
@@
   # Fit the regression model
-  model <- lm((WIS_sovi$AUTO_ADJ) ~ (WIS_sovi$weighted_mean))
+  model <- lm((WIS_sovi$AUTO_AVG) ~ (WIS_sovi$weighted_mean))
   # View the model summary
@@
 
-AUTO_ADJ WIS x BRIC SOCIAL
+AUTO_AVG WIS x BRIC SOCIAL
 
@@
   # Fit the regression model  
-  model <- lm((WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.social))
+  model <- lm((WIS_bric$AUTO_AVG) ~ (WIS_bric$weighted_mean_z_bric.social))
   # View the model summary
@@
 
-AUTO_ADJ WIS x BRIC ECONOMIC
+AUTO_AVG WIS x BRIC ECONOMIC
 
@@
   # Fit the regression model
-  model <- lm((WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.economic))
+  model <- lm((WIS_bric$AUTO_AVG) ~ (WIS_bric$weighted_mean_z_bric.economic))
   # View the model summary
@@
 
-AUTO_ADJ WIS x BRIC Infrastructure
+AUTO_AVG WIS x BRIC Infrastructure
 
@@
   # Fit the regression model
-  model <- lm((WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.infrastructure))
+  model <- lm((WIS_bric$AUTO_AVG) ~ (WIS_bric$weighted_mean_z_bric.infrastructure))
   # View the model summary
@@
 ```
-AUTO_ADJ WIS x BRIC institutional
+
+AUTO_AVG WIS x BRIC institutional
 
@@
   # Fit the regression model
-  model <- lm((WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.institutional))
+  model <- lm((WIS_bric$AUTO_AVG) ~ (WIS_bric$weighted_mean_z_bric.institutional))
   # View the model summary
@@
 
-AUTO_ADJ WIS x BRIC community
+AUTO_AVG WIS x BRIC community
 
@@
   # Fit the regression model
-  model <- lm((WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.community))
+  model <- lm((WIS_bric$AUTO_AVG) ~ (WIS_bric$weighted_mean_z_bric.community))
   # View the model summary
@@
 
-AUTO_ADJ WIS x BRIC environment
+AUTO_AVG WIS x BRIC environment
 
@@
   # Fit the regression model
-  model <- lm((WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.environment))
+  model <- lm((WIS_bric$AUTO_AVG) ~ (WIS_bric$weighted_mean_z_bric.environment))
   # View the model summary
@@
 ```
-AUTO_ADJ WIS x BRIC total
+
+AUTO_AVG WIS x BRIC total
 
@@
   # Fit the regression model
-  model <- lm((WIS_bric$AUTO_ADJ) ~ (WIS_bric$weighted_mean_z_bric.total))
+  model <- lm((WIS_bric$AUTO_AVG) ~ (WIS_bric$weighted_mean_z_bric.total))
   # View the model summary
@@
 
-AUTO_ADJ WIS x Temperature ERA5 Data
+AUTO_AVG WIS x Temperature ERA5 Data
 
@@
   # Fit the regression model
-  model <- lm((WIS_temp$AUTO_ADJ) ~ (WIS_temp$mean))
+  model <- lm((WIS_temp$AUTO_AVG) ~ (WIS_temp$mean))
   # View the model summary
@@
 
-AUTO_ADJ WIS x Specific Humidity ERA5 Data
+AUTO_AVG WIS x Specific Humidity ERA5 Data
 
@@
   # Fit the regression model
-  model <- lm((WIS_humidity$AUTO_ADJ) ~ (WIS_humidity$mean))
+  model <- lm((WIS_humidity$AUTO_AVG) ~ (WIS_humidity$mean))
   # View the model summary
@@
 
-Let's plot some maps with the with the mean (%) of difference between the AUTO_ADJ and the AUTO ARIMA models for the same state.
+Let's plot some maps with the with the mean (%) of difference between the AUTO_AVG and the AUTO ARIMA models for the same state.
 
@@
   geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
-  scale_fill_gradient2(high = "darkred", mid = "lightyellow", low = "blue", midpoint = 0, limits = c(-120, 10)) +  
+  scale_fill_gradient2(low = "darkred", mid = "lightyellow", high = "skyblue", midpoint = 0, limits = c(-30, 60)) +  
   ggtitle("Mean WIS improvement in percentage (W1)") +
@@
     fill = "%",  # Label for the legend
-    subtitle = "(Baseline - AUTO_ADJ_LB) / (Baseline*100)"
+    subtitle = "(1-(Baseline/AUTO_AVG_LB))*100"
             , 
@@
 ES_1WEEK
+
 ggsave(ES_1WEEK,height = 4.5, width = 6.5,  filename="Fig5A.jpg")  
 ```
-
 2 weeks ahead
@@
   geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
-  scale_fill_gradient2(high = "darkred", mid = "lightyellow", low = "blue", midpoint = 0, limits = c(-120, 10)) +  
+  scale_fill_gradient2(low = "darkred", mid = "lightyellow", high = "skyblue", midpoint = 0, limits = c(-30, 60)) +  
   ggtitle("Mean WIS improvement in percentage (W2)") +
@@
 ES_2WEEK
+
 ggsave(ES_2WEEK, height = 4.5, width = 6.5, filename="Fig5B.jpg")  
@@
 # Assuming merged_data already includes the necessary columns: model and percentage_improveme
-# Create the map plot for 1 Week Ahead with best models and percentage improvement
-# Create the map plot for 1 Week Ahead with best models and percentage improvement
+
 ES_3WEEK <- ggplot(W3_map) +
   geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
-  scale_fill_gradient2(high = "darkred", mid = "lightyellow", low = "blue", midpoint = 0, limits = c(-120, 10)) +  
+  scale_fill_gradient2(low = "darkred", mid = "lightyellow", high = "skyblue", midpoint = 0, limits = c(-30, 60)) +  
   ggtitle("Mean WIS improvement in percentage (W3)") +
@@
 ES_3WEEK
+
 ggsave(ES_3WEEK, height = 4.5, width = 6.5, filename="Fig5C.jpg")  
@@
 # Assuming merged_data already includes the necessary columns: model and percentage_improveme
-# Create the map plot for 1 Week Ahead with best models and percentage improvement
+
 ES_4WEEK <- ggplot(W4_map) +
   geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
-  scale_fill_gradient2(high = "darkred", mid = "lightyellow", low = "blue", midpoint = 0, limits = c(-120, 10)) +  
+  scale_fill_gradient2(low = "darkred", mid = "lightyellow", high = "skyblue", midpoint = 0, limits = c(-30, 60)) +  
   ggtitle("Mean WIS improvement in percentage (W4)") +
@@
 ES_4WEEK
+
 ggsave(ES_4WEEK, height = 4.5, width = 6.5,  filename="Fig5D.jpg")  
@@
 
-AUTO_ADJ WIS x RESIDENT POPULATION 2020
+AUTO_AVG WIS x RESIDENT POPULATION 2020
 
@@
 
-AUTO_ADJ WIS x POPULATION DENSITY
+AUTO_AVG WIS x POPULATION DENSITY
 
@@
 
-AUTO_ADJ WIS x BRIC institutional
+AUTO_AVG WIS x BRIC institutional
 
@@
 
-AUTO_ADJ WIS x Specific Humidity ERA5 Data
-
-```{r}
-library(ggpubr)
+AUTO_AVG WIS x Specific Humidity ERA5 Data
+
+```{r}
 
@@
 ```{r}
-
-# WEEK 1
-W1_AUTO_AR_NoLg_WIS_BY_EPI <- filtered_df_W1_NoLg %>%
-  group_by(epiweek) %>%
-  summarize(AUTO_AR_NoLg = mean(AUTO_AR_NoLg, na.rm = TRUE))
-
-W1_AUTO_ADJ_WIS_BY_EPI <- filtered_df_W1 %>%
-  group_by(epiweek) %>%
-  summarize(AUTO_ADJ = mean(AUTO_ADJ, na.rm = TRUE))
-# WEEK 2
-W2_AUTO_AR_NoLg_WIS_BY_EPI <- filtered_df_W2_NoLg %>%
-  group_by(epiweek) %>%
-  summarize(AUTO_AR_NoLg = mean(AUTO_AR_NoLg, na.rm = TRUE))
-
-W2_AUTO_ADJ_WIS_BY_EPI <- filtered_df_W2 %>%
-  group_by(epiweek) %>%
-  summarize(AUTO_ADJ = mean(AUTO_ADJ, na.rm = TRUE))
-
-# WEEK 3
-W3_AUTO_AR_NoLg_WIS_BY_EPI <- filtered_df_W3_NoLg %>%
-  group_by(epiweek) %>%
-  summarize(AUTO_AR_NoLg = mean(AUTO_AR_NoLg, na.rm = TRUE))
-
-W3_AUTO_ADJ_WIS_BY_EPI <- filtered_df_W3 %>%
-  group_by(epiweek) %>%
-  summarize(AUTO_ADJ = mean(AUTO_ADJ, na.rm = TRUE))
-
-# WEEK 4
-W4_AUTO_AR_NoLg_WIS_BY_EPI <- filtered_df_W4_NoLg %>%
-  group_by(epiweek) %>%
-  summarize(AUTO_AR_NoLg = mean(AUTO_AR_NoLg, na.rm = TRUE))
-
-W4_AUTO_ADJ_WIS_BY_EPI <- filtered_df_W4 %>%
-  group_by(epiweek) %>%
-  summarize(AUTO_ADJ = mean(AUTO_ADJ, na.rm = TRUE))
-
-filtered_df_W1
-filtered_df_W2
-
-```
-
-Absolute WIS difference by epidemiological week
-
-```{r}
-
-# Define pairs of data frames
-auto_df <- list(W1_AUTO_AR_NoLg_WIS_BY_EPI, W2_AUTO_AR_NoLg_WIS_BY_EPI, W3_AUTO_AR_NoLg_WIS_BY_EPI, W4_AUTO_AR_NoLg_WIS_BY_EPI)
-adj_df <- list(W1_AUTO_ADJ_WIS_BY_EPI, W2_AUTO_ADJ_WIS_BY_EPI, W3_AUTO_ADJ_WIS_BY_EPI, W4_AUTO_ADJ_WIS_BY_EPI)
-week_labels <- c("W1", "W2", "W3", "W4")  # Labels for plots
-
-# Function to process each pair
-for (i in 1:4) {
-  merged_df <- merge(auto_df[[i]], adj_df[[i]], by = "epiweek") %>%
-    mutate(percentage_diff = AUTO_AR_NoLg - AUTO_ADJ) %>%
-    filter((epiweek >= 40 & epiweek <= 53) | (epiweek >= 1 & epiweek <= 20)) %>%
-    mutate(epiweek_order = factor(epiweek, levels = c(40:53, 1:20)))
-
-  mean_value <- mean(merged_df$percentage_diff, na.rm = TRUE)
-
-  plots <- ggplot(merged_df, aes(x = epiweek_order, y = percentage_diff, fill = percentage_diff > 0)) +
-    geom_bar(stat = "identity") +
-    scale_fill_manual(values = c("TRUE" = "red", "FALSE" = "blue"), guide = "none") +  # Blue for positive, Red for negative
-    geom_hline(yintercept = mean_value, linetype = "dashed", color = "black") +
-    labs(x = "Epiweek", y = "Absolute WIS difference",
-         title = paste("Mean absolute WIS improvement by epidemiological week across 48 states (", week_labels[i],")"),
-         subtitle = "Baseline - AUTO_ADJ_LB") +
-    theme_minimal() +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-    annotate("text", x = 1, y = mean_value, label = paste("Mean:", round(mean_value, 2)), color = "black", vjust = -1)
-
-  print(plots)  # Print each plot inside the loop
-
-    # Save the plot with a unique filename
-  ggsave(filename = paste0("Fig6_", week_labels[i], ".png"),
-         plot = plots, width = 8, height = 6, dpi = 300)
-}
-
-```
-
-Absolute WIS difference by epidemiological week
-
-```{r}
-library(dplyr)
-library(ggplot2)
-
-# Define lists of data frames
-auto_df <- list(filtered_df_W1_NoLg, filtered_df_W2_NoLg, filtered_df_W3_NoLg, filtered_df_W4_NoLg)
-adj_df <- list(filtered_df_W1, filtered_df_W2, filtered_df_W3, filtered_df_W4)
-
-data__<-merge(filtered_df_W1_NoLg,filtered_df_W1)
-
-library(dplyr)
-
-# Summing values by Julian date across all states
-summed_data <- data__ %>%
-  group_by(Julian_date) %>%  # Group only by Julian date
-  summarize(
-    AUTO_AR_NoLg = sum(AUTO_AR_NoLg, na.rm = TRUE),  # Sum AUTO_AR_NoLg across all states
-    AUTO_ADJ = sum(AUTO_ADJ, na.rm = TRUE)           # Sum ES27_EPI across all states
-  )
-
-```
-
-```{r}
 library(dplyr)
@@
       AUTO_AR_NoLg = sum(AUTO_AR_NoLg, na.rm = TRUE),  # Sum across all states
-      AUTO_ADJ = sum(AUTO_ADJ, na.rm = TRUE)           # Sum across all states
+      AUTO_AVG = sum(AUTO_AVG, na.rm = TRUE)           # Sum across all states
     ) %>%
     # Calculate the difference ratio
-    mutate(Difference = 1 - (AUTO_AR_NoLg / AUTO_ADJ),
+    mutate(Difference = 1 - (AUTO_AVG / AUTO_AR_NoLg),
            Week = paste0("Week ", week))
@@
 # Plot the difference using ggplot2
-ggplot(summed_data_all, aes(x = Julian_date, y = Difference, color = Week)) +
-  geom_line(size = 1) +        # Line plot
+ggplot(summed_data_all, aes(x = Julian_date, y = (Difference)*100, color = Week)) +
+#  geom_line(size = 1) +        # Line plot
   geom_point(size = 2) +       # Points on the line
-  geom_hline(yintercept = 0, linetype = "dashed", color = "red", size = 1) +  # Reference line at 0
+#  geom_hline(yintercept = 0, linetype = "dashed", color = "red", size = 1) +  # Reference line at 0
   theme_minimal() +
   labs(
-    title = "Difference between AUTO_AR_NoLg and AUTO_ADJ across States for 4 Weeks",
+    title = "Percentage of WIS difference across all states AUTO_AVG_LB compared to baseline",
     x = "Julian Date",
-    y = "Difference (1 - AUTO_AR_NoLg / AUTO_ADJ)"
+    y = "Difference (1 - AUTO_AR_NoLg / AUTO_AVG)"
   ) +

### 2025-05-05 manuscript_results.Rmd
update

@@
 ggplot(summed_data_all, aes(x = Julian_date, y = (Difference)*100, color = Week)) +
-#  geom_line(size = 1) +        # Line plot
-  geom_point(size = 2) +       # Points on the line
-#  geom_hline(yintercept = 0, linetype = "dashed", color = "red", size = 1) +  # Reference line at 0
+  geom_point(size = 2) +
+  geom_hline(yintercept = 0, color = "black", linetype = "dashed", size = 0.8) +
   theme_minimal() +
   labs(
-    title = "Percentage of WIS difference across all states AUTO_AVG_LB compared to baseline",
+    title = "Mean percentage of WIS difference across all states",
+    subtitle =  "AUTO_AVG_LB compared to baseline",
     x = "Julian Date",
-    y = "Difference (1 - AUTO_AR_NoLg / AUTO_AVG)"
+    y = "Difference (%)"
   ) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
+  scale_y_continuous(
+    limits = c(-100, 100),  # adjust if necessary
+    breaks = seq(-100, 100, by = 20)
+  ) +
+  scale_x_date(
+    date_breaks = "1 month",
+    date_labels = "%b %y"
+  ) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_color_manual(
+    values = c("Week 1" = "#ffcccc",
+               "Week 2" = "#ff6666",
+               "Week 3" = "#cc0000",
+               "Week 4" = "#990000")
+  )
+```

### 2025-05-05 manuscript_results.Rmd
update

--- 2025-04-25 manuscript_results.Rmd
@@
        y = "Independent Variables",
-       fill = "Significance") +  
+       fill = "Significance (p < 0.05)") +  
   theme_minimal() +  
@@
        y = "Independent Variables",
-       fill = "Significance") +    
+       fill = "Significance (p < 0.05)") +    
   theme_minimal() +  
@@
 ggplot(summed_data_all, aes(x = Julian_date, y = (Difference)*100, color = Week)) +
-#  geom_line(size = 1) +        # Line plot
-  geom_point(size = 2) +       # Points on the line
-#  geom_hline(yintercept = 0, linetype = "dashed", color = "red", size = 1) +  # Reference line at 0
+  geom_point(size = 3) +
+  geom_hline(yintercept = 0, color = "black", linetype = "dashed", size = 0.8) +
   theme_minimal() +
   labs(
-    title = "Percentage of WIS difference across all states AUTO_AVG_LB compared to baseline",
-    x = "Julian Date",
-    y = "Difference (1 - AUTO_AR_NoLg / AUTO_AVG)"
+    title = "Mean percentage of WIS differences across all states",
+    subtitle =  "AUTO_AVG_LB compared to baseline",
+    x = "",
+    y = "Difference (%)"
   ) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
+  scale_y_continuous(
+    limits = c(-100, 100),  # adjust if necessary
+    breaks = seq(-100, 100, by = 20)
+  ) +
+  scale_x_date(
+    date_breaks = "1 month",
+    date_labels = "%b %y"
+  ) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_color_manual(
+    values = c("Week 1" = "#ffcccc",
+               "Week 2" = "#ff6666",
+               "Week 3" = "#cc0000",
+               "Week 4" = "#990000")
+  )
+```

### 2025-05-05 manuscript_results.Rmd
update

@@
                       name = "Percentage (%)", limits = c(-35, 35)) +  # Fixed color scale from -30 to 30
-  labs(title = "Mean model improvement and standard deviation comparing each model' WIS with the baseline across 48 states",
-       subtitle = "Gray boxes indicate models not significantly different from the baseline model based on Wilcoxon test with Holm's adjusted p-values",
+  labs(title = "Figure X: Mean differences between models' WIS and the baseline across 48 states",
+       subtitle = "Gray boxes indicate models that are not significantly different from the baseline (p>0.05)",
+         caption = "m = mean difference, sd = standard deviation",
        x = "Week Ahead",
@@
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+  plot.caption = element_text(size = 12),
+    plot.title = element_text(size = 18, face = "bold"),
+    plot.subtitle = element_text(size = 14),) +
   facet_wrap(~ Model_Type, scales = "free_y")  # Separate NoLg and other models
@@
 # Plot the difference using ggplot2
-ggplot(summed_data_all, aes(x = Julian_date, y = (Difference)*100, color = Week)) +
+model_improv<-ggplot(summed_data_all, aes(x = Julian_date, y = (Difference)*100, color = Week)) +
   geom_point(size = 3) +
@@
   labs(
-    title = "Mean percentage of WIS differences across all states",
+    title = "Figure X: Mean percentage of WIS differences across all states",
     subtitle = "AUTO_AVG_LB compared to baseline",
@@
   scale_y_continuous(
-    limits = c(-100, 100),  # adjust if necessary
+    limits = c(-100, 100),
     breaks = seq(-100, 100, by = 20)
@@
   ) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
   scale_color_manual(
@@
                "Week 4" = "#990000")
+  ) +
+  theme(
+    plot.title = element_text(size = 18, face = "bold"),
+    plot.subtitle = element_text(size = 14),
+    axis.text.x = element_text(angle = 45, hjust = 1)
   )
-```
+
+model_improv
+
+ggsave(model_improv, height = 4.5, width = 9,  filename="Fig6.jpg")  
+```

### 2025-05-05 manuscript_results.Rmd
update

@@
 
-  Models = rep(c("AUTO_AR_Lg", "ES27_AR_Lg", "ES64_AR_Lg", "AUTO_ADJ_Lg", "ES27_ADJ_Lg", "ES64_ADJ_Lg", "AUTO_TMP_Lg", "ES27_TMP_Lg", "ES64_TMP_Lg", "AUTO_EPI_Lg", "ES27_EPI_Lg", "ES64_EPI_Lg", "AUTO_AVG_Lg", "ES27_AVG_Lg", "ES64_AVG_Lg"), each=48, times = 4),
+  Models = rep(c("AUTO_AR_LB", "ES27_AR_LB", "ES64_AR_LB", "AUTO_ADJ_LB", "ES27_ADJ_LB", "ES64_ADJ_LB", "AUTO_TMP_LB", "ES27_TMP_LB", "ES64_TMP_LB", "AUTO_EPI_LB", "ES27_EPI_LB", "ES64_EPI_LB", "AUTO_AVG_LB", "ES27_AVG_LB", "ES64_AVG_LB"), each=48, times = 4),
 
@@
             W4_Lg$AUTO_AVG_Lg, W4_Lg$ES27_AVG_Lg, W4_Lg$ES64_AVG_Lg)
-
 )
@@
 custom_colors <- c(
-  "AUTO_AR_Lg" = colors_ar[1], "ES27_AR_Lg" = colors_ar[2], "ES64_AR_Lg" = colors_ar[3],
-  "AUTO_EPI_Lg" = colors_epi[1], "ES27_EPI_Lg" = colors_epi[2], "ES64_EPI_Lg" = colors_epi[3],
-  "AUTO_TMP_Lg" = colors_tmp[1], "ES27_TMP_Lg" = colors_tmp[2], "ES64_TMP_Lg" = colors_tmp[3],
-  "AUTO_ADJ_Lg" = colors_adj[1], "ES27_ADJ_Lg" = colors_adj[2], "ES64_ADJ_Lg" = colors_adj[3],
-  "AUTO_AVG_Lg" = colors_avg[1], "ES27_AVG_Lg" = colors_avg[2], "ES64_AVG_Lg" = colors_avg[3]
+  "AUTO_AR_LB" = colors_ar[1], "ES27_AR_LB" = colors_ar[2], "ES64_AR_LB" = colors_ar[3],
+  "AUTO_EPI_LB" = colors_epi[1], "ES27_EPI_LB" = colors_epi[2], "ES64_EPI_LB" = colors_epi[3],
+  "AUTO_TMP_LB" = colors_tmp[1], "ES27_TMP_LB" = colors_tmp[2], "ES64_TMP_LB" = colors_tmp[3],
+  "AUTO_ADJ_LB" = colors_adj[1], "ES27_ADJ_LB" = colors_adj[2], "ES64_ADJ_LB" = colors_adj[3],
+  "AUTO_AVG_LB" = colors_avg[1], "ES27_AVG_LB" = colors_avg[2], "ES64_AVG_LB" = colors_avg[3]
 )

### 2025-05-05 manuscript_results.Rmd
figures update

@@
   epiweek = AUTO_ARIMA_WEEK1$epiweek,
-  AUTO_AR_NoLg=AUTO_ARIMA_WEEK1$WIS, 
-  AUTO_ADJ_NoLg=AUTO_ADJACENT_WEEK1$WIS, 
-  AUTO_EPI_NoLg=AUTO_EPIWEEK_WEEK1$WIS,
-  AUTO_TEMP_NoLg=AUTO_TEMPERATURE_WEEK1$WIS,
-  AUTO_AVG_NoLg=AUTO_AVERAGE_WEEK1$WIS,
-
-  ES27_AR_NoLg=ES27_ARIMA_WEEK1$WIS, 
-  ES27_ADJ_NoLg=ES27_ADJACENT_WEEK1$WIS, 
-  ES27_EPI_NoLg=ES27_EPIWEEK_WEEK1$WIS,
-  ES27_TEMP_NoLg=ES27_TEMPERATURE_WEEK1$WIS,
-  ES27_AVG_NoLg=ES27_AVERAGE_WEEK1$WIS,
-
-  ES64_AR_NoLg=ES64_ARIMA_WEEK1$WIS, 
-  ES64_ADJ_NoLg=ES64_ADJACENT_WEEK1$WIS, 
-  ES64_EPI_NoLg=ES64_EPIWEEK_WEEK1$WIS,
-  ES64_TEMP_NoLg=ES64_TEMPERATURE_WEEK1$WIS,
-  ES64_AVG_NoLg=ES64_AVERAGE_WEEK1$WIS
+  AUTO_AR=AUTO_ARIMA_WEEK1$WIS, 
+  AUTO_ADJ=AUTO_ADJACENT_WEEK1$WIS, 
+  AUTO_EPI=AUTO_EPIWEEK_WEEK1$WIS,
+  AUTO_TEMP=AUTO_TEMPERATURE_WEEK1$WIS,
+  AUTO_AVG=AUTO_AVERAGE_WEEK1$WIS,
+
+  ES27_AR=ES27_ARIMA_WEEK1$WIS, 
+  ES27_ADJ=ES27_ADJACENT_WEEK1$WIS, 
+  ES27_EPI=ES27_EPIWEEK_WEEK1$WIS,
+  ES27_TEMP=ES27_TEMPERATURE_WEEK1$WIS,
+  ES27_AVG=ES27_AVERAGE_WEEK1$WIS,
+
+  ES64_AR=ES64_ARIMA_WEEK1$WIS, 
+  ES64_ADJ=ES64_ADJACENT_WEEK1$WIS, 
+  ES64_EPI=ES64_EPIWEEK_WEEK1$WIS,
+  ES64_TEMP=ES64_TEMPERATURE_WEEK1$WIS,
+  ES64_AVG=ES64_AVERAGE_WEEK1$WIS
 )
@@
   epiweek = AUTO_ARIMA_WEEK2$epiweek,
-  AUTO_AR_NoLg=AUTO_ARIMA_WEEK2$WIS, 
-  AUTO_ADJ_NoLg=AUTO_ADJACENT_WEEK2$WIS, 
-  AUTO_EPI_NoLg=AUTO_EPIWEEK_WEEK2$WIS,
-  AUTO_TEMP_NoLg=AUTO_TEMPERATURE_WEEK2$WIS,
-  AUTO_AVG_NoLg=AUTO_AVERAGE_WEEK2$WIS,
-
-  ES27_AR_NoLg=ES27_ARIMA_WEEK2$WIS, 
-  ES27_ADJ_NoLg=ES27_ADJACENT_WEEK2$WIS, 
-  ES27_EPI_NoLg=ES27_EPIWEEK_WEEK2$WIS,
-  ES27_TEMP_NoLg=ES27_TEMPERATURE_WEEK2$WIS,
-  ES27_AVG_NoLg=ES27_AVERAGE_WEEK2$WIS,
-
-  ES64_AR_NoLg=ES64_ARIMA_WEEK2$WIS, 
-  ES64_ADJ_NoLg=ES64_ADJACENT_WEEK2$WIS, 
-  ES64_EPI_NoLg=ES64_EPIWEEK_WEEK2$WIS,
-  ES64_TEMP_NoLg=ES64_TEMPERATURE_WEEK2$WIS,
-  ES64_AVG_NoLg=ES64_AVERAGE_WEEK2$WIS
+  AUTO_AR=AUTO_ARIMA_WEEK2$WIS, 
+  AUTO_ADJ=AUTO_ADJACENT_WEEK2$WIS, 
+  AUTO_EPI=AUTO_EPIWEEK_WEEK2$WIS,
+  AUTO_TEMP=AUTO_TEMPERATURE_WEEK2$WIS,
+  AUTO_AVG=AUTO_AVERAGE_WEEK2$WIS,
+
+  ES27_AR=ES27_ARIMA_WEEK2$WIS, 
+  ES27_ADJ=ES27_ADJACENT_WEEK2$WIS, 
+  ES27_EPI=ES27_EPIWEEK_WEEK2$WIS,
+  ES27_TEMP=ES27_TEMPERATURE_WEEK2$WIS,
+  ES27_AVG=ES27_AVERAGE_WEEK2$WIS,
+
+  ES64_AR=ES64_ARIMA_WEEK2$WIS, 
+  ES64_ADJ=ES64_ADJACENT_WEEK2$WIS, 
+  ES64_EPI=ES64_EPIWEEK_WEEK2$WIS,
+  ES64_TEMP=ES64_TEMPERATURE_WEEK2$WIS,
+  ES64_AVG=ES64_AVERAGE_WEEK2$WIS
 
@@
   epiweek = AUTO_ARIMA_WEEK3$epiweek,
-  AUTO_AR_NoLg=AUTO_ARIMA_WEEK3$WIS, 
-  AUTO_ADJ_NoLg=AUTO_ADJACENT_WEEK3$WIS, 
-  AUTO_EPI_NoLg=AUTO_EPIWEEK_WEEK3$WIS,
-  AUTO_TEMP_NoLg=AUTO_TEMPERATURE_WEEK3$WIS,
-  AUTO_AVG_NoLg=AUTO_AVERAGE_WEEK3$WIS,
-
-  ES27_AR_NoLg=ES27_ARIMA_WEEK3$WIS, 
-  ES27_ADJ_NoLg=ES27_ADJACENT_WEEK3$WIS, 
-  ES27_EPI_NoLg=ES27_EPIWEEK_WEEK3$WIS,
-  ES27_TEMP_NoLg=ES27_TEMPERATURE_WEEK3$WIS,
-  ES27_AVG_NoLg=ES27_AVERAGE_WEEK3$WIS,
-
-  ES64_AR_NoLg=ES64_ARIMA_WEEK3$WIS, 
-  ES64_ADJ_NoLg=ES64_ADJACENT_WEEK3$WIS, 
-  ES64_EPI_NoLg=ES64_EPIWEEK_WEEK3$WIS,
-  ES64_TEMP_NoLg=ES64_TEMPERATURE_WEEK3$WIS,
-  ES64_AVG_NoLg=ES64_AVERAGE_WEEK3$WIS
+  AUTO_AR=AUTO_ARIMA_WEEK3$WIS, 
+  AUTO_ADJ=AUTO_ADJACENT_WEEK3$WIS, 
+  AUTO_EPI=AUTO_EPIWEEK_WEEK3$WIS,
+  AUTO_TEMP=AUTO_TEMPERATURE_WEEK3$WIS,
+  AUTO_AVG=AUTO_AVERAGE_WEEK3$WIS,
+
+  ES27_AR=ES27_ARIMA_WEEK3$WIS, 
+  ES27_ADJ=ES27_ADJACENT_WEEK3$WIS, 
+  ES27_EPI=ES27_EPIWEEK_WEEK3$WIS,
+  ES27_TEMP=ES27_TEMPERATURE_WEEK3$WIS,
+  ES27_AVG=ES27_AVERAGE_WEEK3$WIS,
+
+  ES64_AR=ES64_ARIMA_WEEK3$WIS, 
+  ES64_ADJ=ES64_ADJACENT_WEEK3$WIS, 
+  ES64_EPI=ES64_EPIWEEK_WEEK3$WIS,
+  ES64_TEMP=ES64_TEMPERATURE_WEEK3$WIS,
+  ES64_AVG=ES64_AVERAGE_WEEK3$WIS
 )
@@
   epiweek = AUTO_ARIMA_WEEK4$epiweek,
-  AUTO_AR_NoLg=AUTO_ARIMA_WEEK4$WIS, 
-  AUTO_ADJ_NoLg=AUTO_ADJACENT_WEEK4$WIS, 
-  AUTO_EPI_NoLg=AUTO_EPIWEEK_WEEK4$WIS,
-  AUTO_TEMP_NoLg=AUTO_TEMPERATURE_WEEK4$WIS,
-  AUTO_AVG_NoLg=AUTO_AVERAGE_WEEK4$WIS,
-
-  ES27_AR_NoLg=ES27_ARIMA_WEEK4$WIS, 
-  ES27_ADJ_NoLg=ES27_ADJACENT_WEEK4$WIS, 
-  ES27_EPI_NoLg=ES27_EPIWEEK_WEEK4$WIS,
-  ES27_TEMP_NoLg=ES27_TEMPERATURE_WEEK4$WIS,
-  ES27_AVG_NoLg=ES27_AVERAGE_WEEK4$WIS,
-
-  ES64_AR_NoLg=ES64_ARIMA_WEEK4$WIS, 
-  ES64_ADJ_NoLg=ES64_ADJACENT_WEEK4$WIS, 
-  ES64_EPI_NoLg=ES64_EPIWEEK_WEEK4$WIS,
-  ES64_TEMP_NoLg=ES64_TEMPERATURE_WEEK4$WIS,
-  ES64_AVG_NoLg=ES64_AVERAGE_WEEK4$WIS  
+  AUTO_AR=AUTO_ARIMA_WEEK4$WIS, 
+  AUTO_ADJ=AUTO_ADJACENT_WEEK4$WIS, 
+  AUTO_EPI=AUTO_EPIWEEK_WEEK4$WIS,
+  AUTO_TEMP=AUTO_TEMPERATURE_WEEK4$WIS,
+  AUTO_AVG=AUTO_AVERAGE_WEEK4$WIS,
+
+  ES27_AR=ES27_ARIMA_WEEK4$WIS, 
+  ES27_ADJ=ES27_ADJACENT_WEEK4$WIS, 
+  ES27_EPI=ES27_EPIWEEK_WEEK4$WIS,
+  ES27_TEMP=ES27_TEMPERATURE_WEEK4$WIS,
+  ES27_AVG=ES27_AVERAGE_WEEK4$WIS,
+
+  ES64_AR=ES64_ARIMA_WEEK4$WIS, 
+  ES64_ADJ=ES64_ADJACENT_WEEK4$WIS, 
+  ES64_EPI=ES64_EPIWEEK_WEEK4$WIS,
+  ES64_TEMP=ES64_TEMPERATURE_WEEK4$WIS,
+  ES64_AVG=ES64_AVERAGE_WEEK4$WIS  
 )
@@
 ) %>%
-  select(STATE, epiweek, AUTO_AR_NoLg, Dataset, Julian_date) %>%
-  rename(WIS = AUTO_AR_NoLg)  # Rename for consistency
+  select(STATE, epiweek, AUTO_AR, Dataset, Julian_date) %>%
+  rename(WIS = AUTO_AR)  # Rename for consistency
 
@@
     summarize(
-      AUTO_AR_NoLg = mean(AUTO_AR_NoLg, na.rm = TRUE),
-      AUTO_ADJ_NoLg = mean(AUTO_ADJ_NoLg, na.rm = TRUE),
-      AUTO_EPI_NoLg = mean(AUTO_EPI_NoLg, na.rm = TRUE),
-      AUTO_TMP_NoLg = mean(AUTO_TEMP_NoLg, na.rm = TRUE),
-      AUTO_AVG_NoLg = mean(AUTO_AVG_NoLg, na.rm = TRUE),
-
-      ES27_AR_NoLg = mean(ES27_AR_NoLg, na.rm = TRUE),
-      ES27_ADJ_NoLg = mean(ES27_ADJ_NoLg, na.rm = TRUE),
-      ES27_EPI_NoLg = mean(ES27_EPI_NoLg, na.rm = TRUE),
-      ES27_TMP_NoLg = mean(ES27_TEMP_NoLg, na.rm = TRUE),
-      ES27_AVG_NoLg = mean(ES27_AVG_NoLg, na.rm = TRUE),
-
-      ES64_AR_NoLg = mean(ES64_AR_NoLg, na.rm = TRUE),
-      ES64_ADJ_NoLg = mean(ES64_ADJ_NoLg, na.rm = TRUE),
-      ES64_EPI_NoLg = mean(ES64_EPI_NoLg, na.rm = TRUE),
-      ES64_TMP_NoLg = mean(ES64_TEMP_NoLg, na.rm = TRUE),
-      ES64_AVG_NoLg = mean(ES64_AVG_NoLg, na.rm = TRUE)
+      AUTO_AR = mean(AUTO_AR, na.rm = TRUE),
+      AUTO_ADJ = mean(AUTO_ADJ, na.rm = TRUE),
+      AUTO_EPI = mean(AUTO_EPI, na.rm = TRUE),
+      AUTO_TMP = mean(AUTO_TEMP, na.rm = TRUE),
+      AUTO_AVG = mean(AUTO_AVG, na.rm = TRUE),
+
+      ES27_AR = mean(ES27_AR, na.rm = TRUE),
+      ES27_ADJ = mean(ES27_ADJ, na.rm = TRUE),
+      ES27_EPI = mean(ES27_EPI, na.rm = TRUE),
+      ES27_TMP = mean(ES27_TEMP, na.rm = TRUE),
+      ES27_AVG = mean(ES27_AVG, na.rm = TRUE),
+
+      ES64_AR = mean(ES64_AR, na.rm = TRUE),
+      ES64_ADJ = mean(ES64_ADJ, na.rm = TRUE),
+      ES64_EPI = mean(ES64_EPI, na.rm = TRUE),
+      ES64_TMP = mean(ES64_TEMP, na.rm = TRUE),
+      ES64_AVG = mean(ES64_AVG, na.rm = TRUE)
     )
@@
 
-  value = c(W1_NoLg$AUTO_AR_NoLg, W1_NoLg$ES27_AR_NoLg, W1_NoLg$ES64_AR_NoLg, 
-            W1_NoLg$AUTO_ADJ_NoLg, W1_NoLg$ES27_ADJ_NoLg, W1_NoLg$ES64_ADJ_NoLg, 
-            W1_NoLg$AUTO_TMP_NoLg, W1_NoLg$ES27_TMP_NoLg, W1_NoLg$ES64_TMP_NoLg, 
-            W1_NoLg$AUTO_EPI_NoLg, W1_NoLg$ES27_EPI_NoLg, W1_NoLg$ES64_EPI_NoLg,
-            W1_NoLg$AUTO_AVG_NoLg, W1_NoLg$ES27_AVG_NoLg, W1_NoLg$ES64_AVG_NoLg,
-
-            W2_NoLg$AUTO_AR_NoLg, W2_NoLg$ES27_AR_NoLg, W2_NoLg$ES64_AR_NoLg, 
-            W2_NoLg$AUTO_ADJ_NoLg, W2_NoLg$ES27_ADJ_NoLg, W2_NoLg$ES64_ADJ_NoLg, 
-            W2_NoLg$AUTO_TMP_NoLg, W2_NoLg$ES27_TMP_NoLg, W2_NoLg$ES64_TMP_NoLg, 
-            W2_NoLg$AUTO_EPI_NoLg, W2_NoLg$ES27_EPI_NoLg, W2_NoLg$ES64_EPI_NoLg,
-            W2_NoLg$AUTO_AVG_NoLg, W2_NoLg$ES27_AVG_NoLg, W2_NoLg$ES64_AVG_NoLg,
-
-            W3_NoLg$AUTO_AR_NoLg, W3_NoLg$ES27_AR_NoLg, W3_NoLg$ES64_AR_NoLg,
-            W3_NoLg$AUTO_ADJ_NoLg, W3_NoLg$ES27_ADJ_NoLg, W3_NoLg$ES64_ADJ_NoLg, 
-            W3_NoLg$AUTO_TMP_NoLg, W3_NoLg$ES27_TMP_NoLg, W3_NoLg$ES64_TMP_NoLg, 
-            W3_NoLg$AUTO_EPI_NoLg, W3_NoLg$ES27_EPI_NoLg, W3_NoLg$ES64_EPI_NoLg,
-            W3_NoLg$AUTO_AVG_NoLg, W3_NoLg$ES27_AVG_NoLg, W3_NoLg$ES64_AVG_NoLg,
-
-            W4_NoLg$AUTO_AR_NoLg, W4_NoLg$ES27_AR_NoLg, W4_NoLg$ES64_AR_NoLg,
-            W4_NoLg$AUTO_ADJ_NoLg, W4_NoLg$ES27_ADJ_NoLg, W4_NoLg$ES64_ADJ_NoLg, 
-            W4_NoLg$AUTO_TMP_NoLg, W4_NoLg$ES27_TMP_NoLg, W4_NoLg$ES64_TMP_NoLg, 
-            W4_NoLg$AUTO_EPI_NoLg, W4_NoLg$ES27_EPI_NoLg, W4_NoLg$ES64_EPI_NoLg,
-            W4_NoLg$AUTO_AVG_NoLg, W4_NoLg$ES27_AVG_NoLg, W4_NoLg$ES64_AVG_NoLg
+  value = c(W1_NoLg$AUTO_AR, W1_NoLg$ES27_AR, W1_NoLg$ES64_AR, 
+            W1_NoLg$AUTO_ADJ, W1_NoLg$ES27_ADJ, W1_NoLg$ES64_ADJ, 
+            W1_NoLg$AUTO_TMP, W1_NoLg$ES27_TMP, W1_NoLg$ES64_TMP, 
+            W1_NoLg$AUTO_EPI, W1_NoLg$ES27_EPI, W1_NoLg$ES64_EPI,
+            W1_NoLg$AUTO_AVG, W1_NoLg$ES27_AVG, W1_NoLg$ES64_AVG,
+
+            W2_NoLg$AUTO_AR, W2_NoLg$ES27_AR, W2_NoLg$ES64_AR, 
+            W2_NoLg$AUTO_ADJ, W2_NoLg$ES27_ADJ, W2_NoLg$ES64_ADJ, 
+            W2_NoLg$AUTO_TMP, W2_NoLg$ES27_TMP, W2_NoLg$ES64_TMP, 
+            W2_NoLg$AUTO_EPI, W2_NoLg$ES27_EPI, W2_NoLg$ES64_EPI,
+            W2_NoLg$AUTO_AVG, W2_NoLg$ES27_AVG, W2_NoLg$ES64_AVG,
+
+            W3_NoLg$AUTO_AR, W3_NoLg$ES27_AR, W3_NoLg$ES64_AR,
+            W3_NoLg$AUTO_ADJ, W3_NoLg$ES27_ADJ, W3_NoLg$ES64_ADJ, 
+            W3_NoLg$AUTO_TMP, W3_NoLg$ES27_TMP, W3_NoLg$ES64_TMP, 
+            W3_NoLg$AUTO_EPI, W3_NoLg$ES27_EPI, W3_NoLg$ES64_EPI,
+            W3_NoLg$AUTO_AVG, W3_NoLg$ES27_AVG, W3_NoLg$ES64_AVG,
+
+            W4_NoLg$AUTO_AR, W4_NoLg$ES27_AR, W4_NoLg$ES64_AR,
+            W4_NoLg$AUTO_ADJ, W4_NoLg$ES27_ADJ, W4_NoLg$ES64_ADJ, 
+            W4_NoLg$AUTO_TMP, W4_NoLg$ES27_TMP, W4_NoLg$ES64_TMP, 
+            W4_NoLg$AUTO_EPI, W4_NoLg$ES27_EPI, W4_NoLg$ES64_EPI,
+            W4_NoLg$AUTO_AVG, W4_NoLg$ES27_AVG, W4_NoLg$ES64_AVG
             )
@@
   epiweek = AUTO_ARIMA_WEEK1$epiweek,
-  AUTO_AR=AUTO_ARIMA_WEEK1$WIS, 
-  AUTO_ADJ=AUTO_ADJACENT_WEEK1$WIS, 
-  AUTO_EPI=AUTO_EPIWEEK_WEEK1$WIS,
-  AUTO_TEMP=AUTO_TEMPERATURE_WEEK1$WIS,
-  AUTO_AVG=AUTO_AVERAGE_WEEK1$WIS,
-
-  ES27_AR=ES27_ARIMA_WEEK1$WIS, 
-  ES27_ADJ=ES27_ADJACENT_WEEK1$WIS, 
-  ES27_EPI=ES27_EPIWEEK_WEEK1$WIS,
-  ES27_TEMP=ES27_TEMPERATURE_WEEK1$WIS,
-  ES27_AVG=ES27_AVERAGE_WEEK1$WIS,
-
-  ES64_AR=ES64_ARIMA_WEEK1$WIS, 
-  ES64_ADJ=ES64_ADJACENT_WEEK1$WIS, 
-  ES64_EPI=ES64_EPIWEEK_WEEK1$WIS,
-  ES64_TEMP=ES64_TEMPERATURE_WEEK1$WIS,
-  ES64_AVG=ES64_AVERAGE_WEEK1$WIS  
+  AUTO_AR_LB=AUTO_ARIMA_WEEK1$WIS, 
+  AUTO_ADJ_LB=AUTO_ADJACENT_WEEK1$WIS, 
+  AUTO_EPI_LB=AUTO_EPIWEEK_WEEK1$WIS,
+  AUTO_TEMP_LB=AUTO_TEMPERATURE_WEEK1$WIS,
+  AUTO_AVG_LB=AUTO_AVERAGE_WEEK1$WIS,
+
+  ES27_AR_LB=ES27_ARIMA_WEEK1$WIS, 
+  ES27_ADJ_LB=ES27_ADJACENT_WEEK1$WIS, 
+  ES27_EPI_LB=ES27_EPIWEEK_WEEK1$WIS,
+  ES27_TEMP_LB=ES27_TEMPERATURE_WEEK1$WIS,
+  ES27_AVG_LB=ES27_AVERAGE_WEEK1$WIS,
+
+  ES64_AR_LB=ES64_ARIMA_WEEK1$WIS, 
+  ES64_ADJ_LB=ES64_ADJACENT_WEEK1$WIS, 
+  ES64_EPI_LB=ES64_EPIWEEK_WEEK1$WIS,
+  ES64_TEMP_LB=ES64_TEMPERATURE_WEEK1$WIS,
+  ES64_AVG_LB=ES64_AVERAGE_WEEK1$WIS  
 )
@@
   epiweek = AUTO_ARIMA_WEEK2$epiweek,
-  AUTO_AR=AUTO_ARIMA_WEEK2$WIS, 
-  AUTO_ADJ=AUTO_ADJACENT_WEEK2$WIS, 
-  AUTO_EPI=AUTO_EPIWEEK_WEEK2$WIS,
-  AUTO_TEMP=AUTO_TEMPERATURE_WEEK2$WIS,
-  AUTO_AVG=AUTO_AVERAGE_WEEK2$WIS,
-
-  ES27_AR=ES27_ARIMA_WEEK2$WIS, 
-  ES27_ADJ=ES27_ADJACENT_WEEK2$WIS, 
-  ES27_EPI=ES27_EPIWEEK_WEEK2$WIS,
-  ES27_TEMP=ES27_TEMPERATURE_WEEK2$WIS,
-  ES27_AVG=ES27_AVERAGE_WEEK2$WIS,
-
-  ES64_AR=ES64_ARIMA_WEEK2$WIS, 
-  ES64_ADJ=ES64_ADJACENT_WEEK2$WIS, 
-  ES64_EPI=ES64_EPIWEEK_WEEK2$WIS,
-  ES64_TEMP=ES64_TEMPERATURE_WEEK2$WIS,
-  ES64_AVG=ES64_AVERAGE_WEEK2$WIS  
+  AUTO_AR_LB=AUTO_ARIMA_WEEK2$WIS, 
+  AUTO_ADJ_LB=AUTO_ADJACENT_WEEK2$WIS, 
+  AUTO_EPI_LB=AUTO_EPIWEEK_WEEK2$WIS,
+  AUTO_TEMP_LB=AUTO_TEMPERATURE_WEEK2$WIS,
+  AUTO_AVG_LB=AUTO_AVERAGE_WEEK2$WIS,
+
+  ES27_AR_LB=ES27_ARIMA_WEEK2$WIS, 
+  ES27_ADJ_LB=ES27_ADJACENT_WEEK2$WIS, 
+  ES27_EPI_LB=ES27_EPIWEEK_WEEK2$WIS,
+  ES27_TEMP_LB=ES27_TEMPERATURE_WEEK2$WIS,
+  ES27_AVG_LB=ES27_AVERAGE_WEEK2$WIS,
+
+  ES64_AR_LB=ES64_ARIMA_WEEK2$WIS, 
+  ES64_ADJ_LB=ES64_ADJACENT_WEEK2$WIS, 
+  ES64_EPI_LB=ES64_EPIWEEK_WEEK2$WIS,
+  ES64_TEMP_LB=ES64_TEMPERATURE_WEEK2$WIS,
+  ES64_AVG_LB=ES64_AVERAGE_WEEK2$WIS  
 )
@@
   epiweek = AUTO_ARIMA_WEEK3$epiweek,
-  AUTO_AR=AUTO_ARIMA_WEEK3$WIS, 
-  AUTO_ADJ=AUTO_ADJACENT_WEEK3$WIS, 
-  AUTO_EPI=AUTO_EPIWEEK_WEEK3$WIS,
-  AUTO_TEMP=AUTO_TEMPERATURE_WEEK3$WIS,
-  AUTO_AVG=AUTO_AVERAGE_WEEK3$WIS,
-
-  ES27_AR=ES27_ARIMA_WEEK3$WIS, 
-  ES27_ADJ=ES27_ADJACENT_WEEK3$WIS, 
-  ES27_EPI=ES27_EPIWEEK_WEEK3$WIS,
-  ES27_TEMP=ES27_TEMPERATURE_WEEK3$WIS,
-  ES27_AVG=ES27_AVERAGE_WEEK3$WIS,
-
-  ES64_AR=ES64_ARIMA_WEEK3$WIS, 
-  ES64_ADJ=ES64_ADJACENT_WEEK3$WIS, 
-  ES64_EPI=ES64_EPIWEEK_WEEK3$WIS,
-  ES64_TEMP=ES64_TEMPERATURE_WEEK3$WIS,
-  ES64_AVG=ES64_AVERAGE_WEEK3$WIS  
+  AUTO_AR_LB=AUTO_ARIMA_WEEK3$WIS, 
+  AUTO_ADJ_LB=AUTO_ADJACENT_WEEK3$WIS, 
+  AUTO_EPI_LB=AUTO_EPIWEEK_WEEK3$WIS,
+  AUTO_TEMP_LB=AUTO_TEMPERATURE_WEEK3$WIS,
+  AUTO_AVG_LB=AUTO_AVERAGE_WEEK3$WIS,
+
+  ES27_AR_LB=ES27_ARIMA_WEEK3$WIS, 
+  ES27_ADJ_LB=ES27_ADJACENT_WEEK3$WIS, 
+  ES27_EPI_LB=ES27_EPIWEEK_WEEK3$WIS,
+  ES27_TEMP_LB=ES27_TEMPERATURE_WEEK3$WIS,
+  ES27_AVG_LB=ES27_AVERAGE_WEEK3$WIS,
+
+  ES64_AR_LB=ES64_ARIMA_WEEK3$WIS, 
+  ES64_ADJ_LB=ES64_ADJACENT_WEEK3$WIS, 
+  ES64_EPI_LB=ES64_EPIWEEK_WEEK3$WIS,
+  ES64_TEMP_LB=ES64_TEMPERATURE_WEEK3$WIS,
+  ES64_AVG_LB=ES64_AVERAGE_WEEK3$WIS  
 )
@@
   epiweek = AUTO_ARIMA_WEEK4$epiweek,
-  AUTO_AR=AUTO_ARIMA_WEEK4$WIS, 
-  AUTO_ADJ=AUTO_ADJACENT_WEEK4$WIS, 
-  AUTO_EPI=AUTO_EPIWEEK_WEEK4$WIS,
-  AUTO_TEMP=AUTO_TEMPERATURE_WEEK4$WIS,
-  AUTO_AVG=AUTO_AVERAGE_WEEK4$WIS,
-
-  ES27_AR=ES27_ARIMA_WEEK4$WIS, 
-  ES27_ADJ=ES27_ADJACENT_WEEK4$WIS, 
-  ES27_EPI=ES27_EPIWEEK_WEEK4$WIS,
-  ES27_TEMP=ES27_TEMPERATURE_WEEK4$WIS,
-  ES27_AVG=ES27_AVERAGE_WEEK4$WIS,
-
-  ES64_AR=ES64_ARIMA_WEEK4$WIS, 
-  ES64_ADJ=ES64_ADJACENT_WEEK4$WIS, 
-  ES64_EPI=ES64_EPIWEEK_WEEK4$WIS,
-  ES64_TEMP=ES64_TEMPERATURE_WEEK4$WIS,
-  ES64_AVG=ES64_AVERAGE_WEEK4$WIS  
+  AUTO_AR_LB=AUTO_ARIMA_WEEK4$WIS, 
+  AUTO_ADJ_LB=AUTO_ADJACENT_WEEK4$WIS, 
+  AUTO_EPI_LB=AUTO_EPIWEEK_WEEK4$WIS,
+  AUTO_TEMP_LB=AUTO_TEMPERATURE_WEEK4$WIS,
+  AUTO_AVG_LB=AUTO_AVERAGE_WEEK4$WIS,
+
+  ES27_AR_LB=ES27_ARIMA_WEEK4$WIS, 
+  ES27_ADJ_LB=ES27_ADJACENT_WEEK4$WIS, 
+  ES27_EPI_LB=ES27_EPIWEEK_WEEK4$WIS,
+  ES27_TEMP_LB=ES27_TEMPERATURE_WEEK4$WIS,
+  ES27_AVG_LB=ES27_AVERAGE_WEEK4$WIS,
+
+  ES64_AR_LB=ES64_ARIMA_WEEK4$WIS, 
+  ES64_ADJ_LB=ES64_ADJACENT_WEEK4$WIS, 
+  ES64_EPI_LB=ES64_EPIWEEK_WEEK4$WIS,
+  ES64_TEMP_LB=ES64_TEMPERATURE_WEEK4$WIS,
+  ES64_AVG_LB=ES64_AVERAGE_WEEK4$WIS  
 )
@@
     summarize(
-      AUTO_AR_Lg = mean(AUTO_AR, na.rm = TRUE),
-      AUTO_ADJ_Lg = mean(AUTO_ADJ, na.rm = TRUE),
-      AUTO_EPI_Lg = mean(AUTO_EPI, na.rm = TRUE),
-      AUTO_TMP_Lg = mean(AUTO_TEMP, na.rm = TRUE),
-      AUTO_AVG_Lg = mean(AUTO_AVG, na.rm = TRUE),
-
-      ES27_AR_Lg = mean(ES27_AR, na.rm = TRUE),
-      ES27_ADJ_Lg = mean(ES27_ADJ, na.rm = TRUE),
-      ES27_EPI_Lg = mean(ES27_EPI, na.rm = TRUE),
-      ES27_TMP_Lg = mean(ES27_TEMP, na.rm = TRUE),
-      ES27_AVG_Lg = mean(ES27_AVG, na.rm = TRUE),
-
-      ES64_AR_Lg = mean(ES64_AR, na.rm = TRUE),
-      ES64_ADJ_Lg = mean(ES64_ADJ, na.rm = TRUE),
-      ES64_EPI_Lg = mean(ES64_EPI, na.rm = TRUE),
-      ES64_TMP_Lg = mean(ES64_TEMP, na.rm = TRUE),
-      ES64_AVG_Lg = mean(ES64_AVG, na.rm = TRUE),
+      AUTO_AR_LB = mean(AUTO_AR_LB, na.rm = TRUE),
+      AUTO_ADJ_LB = mean(AUTO_ADJ_LB, na.rm = TRUE),
+      AUTO_EPI_LB = mean(AUTO_EPI_LB, na.rm = TRUE),
+      AUTO_TMP_LB = mean(AUTO_TEMP_LB, na.rm = TRUE),
+      AUTO_AVG_LB = mean(AUTO_AVG_LB, na.rm = TRUE),
+
+      ES27_AR_LB = mean(ES27_AR_LB, na.rm = TRUE),
+      ES27_ADJ_LB = mean(ES27_ADJ_LB, na.rm = TRUE),
+      ES27_EPI_LB = mean(ES27_EPI_LB, na.rm = TRUE),
+      ES27_TMP_LB = mean(ES27_TEMP_LB, na.rm = TRUE),
+      ES27_AVG_LB = mean(ES27_AVG_LB, na.rm = TRUE),
+
+      ES64_AR_LB = mean(ES64_AR_LB, na.rm = TRUE),
+      ES64_ADJ_LB = mean(ES64_ADJ_LB, na.rm = TRUE),
+      ES64_EPI_LB = mean(ES64_EPI_LB, na.rm = TRUE),
+      ES64_TMP_LB = mean(ES64_TEMP_LB, na.rm = TRUE),
+      ES64_AVG_LB = mean(ES64_AVG_LB, na.rm = TRUE),
 
@@
 
-  value = c(W1_Lg$AUTO_AR_Lg, W1_Lg$ES27_AR_Lg, W1_Lg$ES64_AR_Lg, 
-            W1_Lg$AUTO_ADJ_Lg, W1_Lg$ES27_ADJ_Lg, W1_Lg$ES64_ADJ_Lg, 
-            W1_Lg$AUTO_TMP_Lg, W1_Lg$ES27_TMP_Lg, W1_Lg$ES64_TMP_Lg, 
-            W1_Lg$AUTO_EPI_Lg, W1_Lg$ES27_EPI_Lg, W1_Lg$ES64_EPI_Lg,
-            W1_Lg$AUTO_AVG_Lg, W1_Lg$ES27_AVG_Lg, W1_Lg$ES64_AVG_Lg,
-
-            W2_Lg$AUTO_AR_Lg, W2_Lg$ES27_AR_Lg, W2_Lg$ES64_AR_Lg, 
-            W2_Lg$AUTO_ADJ_Lg, W2_Lg$ES27_ADJ_Lg, W2_Lg$ES64_ADJ_Lg, 
-            W2_Lg$AUTO_TMP_Lg, W2_Lg$ES27_TMP_Lg, W2_Lg$ES64_TMP_Lg, 
-            W2_Lg$AUTO_EPI_Lg, W2_Lg$ES27_EPI_Lg, W2_Lg$ES64_EPI_Lg,
-            W2_Lg$AUTO_AVG_Lg, W2_Lg$ES27_AVG_Lg, W2_Lg$ES64_AVG_Lg,
-
-            W3_Lg$AUTO_AR_Lg, W3_Lg$ES27_AR_Lg, W3_Lg$ES64_AR_Lg, 
-            W3_Lg$AUTO_ADJ_Lg, W3_Lg$ES27_ADJ_Lg, W3_Lg$ES64_ADJ_Lg, 
-            W3_Lg$AUTO_TMP_Lg, W3_Lg$ES27_TMP_Lg, W3_Lg$ES64_TMP_Lg, 
-            W3_Lg$AUTO_EPI_Lg, W3_Lg$ES27_EPI_Lg, W3_Lg$ES64_EPI_Lg,
-            W3_Lg$AUTO_AVG_Lg, W3_Lg$ES27_AVG_Lg, W3_Lg$ES64_AVG_Lg,
-
-            W4_Lg$AUTO_AR_Lg, W4_Lg$ES27_AR_Lg, W4_Lg$ES64_AR_Lg, 
-            W4_Lg$AUTO_ADJ_Lg, W4_Lg$ES27_ADJ_Lg, W4_Lg$ES64_ADJ_Lg, 
-            W4_Lg$AUTO_TMP_Lg, W4_Lg$ES27_TMP_Lg, W4_Lg$ES64_TMP_Lg, 
-            W4_Lg$AUTO_EPI_Lg, W4_Lg$ES27_EPI_Lg, W4_Lg$ES64_EPI_Lg,
-            W4_Lg$AUTO_AVG_Lg, W4_Lg$ES27_AVG_Lg, W4_Lg$ES64_AVG_Lg)
+  value = c(W1_Lg$AUTO_AR_LB, W1_Lg$ES27_AR_LB, W1_Lg$ES64_AR_LB, 
+            W1_Lg$AUTO_ADJ_LB, W1_Lg$ES27_ADJ_LB, W1_Lg$ES64_ADJ_LB, 
+            W1_Lg$AUTO_TMP_LB, W1_Lg$ES27_TMP_LB, W1_Lg$ES64_TMP_LB, 
+            W1_Lg$AUTO_EPI_LB, W1_Lg$ES27_EPI_LB, W1_Lg$ES64_EPI_LB,
+            W1_Lg$AUTO_AVG_LB, W1_Lg$ES27_AVG_LB, W1_Lg$ES64_AVG_LB,
+
+            W2_Lg$AUTO_AR_LB, W2_Lg$ES27_AR_LB, W2_Lg$ES64_AR_LB, 
+            W2_Lg$AUTO_ADJ_LB, W2_Lg$ES27_ADJ_LB, W2_Lg$ES64_ADJ_LB, 
+            W2_Lg$AUTO_TMP_LB, W2_Lg$ES27_TMP_LB, W2_Lg$ES64_TMP_LB, 
+            W2_Lg$AUTO_EPI_LB, W2_Lg$ES27_EPI_LB, W2_Lg$ES64_EPI_LB,
+            W2_Lg$AUTO_AVG_LB, W2_Lg$ES27_AVG_LB, W2_Lg$ES64_AVG_LB,
+
+            W3_Lg$AUTO_AR_LB, W3_Lg$ES27_AR_LB, W3_Lg$ES64_AR_LB, 
+            W3_Lg$AUTO_ADJ_LB, W3_Lg$ES27_ADJ_LB, W3_Lg$ES64_ADJ_LB, 
+            W3_Lg$AUTO_TMP_LB, W3_Lg$ES27_TMP_LB, W3_Lg$ES64_TMP_LB, 
+            W3_Lg$AUTO_EPI_LB, W3_Lg$ES27_EPI_LB, W3_Lg$ES64_EPI_LB,
+            W3_Lg$AUTO_AVG_LB, W3_Lg$ES27_AVG_LB, W3_Lg$ES64_AVG_LB,
+
+            W4_Lg$AUTO_AR_LB, W4_Lg$ES27_AR_LB, W4_Lg$ES64_AR_LB, 
+            W4_Lg$AUTO_ADJ_LB, W4_Lg$ES27_ADJ_LB, W4_Lg$ES64_ADJ_LB, 
+            W4_Lg$AUTO_TMP_LB, W4_Lg$ES27_TMP_LB, W4_Lg$ES64_TMP_LB, 
+            W4_Lg$AUTO_EPI_LB, W4_Lg$ES27_EPI_LB, W4_Lg$ES64_EPI_LB,
+            W4_Lg$AUTO_AVG_LB, W4_Lg$ES27_AVG_LB, W4_Lg$ES64_AVG_LB)
+
 )
@@
 # Create a new column to classify the model type
-W1$Model_Type <- ifelse(grepl("_NoLg$", W1$Best_Result), "Without log-back transformation", "With log-back transformation")
+W1$Model_Type <- ifelse(grepl("_LB$", W1$Best_Result), "With log-back transformation", "Without log-back transformation")
 
@@
 # Create a new column to classify the model type
-W2$Model_Type <- ifelse(grepl("_NoLg$", W2$Best_Result), "Without log-back transformation", "With log-back transformation")
+W2$Model_Type <- ifelse(grepl("_LB$", W2$Best_Result),  "With log-back transformation","Without log-back transformation")
 
@@
 # Create a new column to classify the model type
-W3$Model_Type <- ifelse(grepl("_NoLg$", W3$Best_Result), "Without log-back transformation", "With log-back transformation")
+W3$Model_Type <- ifelse(grepl("_LB$", W3$Best_Result),  "With log-back transformation","Without log-back transformation")
 
@@
 # Create a new column to classify the model type
-W4$Model_Type <- ifelse(grepl("_NoLg$", W4$Best_Result), "Without log-back transformation", "With log-back transformation")
+W4$Model_Type <- ifelse(grepl("_LB$", W4$Best_Result),  "With log-back transformation","Without log-back transformation")
 # Print merged results
@@
 custom_colors <- c(
-  "AUTO_AR_Lg" = colors_ar[1], "ES27_AR_Lg" = colors_ar[2], "ES64_AR_Lg" = colors_ar[3],
-  "AUTO_EPI_Lg" = colors_epi[1], "ES27_EPI_Lg" = colors_epi[2], "ES64_EPI_Lg" = colors_epi[3],
-  "AUTO_TMP_Lg" = colors_tmp[1], "ES27_TMP_Lg" = colors_tmp[2], "ES64_TMP_Lg" = colors_tmp[3],
-  "AUTO_ADJ_Lg" = colors_adj[1], "ES27_ADJ_Lg" = colors_adj[2], "ES64_ADJ_Lg" = colors_adj[3],
-  "AUTO_AVG_Lg" = colors_avg[1], "ES27_AVG_Lg" = colors_avg[2], "ES64_AVG_Lg" = colors_avg[3],
-
-  "AUTO_AR_NoLg" = colors_ar[1], "ES27_AR_NoLg" = colors_ar[2], "ES64_AR_NoLg" = colors_ar[3],
-  "AUTO_EPI_NoLg" = colors_epi[1], "ES27_EPI_NoLg" = colors_epi[2], "ES64_EPI_NoLg" = colors_epi[3],
-  "AUTO_TMP_NoLg" = colors_tmp[1], "ES27_TMP_NoLg" = colors_tmp[2], "ES64_TMP_NoLg" = colors_tmp[3],
-  "AUTO_ADJ_NoLg" = colors_adj[1], "ES27_ADJ_NoLg" = colors_adj[2], "ES64_ADJ_NoLg" = colors_adj[3],
-  "AUTO_AVG_NoLg" = colors_avg[1], "ES27_AVG_NoLg" = colors_avg[2], "ES64_AVG_NoLg" = colors_avg[3]
+  "AUTO_AR_LB" = colors_ar[1], "ES27_AR_LB" = colors_ar[2], "ES64_AR_LB" = colors_ar[3],
+  "AUTO_EPI_LB" = colors_epi[1], "ES27_EPI_LB" = colors_epi[2], "ES64_EPI_LB" = colors_epi[3],
+  "AUTO_TMP_LB" = colors_tmp[1], "ES27_TMP_LB" = colors_tmp[2], "ES64_TMP_LB" = colors_tmp[3],
+  "AUTO_ADJ_LB" = colors_adj[1], "ES27_ADJ_LB" = colors_adj[2], "ES64_ADJ_LB" = colors_adj[3],
+  "AUTO_AVG_LB" = colors_avg[1], "ES27_AVG_LB" = colors_avg[2], "ES64_AVG_LB" = colors_avg[3],
+
+  "AUTO_AR" = colors_ar[1], "ES27_AR" = colors_ar[2], "ES64_AR" = colors_ar[3],
+  "AUTO_EPI" = colors_epi[1], "ES27_EPI" = colors_epi[2], "ES64_EPI" = colors_epi[3],
+  "AUTO_TMP" = colors_tmp[1], "ES27_TMP" = colors_tmp[2], "ES64_TMP" = colors_tmp[3],
+  "AUTO_ADJ" = colors_adj[1], "ES27_ADJ" = colors_adj[2], "ES64_ADJ" = colors_adj[3],
+  "AUTO_AVG" = colors_avg[1], "ES27_AVG" = colors_avg[2], "ES64_AVG" = colors_avg[3]
 )
@@
 # Get all model types except "AUTO_AAR"
-model_types <- setdiff(names(wide_data), c("STATE", "AUTO_AR_NoLg","Best_Result","Model_Type","Week_Ahead"))
+model_types <- setdiff(names(wide_data), c("STATE", "AUTO_AR","Best_Result","Model_Type","Week_Ahead"))
 
@@
 wilcox_results <- map_df(model_types, function(model) {
-  test <- wilcox.test(wide_data[[model]], wide_data$AUTO_AR_NoLg, paired = TRUE)
+  test <- wilcox.test(wide_data[[model]], wide_data$AUTO_AR, paired = TRUE)
   data.frame(Model = model, p_value = test$p.value, W = test$statistic)
@@
   return(data.frame(
-    AUTO_AR_NoLg = (data$AUTO_AR_NoLg / data$AUTO_AR_NoLg),
-
-    ES27_AR_NoLg = ((1 - (data$ES27_AR_NoLg/data$AUTO_AR_NoLg)) * 100),
-    ES64_AR_NoLg = ((1 - (data$ES64_AR_NoLg/data$AUTO_AR_NoLg )) * 100),
-
-    AUTO_ADJ_NoLg = ((1 - (data$AUTO_ADJ_NoLg/data$AUTO_AR_NoLg)) * 100),
-    ES27_ADJ_NoLg = ((1 - (data$ES27_ADJ_NoLg/data$AUTO_AR_NoLg)) * 100),
-    ES64_ADJ_NoLg = ((1 - (data$ES64_ADJ_NoLg/data$AUTO_AR_NoLg)) * 100),
-
-    AUTO_TMP_NoLg = ((1 - (data$AUTO_TMP_NoLg/data$AUTO_AR_NoLg)) * 100),
-    ES27_TMP_NoLg = ((1 - (data$ES27_TMP_NoLg/data$AUTO_AR_NoLg)) * 100),
-    ES64_TMP_NoLg = ((1 - (data$ES64_TMP_NoLg/data$AUTO_AR_NoLg)) * 100),
-
-    AUTO_EPI_NoLg = ((1 - (data$AUTO_EPI_NoLg/data$AUTO_AR_NoLg)) * 100),
-    ES27_EPI_NoLg = ((1 - (data$ES27_EPI_NoLg/data$AUTO_AR_NoLg)) * 100),
-    ES64_EPI_NoLg = ((1 - (data$ES64_EPI_NoLg/data$AUTO_AR_NoLg)) * 100),
-
-    AUTO_AVG_NoLg = ((1 - (data$AUTO_AVG_NoLg/data$AUTO_AR_NoLg)) * 100),
-    ES27_AVG_NoLg = ((1 - (data$ES27_AVG_NoLg/data$AUTO_AR_NoLg)) * 100),
-    ES64_AVG_NoLg = ((1 - (data$ES64_AVG_NoLg/data$AUTO_AR_NoLg)) * 100),
-
-    AUTO_AR_Lg = ((1 - (data$AUTO_AR_Lg/data$AUTO_AR_NoLg)) * 100),
-    ES27_AR_Lg = ((1 - (data$ES27_AR_Lg/data$AUTO_AR_NoLg)) * 100),
-    ES64_AR_Lg = ((1 - (data$ES64_AR_Lg/data$AUTO_AR_NoLg)) * 100),
-
-    AUTO_ADJ_Lg = ((1 - (data$AUTO_ADJ_Lg/data$AUTO_AR_NoLg)) * 100),
-    ES27_ADJ_Lg = ((1 - (data$ES27_ADJ_Lg/data$AUTO_AR_NoLg)) * 100),
-    ES64_ADJ_Lg = ((1 - (data$ES64_ADJ_Lg/data$AUTO_AR_NoLg)) * 100),
-
-    AUTO_TMP_Lg = ((1 - (data$AUTO_TMP_Lg/data$AUTO_AR_NoLg)) * 100),
-    ES27_TMP_Lg = ((1 - (data$ES27_TMP_Lg/data$AUTO_AR_NoLg)) * 100),
-    ES64_TMP_Lg = ((1 - (data$ES64_TMP_Lg/data$AUTO_AR_NoLg)) * 100),
-
-    AUTO_EPI_Lg = ((1 - (data$AUTO_EPI_Lg/data$AUTO_AR_NoLg)) * 100),
-    ES27_EPI_Lg = ((1 - (data$ES27_EPI_Lg/data$AUTO_AR_NoLg)) * 100),
-    ES64_EPI_Lg = ((1 - (data$ES64_EPI_Lg/data$AUTO_AR_NoLg)) * 100),
-
-    AUTO_AVG_Lg = ((1 - (data$AUTO_AVG_Lg/data$AUTO_AR_NoLg)) * 100),
-    ES27_AVG_Lg = ((1 - (data$ES27_AVG_Lg/data$AUTO_AR_NoLg)) * 100),
-    ES64_AVG_Lg = ((1 - (data$ES64_AVG_Lg/data$AUTO_AR_NoLg)) * 100)    
+    AUTO_AR = (data$AUTO_AR / data$AUTO_AR),
+
+    ES27_AR = ((1 - (data$ES27_AR/data$AUTO_AR)) * 100),
+    ES64_AR = ((1 - (data$ES64_AR/data$AUTO_AR )) * 100),
+
+    AUTO_ADJ = ((1 - (data$AUTO_ADJ/data$AUTO_AR)) * 100),
+    ES27_ADJ = ((1 - (data$ES27_ADJ/data$AUTO_AR)) * 100),
+    ES64_ADJ = ((1 - (data$ES64_ADJ/data$AUTO_AR)) * 100),
+
+    AUTO_TMP = ((1 - (data$AUTO_TMP/data$AUTO_AR)) * 100),
+    ES27_TMP = ((1 - (data$ES27_TMP/data$AUTO_AR)) * 100),
+    ES64_TMP = ((1 - (data$ES64_TMP/data$AUTO_AR)) * 100),
+
+    AUTO_EPI = ((1 - (data$AUTO_EPI/data$AUTO_AR)) * 100),
+    ES27_EPI = ((1 - (data$ES27_EPI/data$AUTO_AR)) * 100),
+    ES64_EPI = ((1 - (data$ES64_EPI/data$AUTO_AR)) * 100),
+
+    AUTO_AVG = ((1 - (data$AUTO_AVG/data$AUTO_AR)) * 100),
+    ES27_AVG = ((1 - (data$ES27_AVG/data$AUTO_AR)) * 100),
+    ES64_AVG = ((1 - (data$ES64_AVG/data$AUTO_AR)) * 100),
+
+    AUTO_AR_LB = ((1 - (data$AUTO_AR_LB/data$AUTO_AR)) * 100),
+    ES27_AR_LB = ((1 - (data$ES27_AR_LB/data$AUTO_AR)) * 100),
+    ES64_AR_LB = ((1 - (data$ES64_AR_LB/data$AUTO_AR)) * 100),
+
+    AUTO_ADJ_LB = ((1 - (data$AUTO_ADJ_LB/data$AUTO_AR)) * 100),
+    ES27_ADJ_LB = ((1 - (data$ES27_ADJ_LB/data$AUTO_AR)) * 100),
+    ES64_ADJ_LB = ((1 - (data$ES64_ADJ_LB/data$AUTO_AR)) * 100),
+
+    AUTO_TMP_LB = ((1 - (data$AUTO_TMP_LB/data$AUTO_AR)) * 100),
+    ES27_TMP_LB = ((1 - (data$ES27_TMP_LB/data$AUTO_AR)) * 100),
+    ES64_TMP_LB = ((1 - (data$ES64_TMP_LB/data$AUTO_AR)) * 100),
+
+    AUTO_EPI_LB = ((1 - (data$AUTO_EPI_LB/data$AUTO_AR)) * 100),
+    ES27_EPI_LB = ((1 - (data$ES27_EPI_LB/data$AUTO_AR)) * 100),
+    ES64_EPI_LB = ((1 - (data$ES64_EPI_LB/data$AUTO_AR)) * 100),
+
+    AUTO_AVG_LB = ((1 - (data$AUTO_AVG_LB/data$AUTO_AR)) * 100),
+    ES27_AVG_LB = ((1 - (data$ES27_AVG_LB/data$AUTO_AR)) * 100),
+    ES64_AVG_LB = ((1 - (data$ES64_AVG_LB/data$AUTO_AR)) * 100)    
   ))
@@
 # Create a new column "Model_Type" based on whether "NoLg" is in the Model name
-p_values_and_impr$Model_Type <- ifelse(grepl("NoLg", p_values_and_impr$Model), "NoLog", "Log")
+p_values_and_impr$Model_Type <- ifelse(grepl("_LB$", p_values_and_impr$Model), "LB", "no_LB")
 
@@
   mutate(Significance = ifelse(p_values < alpha, "Significant", "Not Significant"),
-         Model_Type = ifelse(grepl("NoLg", Model), "Without log-back transformation", "With log-back transformation"),
+         Model_Type = ifelse(grepl("_LB$", Model), "With log-back transformation", "Without log-back transformation"),
          Model = factor(Model, levels = model_order))  # Apply ordered factor
@@
                       name = "Percentage (%)", limits = c(-35, 35)) +  # Fixed color scale from -30 to 30
-  labs(title = "Figure X: Mean differences between models' WIS and the baseline across 48 states",
+  labs(title = "Figure X:Mean differences between models' WIS and the baseline across 48 states",
        subtitle = "Gray boxes indicate models that are not significantly different from the baseline (p>0.05)",
@@
   STATE = W1$STATE,
-  percentage_improvement = W1_percentage_of_improvement$AUTO_AVG_Lg,
-  AUTO_AVG=W1$AUTO_AVG_Lg
+  percentage_improvement = W1_percentage_of_improvement$AUTO_AVG_LB,
+  AUTO_AVG=W1$AUTO_AVG_LB
 )
@@
   STATE = W2$STATE,
-  percentage_improvement = W2_percentage_of_improvement$AUTO_AVG_Lg,
-  AUTO_AVG=W2$AUTO_AVG_Lg
+  percentage_improvement = W2_percentage_of_improvement$AUTO_AVG_LB,
+  AUTO_AVG=W2$AUTO_AVG_LB
 )
@@
   STATE = W3$STATE,
-  percentage_improvement = W3_percentage_of_improvement$AUTO_AVG_Lg,
-  AUTO_AVG=W3$AUTO_AVG_Lg
+  percentage_improvement = W3_percentage_of_improvement$AUTO_AVG_LB,
+  AUTO_AVG=W3$AUTO_AVG_LB
 )
@@
   STATE = W4$STATE,
-  percentage_improvement = W4_percentage_of_improvement$AUTO_AVG_Lg,
-  AUTO_AVG=W4$AUTO_AVG_Lg
+  percentage_improvement = W4_percentage_of_improvement$AUTO_AVG_LB,
+  AUTO_AVG=W4$AUTO_AVG_LB
 )
@@
     summarize(
-      AUTO_AR_NoLg = sum(AUTO_AR_NoLg, na.rm = TRUE),  # Sum across all states
-      AUTO_AVG = sum(AUTO_AVG, na.rm = TRUE)           # Sum across all states
+      AUTO_AR = sum(AUTO_AR, na.rm = TRUE),  # Sum across all states
+      AUTO_AVG_LB = sum(AUTO_AVG_LB, na.rm = TRUE)           # Sum across all states
     ) %>%
     # Calculate the difference ratio
-    mutate(Difference = 1 - (AUTO_AVG / AUTO_AR_NoLg),
+    mutate(Difference = 1 - (AUTO_AVG_LB / AUTO_AR),
            Week = paste0("Week ", week))

### 2025-05-06 manuscript_results.Rmd
update

@@
               fill = "gray", alpha = 0.25) +
-  labs(title = "Example of influenza hospitalizaions forecasts with 95% confidence intervals",
+  labs(title = "Figure X: Example of influenza hospitalizaions forecasts",
+       subtitle = "The gray ribbon indicates the 95% confidence interval",
        x = "Target Weeks",
        y = "Forecast Values") +
-  theme_minimal()
+  theme_minimal() +
+  theme(plot.title = element_text(size = 16, face = "bold"),
+        plot.subtitle = element_text(size = 14))
 
@@
 
-ggsave(plt00, height = 5, width = 6, filename="fig00.jpg") 
+ggsave(plt00, height = 5, width = 10, filename="fig00.jpg") 
 ```
@@
   theme_minimal() +  
-  labs(title = "Baseline Model WIS for 48 states on the contiguous U.S.",#,
+  labs(title = "Figure X: Baseline model WIS for 48 states on the contiguous U.S.",#,
 #       subtitle = "Dashed red lines indicates epidemiological weeks outsise of the influenza season",
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        plot.title = element_text(size = 16, face = "bold"),
         legend.position = "none") + 
@@
 
-ggsave(plt0, height = 5, width = 6, filename="fig0.jpg")  
+ggsave(plt0, height = 5, width = 8, filename="fig0.jpg")  
 ```
@@
   scale_fill_manual(values = custom_colors) +
-  labs(title = "Best model based on the lowest mean WIS on 48 U.S. states",
-       x = "", y = "Number of states") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  labs(title = "Figure X: Best model based on the lowest mean WIS",
+       subtitle = "Across the 48 states on the contiguous U.S.",
+       x = "", y = "Number of states",
+       fill = "Best model" ) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        plot.title = element_text(size = 16, face = "bold")) +
   facet_grid(Week_Ahead ~ Model_Type, scales = "free_x")
@@
 # Save and display the plot
-ggsave("Fig2.jpg", combined_plot, width = 10, height = 8)
+ggsave("Fig2.jpg", combined_plot, width =8, height =5)
 print(combined_plot)
@@
 
-    ES27_AR = ((1 - (data$ES27_AR/data$AUTO_AR)) * 100),
-    ES64_AR = ((1 - (data$ES64_AR/data$AUTO_AR )) * 100),
-
-    AUTO_ADJ = ((1 - (data$AUTO_ADJ/data$AUTO_AR)) * 100),
-    ES27_ADJ = ((1 - (data$ES27_ADJ/data$AUTO_AR)) * 100),
-    ES64_ADJ = ((1 - (data$ES64_ADJ/data$AUTO_AR)) * 100),
-
-    AUTO_TMP = ((1 - (data$AUTO_TMP/data$AUTO_AR)) * 100),
-    ES27_TMP = ((1 - (data$ES27_TMP/data$AUTO_AR)) * 100),
-    ES64_TMP = ((1 - (data$ES64_TMP/data$AUTO_AR)) * 100),
-
-    AUTO_EPI = ((1 - (data$AUTO_EPI/data$AUTO_AR)) * 100),
-    ES27_EPI = ((1 - (data$ES27_EPI/data$AUTO_AR)) * 100),
-    ES64_EPI = ((1 - (data$ES64_EPI/data$AUTO_AR)) * 100),
-
-    AUTO_AVG = ((1 - (data$AUTO_AVG/data$AUTO_AR)) * 100),
-    ES27_AVG = ((1 - (data$ES27_AVG/data$AUTO_AR)) * 100),
-    ES64_AVG = ((1 - (data$ES64_AVG/data$AUTO_AR)) * 100),
-
-    AUTO_AR_LB = ((1 - (data$AUTO_AR_LB/data$AUTO_AR)) * 100),
-    ES27_AR_LB = ((1 - (data$ES27_AR_LB/data$AUTO_AR)) * 100),
-    ES64_AR_LB = ((1 - (data$ES64_AR_LB/data$AUTO_AR)) * 100),
-
-    AUTO_ADJ_LB = ((1 - (data$AUTO_ADJ_LB/data$AUTO_AR)) * 100),
-    ES27_ADJ_LB = ((1 - (data$ES27_ADJ_LB/data$AUTO_AR)) * 100),
-    ES64_ADJ_LB = ((1 - (data$ES64_ADJ_LB/data$AUTO_AR)) * 100),
-
-    AUTO_TMP_LB = ((1 - (data$AUTO_TMP_LB/data$AUTO_AR)) * 100),
-    ES27_TMP_LB = ((1 - (data$ES27_TMP_LB/data$AUTO_AR)) * 100),
-    ES64_TMP_LB = ((1 - (data$ES64_TMP_LB/data$AUTO_AR)) * 100),
-
-    AUTO_EPI_LB = ((1 - (data$AUTO_EPI_LB/data$AUTO_AR)) * 100),
-    ES27_EPI_LB = ((1 - (data$ES27_EPI_LB/data$AUTO_AR)) * 100),
-    ES64_EPI_LB = ((1 - (data$ES64_EPI_LB/data$AUTO_AR)) * 100),
-
-    AUTO_AVG_LB = ((1 - (data$AUTO_AVG_LB/data$AUTO_AR)) * 100),
-    ES27_AVG_LB = ((1 - (data$ES27_AVG_LB/data$AUTO_AR)) * 100),
-    ES64_AVG_LB = ((1 - (data$ES64_AVG_LB/data$AUTO_AR)) * 100)    
+    ES27_AR = (-(1 - (data$ES27_AR/data$AUTO_AR)) * 100),
+    ES64_AR = (-(1 - (data$ES64_AR/data$AUTO_AR )) * 100),
+
+    AUTO_ADJ = (-(1 - (data$AUTO_ADJ/data$AUTO_AR)) * 100),
+    ES27_ADJ = (-(1 - (data$ES27_ADJ/data$AUTO_AR)) * 100),
+    ES64_ADJ = (-(1 - (data$ES64_ADJ/data$AUTO_AR)) * 100),
+
+    AUTO_TMP = (-(1 - (data$AUTO_TMP/data$AUTO_AR)) * 100),
+    ES27_TMP = (-(1 - (data$ES27_TMP/data$AUTO_AR)) * 100),
+    ES64_TMP = (-(1 - (data$ES64_TMP/data$AUTO_AR)) * 100),
+
+    AUTO_EPI = (-(1 - (data$AUTO_EPI/data$AUTO_AR)) * 100),
+    ES27_EPI = (-(1 - (data$ES27_EPI/data$AUTO_AR)) * 100),
+    ES64_EPI = (-(1 - (data$ES64_EPI/data$AUTO_AR)) * 100),
+
+    AUTO_AVG = (-(1 - (data$AUTO_AVG/data$AUTO_AR)) * 100),
+    ES27_AVG = (-(1 - (data$ES27_AVG/data$AUTO_AR)) * 100),
+    ES64_AVG = (-(1 - (data$ES64_AVG/data$AUTO_AR)) * 100),
+
+    AUTO_AR_LB = (-(1 - (data$AUTO_AR_LB/data$AUTO_AR)) * 100),
+    ES27_AR_LB = (-(1 - (data$ES27_AR_LB/data$AUTO_AR)) * 100),
+    ES64_AR_LB = (-(1 - (data$ES64_AR_LB/data$AUTO_AR)) * 100),
+
+    AUTO_ADJ_LB = (-(1 - (data$AUTO_ADJ_LB/data$AUTO_AR)) * 100),
+    ES27_ADJ_LB = (-(1 - (data$ES27_ADJ_LB/data$AUTO_AR)) * 100),
+    ES64_ADJ_LB = (-(1 - (data$ES64_ADJ_LB/data$AUTO_AR)) * 100),
+
+    AUTO_TMP_LB = (-(1 - (data$AUTO_TMP_LB/data$AUTO_AR)) * 100),
+    ES27_TMP_LB = (-(1 - (data$ES27_TMP_LB/data$AUTO_AR)) * 100),
+    ES64_TMP_LB = (-(1 - (data$ES64_TMP_LB/data$AUTO_AR)) * 100),
+
+    AUTO_EPI_LB = (-(1 - (data$AUTO_EPI_LB/data$AUTO_AR)) * 100),
+    ES27_EPI_LB = (-(1 - (data$ES27_EPI_LB/data$AUTO_AR)) * 100),
+    ES64_EPI_LB = (-(1 - (data$ES64_EPI_LB/data$AUTO_AR)) * 100),
+
+    AUTO_AVG_LB = (-(1 - (data$AUTO_AVG_LB/data$AUTO_AR)) * 100),
+    ES27_AVG_LB = (-(1 - (data$ES27_AVG_LB/data$AUTO_AR)) * 100),
+    ES64_AVG_LB = (-(1 - (data$ES64_AVG_LB/data$AUTO_AR)) * 100)    
   ))
@@
             color = "black", size = 4) +  # Add text labels
-  scale_fill_gradient(low = "red", high = "cyan", na.value = "gray80", 
+  scale_fill_gradient(low = "cyan", high = "red", na.value = "gray80", 
                       name = "Percentage (%)", limits = c(-35, 35)) +  # Fixed color scale from -30 to 30
-  labs(title = "Figure X:Mean differences between models' WIS and the baseline across 48 states",
+  labs(title = "Figure X: Mean differences between models' WIS and the baseline across 48 states",
        subtitle = "Gray boxes indicate models that are not significantly different from the baseline (p>0.05)",
-         caption = "m = mean difference, sd = standard deviation",
+         caption = "m = mean difference (%), sd = standard deviation",
        x = "Week Ahead",
@@
   geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
-  scale_fill_gradient2(low = "darkred", mid = "lightyellow", high = "skyblue", midpoint = 0, limits = c(-30, 60)) +  
+  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
   ggtitle("Mean WIS improvement in percentage (W1)") +
@@
   theme(legend.position = "right") +
-  geom_sf_text(data = W1_map, aes(label = round(percentage_improvement,1)),  # Round to whole numbers
+  geom_sf_text(data = W1_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
                size = 3,
@@
 
-ggsave(ES_1WEEK,height = 4.5, width = 6.5,  filename="Fig5A.jpg")  
+#ggsave(ES_1WEEK,height = 4.5, width = 6.5,  filename="Fig5A.jpg")  
 ```
@@
   geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
-  scale_fill_gradient2(low = "darkred", mid = "lightyellow", high = "skyblue", midpoint = 0, limits = c(-30, 60)) +  
+  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
   ggtitle("Mean WIS improvement in percentage (W2)") +
@@
   theme(legend.position = "right") +
-  geom_sf_text(data = W2_map, aes(label = round(percentage_improvement,1)),  # Round to whole numbers
+  geom_sf_text(data = W2_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
                size = 3,
@@
 
-ggsave(ES_2WEEK, height = 4.5, width = 6.5, filename="Fig5B.jpg")  
+#ggsave(ES_2WEEK, height = 4.5, width = 6.5, filename="Fig5B.jpg")  
 ```
@@
   geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
-  scale_fill_gradient2(low = "darkred", mid = "lightyellow", high = "skyblue", midpoint = 0, limits = c(-30, 60)) +  
+  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
   ggtitle("Mean WIS improvement in percentage (W3)") +
@@
   theme(legend.position = "right") +
-  geom_sf_text(data = W3_map, aes(label = round(percentage_improvement,1)),  # Round to whole numbers
+  geom_sf_text(data = W3_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
                size = 3,
@@
 
-ggsave(ES_3WEEK, height = 4.5, width = 6.5, filename="Fig5C.jpg")  
+#ggsave(ES_3WEEK, height = 4.5, width = 6.5, filename="Fig5C.jpg")  
 ```
@@
   geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
-  scale_fill_gradient2(low = "darkred", mid = "lightyellow", high = "skyblue", midpoint = 0, limits = c(-30, 60)) +  
+  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
   ggtitle("Mean WIS improvement in percentage (W4)") +
@@
   theme(legend.position = "right") +
-  geom_sf_text(data = W4_map, aes(label = round(percentage_improvement,1)),  # Round to whole numbers
+  geom_sf_text(data = W4_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
                size = 3,
@@
 
-ggsave(ES_4WEEK, height = 4.5, width = 6.5,  filename="Fig5D.jpg")  
-```
+#ggsave(ES_4WEEK, height = 4.5, width = 6.5,  filename="Fig5D.jpg")  
+```
+
+```{r}
+
+library(patchwork)
+
+# Ensure each plot has coord_sf applied
+ES_1WEEK <- ES_1WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
+ES_2WEEK <- ES_2WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
+ES_3WEEK <- ES_3WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
+ES_4WEEK <- ES_4WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
+
+# Combine in a 2x2 grid
+combined_plot <- (ES_1WEEK | ES_2WEEK) /
+                 (ES_3WEEK | ES_4WEEK)
+
+# Print
+print(combined_plot)
+
+# Save to file
+ggsave("Fig5_2x2_combined.jpg", combined_plot, width = 15, height = 8)
+```
+
 REGRESSION MODELS
@@
     # Calculate the difference ratio
-    mutate(Difference = 1 - (AUTO_AVG_LB / AUTO_AR),
+    mutate(Difference =  ((AUTO_AVG_LB / AUTO_AR)-1)*100,
            Week = paste0("Week ", week))
@@
 # Plot the difference using ggplot2
-model_improv<-ggplot(summed_data_all, aes(x = Julian_date, y = (Difference)*100, color = Week)) +
+model_improv<-ggplot(summed_data_all, aes(x = Julian_date, y = (Difference), color = Week)) +
   geom_point(size = 3) +

### 2025-05-06 manuscript_results.Rmd
update

@@
 library("ggpubr")
+library("patchwork")
+
 ```
@@
 
-    ES27_AR = (-(1 - (data$ES27_AR/data$AUTO_AR)) * 100),
-    ES64_AR = (-(1 - (data$ES64_AR/data$AUTO_AR )) * 100),
-
-    AUTO_ADJ = (-(1 - (data$AUTO_ADJ/data$AUTO_AR)) * 100),
-    ES27_ADJ = (-(1 - (data$ES27_ADJ/data$AUTO_AR)) * 100),
-    ES64_ADJ = (-(1 - (data$ES64_ADJ/data$AUTO_AR)) * 100),
-
-    AUTO_TMP = (-(1 - (data$AUTO_TMP/data$AUTO_AR)) * 100),
-    ES27_TMP = (-(1 - (data$ES27_TMP/data$AUTO_AR)) * 100),
-    ES64_TMP = (-(1 - (data$ES64_TMP/data$AUTO_AR)) * 100),
-
-    AUTO_EPI = (-(1 - (data$AUTO_EPI/data$AUTO_AR)) * 100),
-    ES27_EPI = (-(1 - (data$ES27_EPI/data$AUTO_AR)) * 100),
-    ES64_EPI = (-(1 - (data$ES64_EPI/data$AUTO_AR)) * 100),
-
-    AUTO_AVG = (-(1 - (data$AUTO_AVG/data$AUTO_AR)) * 100),
-    ES27_AVG = (-(1 - (data$ES27_AVG/data$AUTO_AR)) * 100),
-    ES64_AVG = (-(1 - (data$ES64_AVG/data$AUTO_AR)) * 100),
-
-    AUTO_AR_LB = (-(1 - (data$AUTO_AR_LB/data$AUTO_AR)) * 100),
-    ES27_AR_LB = (-(1 - (data$ES27_AR_LB/data$AUTO_AR)) * 100),
-    ES64_AR_LB = (-(1 - (data$ES64_AR_LB/data$AUTO_AR)) * 100),
-
-    AUTO_ADJ_LB = (-(1 - (data$AUTO_ADJ_LB/data$AUTO_AR)) * 100),
-    ES27_ADJ_LB = (-(1 - (data$ES27_ADJ_LB/data$AUTO_AR)) * 100),
-    ES64_ADJ_LB = (-(1 - (data$ES64_ADJ_LB/data$AUTO_AR)) * 100),
-
-    AUTO_TMP_LB = (-(1 - (data$AUTO_TMP_LB/data$AUTO_AR)) * 100),
-    ES27_TMP_LB = (-(1 - (data$ES27_TMP_LB/data$AUTO_AR)) * 100),
-    ES64_TMP_LB = (-(1 - (data$ES64_TMP_LB/data$AUTO_AR)) * 100),
-
-    AUTO_EPI_LB = (-(1 - (data$AUTO_EPI_LB/data$AUTO_AR)) * 100),
-    ES27_EPI_LB = (-(1 - (data$ES27_EPI_LB/data$AUTO_AR)) * 100),
-    ES64_EPI_LB = (-(1 - (data$ES64_EPI_LB/data$AUTO_AR)) * 100),
-
-    AUTO_AVG_LB = (-(1 - (data$AUTO_AVG_LB/data$AUTO_AR)) * 100),
-    ES27_AVG_LB = (-(1 - (data$ES27_AVG_LB/data$AUTO_AR)) * 100),
-    ES64_AVG_LB = (-(1 - (data$ES64_AVG_LB/data$AUTO_AR)) * 100)    
+    ES27_AR = (((data$ES27_AR/data$AUTO_AR)-1) * 100),
+    ES64_AR = (((data$ES64_AR/data$AUTO_AR )-1) * 100),
+
+    AUTO_ADJ = (((data$AUTO_ADJ/data$AUTO_AR)-1) * 100),
+    ES27_ADJ = (((data$ES27_ADJ/data$AUTO_AR)-1) * 100),
+    ES64_ADJ = (((data$ES64_ADJ/data$AUTO_AR)-1) * 100),
+
+    AUTO_TMP = (((data$AUTO_TMP/data$AUTO_AR)-1) * 100),
+    ES27_TMP = (((data$ES27_TMP/data$AUTO_AR)-1) * 100),
+    ES64_TMP = (((data$ES64_TMP/data$AUTO_AR)-1) * 100),
+
+    AUTO_EPI = (((data$AUTO_EPI/data$AUTO_AR)-1) * 100),
+    ES27_EPI = (((data$ES27_EPI/data$AUTO_AR)-1) * 100),
+    ES64_EPI = (((data$ES64_EPI/data$AUTO_AR)-1) * 100),
+
+    AUTO_AVG = (((data$AUTO_AVG/data$AUTO_AR)-1) * 100),
+    ES27_AVG = (((data$ES27_AVG/data$AUTO_AR)-1) * 100),
+    ES64_AVG = (((data$ES64_AVG/data$AUTO_AR)-1) * 100),
+
+    AUTO_AR_LB = (((data$AUTO_AR_LB/data$AUTO_AR)-1) * 100),
+    ES27_AR_LB = (((data$ES27_AR_LB/data$AUTO_AR)-1) * 100),
+    ES64_AR_LB = (((data$ES64_AR_LB/data$AUTO_AR)-1) * 100),
+
+    AUTO_ADJ_LB = (((data$AUTO_ADJ_LB/data$AUTO_AR)-1) * 100),
+    ES27_ADJ_LB = (((data$ES27_ADJ_LB/data$AUTO_AR)-1) * 100),
+    ES64_ADJ_LB = (((data$ES64_ADJ_LB/data$AUTO_AR)-1) * 100),
+
+    AUTO_TMP_LB = (((data$AUTO_TMP_LB/data$AUTO_AR)-1) * 100),
+    ES27_TMP_LB = (((data$ES27_TMP_LB/data$AUTO_AR)-1) * 100),
+    ES64_TMP_LB = (((data$ES64_TMP_LB/data$AUTO_AR)-1) * 100),
+
+    AUTO_EPI_LB = (((data$AUTO_EPI_LB/data$AUTO_AR)-1) * 100),
+    ES27_EPI_LB = (((data$ES27_EPI_LB/data$AUTO_AR)-1) * 100),
+    ES64_EPI_LB = (((data$ES64_EPI_LB/data$AUTO_AR)-1) * 100),
+
+    AUTO_AVG_LB = (((data$AUTO_AVG_LB/data$AUTO_AR)-1) * 100),
+    ES27_AVG_LB = (((data$ES27_AVG_LB/data$AUTO_AR)-1) * 100),
+    ES64_AVG_LB = (((data$ES64_AVG_LB/data$AUTO_AR)-1) * 100)    
   ))
@@
             aes(x = factor(WeekAhead), y = Model), fill = "gray80", color = "black") +  # Gray for non-significant
-  geom_text(aes(label = paste0("m:", round(m, 2),", sd:", round(sd, 2))), 
+  geom_text(aes(label = paste0("m:", round(m, 1),", sd:", round(sd, 1))), 
             color = "black", size = 4) +  # Add text labels
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
-  plot.caption = element_text(size = 12),
+  plot.caption = element_text(size = 16),
     plot.title = element_text(size = 18, face = "bold"),
@@
 print(plot_m_heatmap)
-ggsave("Fig3.jpg", plot_m_heatmap, width = 15, height = 7 )
+ggsave("Fig3.jpg", plot_m_heatmap, width = 14, height = 7 )
 
@@
   scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
-  ggtitle("Mean WIS improvement in percentage (W1)") +
+  ggtitle("1 Week Ahead") +
   theme_light() +
-  theme(legend.position = "right") +
+  theme(legend.position = "right",
+    plot.title = element_text(hjust = 0.5)) +
   geom_sf_text(data = W1_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
-               size = 3,
+               size = 2.5,
                color = "black",
-               check_overlap = TRUE) +  # Display percentage improvement in each state
+               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
   labs(
     fill = "%",  # Label for the legend
-    subtitle = "(1-(Baseline/AUTO_AVG_LB))*100"
-            , 
        x = "",
@@
 ES_1WEEK<-ES_1WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
-ES_1WEEK
 
@@
   scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
-  ggtitle("Mean WIS improvement in percentage (W2)") +
+  ggtitle("2 Weeks Ahead") +
   theme_light() +
-  theme(legend.position = "right") +
+  theme(legend.position = "right",
+    plot.title = element_text(hjust = 0.5)) +
   geom_sf_text(data = W2_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
-               size = 3,
+               size = 2.5,
                color = "black",
-               check_overlap = TRUE) +  # Display percentage improvement in each state
+               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
   labs(
     fill = "%",  # Label for the legend
-    subtitle = "(1-(Baseline/AUTO_AVG_LB))*100"
-            , 
        x = "",
@@
 ES_2WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
-ES_2WEEK
 
@@
   scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
-  ggtitle("Mean WIS improvement in percentage (W3)") +
+  ggtitle("3 Weeks Ahead") +
   theme_light() +
-  theme(legend.position = "right") +
+  theme(legend.position = "right",
+    plot.title = element_text(hjust = 0.5)) +
   geom_sf_text(data = W3_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
-               size = 3,
+               size = 2.5,
                color = "black",
-               check_overlap = TRUE) +  # Display percentage improvement in each state
+               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
   labs(
     fill = "%",  # Label for the legend
-    subtitle = "(1-(Baseline/AUTO_AVG_LB))*100"
-            , 
        x = "",
@@
 ES_3WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
-ES_3WEEK
 
@@
   scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
-  ggtitle("Mean WIS improvement in percentage (W4)") +
+  ggtitle("4 Weeks Ahead") +
   theme_light() +
-  theme(legend.position = "right") +
+  theme(legend.position = "right",
+    plot.title = element_text(hjust = 0.5)) +
   geom_sf_text(data = W4_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
-               size = 3,
+               size = 2.5,
                color = "black",
-               check_overlap = TRUE) +  # Display percentage improvement in each state
+               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
   labs(
     fill = "%",  # Label for the legend
-    subtitle = "(1-(Baseline/AUTO_AVG_LB))*100"
-            , 
        x = "",
@@
 ES_4WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
-ES_4WEEK
 
@@
 ```{r}
-
-library(patchwork)
-
 # Ensure each plot has coord_sf applied
@@
 combined_plot <- (ES_1WEEK | ES_2WEEK) /
-                 (ES_3WEEK | ES_4WEEK)
-
+                 (ES_3WEEK | ES_4WEEK) + 
+  plot_annotation(
+    title = "Figure X: Mean WIS differences (%) compared to the baseline",
+    subtitle = "Results by forecast horizon",
+    theme = theme(plot.title = element_text(size = 16, face = "bold"),
+                  plot.subtitle = element_text(size = 14))
+  )
 # Print
@@
 # Save to file
-ggsave("Fig5_2x2_combined.jpg", combined_plot, width = 15, height = 8)
+ggsave("Fig5.jpg", combined_plot, width =12, height =7)
 ```

### 2025-05-06 manuscript_results.Rmd
update

@@
 
-# Assuming merged_data already includes the necessary columns: model and percentage_improveme
-# Create the map plot for 1 Week Ahead with best models and percentage improvement
 ES_1WEEK <- ggplot(W1_map) +
@@
   scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
-  ggtitle("Mean WIS on each state (W1)") +
+  ggtitle("1 Week Ahead") +
   theme_light() +
-  theme(legend.position = "right") +
-  geom_sf_text(data = W1_map, aes(label = round(AUTO_AVG,1)),  # Round to whole numbers
+  theme(legend.position = "right",
+    plot.title = element_text(hjust = 0.5)) +
+  geom_sf_text(data = W1_map, aes(label = round(AUTO_AVG,0)),  # Round to whole numbers
                size = 3,
                color = "black",
-               check_overlap = TRUE) +  # Display percentage improvement in each state
+               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
   labs(
-    fill = "Value",  # Label for the legend
-    subtitle = "Based on 2023 and 2024 influenza seasons"
-    , 
+    fill = "mean WIS",  
        x = "",
@@
 ES_1WEEK<- ES_1WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
-
 ES_1WEEK
-ggsave(ES_1WEEK, height = 4.5, width = 6.5,  filename="Fig4A.jpg")  
-```
+#ggsave(ES_1WEEK, height = 4.5, width = 6.5,  filename="Fig4A.jpg")  
+```
+
 2 weeks ahead
@@
 
-# Assuming merged_data already includes the necessary columns: model and percentage_improveme
-# Create the map plot for 1 Week Ahead with best models and percentage improvement
 ES_2WEEK <- ggplot(W2_map) +
@@
   scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
-  ggtitle("Mean WIS on each state (W2)") +
+  ggtitle("2 Weeks Ahead") +
   theme_light() +
-  theme(legend.position = "right") +
-  geom_sf_text(data = W2_map, aes(label = round(AUTO_AVG,1)),  # Round to whole numbers
+  theme(legend.position = "right",
+    plot.title = element_text(hjust = 0.5)) +
+  geom_sf_text(data = W2_map, aes(label = round(AUTO_AVG,0)),  # Round to whole numbers
                size = 3,
                color = "black",
-               check_overlap = TRUE) +  # Display percentage improvement in each state
+               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
   labs(
-    fill = "Value",  # Label for the legend
-    subtitle = "Based on 2023 and 2024 influenza seasons"
-        , 
+    fill = "mean WIS",
        x = "",
@@
 ES_2WEEK <-ES_2WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
-
 ES_2WEEK
-ggsave(ES_2WEEK, height = 4.5, width = 6.5,  filename="Fig4B.jpg")  
-```
+#ggsave(ES_2WEEK, height = 4.5, width = 6.5,  filename="Fig4B.jpg")  
+```
+
 3 weeks ahead
@@
 
-# Assuming merged_data already includes the necessary columns: model and percentage_improveme
-# Create the map plot for 1 Week Ahead with best models and percentage improvement
-# Create the map plot for 1 Week Ahead with best models and percentage improvement
 ES_3WEEK <- ggplot(W3_map) +
@@
   scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
-  ggtitle("Mean WIS on each state (W3)") +
+  ggtitle("3 Weeks Ahead") +
   theme_light() +
-  theme(legend.position = "right") +
-  geom_sf_text(data = W3_map, aes(label = round(AUTO_AVG,1)),  # Round to whole numbers
+  theme(legend.position = "right",
+    plot.title = element_text(hjust = 0.5)) +
+  geom_sf_text(data = W3_map, aes(label = round(AUTO_AVG,0)),  # Round to whole numbers
                size = 3,
                color = "black",
-               check_overlap = TRUE) +  # Display percentage improvement in each state
+               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
   labs(
-    fill = "Value",  # Label for the legend
-    subtitle = "Based on 2023 and 2024 influenza seasons"
-        , 
+    fill = "mean WIS",
        x = "",
@@
 ES_3WEEK<- ES_3WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
-
 ES_3WEEK
-
-ggsave(ES_3WEEK, height = 4.5, width = 6.5,  filename="Fig4C.jpg")  
+#ggsave(ES_3WEEK, height = 4.5, width = 6.5,  filename="Fig4C.jpg")  
 ```
@@
                        midpoint = 100, limits = c(0, 352)) +  
-  ggtitle("Mean WIS on each state (W4)") +
+  ggtitle("4 Weeks Ahead") +
   theme_light() +
-  theme(legend.position = "right") +
-  geom_sf_text(data = W4_map, aes(label = round(AUTO_AVG,1)),  
+  theme(legend.position = "right",
+    plot.title = element_text(hjust = 0.5)) +
+  geom_sf_text(data = W4_map, aes(label = round(AUTO_AVG,0)),  
                size = 3,
                color = "black",
-               check_overlap = TRUE) +
+               check_overlap = TRUE,fontface = "bold") +
   labs(
-    fill = "Value",  
-    subtitle = "Based on 2023 and 2024 influenza seasons"
-        , 
+    fill = "mean WIS", 
        x = "",
@@
 ES_4WEEK<-ES_4WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
-
 ES_4WEEK
-ggsave(ES_4WEEK,height = 4.5, width = 6.5,  filename="Fig4D.jpg")  
-```
+#ggsave(ES_4WEEK,height = 4.5, width = 6.5,  filename="Fig4D.jpg")  
+```
+
+Combine in a 2x2 grid
+
+```{r}
+# Combine in a 2x2 grid
+combined_plot <- (ES_1WEEK | ES_2WEEK) /
+                 (ES_3WEEK | ES_4WEEK) + 
+  plot_annotation(
+    title = "Figure X: Mean WIS for the 48 states in the contiguous U.S.",
+    subtitle = "Results for the the 2022/2023 and 2023/2024 influenza seasons",
+    theme = theme(plot.title = element_text(size = 16, face = "bold"),
+                  plot.subtitle = element_text(size = 14))
+  )
+# Print
+print(combined_plot)
+
+# Save to file
+ggsave("Fig4.jpg", combined_plot, width =12, height =7)
+
+```
+
 Loading datasets for regression analysis

### 2025-05-06 manuscript_results.Rmd
update

@@
 ```{r}
-# Assuming merged_data already includes the necessary columns: model and percentage_improveme
-# Create the map plot for 1 Week Ahead with best models and percentage improvement
+
 ES_1WEEK <- ggplot(W1_map) +
@@
   geom_sf_text(data = W1_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
-               size = 2.5,
+               size = 3,
                color = "black",
@@
   geom_sf_text(data = W2_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
-               size = 2.5,
+               size = 3,
                color = "black",
@@
 ```{r}
-
-# Assuming merged_data already includes the necessary columns: model and percentage_improveme
 
@@
   geom_sf_text(data = W3_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
-               size = 2.5,
+               size = 3,
                color = "black",
@@
 ```{r}
-
-# Assuming merged_data already includes the necessary columns: model and percentage_improveme
-
 ES_4WEEK <- ggplot(W4_map) +
@@
   geom_sf_text(data = W4_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
-               size = 2.5,
+               size = 3,
                color = "black",

### 2025-05-06 manuscript_results.Rmd
update

@@
 plt1
-ggsave(plt1, height = 8, width = 10, filename="fig1A.jpg")  
+#ggsave(plt1, height = 8, width = 10, filename="fig1A.jpg")  
 ```
@@
 plt2
-ggsave(plt2, height = 8, width = 10, filename="Fig1B.jpg")  
+#ggsave(plt2, height = 8, width = 10, filename="Fig1B.jpg")  
+```
+
+```{r}
+
+# Combine in a 2x2 grid
+combined_plot <- (plt1 | plt2) + 
+  plot_annotation(
+    title = "Figure X: Mean WIS for our 24 models",
+    subtitle = "Results by forecast horizon",
+    theme = theme(plot.title = element_text(size = 16, face = "bold"),
+                  plot.subtitle = element_text(size = 14))
+  )
+# Print
+print(combined_plot)
+
+# Save to file
+ggsave("Fig1.jpg", combined_plot, width =13, height =7)
 ```
@@
   labs(
-    fill = "%",  # Label for the legend
+    fill = "Percentage (%)",  # Label for the legend
        x = "",
@@
   labs(
-    fill = "%",  # Label for the legend
+    fill =  "Percentage (%)",  # Label for the legend
        x = "",

### 2025-05-06 manuscript_results.Rmd
update

@@
               fill = "gray", alpha = 0.25) +
-  labs(title = "Figure X: Example of influenza hospitalizaions forecasts",
+  labs(title = "Figure 1: Example of probabilistic forecasts",
        subtitle = "The gray ribbon indicates the 95% confidence interval",
@@
 
-ggsave(plt00, height = 5, width = 10, filename="fig00.jpg") 
+ggsave(plt00, height = 4, width =7, filename="Fig1.jpg") 
 ```
@@
   geom_vline(data = highlight_dates, aes(xintercept = Julian_date), color = "red", linetype = "dashed", size = 1) +  
-  geom_line(color = "black", size = 0.2) +  
+  geom_line(color = "black", size = 0.1) +  
   facet_wrap(~ Dataset, ncol = 2, scales = "free_y") +  
   theme_minimal() +  
-  labs(title = "Figure X: Baseline model WIS for 48 states on the contiguous U.S.",#,
-#       subtitle = "Dashed red lines indicates epidemiological weeks outsise of the influenza season",
-       x = "Month", 
+  labs(title = "Figure 2: Baseline model WIS for 48 states on the contiguous U.S.",#,
+       subtitle = "Weeks between the dashed red lines are outside of the influenza season",
+       x = "", 
        y = "WIS") +  
@@
         legend.position = "none") +
-  scale_x_date(labels = scales::date_format("%b"), breaks = "2 month")  # Format x-axis to show months and set monthly breaks
+  scale_x_date(
+    date_breaks = "1 month",
+    date_labels = "%b %y"
+  ) # Format x-axis to show months and set monthly breaks
 
@@
 
-ggsave(plt0, height = 5, width = 8, filename="fig0.jpg")  
+ggsave(plt0, height = 5, width = 8, filename="Fig2.jpg")  
 ```
@@
   plot_annotation(
-    title = "Figure X: Mean WIS for our 24 models",
-    subtitle = "Results by forecast horizon",
+    title = "Figure 3: Mean WIS for our 24 models",
+    subtitle = "Results by target week",
     theme = theme(plot.title = element_text(size = 16, face = "bold"),
@@
 # Save to file
-ggsave("Fig1.jpg", combined_plot, width =13, height =7)
+ggsave("Fig3.jpg", combined_plot, width =13, height =7)
 ```
@@
   scale_fill_manual(values = custom_colors) +
-  labs(title = "Figure X: Best model based on the lowest mean WIS",
-       subtitle = "Across the 48 states on the contiguous U.S.",
+  labs(title = "Figure 4: Best model based on the lowest mean WIS",
+       subtitle = "Results for the 48 states on the contiguous U.S. by target week",
        x = "", y = "Number of states",
@@
 # Save and display the plot
-ggsave("Fig2.jpg", combined_plot, width =8, height =5)
+ggsave("Fig4.jpg", combined_plot, width =8, height =5)
 print(combined_plot)
@@
                       name = "Percentage (%)", limits = c(-35, 35)) +  # Fixed color scale from -30 to 30
-  labs(title = "Figure X: Mean differences between models' WIS and the baseline across 48 states",
+  labs(title = "Figure 5: Mean differences between models' WIS and the baseline across 48 states",
        subtitle = "Gray boxes indicate models that are not significantly different from the baseline (p>0.05)",
          caption = "m = mean difference (%), sd = standard deviation",
-       x = "Week Ahead",
+       x = "Target Ahead",
        y = "") + 
@@
 print(plot_m_heatmap)
-ggsave("Fig3.jpg", plot_m_heatmap, width = 14, height = 7 )
+ggsave("Fig5.jpg", plot_m_heatmap, width = 14, height = 7 )
 
@@
   scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
-  ggtitle("1 Week Ahead") +
+  ggtitle("1 weeks ahead") +
   theme_light() +
@@
   scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
-  ggtitle("2 Weeks Ahead") +
+  ggtitle("2 weeks ahead") +
   theme_light() +
@@
   scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
-  ggtitle("3 Weeks Ahead") +
+  ggtitle("3 weeks ahead") +
   theme_light() +
@@
                        midpoint = 100, limits = c(0, 352)) +  
-  ggtitle("4 Weeks Ahead") +
+  ggtitle("4 weeks ahead") +
   theme_light() +
@@
   plot_annotation(
-    title = "Figure X: Mean WIS for the 48 states in the contiguous U.S.",
+    title = "Figure 6: Mean WIS for the 48 states in the contiguous U.S.",
     subtitle = "Results for the the 2022/2023 and 2023/2024 influenza seasons",
@@
 # Save to file
-ggsave("Fig4.jpg", combined_plot, width =12, height =7)
+ggsave("Fig6.jpg", combined_plot, width =12, height =7)
 
@@
   scale_fill_manual(values = c("Significant" = "cyan", "Not Significant" = "gray")) +  
-  labs(title = "Regression models",
+  labs(title = "Figure 7: Regression models",
        subtitle = "Mean WIS as dependent variable",
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1), 
-        axis.text.y = element_text(size = 10))
+        axis.text.y = element_text(size = 10),
+        plot.title = element_text(size = 16, face = "bold"),
+                  plot.subtitle = element_text(size = 14))
 
@@
 
-ggsave(reg2, height = 8, width = 10, filename="Fig4E.jpg")  
+ggsave(reg2, height = 6, width = 8, filename="Fig7.jpg")  
 #ggsave(reg2, filename="reg2.jpg",height = 8, width = 10)
@@
   scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
-  ggtitle("1 Week Ahead") +
+  ggtitle("1 weeks ahead") +
   theme_light() +
@@
   scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
-  ggtitle("2 Weeks Ahead") +
+  ggtitle("2 weeks ahead") +
   theme_light() +
@@
   scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
-  ggtitle("3 Weeks Ahead") +
+  ggtitle("3 weeks ahead") +
   theme_light() +
@@
   scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
-  ggtitle("4 Weeks Ahead") +
+  ggtitle("4 weeks ahead") +
   theme_light() +
@@
   plot_annotation(
-    title = "Figure X: Mean WIS differences (%) compared to the baseline",
+    title = "Figure 8: Mean WIS differences (%) compared to the baseline",
     subtitle = "Results by forecast horizon",
@@
 # Save to file
-ggsave("Fig5.jpg", combined_plot, width =12, height =7)
+ggsave("Fig8.jpg", combined_plot, width =12, height =7)
 ```
@@
   scale_fill_manual(values = c("Significant" = "cyan", "Not Significant" = "gray")) +  
-  labs(title = "Regression models",
-       subtitle = "Percentage of WIS improvement as dependent variable",
-       x = "Week Ahead",
+  labs(title = "Figure 9: Regression models",
+       subtitle = "Mean WIS differences (%) as dependent variable",
+       x = "Target Ahead",
        y = "Independent Variables",
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1), 
-        axis.text.y = element_text(size = 10))
+        axis.text.y = element_text(size = 10),
+        plot.title = element_text(size = 16, face = "bold"),
+                  plot.subtitle = element_text(size = 14))
 
 reg2
-ggsave(reg2, filename="Fig5E.jpg",height = 8, width = 10)
+ggsave(reg2, filename="Fig9.jpg",height = 6, width = 8)
 ```
@@
 ```{r}
-library(dplyr)
-library(ggplot2)
-
 # Create a list of data frames for the four weeks
@@
   labs(
-    title = "Figure X: Mean percentage of WIS differences across all states",
+    title = "Figure 10: Mean percentage of WIS differences across all states",
     subtitle = "AUTO_AVG_LB compared to baseline",
@@
 
-ggsave(model_improv, height = 4.5, width = 9,  filename="Fig6.jpg")  
-```
+ggsave(model_improv, height = 4.5, width = 9,  filename="Fig10.jpg")  
+```
+Talvez incluir standar deviation das diferenceas
+
+```{r}
+library(dplyr)
+library(ggplot2)
+
+# Create a list of data frames for the four weeks
+weeks <- 1:4
+summed_data_list <- list()
+
+for (week in weeks) {
+  # Merge the data frames for the current week
+  data__ <- merge(get(paste0("filtered_df_W", week, "_NoLg")),
+                  get(paste0("filtered_df_W", week)))
+
+  # Calculate the percentage difference at the row level (i.e., per state-date)
+  data__ <- data__ %>%
+    mutate(Difference = ((AUTO_AVG_LB / AUTO_AR) - 1) * 100)
+
+  # Group by Julian date and calculate mean and standard deviation of Difference
+  summarized <- data__ %>%
+    group_by(Julian_date) %>%
+    summarize(
+      Mean_Diff = Difference,
+      SD_Diff = sd(Difference, na.rm = TRUE)
+    ) %>%
+    mutate(Week = paste0("Week ", week))
+
+  # Store the summarized data
+  summed_data_list[[week]] <- summarized
+}
+
+# Combine all weeks into a single data frame
+summed_data_all <- bind_rows(summed_data_list)
+
+# Plot the mean difference with error bars for standard deviation
+model_improv <- ggplot(summed_data_all, aes(x = Julian_date, y = Mean_Diff, color = Week)) +
+  geom_point(size = 3) +
+  geom_errorbar(aes(ymin = Mean_Diff - SD_Diff, ymax = Mean_Diff + SD_Diff), width = 0.5) +
+  geom_hline(yintercept = 0, color = "black", linetype = "dashed", size = 0.8) +
+  theme_minimal() +
+  labs(
+    title = "Figure 10: Mean percentage of WIS differences across all states",
+    subtitle = "AUTO_AVG_LB compared to baseline",
+    x = "",
+    y = "Difference (%)"
+  ) +
+  scale_y_continuous(
+    limits = c(-100, 100),
+    breaks = seq(-100, 100, by = 20)
+  ) +
+  scale_x_date(
+    date_breaks = "1 month",
+    date_labels = "%b %y"
+  ) +
+  scale_color_manual(
+    values = c("Week 1" = "#ffcccc",
+               "Week 2" = "#ff6666",
+               "Week 3" = "#cc0000",
+               "Week 4" = "#990000")
+  ) +
+  theme(
+    plot.title = element_text(size = 18, face = "bold"),
+    plot.subtitle = element_text(size = 14),
+    axis.text.x = element_text(angle = 45, hjust = 1)
+  )
+
+model_improv
+
+#ggsave(model_improv, height = 4.5, width = 9, filename = "Fig11.jpg")
+
+```

### 2025-05-07 manuscript_results.Rmd
update

@@
 ```{r}
-library(dplyr)
-library(ggplot2)
-
-# Create a list of data frames for the four weeks
 weeks <- 1:4
@@
 for (week in weeks) {
-  # Merge the data frames for the current week
+  # Merge the data frames
   data__ <- merge(get(paste0("filtered_df_W", week, "_NoLg")),
@@
 
-  # Calculate the percentage difference at the row level (i.e., per state-date)
+  # Step 1: Calculate % difference at state level
   data__ <- data__ %>%
-    mutate(Difference = ((AUTO_AVG_LB / AUTO_AR) - 1) * 100)
-
-  # Group by Julian date and calculate mean and standard deviation of Difference
-  summarized <- data__ %>%
+    mutate(Diff_state = ((AUTO_AVG_LB / AUTO_AR) - 1) * 100)
+
+  # Step 2: Aggregate sum of components across states for mean difference
+  summed_data <- data__ %>%
     group_by(Julian_date) %>%
     summarize(
-      Mean_Diff = Difference,
-      SD_Diff = sd(Difference, na.rm = TRUE)
+      AUTO_AR = sum(AUTO_AR, na.rm = TRUE),
+      AUTO_AVG_LB = sum(AUTO_AVG_LB, na.rm = TRUE),
+      SD_Diff = sd(Diff_state, na.rm = TRUE)  # SD of raw state-level % differences
     ) %>%
-    mutate(Week = paste0("Week ", week))
-
-  # Store the summarized data
-  summed_data_list[[week]] <- summarized
+    mutate(
+      Difference = ((AUTO_AVG_LB / AUTO_AR) - 1) * 100,  # Mean from totals
+      Week = paste0("Week ", week)
+    )
+
+  summed_data_list[[week]] <- summed_data
 }
@@
 
-# Plot the mean difference with error bars for standard deviation
-model_improv <- ggplot(summed_data_all, aes(x = Julian_date, y = Mean_Diff, color = Week)) +
+model_improv <- ggplot(summed_data_all, aes(x = Julian_date, y = Difference, color = Week)) +
   geom_point(size = 3) +
-  geom_errorbar(aes(ymin = Mean_Diff - SD_Diff, ymax = Mean_Diff + SD_Diff), width = 0.5) +
+  geom_errorbar(aes(ymin = Difference - SD_Diff, ymax = Difference + SD_Diff), width = 0.5) +
   geom_hline(yintercept = 0, color = "black", linetype = "dashed", size = 0.8) +
@@
   scale_y_continuous(
-    limits = c(-100, 100),
+    limits = c(-400, 400),
     breaks = seq(-100, 100, by = 20)
@@
 model_improv
-
-#ggsave(model_improv, height = 4.5, width = 9, filename = "Fig11.jpg")
-
-```
+```

### 2025-05-07 ERA5 data/1990-2020/ERA5_1990_2020.Rmd
era5 data

---
title: "ERA5 Temperature and Humidity climatologies for regression analysis"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document converts ERA5 NetCDF's with daily data into weekly data for each state on the contiguous U.S.

```{r}
library(terra)
library(dplyr)
library(lubridate)
library(tidyr)
library(parallel)
library(ncdf4)
library(MMWRweek)
```

Downloading temperature and humidity data from the cloud.

```{python}
# Importando a biblioteca do xarray e outras auxiliares
import xarray as xr
import numpy as np
import os
import matplotlib.pyplot as plt
from mpl_toolkits.axes_grid1.inset_locator import inset_axes
import datetime as dt
import pandas as pd
import geopandas
import regionmask
import seaborn as sns
import pytz
import warnings
warnings.filterwarnings("ignore")

url='http://apdrc.soest.hawaii.edu/dods/public_data/Reanalysis_Data/ERA5/daily_3d/'
temperature = xr.open_dataset(url+'Air_Temperature')
temperature =temperature.sel(time=slice("1990-01-01","2020-12-31"))
temperature = temperature.assign_coords(lon=(((temperature.lon + 180) % 360) - 180))
temperature
# xarray dataset
temperature=temperature.sel(lev=1000, lat=slice(20,60),lon=slice(-130,-60))
temperature.to_netcdf(path='ERA5_temperature_data_US_2010_2024.nc')

url='http://apdrc.soest.hawaii.edu/dods/public_data/Reanalysis_Data/ERA5/daily_3d/'
humidity= xr.open_dataset(url+'Specific_humidity')
humidity=humidity.sel(time=slice("1990-01-01","2020-12-31"))
humidity= humidity.assign_coords(lon=(((humidity.lon + 180) % 360) - 180))
humidity=humidity.sel(lev=1000, lat=slice(20,60),lon=slice(-130,-60))
# xarray dataset
humidity.to_netcdf(path='ERA5_humidity_data_US_2010_2024.nc')
```
Plotting temperature and humidity NetCDFs.

```{r}
temp <- rast('ERA5_temperature_US_1990_2020.nc')  # Temperature
humi <- rast('ERA5_humidity_data_US_1990_2020.nc')  # Temperature

# Calculate the mean over the time dimension
temp_mean <- mean(temp, na.rm = TRUE)

# Calculate the mean over the time dimension
humi_mean <- mean(humi, na.rm = TRUE)

# Plot
plot(temp_mean)
plot(humi_mean)
```

Here I will load and inspect the states shapefiles. I will use them later in order to calculate weekly temperature and humidity for each state.

```{r}
# Codes and abbreviations of each U.S. state
state_code<-read.csv("State_Codes.csv")

# Read the shapefile
US <- vect("states_shapefile/States_shapefile.shp")

plot(US)

# Loop through the first 48 state codes
for (i in 1:48) {
    code <- state_code$abbreviation[i]

    # Filter the SpatVector based on the state code
    state_shapefile <- US[US$State_Code == code, ]

    # Plot the filtered state shapefile
    plot(state_shapefile, main = paste("State:", code))
}

state_code
```

Here I develop a function that take states codes (for shapefiles selection) and NetCDFs with daily resolution (temperature or humidity) as inputs, and creates a dataframe with weekly values for each U.S. state.

```{r}
ERA5_mean_by_state<-function(state_code=state_code, my_netcdf) {
  # Filter the state shapefile for the current state
  state_shapefile <- US[US$State_Code == state_code$abbreviation, ]
  # Create a mask for the state using terra's mask function
  netcdf_state_daily <- mask(my_netcdf, state_shapefile)
  # Calculate the mean for a given state
  temp_mean <- mean(netcdf_state_daily, na.rm = TRUE)  
  # Calculate the mean across the spatial dimensions (lat, lon)
  state_mean <- global(temp_mean, mean, na.rm = TRUE)
    # Rename the column to the state's location name
  colnames(state_mean) <- state_code$location_name
  return(state_mean)
}
```

Here I select only the abbreviations and state codes of states on the contiguous U.S.

```{r}
# Split dataframe into a list of dataframes by row
states_codes <- split(state_code, seq(nrow(state_code)))
states_codes<-states_codes[1:48] # select only states up to Florida

```

Here I run the function to calculate mean weekly TEMPERATURE for all states on the contiguous U.S.

```{r}
# Run using mclapply

mean_temperature_each_state_1990_2020 <- mclapply(states_codes, function(state_code) {
  ERA5_mean_by_state(state_code = state_code, my_netcdf = temp)
}, mc.cores = 1)

mean_temperature_each_state_1990_2020_df<-bind_cols(mean_temperature_each_state_1990_2020)
mean_temperature_each_state_1990_2020_df<-data.frame(t(mean_temperature_each_state_1990_2020_df))
mean_temperature_each_state_1990_2020_df$STATE <-rownames(mean_temperature_each_state_1990_2020_df)
rownames(mean_temperature_each_state_1990_2020_df) <- NULL
mean_temperature_each_state_1990_2020_df
write.csv(mean_temperature_each_state_1990_2020_df, file = "temperature_climatology_1990_2020.csv")
```

Here I run the function to calculate mean weekly SPECIFIC HUMIDITY for all states on the contiguous U.S.

```{r}
mean_humidity_each_state_1990_2020 <- mclapply(states_codes, function(state_code) {
  ERA5_mean_by_state(state_code = state_code, my_netcdf = humi)
}, mc.cores = 1)

mean_humidity_each_state_1990_2020_df<-bind_cols(mean_humidity_each_state_1990_2020)
mean_humidity_each_state_1990_2020_df<-data.frame(t(mean_humidity_each_state_1990_2020_df))
mean_humidity_each_state_1990_2020_df$STATE <-rownames(mean_humidity_each_state_1990_2020_df)
rownames(mean_humidity_each_state_1990_2020_df) <- NULL
mean_humidity_each_state_1990_2020_df
write.csv(mean_humidity_each_state_1990_2020_df, file = "humidity_climatology_1990_2020.csv")
```

### 2025-05-07 ERA5 data/2010-2024/ERA5_2010_2024.Rmd
era5 data

---
title: "ERA5 Temperature and Humidity weekly data 2010-2024"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document converts ERA5 NetCDF's with daily data into weekly data for each state on the contiguous U.S.

```{r}
library(terra)
library(dplyr)
library(lubridate)
library(tidyr)
library(parallel)
library(ncdf4)
library(MMWRweek)
```

Downloading temperature and humidity data from the cloud.

```{python}
# Importando a biblioteca do xarray e outras auxiliares
import xarray as xr
import numpy as np
import os
import matplotlib.pyplot as plt
from mpl_toolkits.axes_grid1.inset_locator import inset_axes
import datetime as dt
import pandas as pd
import geopandas
import regionmask
import seaborn as sns
import pytz
import warnings
warnings.filterwarnings("ignore")

url='http://apdrc.soest.hawaii.edu/dods/public_data/Reanalysis_Data/ERA5/daily_3d/'
temperature = xr.open_dataset(url+'Air_Temperature')
temperature =temperature.sel(time=slice("2010-01-01","2024-07-12"))
temperature = temperature.assign_coords(lon=(((temperature.lon + 180) % 360) - 180))
temperature
# xarray dataset
temperature=temperature.sel(lev=1000, lat=slice(20,60),lon=slice(-130,-60))
temperature.to_netcdf(path='ERA5_temperature_data_US_2010_2024.nc')

url='http://apdrc.soest.hawaii.edu/dods/public_data/Reanalysis_Data/ERA5/daily_3d/'
humidity= xr.open_dataset(url+'Specific_humidity')
humidity=humidity.sel(time=slice("2010-01-01","2024-07-12"))
humidity= humidity.assign_coords(lon=(((humidity.lon + 180) % 360) - 180))
humidity=humidity.sel(lev=1000, lat=slice(20,60),lon=slice(-130,-60))
# xarray dataset
humidity.to_netcdf(path='ERA5_humidity_data_US_2010_2024.nc')
```

Plotting temperature and humidity NetCDFs.

```{r}
temp <- rast('ERA5_temperature_US_2010_2024.nc')  # Temperature
humi <- rast('ERA5_humidity_data_US_2010_2024.nc')  # Humidity

# Calculate the mean over the time dimension
temp_mean <- mean(temp, na.rm = TRUE)

# Calculate the mean over the time dimension
humi_mean <- mean(humi, na.rm = TRUE)

# Plot
plot(temp_mean)
plot(humi_mean)
```

Here I will load and inspect the states' shapefiles. I will use them later in order to calculate weekly temperature and humidity for each state.

```{r}
# Codes and abbreviations of each U.S. state
state_code<-read.csv("State_Codes.csv")

# Read the shapefile
US <- vect("states_shapefile/States_shapefile.shp")

plot(US)

# Loop through the first 48 state codes
for (i in 1:48) {
    code <- state_code$abbreviation[i]

    # Filter the SpatVector based on the state code
    state_shapefile <- US[US$State_Code == code, ]

    # Plot the filtered state shapefile
    plot(state_shapefile, main = paste("State:", code))
}

state_code
```

Here I develop a function that take states codes for shapefiles selection and NetCDFs with daily resolution (temperature or humidity) as inputs. It then creates a dataframe with weekly values for each U.S. state that will be utilized as covariates.

```{r}
ERA5_weekly_dataframe<-function(state_code=state_code, my_netcdf) {
  # Filter the state shapefile for the current state
  state_shapefile <- US[US$State_Code == state_code$abbreviation, ]
  # Create a mask for the state using terra's mask function
  netcdf_state_daily <- mask(my_netcdf, state_shapefile)
  # Extract the time dimension (dates)
  dates <- time(netcdf_state_daily)
  # Calculate the mean for each day across the spatial dimensions (lat, lon)
  daily_state_mean_ts <- global(netcdf_state_daily, mean, na.rm = TRUE)
  # Convert the daily time series to a data frame
  daily_state_df <- as.data.frame(daily_state_mean_ts, xy = FALSE, na.rm = TRUE)
  # Add the time column to the data frame
  daily_state_df$time <- as.Date(dates)
  # Calculate the epidemiological week and include as a column
  daily_state_df$epi_week <- MMWRweek(daily_state_df$time)$MMWRweek
  # Get the year and include as a column
  daily_state_df$year <- MMWRweek(daily_state_df$time)$MMWRyear
  # Calculate the mean temperature for each epidemiological week and year
  weekly_mean_value_by_state <- daily_state_df %>%
    group_by(year, epi_week) %>%
    summarise(weekly_value = mean(mean, na.rm = TRUE), .groups = "drop") 
  # Rename the column to the state's location name
  colnames(weekly_mean_value_by_state)[3] <- state_code$location_name
  # Return the weekly data for the current state
  return(weekly_mean_value_by_state)
}
```

Here I select only the abbreviations and state codes of states on the contiguous U.S.

```{r}
# Split dataframe into a list of dataframes by row
states_codes <- split(state_code, seq(nrow(state_code)))
states_codes<-states_codes[1:48] # select only states up to Florida
```

Here I run the function to calculate mean weekly TEMPERATURE for all states on the contiguous U.S.

```{r}
# Run using mclapply

temperature_list_of_dataframes <- mclapply(states_codes, function(state_code) {
  ERA5_weekly_dataframe(state_code = state_code, my_netcdf = temp)
}, mc.cores = 1)

final_temperature_dataframe <- bind_rows(temperature_list_of_dataframes) %>%
   mutate(date = MMWRweek2Date(MMWRyear = year, MMWRweek = epi_week)+6)%>% # I need to add 6 to get the end_week_day
   group_by(date) %>%
   summarize(across(everything(), mean, na.rm = TRUE))

final_temperature_dataframe

write.csv(final_temperature_dataframe, file = "final_temperature_data_2010_2024.csv")
```

Here I run the function to calculate mean weekly SPECIFIC HUMIDITY for all states on the contiguous U.S.

```{r}
# Run using mclapply
humidity_df <- mclapply(states_codes, function(state_code) {
  ERA5_weekly_dataframe(state_code = state_code, my_netcdf = humi)
}, mc.cores = 1)

final_humidity_dataframe <- bind_rows(humidity_df) %>%
   mutate(date = MMWRweek2Date(MMWRyear = year, MMWRweek = epi_week)+6)%>%
   group_by(date) %>%
   summarize(across(everything(), mean, na.rm = TRUE))

write.csv(final_humidity_dataframe , file = "final_humidity_data_2010_2024.csv")
```

### 2025-05-21 manuscript_results.Rmd
update

--- 2025-05-06 manuscript_results.Rmd
@@
   theme_minimal() +  
-  labs(title = "Figure 2: Baseline model WIS for 48 states on the contiguous U.S.",#,
+  labs(title = "Baseline model WIS for 48 states on the contiguous U.S.",#,
        subtitle = "Weeks between the dashed red lines are outside of the influenza season",
@@
     scale_fill_manual(values = custom_colors) +
-    facet_wrap(category ~ Models , nrow = 3) +
+    facet_wrap(category ~ Models , nrow = 5) +
     labs(
       title = "A) Models without log-back transformations",
-      x = "Target Week",
+      x = "",
       y = "Mean Weighted Interval Score (WIS)"
@@
   scale_y_continuous(limits = c(0, 650)) +  # Set y-axis range
-    theme_minimal()
+    theme_minimal()+
+  theme(legend.position = "none")
 plt1
-#ggsave(plt1, height = 8, width = 10, filename="fig1A.jpg")  
+#ggsave(plt1, height = 12, width = 8, filename="fig1A.jpg")  
 ```
@@
     scale_fill_manual(values = custom_colors) +
-    facet_wrap(category ~ Models , nrow = 3) +
+    facet_wrap(category ~ Models , nrow = 5) +
     labs(
@@
   scale_y_continuous(limits = c(0, 650)) +  # Set y-axis range
-    theme_minimal()
+    theme_minimal()+
+  theme(legend.position = "none")
 plt2
@@
   plot_annotation(
-    title = "Figure 3: Mean WIS for our 24 models",
+    title = "Mean WIS for the 24 models",
     subtitle = "Results by target week",
@@
 # Save to file
-ggsave("Fig3.jpg", combined_plot, width =13, height =7)
+ggsave("Fig3.jpg", combined_plot, width =9, height =10)
 ```
@@
   scale_fill_manual(values = custom_colors) +
-  labs(title = "Figure 4: Best model based on the lowest mean WIS",
+  labs(title = "Best model based on the lowest mean WIS",
        subtitle = "Results for the 48 states on the contiguous U.S. by target week",
@@
                       name = "Percentage (%)", limits = c(-35, 35)) +  # Fixed color scale from -30 to 30
-  labs(title = "Figure 5: Mean differences between models' WIS and the baseline across 48 states",
+  labs(title = "Mean differences between models' WIS and the baseline across 48 states",
        subtitle = "Gray boxes indicate models that are not significantly different from the baseline (p>0.05)",
@@
   plot_annotation(
-    title = "Figure 6: Mean WIS for the 48 states in the contiguous U.S.",
+    title = "Mean WIS for the 48 states in the contiguous U.S.",
     subtitle = "Results for the the 2022/2023 and 2023/2024 influenza seasons",
@@
   scale_fill_manual(values = c("Significant" = "cyan", "Not Significant" = "gray")) +  
-  labs(title = "Figure 7: Regression models",
+  labs(title = "Regression models R²",
        subtitle = "Mean WIS as dependent variable",
-       x = "Week Ahead",
+       x = "Target Week",
        y = "Independent Variables",
@@
   plot_annotation(
-    title = "Figure 8: Mean WIS differences (%) compared to the baseline",
+    title = "Mean WIS differences compared to the baseline",
     subtitle = "Results by forecast horizon",
@@
   scale_fill_manual(values = c("Significant" = "cyan", "Not Significant" = "gray")) +  
-  labs(title = "Figure 9: Regression models",
+  labs(title = "Regression models R²",
        subtitle = "Mean WIS differences (%) as dependent variable",
-       x = "Target Ahead",
+       x = "Target Week",
        y = "Independent Variables",
@@
 
-Now let's evaluate which was the mean WIS improvement by epidemiological week
-
 ```{r}
@@
     summarize(
-      AUTO_AR = sum(AUTO_AR, na.rm = TRUE),  # Sum across all states
-      AUTO_AVG_LB = sum(AUTO_AVG_LB, na.rm = TRUE)           # Sum across all states
+      AUTO_AR = mean(AUTO_AR, na.rm = TRUE),  # Sum across all states
+      AUTO_AVG_LB = mean(AUTO_AVG_LB, na.rm = TRUE)           # Sum across all states
     ) %>%
@@
   labs(
-    title = "Figure 10: Mean percentage of WIS differences across all states",
+    title = "Mean percentage of WIS differences across all states",
     subtitle = "AUTO_AVG_LB compared to baseline",
@@
 ```
-Talvez incluir standar deviation das diferenceas
-
-```{r}
-library(dplyr)
-library(ggplot2)
-
-# Create a list of data frames for the four weeks
-weeks <- 1:4
-summed_data_list <- list()
-
-for (week in weeks) {
-  # Merge the data frames for the current week
-  data__ <- merge(get(paste0("filtered_df_W", week, "_NoLg")),
-                  get(paste0("filtered_df_W", week)))
-
-  # Calculate the percentage difference at the row level (i.e., per state-date)
-  data__ <- data__ %>%
-    mutate(Difference = ((AUTO_AVG_LB / AUTO_AR) - 1) * 100)
-
-  # Group by Julian date and calculate mean and standard deviation of Difference
-  summarized <- data__ %>%
-    group_by(Julian_date) %>%
-    summarize(
-      Mean_Diff = Difference,
-      SD_Diff = sd(Difference, na.rm = TRUE)
-    ) %>%
-    mutate(Week = paste0("Week ", week))
-
-  # Store the summarized data
-  summed_data_list[[week]] <- summarized
-}
-
-# Combine all weeks into a single data frame
-summed_data_all <- bind_rows(summed_data_list)
-
-# Plot the mean difference with error bars for standard deviation
-model_improv <- ggplot(summed_data_all, aes(x = Julian_date, y = Mean_Diff, color = Week)) +
-  geom_point(size = 3) +
-  geom_errorbar(aes(ymin = Mean_Diff - SD_Diff, ymax = Mean_Diff + SD_Diff), width = 0.5) +
-  geom_hline(yintercept = 0, color = "black", linetype = "dashed", size = 0.8) +
-  theme_minimal() +
-  labs(
-    title = "Figure 10: Mean percentage of WIS differences across all states",
-    subtitle = "AUTO_AVG_LB compared to baseline",
-    x = "",
-    y = "Difference (%)"
-  ) +
-  scale_y_continuous(
-    limits = c(-100, 100),
-    breaks = seq(-100, 100, by = 20)
-  ) +
-  scale_x_date(
-    date_breaks = "1 month",
-    date_labels = "%b %y"
-  ) +
-  scale_color_manual(
-    values = c("Week 1" = "#ffcccc",
-               "Week 2" = "#ff6666",
-               "Week 3" = "#cc0000",
-               "Week 4" = "#990000")
-  ) +
-  theme(
-    plot.title = element_text(size = 18, face = "bold"),
-    plot.subtitle = element_text(size = 14),
-    axis.text.x = element_text(angle = 45, hjust = 1)
-  )
-
-model_improv
-
-#ggsave(model_improv, height = 4.5, width = 9, filename = "Fig11.jpg")
-
-```

### 2025-05-23 manuscript_results.Rmd
update

@@
             color = "black", size = 4) +  # Add text labels
-  scale_fill_gradient(low = "cyan", high = "red", na.value = "gray80", 
-                      name = "Percentage (%)", limits = c(-35, 35)) +  # Fixed color scale from -30 to 30
+  scale_fill_gradient2(
+  low = "cyan", 
+  mid = "white", 
+  high = "red", 
+  midpoint = 0,     # <-- Sets 0 as the midpoint
+  na.value = "gray80", 
+  name = "Mean difference", 
+  limits = c(-35, 35)       
+)+
+
   labs(title = "Mean differences between models' WIS and the baseline across 48 states",
        subtitle = "Gray boxes indicate models that are not significantly different from the baseline (p>0.05)",
-         caption = "m = mean difference (%), sd = standard deviation",
+         caption = "m = mean difference, sd = standard deviation",
        x = "Target Ahead",
@@
 regression_models
+
+```
+
+best regression plot
+
+```{r}
+plot_list <- list()  # store plots
+
+for(i in 1:4) {
+
+  # Combine WIS data and population
+  WIS_pop_data <- inner_join(WIS_dataframes[[i]], pop_data, by = "STATE")
+
+  # Fit linear model
+  model <- lm(AUTO_AVG ~ Resident_population_2020, data = WIS_pop_data)
+  model_summary <- summary(model)
+
+  # Save R² and p-values
+  r_squared <- round(model_summary$r.squared, 3)
+  p_value <- signif(model_summary$coefficients[2, 4], 3)
+
+  # Create plot
+  p <- ggplot(WIS_pop_data, aes(x = Resident_population_2020, y = AUTO_AVG)) +
+    geom_point(color = "steelblue", size = 2) +
+    geom_smooth(method = "lm", se = FALSE, color = "darkred", linetype = "dashed") +
+    labs(
+      title = paste("Regression","Target Week", i),
+      subtitle = paste("R² =", r_squared, "| p =", p_value),
+      x = "Resident Population (2020)",
+      y = "AUTO_AVG"
+    ) +
+    theme_minimal()
+
+  # Store the plot
+  plot_list[[i]] <- p
+}
+
+# Combine the plots vertically
+combined_regressions <- wrap_plots(plot_list, ncol = 1)
+
+# Display (optional)
+combined_regressions
+
+# Save the combined plot
+#ggsave("Figx.png", combined_plot, width = 3, height = 5, dpi = 600)
 ```
@@
 
-ggsave(reg2, height = 6, width = 8, filename="Fig7.jpg")  
+# Combine the plots vertically
+reg2_scatter <- reg2 | combined_regressions 
+reg2_scatter
+ggsave(reg2_scatter, height = 6, width = 12, filename="Fig7.jpg")  
 #ggsave(reg2, filename="reg2.jpg",height = 8, width = 10)
@@
   labs(
-    fill = "Percentage (%)",  # Label for the legend
+    fill = "Mean difference",  # Label for the legend
        x = "",
@@
   labs(
-    fill =  "Percentage (%)",  # Label for the legend
+    fill =  "Mean difference",  # Label for the legend
        x = "",
@@
 
+best regression plot
+
+```{r}
+
+# Prepare a list to hold your plots
+bric_plots <- list()
+
+for (i in 1:4) {
+  # 1) Join the WIS and BRIC data
+  WIS_bric <- inner_join(
+    WIS_dataframes[[i]],
+    bric_by_state,
+    by = "STATE"
+  )
+
+  # 2) Fit the linear model
+  model <- lm(
+    percentage_improvement ~ weighted_mean_z_bric.infrastructure,
+    data = WIS_bric
+  )
+  ms <- summary(model)
+
+  # 3) Extract R² and p‐value
+  r2 <- round(ms$r.squared, 3)
+  p  <- signif(ms$coefficients[2, 4], 3)
+
+  # 4) Build the ggplot
+  p <- ggplot(WIS_bric,
+              aes(x = weighted_mean_z_bric.infrastructure,
+                  y = percentage_improvement)) +
+    geom_point(color = "steelblue", size = 2) +
+    geom_smooth(method = "lm", se = FALSE, color = "darkred") +
+    labs(
+      title    = paste("Regression","Target Week", i),
+      subtitle = paste0("R² = ", r2, "   p = ", p),
+      x        = "BRIC infrastructure (2015)",
+      y        = "Difference"
+    ) +
+    theme_minimal()
+
+  # 5) Store it
+  bric_plots[[i]] <- p
+}
+
+# Combine the plots vertically
+combined_regressions <- wrap_plots(bric_plots, ncol = 1)
+
+# Display 
+combined_regressions
+
+```
+
 AUTO_AVG WIS x BRIC institutional
@@
 reg2
-ggsave(reg2, filename="Fig9.jpg",height = 6, width = 8)
+
+reg2_scatter <- reg2 | combined_regressions 
+reg2_scatter
+ggsave(reg2_scatter, height = 6, width = 12, filename="Fig9.jpg")  
+#ggsave(reg2, filename="Fig9.jpg",height = 6, width = 8)
 ```
@@
 model_improv<-ggplot(summed_data_all, aes(x = Julian_date, y = (Difference), color = Week)) +
-  geom_point(size = 3) +
+  geom_point(size = 2.5) +
   geom_hline(yintercept = 0, color = "black", linetype = "dashed", size = 0.8) +
@@
   labs(
-    title = "Mean percentage of WIS differences across all states",
+    title = "Weekly mean WIS differences across all states",
     subtitle = "AUTO_AVG_LB compared to baseline",
     x = "",
-    y = "Difference (%)"
+    y = "Mean difference"
   ) +
@@
   scale_color_manual(
-    values = c("Week 1" = "#ffcccc",
-               "Week 2" = "#ff6666",
-               "Week 3" = "#cc0000",
-               "Week 4" = "#990000")
+    values = c(
+    "Week 1" = "#00BFC4",  # Bright cyan
+    "Week 2" = "#F8766D",  # Bright coral red
+    "Week 3" = "#7CAE00",  # Lime green
+    "Week 4" = "#C77CFF"   # Bright purple
+  )
   ) +
@@
 ```
+hospitalizations
+
+```{r}
+
+total_hospitalizations_by_week <- AUTO_ADJACENT_WEEK1 %>%
+  filter(epiweek >= 40 | epiweek <= 20) %>%  # Apply filter AFTER epiweek exists
+  group_by(target_end_date) %>%
+  summarise(mean_cases = sum(cases, na.rm = TRUE))
+
+hospital_plot <-ggplot(total_hospitalizations_by_week, aes(x = target_end_date, y = mean_cases)) +
+  geom_bar(stat = "identity",fill = "steelblue") +
+  labs(title = "Total influenza hospitalizations on the 48 states",
+       x = "",
+       y = "Influenza hospitalizations") +
+  scale_x_date(
+    date_breaks = "1 month",
+    date_labels = "%b %y"
+  ) +
+  theme_minimal()+
+  theme(plot.title = element_text(size = 18, face = "bold"),
+    plot.subtitle = element_text(size = 14),
+    axis.text.x = element_text(angle = 45, hjust = 1))
+
+hospital_plot
+
+```
+
+```{r}
+
+# Combine the plots vertically
+combined_plot <- hospital_plot / model_improv
+
+# Display (optional)
+combined_plot
+
+# Save the combined plot
+ggsave("Fig10.png", combined_plot, width = 7, height = 7, dpi = 600)
+
+```

### 2025-05-23 manuscript_results.Rmd
update

@@
 
-ggsave(model_improv, height = 4.5, width = 9,  filename="Fig10.jpg")  
+#ggsave(model_improv, height = 4.5, width = 9,  filename="Fig10.jpg")  
 ```
@@
 # Save the combined plot
-ggsave("Fig10.png", combined_plot, width = 7, height = 7, dpi = 600)
-
-```
+ggsave("Fig10.jpg", combined_plot, width = 7, height = 7, dpi = 600)
+
+```

### 2025-05-29 manuscript_results.Rmd
update

@@
 
-Example of influenza hospitalizations forecasts and 95% confidence interval.
-
-```{r}
-
-# Create a data frame with forecasts for plotting
-forecast_df <- data.frame(
-  Week = factor(1:4, labels = paste0("Week ", 1:4)),  # Label weeks
-  Value = c(AUTO_ARIMA_WEEK1_list[[1]]$forecasts[5], 
-            AUTO_ARIMA_WEEK2_list[[1]]$forecasts[5], 
-            AUTO_ARIMA_WEEK3_list[[1]]$forecasts[5], 
-            AUTO_ARIMA_WEEK4_list[[1]]$forecasts[5]),
-  Lower = c(AUTO_ARIMA_WEEK1_list[[1]][["0.025"]][5], 
-            AUTO_ARIMA_WEEK2_list[[1]][["0.025"]][5], 
-            AUTO_ARIMA_WEEK3_list[[1]][["0.025"]][5], 
-            AUTO_ARIMA_WEEK4_list[[1]][["0.025"]][5]),
-  Upper = c(AUTO_ARIMA_WEEK1_list[[1]][["0.975"]][5], 
-            AUTO_ARIMA_WEEK2_list[[1]][["0.975"]][5], 
-            AUTO_ARIMA_WEEK3_list[[1]][["0.975"]][5], 
-            AUTO_ARIMA_WEEK4_list[[1]][["0.975"]][5])
-)
-
-# Plotting forecasts and ribbon
-plt00 <- ggplot(forecast_df, aes(x = Week, y = Value, group = 1)) +
-  geom_line(color = "black", size = 1) +          # Line connecting forecast values
-  geom_point(color = "black", size = 3) +         # Points for each forecast value
-  geom_ribbon(aes(ymin = Lower, ymax = Upper),   # Ribbon for confidence intervals
-              fill = "gray", alpha = 0.25) +
-  labs(title = "Figure 1: Example of probabilistic forecasts",
-       subtitle = "The gray ribbon indicates the 95% confidence interval",
-       x = "Target Weeks",
-       y = "Forecast Values") +
-  theme_minimal() +
-  theme(plot.title = element_text(size = 16, face = "bold"),
-        plot.subtitle = element_text(size = 14))
-
-plt00
-
-ggsave(plt00, height = 4, width =7, filename="Fig1.jpg") 
-```
-
 Weighted interval scores for each state and each target week.
@@
 
+Fitting data.
+
+```{r}
+i  <- 57
+ii <- 8 # state number
+
+my_data = read.csv("NHSH_hospitalizations_seed/treated_influenza_hosp_dataframe_v2.csv")
+my_data$target_end_date<-as.Date(my_data$target_end_date) # set the dates as dates
+
+list_of_states <- split(my_data, my_data$state_name)
+
+# Convert to Date if not already
+list_of_states[[ii]]$target_end_date <- as.Date(list_of_states[[ii]]$target_end_date)
+
+# Define the date range
+end_date <- as.Date(AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i]-7)
+start_date <- end_date - 365*2  # 2 years before
+
+# Subset the data
+subset_data <- subset(list_of_states[[ii]], 
+                      target_end_date >= start_date & target_end_date <= end_date)
+
+# Plot
+plot(subset_data$target_end_date, subset_data$cases, type = "l",
+     xlab = "Date", ylab = "Cases", main = "Cases - Last 2 Years")
+```
+
+```{r}
+# 2) build your forecast_df for i=55, ii=8
+
+i  <- 57
+ii <- 8 # state number
+
+forecast_df <- data.frame(
+  Dates = c(
+    AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i],
+    AUTO_AVERAGE_WEEK2_list[[ii]]$target_end_date[i],
+    AUTO_AVERAGE_WEEK3_list[[ii]]$target_end_date[i],
+    AUTO_AVERAGE_WEEK4_list[[ii]]$target_end_date[i]
+  ),
+  Hosp  = c(
+    AUTO_AVERAGE_WEEK1_list[[ii]]$cases[i],
+    AUTO_AVERAGE_WEEK2_list[[ii]]$cases[i],
+    AUTO_AVERAGE_WEEK3_list[[ii]]$cases[i],
+    AUTO_AVERAGE_WEEK4_list[[ii]]$cases[i]
+  ),
+  Value = c(
+    AUTO_AVERAGE_WEEK1_list[[ii]]$forecasts[i],
+    AUTO_AVERAGE_WEEK2_list[[ii]]$forecasts[i],
+    AUTO_AVERAGE_WEEK3_list[[ii]]$forecasts[i],
+    AUTO_AVERAGE_WEEK4_list[[ii]]$forecasts[i]
+  ),
+  Lower = c(
+    AUTO_AVERAGE_WEEK1_list[[ii]][["0.025"]][i],
+    AUTO_AVERAGE_WEEK2_list[[ii]][["0.025"]][i],
+    AUTO_AVERAGE_WEEK3_list[[ii]][["0.025"]][i],
+    AUTO_AVERAGE_WEEK4_list[[ii]][["0.025"]][i]
+  ),
+  Upper = c(
+    AUTO_AVERAGE_WEEK1_list[[ii]][["0.975"]][i],
+    AUTO_AVERAGE_WEEK2_list[[ii]][["0.975"]][i],
+    AUTO_AVERAGE_WEEK3_list[[ii]][["0.975"]][i],
+    AUTO_AVERAGE_WEEK4_list[[ii]][["0.975"]][i]
+  )
+)
+
+conceptual_figure = ggplot() +
+  # 3a) two-year observed history as a black line (mapped to color for legend)
+  geom_point(data = subset_data, 
+            aes(x = target_end_date, y = cases, color = "Fitting data"), 
+            size = 2.5) +
+
+  # 3b) forecast ribbon (not in legend)
+  geom_ribbon(data = forecast_df, 
+              aes(x = Dates, ymin = Lower, ymax = Upper), 
+              fill = "gray", alpha = 0.4) +
+
+  # 3c) forecast mean (dashed blue line + hollow points)
+  geom_point(data = forecast_df, 
+            aes(x = Dates, y = Value, color = "Mean forecast"), 
+             size = 3) +
+
+  geom_point(data = forecast_df, 
+             aes(x = Dates, y = Value, color = "Mean forecast"), 
+             shape = 21, fill = "white", size = 2) +
+
+  # 3d) actual post-cutoff cases (mapped to color for legend)
+  geom_point(data = forecast_df, 
+             aes(x = Dates, y = Hosp, color = "Hospitalizations"), 
+             shape = 16, size = 2.5) +
+
+  # Manually define legend entries
+  scale_color_manual(name = "Legend",
+                     values = c("Fitting data" = "red", 
+                                "Mean forecast" = "cyan3",
+                                "Hospitalizations" = "black")) +
+    scale_x_date(
+    date_breaks = "1 month",
+    date_labels = "%b %y"
+  ) +
+
+  labs(
+    title    = "Example of probabilistic forecasts",
+    subtitle = "Gray ribbon represents 95% confidence interval",
+    x        = "",
+    y        = "Hospitalizations"
+  ) +
+  theme_minimal() +
+  theme(plot.title = element_text(size = 18, face = "bold"),
+    plot.subtitle = element_text(size = 14),
+    axis.text.x = element_text(angle = 45, hjust = 1))
+
+conceptual_figure
+# Save the combined plot
+ggsave("Fig1.jpg", conceptual_figure, width = 7, height = 4, dpi = 600)
+
+```
 Calculate mean weighted interval score for all forecast weeks on each model.

### 2025-05-30 manuscript_results.Rmd
update

@@
 library("patchwork")
+library("ggpattern")
 
@@
 end_date <- as.Date(AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i]-7)
-start_date <- end_date - 365*2  # 2 years before
+start_date <- end_date - (7*103)  # 2 years before
 
@@
 
-i  <- 57
+i  <- 57 # target_end_date
 ii <- 8 # state number
@@
 
+custom_colors <- c(
+  "AUTO_AR_LB" = colors_ar[1], "ES27_AR_LB" = colors_ar[1], "ES64_AR_LB" = colors_ar[1],
+  "AUTO_EPI_LB" = colors_epi[1], "ES27_EPI_LB" = colors_epi[1], "ES64_EPI_LB" = colors_epi[1],
+  "AUTO_TMP_LB" = colors_tmp[1], "ES27_TMP_LB" = colors_tmp[1], "ES64_TMP_LB" = colors_tmp[1],
+  "AUTO_ADJ_LB" = colors_adj[1], "ES27_ADJ_LB" = colors_adj[1], "ES64_ADJ_LB" = colors_adj[1],
+  "AUTO_AVG_LB" = colors_avg[1], "ES27_AVG_LB" = colors_avg[1], "ES64_AVG_LB" = colors_avg[1],
+
+  "AUTO_AR" = colors_ar[1], "ES27_AR" = colors_ar[1], "ES64_AR" = colors_ar[1],
+  "AUTO_EPI" = colors_epi[1], "ES27_EPI" = colors_epi[1], "ES64_EPI" = colors_epi[1],
+  "AUTO_TMP" = colors_tmp[1], "ES27_TMP" = colors_tmp[1], "ES64_TMP" = colors_tmp[1],
+  "AUTO_ADJ" = colors_adj[1], "ES27_ADJ" = colors_adj[1], "ES64_ADJ" = colors_adj[1],
+  "AUTO_AVG" = colors_avg[1], "ES27_AVG" = colors_avg[1], "ES64_AVG" = colors_avg[1]
+)
+
+# define a manual pattern scale
+custom_patterns <- c(
+  # AR
+  "AUTO_AR"      = "circle",      "AUTO_AR_LB"      = "circle",
+  "ES27_AR"      = "stripe",      "ES27_AR_LB"      = "stripe",
+  "ES64_AR"      = "crosshatch",  "ES64_AR_LB"      = "crosshatch",
+  # EPI
+  "AUTO_EPI"     = "circle",      "AUTO_EPI_LB"     = "circle",
+  "ES27_EPI"     = "stripe",      "ES27_EPI_LB"     = "stripe",
+  "ES64_EPI"     = "crosshatch",  "ES64_EPI_LB"     = "crosshatch",
+  # TMP
+  "AUTO_TMP"     = "circle",      "AUTO_TMP_LB"     = "circle",
+  "ES27_TMP"     = "stripe",      "ES27_TMP_LB"     = "stripe",
+  "ES64_TMP"     = "crosshatch",  "ES64_TMP_LB"     = "crosshatch",
+  # ADJ
+  "AUTO_ADJ"     = "circle",      "AUTO_ADJ_LB"     = "circle",
+  "ES27_ADJ"     = "stripe",      "ES27_ADJ_LB"     = "stripe",
+  "ES64_ADJ"     = "crosshatch",  "ES64_ADJ_LB"     = "crosshatch",
+  # AVG
+  "AUTO_AVG"     = "circle",      "AUTO_AVG_LB"     = "circle",
+  "ES27_AVG"     = "stripe",      "ES27_AVG_LB"     = "stripe",
+  "ES64_AVG"     = "crosshatch",  "ES64_AVG_LB"     = "crosshatch"
+)
+
+#custom_patterns <- custom_patterns[custom_patterns != "circle"]
+
+plot1 <- ggplot(W1, aes(x = Best_Result,
+                        pattern = Best_Result,
+                        fill    = Best_Result)) +
+  geom_bar_pattern(
+    colour        = NA,            # no border
+    pattern_fill  = "black",
+    pattern_size  = 0.2,
+    pattern_angle = 45             # set globally, not inside aes()
+  ) +
+  scale_fill_manual(values   = custom_colors) +
+  scale_pattern_manual(values = custom_patterns) +
+  labs(title = "Best model on the 48 U.S. states (W1)",
+       x = NULL, y = "Number of states") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  facet_wrap(~ Model_Type, scales = "free_x") +
+  guides(pattern_angle = "none")  # this line is optional now but safe
+
+plot1
+
+plot2 <- ggplot(W2, aes(x = Best_Result,
+                        pattern = Best_Result,
+                        fill    = Best_Result)) +
+  geom_bar_pattern(
+    colour        = NA,            # no border
+    pattern_fill  = "black",
+    pattern_size  = 0.2,
+    pattern_angle = 45             # set globally, not inside aes()
+  ) +
+  scale_fill_manual(values   = custom_colors) +
+  scale_pattern_manual(values = custom_patterns) +
+  labs(title = "Best model on the 48 U.S. states (W2)",
+       x = NULL, y = "Number of states") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  facet_wrap(~ Model_Type, scales = "free_x") +
+  guides(pattern_angle = "none")  # this line is optional now but safe
+
+plot2
+
+plot3 <- ggplot(W3, aes(x = Best_Result,
+                        pattern = Best_Result,
+                        fill    = Best_Result)) +
+  geom_bar_pattern(
+    colour        = NA,            # no border
+    pattern_fill  = "black",
+    pattern_size  = 0.2,
+    pattern_angle = 45             # set globally, not inside aes()
+  ) +
+  scale_fill_manual(values   = custom_colors) +
+  scale_pattern_manual(values = custom_patterns) +
+  labs(title = "Best model on the 48 U.S. states (W3)",
+       x = NULL, y = "Number of states") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  facet_wrap(~ Model_Type, scales = "free_x") +
+  guides(pattern_angle = "none")  # this line is optional now but safe
+
+plot3
+
+plot4 <- ggplot(W4, aes(x = Best_Result,
+                        pattern = Best_Result,
+                        fill    = Best_Result)) +
+  geom_bar_pattern(
+    colour        = NA,            # no border
+    pattern_fill  = "black",
+    pattern_size  = 0.2,
+    pattern_angle = 45             # set globally, not inside aes()
+  ) +
+  scale_fill_manual(values   = custom_colors) +
+  scale_pattern_manual(values = custom_patterns) +
+  labs(title = "Best model on the 48 U.S. states (W4)",
+       x = NULL, y = "Number of states") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  facet_wrap(~ Model_Type, scales = "free_x") +
+  guides(pattern_angle = "none")  # this line is optional now but safe
+
+plot4
+
+```
+
+```{r}
+
+library(ggpattern)
+library(ggplot2)
+library(dplyr)
+
+# Assign week labels
+W1$Week_Ahead <- "W1"
+W2$Week_Ahead <- "W2"
+W3$Week_Ahead <- "W3"
+W4$Week_Ahead <- "W4"
+
+# Combine all weeks
+all_weeks <- bind_rows(W1, W2, W3, W4)
+
+# Create the patterned plot
+combined_plot <- ggplot(all_weeks, aes(x = Best_Result, 
+                                       fill = Best_Result,
+                                       pattern = Best_Result)) +
+  geom_bar_pattern(
+    pattern_fill  = "black",
+    pattern_size  = 0.1,
+    pattern_angle = 45,  # set globally
+    colour        = NA
+  ) +
+  scale_fill_manual(values = custom_colors) +
+  scale_pattern_manual(values = custom_patterns) +  # define this before
+  labs(title = "Best model based on the lowest mean WIS",
+       subtitle = "Results for the 48 states on the contiguous U.S. by target week",
+       x = "", y = "Number of states",
+       fill = "Best model",
+       pattern = "Best model") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        plot.title = element_text(size = 16, face = "bold"),legend.position = "none") +
+  facet_grid(Week_Ahead ~ Model_Type, scales = "free_x") +
+  guides(pattern_angle = "none")  # just in case
+
+# Save and display the plot
+ggsave("Fig4.jpg", combined_plot, width = 7, height = 6)
+print(combined_plot)
+```
+
+```{r}
 # Create a custom color mapping for each variable
@@
 # Save and display the plot
-ggsave("Fig4.jpg", combined_plot, width =8, height =5)
+#ggsave("Fig4.jpg", combined_plot, width =8, height =5)
 print(combined_plot)

### 2025-06-03 manuscript_results.Rmd
update

@@
 library("ggpattern")
-
+library("cowplot")
 ```
@@
 ```{r}
-
 custom_colors <- c(
@@
 plot4
-
-```
-
-```{r}
-
-library(ggpattern)
-library(ggplot2)
-library(dplyr)
+```
+
+```{r}
 
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        plot.title = element_text(size = 16, face = "bold"),legend.position = "none") +
-  facet_grid(Week_Ahead ~ Model_Type, scales = "free_x") +
-  guides(pattern_angle = "none")  # just in case
+        plot.title = element_text(size = 16, face = "bold"),legend.position = "left") +
+  facet_grid(Week_Ahead ~ Model_Type, scales = "free_x") #+
+ # guides(pattern_angle = "none")  # just in case
 
 # Save and display the plot
-ggsave("Fig4.jpg", combined_plot, width = 7, height = 6)
+#ggsave("Fig4.jpg", combined_plot, width = 7, height = 6)
 print(combined_plot)
+
+```
+
+```{r}
+
+# Define the custom patterns
+custom_patterns <- c("AUTO" = "circle", "ES27" = "stripe", "ES64" = "crosshatch")
+custom_fills <- c("AUTO" = "white", "ES27" = "white", "ES64" = "white")  # or any colors you like
+
+# Create the custom legend plot
+legend_data <- data.frame(
+  Model_Type = c("AUTO", "ES27", "ES64", "ES27", "ES64"),
+  Model_Family = c("AR","TMP", "EPI", "ADJ","AVG"),
+  x = 1:5,
+  y = 1
+)
+
+# Define patterns for model types
+custom_patterns <- c("AUTO" = "circle", "ES27" = "stripe", "ES64" = "crosshatch")
+
+# Define fill colors for approaches
+custom_fills <- c("AR" = colors_ar[1], "TMP"=colors_tmp[1], "EPI" = colors_epi[1], "ADJ" = colors_adj[1], "AVG" = colors_avg[1])
+
+# Custom legend plot
+custom_legend <- ggplot(legend_data, aes(x = factor(x), y = y, 
+                                         pattern = Model_Type, fill = Model_Family)) +
+  geom_bar_pattern(stat = "identity",
+                   pattern_fill = "black",
+                   pattern_density = 0.3,
+                   pattern_spacing = 0.05,
+                   pattern_size = 0.15,
+                   colour = "black",
+                   width = 0.7) +
+  scale_fill_manual(values = custom_fills, name = "Model Family") +
+  scale_pattern_manual(values = custom_patterns, name = "Model Type") +
+  theme_void() +
+  theme(
+    legend.position = "right",
+    legend.title = element_text(size = 12, face = "bold"),
+    legend.text = element_text(size = 10)
+  ) +
+  guides(
+    fill = guide_legend(
+      override.aes = list(pattern = "none", pattern_fill = NA, colour = NA)
+    ),
+    pattern = guide_legend(
+      override.aes = list(fill = "white", pattern_fill = "black")
+    )
+  )
+custom_legend
+```
+
+```{r}
+
+# Remove legends from combined_plot
+combined_plot_clean <- combined_plot + theme(legend.position = "none")
+
+# Extract only the legend from custom_legend
+
+legend_only <- cowplot::get_legend(custom_legend)
+
+# Combine your main plot with the custom legend
+final_plot <- combined_plot_clean + 
+  patchwork::plot_spacer() + 
+  patchwork::wrap_elements(legend_only) + 
+  plot_layout(widths = c(4, 0.05, 0.4))  # Adjust widths as needed
+
+# Display the result
+final_plot
+
+ggsave("Fig4.jpg", final_plot, width = 10, height = 6)
 ```
@@
   return(data.frame(
-    AUTO_AR = (data$AUTO_AR / data$AUTO_AR),
+    AUTO_AR = (((data$AUTO_AR / data$AUTO_AR)-1) * 100),
 
@@
 p_values_and_impr
+```
+
+```{r}
+
+new_long_data <- summary_impr %>%
+  mutate(
+    LogBack = grepl("_LB$", Model),
+    Family = sub("_.*", "", Model),  # take everything before the first underscore
+    Method = sub("^.*?_(AR|ADJ|EPI|AVG|TMP)(_LB)?$", "\\1", Model)
+  )
+
+new_long_data <- new_long_data %>%
+  mutate(Model = ifelse(Model == "AUTO_AR", "AUTO_AAR", Model))
+
+model <- lm(m ~ Model, data = new_long_data)
+summary(model)
+
+model
+new_long_data$Model
+# Get tidy summary of coefficients
+tidy_model <- broom::tidy(model)
+
+# Filter out intercept for clarity if you want
+tidy_model_no_intercept <- tidy_model %>% filter(term != "(Intercept)")
+
+# Plot coefficients with 95% confidence intervals
+tidy_model_no_intercept <- tidy_model_no_intercept %>%
+  mutate(term = forcats::fct_reorder(term, estimate, .desc = TRUE))
+
+# Plot
+ggplot(tidy_model_no_intercept, aes(x = term, y = estimate)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = estimate - std.error * 1.96,
+                    ymax = estimate + std.error * 1.96),
+                width = 0.2) +
+  coord_flip() +  # flip for better readability
+  labs(
+    title = "Effect Size of Model Types on outcome",
+    subtitle =  "Estimates relative to baseline (intercept)",
+    y = "Estimate",
+    x = ""
+  ) +
+  theme_minimal()
+
+```
+
+```{r}
+
+new_long_data <- summary_impr %>%
+  mutate(
+    LogBack = grepl("_LB$", Model),
+    Type = sub("_.*", "", Model),  # take everything before the first underscore
+    Family = sub("^.*?_(AR|ADJ|EPI|AVG|TMP)(_LB)?$", "\\1", Model)
+  )
+
+new_long_data <- new_long_data %>%
+  mutate(Family = ifelse(Family == "AR", "AAR", Family))
+
+model <- lm(m ~ Type + Family + LogBack, data = new_long_data)
+summary(model)
+
+model
+new_long_data$Model
+# Get tidy summary of coefficients
+tidy_model <- broom::tidy(model)
+
+# Filter out intercept for clarity if you want
+tidy_model_no_intercept <- tidy_model %>% filter(term != "(Intercept)")
+
+# Plot coefficients with 95% confidence intervals
+tidy_model_no_intercept <- tidy_model_no_intercept %>%
+  mutate(term = forcats::fct_reorder(term, estimate, .desc = TRUE))
+
+# Plot
+effect_size<- ggplot(tidy_model_no_intercept, aes(x = term, y = estimate)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = estimate - std.error * 1.96,
+                    ymax = estimate + std.error * 1.96),
+                width = 0.2) +
+  coord_flip() +  # flip for better readability
+  labs(
+    title = "Effect size of model type, family and transformation on WIS difference",
+    subtitle =  "WIS Difference relative to the AUTO type, ARIMA family and LogBack False (intercept)",
+    y = "WIS Difference",
+    x = ""
+  ) +
+  theme_minimal() + theme(plot.title = element_text(size = 16, face = "bold"),
+                  plot.subtitle = element_text(size = 14))
+
+effect_size
+
+ggsave("Fig6.jpg", effect_size, width =10, height =5)
+
 ```
@@
 # Save to file
-ggsave("Fig6.jpg", combined_plot, width =12, height =7)
+ggsave("Fig7.jpg", combined_plot, width =12, height =7)
 
@@
 reg2_scatter
-ggsave(reg2_scatter, height = 6, width = 12, filename="Fig7.jpg")  
+ggsave(reg2_scatter, height = 6, width = 12, filename="Fig8.jpg")  
 #ggsave(reg2, filename="reg2.jpg",height = 8, width = 10)
@@
 # Save to file
-ggsave("Fig8.jpg", combined_plot, width =12, height =7)
+ggsave("Fig9.jpg", combined_plot, width =12, height =7)
 ```
@@
   labs(title = "Regression models R²",
-       subtitle = "Mean WIS differences (%) as dependent variable",
+       subtitle = "Mean WIS differences as dependent variable",
        x = "Target Week",
@@
 reg2_scatter
-ggsave(reg2_scatter, height = 6, width = 12, filename="Fig9.jpg")  
+ggsave(reg2_scatter, height = 6, width = 12, filename="Fig10.jpg")  
 #ggsave(reg2, filename="Fig9.jpg",height = 6, width = 8)
@@
   labs(
-    title = "Weekly mean WIS differences across all states",
+    title = "A) Weekly mean WIS differences across all states",
     subtitle = "AUTO_AVG_LB compared to baseline",
@@
   values = c(
-    "Week 1" = "#00BFC4",  # Bright cyan
-    "Week 2" = "#F8766D",  # Bright coral red
-    "Week 3" = "#7CAE00",  # Lime green
-    "Week 4" = "#C77CFF"   # Bright purple
+    "Week 1" = "#4575B4",  # Blue
+    "Week 2" = "#91BFDB",  # Light blue
+    "Week 3" = "#E89C9C",  # Light red
+    "Week 4" = "#D73027"   # Red
   )
@@
   geom_bar(stat = "identity",fill = "steelblue") +
-  labs(title = "Total influenza hospitalizations on the 48 states",
+  labs(title = "B) Total influenza hospitalizations on the 48 states",
        x = "",
@@
 # Combine the plots vertically
-combined_plot <- hospital_plot / model_improv
+combined_plot <-  model_improv /hospital_plot
 
@@
 # Save the combined plot
-ggsave("Fig10.jpg", combined_plot, width = 7, height = 7, dpi = 600)
-
-```
+ggsave("Fig11.jpg", combined_plot, width = 7, height = 7, dpi = 600)
+
+```

### 2025-06-09 ERA5 data/1990-2020/ERA5_1990_2020.Rmd
update

@@
 Downloading temperature and humidity data from the cloud.
+You may run this code in python.
 
 ```{python}
-# Importando a biblioteca do xarray e outras auxiliares
+# Libraries
 import xarray as xr

### 2025-06-09 ERA5 data/2010-2024/ERA5_2010_2024.Rmd
update

@@
 Downloading temperature and humidity data from the cloud.
+You may run this code in python.
 
 ```{python}
-# Importando a biblioteca do xarray e outras auxiliares
+# libraries
 import xarray as xr

### 2025-06-12 ERA5 data/1990-2020/ERA5_1990_2020.Rmd
update

@@
 ```
+
+After downloading the file you can use the code below to get mean weekly temperature for each state.
+
 Plotting temperature and humidity NetCDFs.

### 2025-06-12 NHSH_hospitalizations_seed/ES_ADJACENT/ES_ADJACENT_48.R
update

@@ -1,5 +1,5 @@
 #' ES_ADJACENT 
-#' This function utilizes ensembles and single automatic ARIMAX models which have mean cases by adjacent states as exogenous variables.
+#' This function use ensembles and single automatic ARIMAX models which have mean cases by adjacent states as exogenous variables.
 #' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
 #' It return some metrics that evaluate the performance of the models:
 #' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  

### 2025-06-17 manuscript_results.Rmd
cleaning code up to line 1600

@@
 library("dplyr")
-library("scales")  # For `breaks_extended`
+library("scales") 
 library("ggplot2")
@@
 library("cowplot")
+library("lme4")
+library("ez")
 ```
@@
 ```{r}
-library(dplyr)
-library(ggplot2)
 
@@
 
-  # Create the box plot with additional facet for categories
+# Create the box plot with facet for categories
 plt1<-  ggplot(combined_data, aes(x = week, y = value, fill = Models)) +
@@
     ) +
-  scale_y_continuous(limits = c(0, 650)) +  # Set y-axis range
+  scale_y_continuous(limits = c(0, 650)) +  
     theme_minimal()+
@@
 ```{r}
-i  <- 57
+i  <- 57 # forecasted week
 ii <- 8 # state number
@@
 list_of_states <- split(my_data, my_data$state_name)
-
-# Convert to Date if not already
-list_of_states[[ii]]$target_end_date <- as.Date(list_of_states[[ii]]$target_end_date)
 
@@
 ```{r}
-# 2) build your forecast_df for i=55, ii=8
-
-i  <- 57 # target_end_date
+i  <- 57 # forecast week
 ii <- 8 # state number
@@
 conceptual_figure = ggplot() +
-  # 3a) two-year observed history as a black line (mapped to color for legend)
+  # two-year fitting data
   geom_point(data = subset_data, 
@@
             size = 2.5) +
-
-  # 3b) forecast ribbon (not in legend)
+  # grey ribbon 
   geom_ribbon(data = forecast_df, 
@@
               fill = "gray", alpha = 0.4) +
-
-  # 3c) forecast mean (dashed blue line + hollow points)
+  # mean forecast 
   geom_point(data = forecast_df, 
@@
              shape = 21, fill = "white", size = 2) +
-
-  # 3d) actual post-cutoff cases (mapped to color for legend)
+  # actual hospitalizations 
   geom_point(data = forecast_df, 
@@
              shape = 16, size = 2.5) +
-
   # Manually define legend entries
@@
 ```
-Calculate mean weighted interval score for all forecast weeks on each model.
+Evaluating other forecasts
+
+```{r}
+plot_forecast <- function(i, ii,
+                          list_of_states,
+                          AUTO_AVERAGE_WEEK1_list,
+                          AUTO_AVERAGE_WEEK2_list,
+                          AUTO_AVERAGE_WEEK3_list,
+                          AUTO_AVERAGE_WEEK4_list) {
+
+  # Define date range for 2-year history
+  end_date <- as.Date(AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i] - 7)
+  start_date <- end_date - (7 * 103)
+
+  # Subset observed data
+  subset_data <- subset(list_of_states[[ii]],
+                        target_end_date >= start_date & target_end_date <= end_date)
+
+  # Build forecast dataframe
+  forecast_df <- data.frame(
+    Dates = c(
+      AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i],
+      AUTO_AVERAGE_WEEK2_list[[ii]]$target_end_date[i],
+      AUTO_AVERAGE_WEEK3_list[[ii]]$target_end_date[i],
+      AUTO_AVERAGE_WEEK4_list[[ii]]$target_end_date[i]
+    ),
+    Hosp = c(
+      AUTO_AVERAGE_WEEK1_list[[ii]]$cases[i],
+      AUTO_AVERAGE_WEEK2_list[[ii]]$cases[i],
+      AUTO_AVERAGE_WEEK3_list[[ii]]$cases[i],
+      AUTO_AVERAGE_WEEK4_list[[ii]]$cases[i]
+    ),
+    Value = c(
+      AUTO_AVERAGE_WEEK1_list[[ii]]$forecasts[i],
+      AUTO_AVERAGE_WEEK2_list[[ii]]$forecasts[i],
+      AUTO_AVERAGE_WEEK3_list[[ii]]$forecasts[i],
+      AUTO_AVERAGE_WEEK4_list[[ii]]$forecasts[i]
+    ),
+    Lower = c(
+      AUTO_AVERAGE_WEEK1_list[[ii]][["0.025"]][i],
+      AUTO_AVERAGE_WEEK2_list[[ii]][["0.025"]][i],
+      AUTO_AVERAGE_WEEK3_list[[ii]][["0.025"]][i],
+      AUTO_AVERAGE_WEEK4_list[[ii]][["0.025"]][i]
+    ),
+    Upper = c(
+      AUTO_AVERAGE_WEEK1_list[[ii]][["0.975"]][i],
+      AUTO_AVERAGE_WEEK2_list[[ii]][["0.975"]][i],
+      AUTO_AVERAGE_WEEK3_list[[ii]][["0.975"]][i],
+      AUTO_AVERAGE_WEEK4_list[[ii]][["0.975"]][i]
+    )
+  )
+
+  # Build the plot
+  ggplot() +
+    geom_point(data = subset_data,
+               aes(x = target_end_date, y = cases, color = "Fitting data"),
+               size = 2.5) +
+    geom_ribbon(data = forecast_df,
+                aes(x = Dates, ymin = Lower, ymax = Upper),
+                fill = "gray", alpha = 0.4) +
+    geom_point(data = forecast_df,
+               aes(x = Dates, y = Value, color = "Mean forecast"),
+               size = 3) +
+    geom_point(data = forecast_df,
+               aes(x = Dates, y = Value, color = "Mean forecast"),
+               shape = 21, fill = "white", size = 2) +
+    geom_point(data = forecast_df,
+               aes(x = Dates, y = Hosp, color = "Hospitalizations"),
+               shape = 16, size = 2.5) +
+    scale_color_manual(name = "Legend",
+                       values = c("Fitting data" = "red",
+                                  "Mean forecast" = "cyan3",
+                                  "Hospitalizations" = "black")) +
+    scale_x_date(date_breaks = "1 month", date_labels = "%b %y") +
+    labs(title = paste("State:", names(list_of_states)[ii], "- Week Index:", i),
+         subtitle = "Gray ribbon represents 95% confidence interval",
+         x = "", y = "Hospitalizations") +
+    theme_minimal() +
+    theme(plot.title = element_text(size = 18, face = "bold"),
+          plot.subtitle = element_text(size = 14),
+          axis.text.x = element_text(angle = 45, hjust = 1))
+}
+
+for (ii in 1:48) {
+  for (i in 60:70) {
+    # Safely check if index exists
+    if (i <= nrow(AUTO_AVERAGE_WEEK1_list[[ii]])) {
+      p <- plot_forecast(i, ii,
+                         list_of_states,
+                         AUTO_AVERAGE_WEEK1_list,
+                         AUTO_AVERAGE_WEEK2_list,
+                         AUTO_AVERAGE_WEEK3_list,
+                         AUTO_AVERAGE_WEEK4_list)
+      # Show the plot
+      print(p)
+    }
+  }
+}
+
+```
+
+Calculate mean weighted interval score for influenza seasons on each model and each state.
 
@@
       ES64_AVG_LB = mean(ES64_AVG_LB, na.rm = TRUE),
-
-    )
-}
-
-# Now you can use the function with any dataframe
+    )
+}
+
+# Now we can use the function with any dataframe
 W1_Lg <- calculate_mean_wis(filtered_df_W1)
@@
 ```{r}
-# Combine the data into one data frame
+# Combine the data into one dataframe
 combined_data <- data.frame(
@@
             W4_Lg$AUTO_AVG_LB, W4_Lg$ES27_AVG_LB, W4_Lg$ES64_AVG_LB)
-
-)
-
-# Create a new column for categories
+)
+
+# Create a new column for model families
 combined_data <- combined_data %>%
@@
 
-# Define color ramps for each category
+# Define color ramps for each model family
 colors_ar <- c("#a6cee3","#226e83", "#08306b")
@@
 
-# Create a custom color mapping for each variable
+# Use a custom color on each model
 custom_colors <- c(
@@
 ```
-
-```{r}
-
-# Combine in a 2x2 grid
-combined_plot <- (plt1 | plt2) + 
+Figure 3
+
+```{r}
+# Combine plot in a 2x2 grid
+figure3 <- (plt1 | plt2) + 
   plot_annotation(
@@
 # Print
-print(combined_plot)
+print(figure3)
 
 # Save to file
-ggsave("Fig3.jpg", combined_plot, width =9, height =10)
-```
-
-COMBINING THE RESULTS WITH LOG AND NO LOG TRASNFORMATION
+ggsave("Fig3.jpg", figure3, width =9, height =10)
+```
+
+COMBINING THE RESULTS WITH LOG-BACK AND NO LOG-BACK TRASNFORMATION
 
@@
 
-Here I include a column with the best model result based on the lowest mean(WIS) of each state.
+Here I include a column with the best model based on the lowest mean(WIS) of each state.
 
@@
 
-# Extract the columns of interest 
-cols <- colnames(W1)[-1]
-# Initialize a vector to store results
-W1$Best_Result <- character(nrow(W1))
-
-# Give me the model with lower WIS value in the best result column
+cols <- colnames(W1)[-1] # get states names
+W1$Best_Result <- character(nrow(W1)) # create an empty column for the best models
+
+# Give me the model with lower WIS value 
 for (i in 1:nrow(W1)) {
-  # Find the column name with the minimum value for each row
+  # Find the model with the minimum value on each row
+  # based on the column name
+  # save it in the best result
   W1$Best_Result[i] <- cols[which.min(W1[i, cols])]
@@
 W1$Best_Result <- fct_infreq(W1$Best_Result)
-# Create a new column to classify the model type
+# Create a new column to indicate if model has a log-back
+# transformation or not
 W1$Model_Type <- ifelse(grepl("_LB$", W1$Best_Result), "With log-back transformation", "Without log-back transformation")
@@
 # Extract the columns of interest 
-cols <- colnames(W2)[-1]
-# Initialize a vector to store results
-W2$Best_Result <- character(nrow(W2))
-# Give me the model with lower WIS value in the best result column
+cols <- colnames(W2)[-1]  # get states names
+W2$Best_Result <- character(nrow(W2)) # create an empty column for the best models
+# Give me the model with lower WIS value 
 for (i in 1:nrow(W2)) {
-  # Find the column name with the minimum value for each row
+  # Find the model with the minimum value on each row
+  # based on the column name
+  # save it in the best result
   W2$Best_Result[i] <- cols[which.min(W2[i, cols])]
@@
 W2$Best_Result <- fct_infreq(W2$Best_Result)
-# Create a new column to classify the model type
+# Create a new column to indicate if model has a log-back
+# transformation or not
 W2$Model_Type <- ifelse(grepl("_LB$", W2$Best_Result),  "With log-back transformation","Without log-back transformation")
@@
 # Extract the columns of interest 
-cols <- colnames(W3)[-1]
-# Initialize a vector to store results
-W3$Best_Result <- character(nrow(W3))
-
-# Give me the model with lower WIS value in the best result column
+cols <- colnames(W3)[-1] # get states names
+W3$Best_Result <- character(nrow(W3))# create an empty column for the best models
+# Give me the model with lower WIS value
 for (i in 1:nrow(W3)) {
-  # Find the column name with the minimum value for each row
+  # Find the model with the minimum value on each row
+  # based on the column name
+  # save it in the best result
   W3$Best_Result[i] <- cols[which.min(W3[i, cols])]
@@
 W3$Best_Result <- fct_infreq(W3$Best_Result)
-# Create a new column to classify the model type
+# Create a new column to indicate if model has a log-back
+# transformation or not
 W3$Model_Type <- ifelse(grepl("_LB$", W3$Best_Result),  "With log-back transformation","Without log-back transformation")
@@
 # Extract the columns of interest 
-cols <- colnames(W4)[-1]
-# Initialize a vector to store results
-W4$Best_Result <- character(nrow(W4))
-# Give me the model with lower WIS value in the best result column
+cols <- colnames(W4)[-1] # get states names
+W4$Best_Result <- character(nrow(W4)) # create an empty column for the best models
+
+# Give me the model with lower WIS value
 for (i in 1:nrow(W4)) {
-  # Find the column name with the minimum value for each row
+  # Find the model with the minimum value on each row
+  # based on the column name
+  # save it in the best result
   W4$Best_Result[i] <- cols[which.min(W4[i, cols])]
@@
 W4$Best_Result <- fct_infreq(W4$Best_Result)
-# Create a new column to classify the model type
+# Create a new column to indicate if model has a log-back
+# transformation or not
 W4$Model_Type <- ifelse(grepl("_LB$", W4$Best_Result),  "With log-back transformation","Without log-back transformation")
@@
 ```{r}
+
+# define colors for the plot
 custom_colors <- c(
@@
 
-# define a manual pattern scale
+# define patterns for the plot
 custom_patterns <- c(
@@
 
-#custom_patterns <- custom_patterns[custom_patterns != "circle"]
-
 plot1 <- ggplot(W1, aes(x = Best_Result,
@@
   geom_bar_pattern(
-    colour        = NA,            # no border
+    colour        = NA,         
     pattern_fill  = "black",
     pattern_size  = 0.2,
-    pattern_angle = 45             # set globally, not inside aes()
+    pattern_angle = 45          
   ) +
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  facet_wrap(~ Model_Type, scales = "free_x") +
-  guides(pattern_angle = "none")  # this line is optional now but safe
+  facet_wrap(~ Model_Type, scales = "free_x")
 
@@
   geom_bar_pattern(
-    colour        = NA,            # no border
+    colour        = NA,           
     pattern_fill  = "black",
     pattern_size  = 0.2,
-    pattern_angle = 45             # set globally, not inside aes()
+    pattern_angle = 45      
   ) +
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  facet_wrap(~ Model_Type, scales = "free_x") +
-  guides(pattern_angle = "none")  # this line is optional now but safe
-
+  facet_wrap(~ Model_Type, scales = "free_x") 
 plot3
@@
   geom_bar_pattern(
-    colour        = NA,            # no border
+    colour        = NA,       
     pattern_fill  = "black",
     pattern_size  = 0.2,
-    pattern_angle = 45             # set globally, not inside aes()
+    pattern_angle = 45          
   ) +
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  facet_wrap(~ Model_Type, scales = "free_x") +
-  guides(pattern_angle = "none")  # this line is optional now but safe
-
+  facet_wrap(~ Model_Type, scales = "free_x") 
 plot4
@@
 
-```{r}
-
-# Assign week labels
+Now I will combine all these plots.
+
+```{r}
+# Include columns with target week labels
 W1$Week_Ahead <- "W1"
@@
 
-# Combine all weeks
+# Combine all dataframes
 all_weeks <- bind_rows(W1, W2, W3, W4)
 
-# Create the patterned plot
+# Create combined plot
 combined_plot <- ggplot(all_weeks, aes(x = Best_Result, 
@@
     pattern_size  = 0.1,
-    pattern_angle = 45,  # set globally
+    pattern_angle = 45,  
     colour        = NA
@@
   scale_fill_manual(values = custom_colors) +
-  scale_pattern_manual(values = custom_patterns) +  # define this before
+  scale_pattern_manual(values = custom_patterns) +
   labs(title = "Best model based on the lowest mean WIS",
@@
         plot.title = element_text(size = 16, face = "bold"),legend.position = "left") +
-  facet_grid(Week_Ahead ~ Model_Type, scales = "free_x") #+
- # guides(pattern_angle = "none")  # just in case
+  facet_grid(Week_Ahead ~ Model_Type, scales = "free_x")
 
@@
 ```
+These legends from ggplot don't work well. I will create an legend for a random plot and I will use this legend in the previous plot.
 
@@
 custom_patterns <- c("AUTO" = "circle", "ES27" = "stripe", "ES64" = "crosshatch")
-custom_fills <- c("AUTO" = "white", "ES27" = "white", "ES64" = "white")  # or any colors you like
-
-# Create the custom legend plot
+custom_fills <- c("AUTO" = "white", "ES27" = "white", "ES64" = "white")
+
+# Create a example of plot
 legend_data <- data.frame(
@@
 
-# Define fill colors for approaches
+# Define colors
 custom_fills <- c("AR" = colors_ar[1], "TMP"=colors_tmp[1], "EPI" = colors_epi[1], "ADJ" = colors_adj[1], "AVG" = colors_avg[1])
 
-# Custom legend plot
+# Visualizing an example
 custom_legend <- ggplot(legend_data, aes(x = factor(x), y = y, 
@@
 ```
-
-```{r}
-
-# Remove legends from combined_plot
+Here is the final plot with custom legend.
+
+```{r}
+
+# Remove legends from initial plot
 combined_plot_clean <- combined_plot + theme(legend.position = "none")
 
-# Extract only the legend from custom_legend
-
+# Extract only the custom legend
 legend_only <- cowplot::get_legend(custom_legend)
@@
 # Combine your main plot with the custom legend
-final_plot <- combined_plot_clean + 
+figure4 <- combined_plot_clean + 
   patchwork::plot_spacer() + 
@@
 # Display the result
-final_plot
-
-ggsave("Fig4.jpg", final_plot, width = 10, height = 6)
-```
+figure4
+
+ggsave("Fig4.jpg", figure4, width = 10, height = 6)
+
+```
+
+ERASE
 
@@
 
-Let's plot all results
-
-```{r}
-library(ggplot2)
-library(dplyr)
-
+ERASE
+
+```{r}
 # Add a column to indicate the week ahead
@@
 
- Now we will utilize a Wilcox test with Holm p-value ajustment for repeated comparassion to evaluate the differences in model performances.
+Now we use a Wilcox test with Holm p-value adjustment for repeated comparisons to evaluate the differences in model performances.
 
@@
 ```{r}
-library(lme4)
-library(ez)
-
-# Reshape the data into long format
-long_data <- W1 %>%
-  pivot_longer(cols = c(-STATE,-Best_Result,-Model_Type,-Week_Ahead) , names_to = "Type", values_to = "Value")#%>%
-
-wide_data <- long_data %>%
-  pivot_wider(names_from = Type, values_from = Value)
-
-# Get all model types except "AUTO_AAR"
-model_types <- setdiff(names(wide_data), c("STATE", "AUTO_AR","Best_Result","Model_Type","Week_Ahead"))
-
-# Perform Wilcoxon test for each model type against AUTO_AAR
-wilcox_results <- map_df(model_types, function(model) {
-  test <- wilcox.test(wide_data[[model]], wide_data$AUTO_AR, paired = TRUE)
+# Get all models names except baseline
+model_types <- setdiff(names(W1), c("STATE", "AUTO_AR","Best_Result","Model_Type","Week_Ahead"))
+
+# Perform Wilcoxon test for each model type against baseline
+wilcox_results1 <- map_df(model_types, function(model) {
+  test <- wilcox.test(W1[[model]], W1$AUTO_AR, paired = TRUE)
   data.frame(Model = model, p_value = test$p.value, W = test$statistic)
@@
 
-# View results
-p_values<-wilcox_results$p_value 
+# p values adjustment
+p_values<-wilcox_results1$p_value 
 p_values<-p.adjust(p_values, method = "holm")
@@
 ```{r}
-# Reshape the data into long format
-long_data <- W2 %>%
-  pivot_longer(cols = c(-STATE,-Best_Result,-Model_Type,-Week_Ahead) , names_to = "Type", values_to = "Value")#%>%
-
-################################
-wide_data <- long_data %>%
-  pivot_wider(names_from = Type, values_from = Value)
-
-# Get all model types except "AUTO_AAR"
-model_types <- setdiff(names(wide_data), c("STATE", "AUTO_AR","Best_Result","Model_Type","Week_Ahead"))
-
-# Perform Wilcoxon test for each model type against AUTO_AAR
-wilcox_results <- map_df(model_types, function(model) {
-  test <- wilcox.test(wide_data[[model]], wide_data$AUTO_AR, paired = TRUE)
+# Get all models names except baseline
+model_types <- setdiff(names(W2), c("STATE", "AUTO_AR","Best_Result","Model_Type","Week_Ahead"))
+
+# Perform Wilcoxon test for each model type against baseline
+wilcox_results2 <- map_df(model_types, function(model) {
+  test <- wilcox.test(W2[[model]], W2$AUTO_AR, paired = TRUE)
   data.frame(Model = model, p_value = test$p.value, W = test$statistic)
@@
 
-# View results
-print(wilcox_results)
-
-p_values<-wilcox_results$p_value 
+# p values adjustment
+p_values<-wilcox_results2$p_value 
 p_values<-p.adjust(p_values, method = "holm")
@@
 p_values_wk2
-
 ```
@@
 ```{r}
-# Reshape the data into long format
-long_data <- W3 %>%
-  pivot_longer(cols = c(-STATE,-Best_Result,-Model_Type,-Week_Ahead) , names_to = "Type", values_to = "Value")#%>%
-
-######################
-wide_data <- long_data %>%
-  pivot_wider(names_from = Type, values_from = Value)
-
-# Get all model types except "AUTO_AAR"
-model_types <- setdiff(names(wide_data), c("STATE", "AUTO_AR","Best_Result","Model_Type","Week_Ahead"))
-
-# Perform Wilcoxon test for each model type against AUTO_AAR
-wilcox_results <- map_df(model_types, function(model) {
-  test <- wilcox.test(wide_data[[model]], wide_data$AUTO_AR, paired = TRUE)
+# Get all models names except baseline
+model_types <- setdiff(names(W3), c("STATE", "AUTO_AR","Best_Result","Model_Type","Week_Ahead"))
+
+# Perform Wilcoxon test for each model type against baseline
+wilcox_results3 <- map_df(model_types, function(model) {
+  test <- wilcox.test(W3[[model]], W3$AUTO_AR, paired = TRUE)
   data.frame(Model = model, p_value = test$p.value, W = test$statistic)
@@
 
-# View results
-print(wilcox_results)
-
-p_values<-wilcox_results$p_value 
+# p values adjustment
+p_values<-wilcox_results3$p_value 
 p_values<-p.adjust(p_values, method = "holm")
@@
 ```{r}
-# Reshape the data into long format
-long_data <- W4 %>%
-  pivot_longer(cols = c(-STATE,-Best_Result,-Model_Type,-Week_Ahead) , names_to = "Type", values_to = "Value")#%>%
-#  mutate(Type = recode(Type, "AUTO_AR" = "AUTO_AR"))
-
-######################
-wide_data <- long_data %>%
-  pivot_wider(names_from = Type, values_from = Value)
-
-# Get all model types except "AUTO_AAR"
-model_types <- setdiff(names(wide_data), c("STATE", "AUTO_AR","Best_Result","Model_Type","Week_Ahead"))
-
-# Perform Wilcoxon test for each model type against AUTO_AAR
-wilcox_results <- map_df(model_types, function(model) {
-  test <- wilcox.test(wide_data[[model]], wide_data$AUTO_AR, paired = TRUE)
+# Get all models names except baseline
+model_types <- setdiff(names(W4), c("STATE", "AUTO_AR","Best_Result","Model_Type","Week_Ahead"))
+
+# Perform Wilcoxon test for each model type against baseline
+wilcox_results4 <- map_df(model_types, function(model) {
+  test <- wilcox.test(W4[[model]], W4$AUTO_AR, paired = TRUE)
   data.frame(Model = model, p_value = test$p.value, W = test$statistic)
@@
 # Perform Wilcoxon test for each model type against AUTO_AAR
-p_values<-wilcox_results$p_value 
+p_values<-wilcox_results4$p_value 
 p_values<-p.adjust(p_values, method = "holm")
@@
 Later we sum the results of these comparisons.
-Positive results indicate that there was a general improvement in the mean(WIS) for a given model type among all states.
-Note that larger values on specific states may override a general trend and make the results less informative.
-So I do not strongly advice this analysis.
+Negative results indicate that there was a general improvement in the mean(WIS) for a given model type among all states.
 
@@
     AUTO_AR = (((data$AUTO_AR / data$AUTO_AR)-1) * 100),
-
     ES27_AR = (((data$ES27_AR/data$AUTO_AR)-1) * 100),
@@
 
-# Calculate percentage of improvemtn 
+# Calculate percentage of improvement 
 W1_percentage_of_improvement <- calculate_percentage_of_improvement(W1)
@@
 
-W4_percentage_of_improvement
-```
-
-Now let's plot a map of percentage of WIS improvement for each state.
-
-Mean and sd percentage of improvement compared to auto_arima
-
-```{r}
-# List of column names
-variables <-c(names(W1_percentage_of_improvement))
+head(W1_percentage_of_improvement)
+head(W2_percentage_of_improvement)
+head(W3_percentage_of_improvement)
+head(W4_percentage_of_improvement)
+```
+
+Now let's plot histograms of percentage of WIS improvement for each state including mean and standard deviation of percentage of improvement compared to auto_arima (baseline).
+
+```{r}
+# List of models names
+models_names <-c(names(W1_percentage_of_improvement))
 
@@
 
-# List of datasets
-datasets <- list("1" = W1_percentage_of_improvement, 
+# List of datasets. One dataset for each target week.
+datasets_list <- list("1" = W1_percentage_of_improvement, 
                  "2" = W2_percentage_of_improvement, 
@@
 
-# Loop through datasets and variables
-for (dataset_name in names(datasets)) {
-  dataset <- datasets[[dataset_name]]  # Get the dataset
-
-  for (var in variables) {
-    data <- dataset[[var]]
+# Here I have 2 for loops. One take into account the target week and the other the model.
+# Loop through target weeks
+for (target_week_ in names(datasets_list)) {
+  dataset <- datasets_list[[target_week_]]  # Get the dataset based on target week
+  # Loop through models
+  for (given_model in models_names) {
+    data <- dataset[[given_model]]
     mean_val <- mean(data, na.rm = TRUE)
@@
     # Append results to dataframe
-    summary_impr <- rbind(summary_impr, data.frame(WeekAhead = dataset_name, Model = var, m = mean_val, sd = sd_val))
+    summary_impr <- rbind(summary_impr, data.frame(WeekAhead = target_week_, Model = given_model, m = mean_val, sd = sd_val))
 
     # Generate histogram
-    p <- ggplot(dataset, aes(x = .data[[var]])) +
+    p <- ggplot(dataset, aes(x = .data[[given_model]])) +
       geom_histogram(color = "black", fill = "lightblue", bins = 30, alpha = 0.7) +

### 2025-06-17 manuscript_results.Rmd
code review line 1690

@@
 
-Now I will combine all these plots.
+Now I will combine all these plots 2x2.
 
@@
       geom_vline(aes(xintercept = mean_val + sd_val), color = "blue", linetype = "dotted", linewidth = 1) +
-      labs(title = paste("Histogram of", var, "in", dataset_name),
-           x = var, 
+      labs(title = paste("Histogram of", given_model, "target week", target_week_),
+           x = given_model, 
            y = "Frequency") +

### 2025-06-18 manuscript_results.Rmd
code review line 1800

@@
 library("sf")
-library("dplyr")
 library("scales") 
@@
 
-Datasets with differences and Wilcoxon Holm adjusted p-values
+Datasets with mean differences, standard deviations and Wilcoxon Holm adjusted p-values
 
@@
 ```
-
-```{r}
-
-new_long_data <- summary_impr %>%
-  mutate(
-    LogBack = grepl("_LB$", Model),
-    Family = sub("_.*", "", Model),  # take everything before the first underscore
-    Method = sub("^.*?_(AR|ADJ|EPI|AVG|TMP)(_LB)?$", "\\1", Model)
-  )
-
-new_long_data <- new_long_data %>%
-  mutate(Model = ifelse(Model == "AUTO_AR", "AUTO_AAR", Model))
-
-model <- lm(m ~ Model, data = new_long_data)
-summary(model)
-
-model
-new_long_data$Model
+ERASE
+
+```{r}
+
+#new_long_data <- summary_impr %>%
+#  mutate(
+#    LogBack = grepl("_LB$", Model),
+#    Family = sub("_.*", "", Model),  # take everything before the first underscore
+#    Method = sub("^.*?_(AR|ADJ|EPI|AVG|TMP)(_LB)?$", "\\1", Model)
+#  )
+
+#new_long_data <- new_long_data %>%
+#  mutate(Model = ifelse(Model == "AUTO_AR", "AUTO_AAR", Model))
+
+#model <- lm(m ~ Model, data = new_long_data)
+#summary(model)
+
+#model
+#new_long_data$Model
 # Get tidy summary of coefficients
-tidy_model <- broom::tidy(model)
+#tidy_model <- broom::tidy(model)
 
 # Filter out intercept for clarity if you want
-tidy_model_no_intercept <- tidy_model %>% filter(term != "(Intercept)")
+#tidy_model_no_intercept <- tidy_model %>% filter(term != "(Intercept)")
 
 # Plot coefficients with 95% confidence intervals
-tidy_model_no_intercept <- tidy_model_no_intercept %>%
-  mutate(term = forcats::fct_reorder(term, estimate, .desc = TRUE))
+#tidy_model_no_intercept <- tidy_model_no_intercept %>%
+#  mutate(term = forcats::fct_reorder(term, estimate, .desc = TRUE))
 
 # Plot
-ggplot(tidy_model_no_intercept, aes(x = term, y = estimate)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = estimate - std.error * 1.96,
-                    ymax = estimate + std.error * 1.96),
-                width = 0.2) +
-  coord_flip() +  # flip for better readability
-  labs(
-    title = "Effect Size of Model Types on outcome",
-    subtitle =  "Estimates relative to baseline (intercept)",
-    y = "Estimate",
-    x = ""
-  ) +
-  theme_minimal()
-
-```
-
-```{r}
-
-new_long_data <- summary_impr %>%
+#ggplot(tidy_model_no_intercept, aes(x = term, y = estimate)) +
+#  geom_point() +
+#  geom_errorbar(aes(ymin = estimate - std.error * 1.96,
+#                    ymax = estimate + std.error * 1.96),
+#                width = 0.2) +
+#  coord_flip() +  # flip for better readability
+#  labs(
+#    title = "Effect Size of Model Types on outcome",
+#    subtitle =  "Estimates relative to baseline (intercept)",
+#    y = "Estimate",
+#    x = ""
+#  ) +
+#  theme_minimal()
+
+```
+
+Now let's use a regression model to evaluate effect size of each modelling approach.
+
+```{r}
+
+model_approaches <- summary_impr %>%
   mutate(
@@
 
-new_long_data <- new_long_data %>%
-  mutate(Family = ifelse(Family == "AR", "AAR", Family))
-
-model <- lm(m ~ Type + Family + LogBack, data = new_long_data)
-summary(model)
-
-model
-new_long_data$Model
-# Get tidy summary of coefficients
+model_approaches <- model_approaches %>%
+  mutate(Family = ifelse(Family == "AR", "AAR", Family)) # rename because the intercept is in alphabetic order and I want to use the baseline
+
+# run the regression model
+model <- lm(m ~ Type + Family + LogBack, data = model_approaches)
+
+# extracting models coefficients
 tidy_model <- broom::tidy(model)
 
-# Filter out intercept for clarity if you want
+# filter out intercept
 tidy_model_no_intercept <- tidy_model %>% filter(term != "(Intercept)")
 
-# Plot coefficients with 95% confidence intervals
+# Reordering for plotting
 tidy_model_no_intercept <- tidy_model_no_intercept %>%
@@
                 width = 0.2) +
-  coord_flip() +  # flip for better readability
+  coord_flip() + 
   labs(
@@
 ```{r}
-
-library(ggplot2)
-library(dplyr)
-library(forcats)
 

### 2025-06-18 manuscript_results.Rmd
code review line 1643

@@
 ```
+Plotting conceptual image
 
@@
 
-Now let's use a regression model to evaluate effect size of each modelling approach.
-
-```{r}
-
-model_approaches <- summary_impr %>%
-  mutate(
-    LogBack = grepl("_LB$", Model),
-    Type = sub("_.*", "", Model),  # take everything before the first underscore
-    Family = sub("^.*?_(AR|ADJ|EPI|AVG|TMP)(_LB)?$", "\\1", Model)
-  )
-
-model_approaches <- model_approaches %>%
-  mutate(Family = ifelse(Family == "AR", "AAR", Family)) # rename because the intercept is in alphabetic order and I want to use the baseline
-
-# run the regression model
-model <- lm(m ~ Type + Family + LogBack, data = model_approaches)
-
-# extracting models coefficients
-tidy_model <- broom::tidy(model)
-
-# filter out intercept
-tidy_model_no_intercept <- tidy_model %>% filter(term != "(Intercept)")
-
-# Reordering for plotting
-tidy_model_no_intercept <- tidy_model_no_intercept %>%
-  mutate(term = forcats::fct_reorder(term, estimate, .desc = TRUE))
-
-# Plot
-effect_size<- ggplot(tidy_model_no_intercept, aes(x = term, y = estimate)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = estimate - std.error * 1.96,
-                    ymax = estimate + std.error * 1.96),
-                width = 0.2) +
-  coord_flip() + 
-  labs(
-    title = "Effect size of model type, family and transformation on WIS difference",
-    subtitle =  "WIS Difference relative to the AUTO type, ARIMA family and LogBack False (intercept)",
-    y = "WIS Difference",
-    x = ""
-  ) +
-  theme_minimal() + theme(plot.title = element_text(size = 16, face = "bold"),
-                  plot.subtitle = element_text(size = 14))
-
-effect_size
-
-ggsave("Fig6.jpg", effect_size, width =10, height =5)
-
-```
-
-Mean difference and standard deviation
-
-```{r}
-
+Here we are plotting the mean differences and standard deviation
+
+```{r}
 # Define significance threshold  
@@
 ggsave("Fig5.jpg", plot_m_heatmap, width = 14, height = 7 )
+
+```
+
+Now let's use a regression model to evaluate effect size of each modelling approach.
+
+```{r}
+
+model_approaches <- summary_impr %>%
+  mutate(
+    LogBack = grepl("_LB$", Model),
+    Type = sub("_.*", "", Model),  # take everything before the first underscore
+    Family = sub("^.*?_(AR|ADJ|EPI|AVG|TMP)(_LB)?$", "\\1", Model)
+  )
+
+model_approaches <- model_approaches %>%
+  mutate(Family = ifelse(Family == "AR", "AAR", Family)) # rename because the intercept is in alphabetic order and I want to use the baseline
+
+# run the regression model
+model <- lm(m ~ Type + Family + LogBack, data = model_approaches)
+
+# extracting models coefficients
+tidy_model <- broom::tidy(model)
+
+# filter out intercept
+tidy_model_no_intercept <- tidy_model %>% filter(term != "(Intercept)")
+
+# Reordering for plotting
+tidy_model_no_intercept <- tidy_model_no_intercept %>%
+  mutate(term = forcats::fct_reorder(term, estimate, .desc = TRUE))
+
+# Plot
+effect_size<- ggplot(tidy_model_no_intercept, aes(x = term, y = estimate)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = estimate - std.error * 1.96,
+                    ymax = estimate + std.error * 1.96),
+                width = 0.2) +
+  coord_flip() + 
+  labs(
+    title = "Effect size of model type, family and transformation on WIS difference",
+    subtitle =  "WIS Difference relative to the AUTO type, ARIMA family and LogBack False (intercept)",
+    y = "WIS Difference",
+    x = ""
+  ) +
+  theme_minimal() + theme(plot.title = element_text(size = 16, face = "bold"),
+                  plot.subtitle = element_text(size = 14))
+
+effect_size
+
+ggsave("Fig6.jpg", effect_size, width =10, height =5)
 

### 2025-06-18 manuscript_results.Rmd
update

@@
 
-Datasets with mean differences, standard deviations and Wilcoxon Holm adjusted p-values
+Let create a combined dataset with mean differences, standard deviations and Wilcoxon Holm adjusted p-values
 
@@
 
-# Create a new column "Model_Type" based on whether "NoLg" is in the Model name
+# Create a new column "Model_Type" based on model with and withouth log-back transformation
 p_values_and_impr$Model_Type <- ifelse(grepl("_LB$", p_values_and_impr$Model), "LB", "no_LB")
@@
 ```{r}
-# Define significance threshold  
-alpha <- 0.05
-
-# Get model order based on WeekAhead == 1  
+
+# Get model order based based on lower values on WeekAhead == 1  
 model_order <- p_values_and_impr %>%
@@
   arrange(desc(m)) %>%
-  pull(Model)  # Extract ordered model names
-
-# Prepare data with significance flag and ordering  
+  pull(Model)
+
+# Prepare data with significance level, ordering, and model name
 p_df <- p_values_and_impr %>%
-  mutate(Significance = ifelse(p_values < alpha, "Significant", "Not Significant"),
+  mutate(Significance = ifelse(p_values < 0.05 , "Significant", "Not Significant"),
          Model_Type = ifelse(grepl("_LB$", Model), "With log-back transformation", "Without log-back transformation"),
-         Model = factor(Model, levels = model_order))  # Apply ordered factor
-
-# Heatmap plot with models ordered by WeekAhead == 1  
+         Model = factor(Model, levels = model_order))
+
+# Heatmap plot  
 plot_m_heatmap <- ggplot(p_df, aes(x = factor(WeekAhead), y = Model, fill = m)) +
@@
 )+
-
   labs(title = "Mean differences between models' WIS and the baseline across 48 states",
@@
     plot.subtitle = element_text(size = 14),) +
-  facet_wrap(~ Model_Type, scales = "free_y")  # Separate NoLg and other models
+  facet_wrap(~ Model_Type, scales = "free_y")
 
@@
 ```{r}
-
+# get model approaches and mean improvement
 model_approaches <- summary_impr %>%
@@
     LogBack = grepl("_LB$", Model),
-    Type = sub("_.*", "", Model),  # take everything before the first underscore
+    Type = sub("_.*", "", Model),  # extract the model type
     Family = sub("^.*?_(AR|ADJ|EPI|AVG|TMP)(_LB)?$", "\\1", Model)
@@
 
+# rename "AR" because the intercept is in alphabetic order and I want to use the baseline
 model_approaches <- model_approaches %>%
-  mutate(Family = ifelse(Family == "AR", "AAR", Family)) # rename because the intercept is in alphabetic order and I want to use the baseline
+  mutate(Family = ifelse(Family == "AR", "AAR", Family))
 
@@
 
-Now let's plot a map of percentage of WIS improvement for each state.
+Now let's organize the data to plot some maps with the best models.
 
@@
 # MAP 1 week ahead
-
 W1_map <- data.frame(
@@
 
-# Merge the best model data with the states map
+# Include geometry
 W1_map <- states %>%
@@
 
-################################### MAP 2 weeka ahead
-
+# MAP 2 weeks a ahead
 W2_map <- data.frame(
@@
 
-# Merge the best model data with the states map
+# Include geometry
 W2_map <- states %>%
@@
 
-##################################### MAP 3 weeka ahead
-
+# MAP 3 weeks a ahead
 W3_map <- data.frame(
@@
 
-# Merge the best model data with the states map
+# Include geometry
 W3_map <- states %>%
@@
 
-############################### MAP 4 weeka ahead
-
+# MAP 4 weeks ahead
 W4_map <- data.frame(
@@
 
-# Merge the best model data with the states map
+# Include geometry
 W4_map <- states %>%
@@
 
-Let's plot some maps with the best models in each state and the percentage of improvement compared to the AUTO ARIMA models for the same state.
-Here I define some colors categories that I will use on the next maps.
+Let's plot some maps with mean WIS for each state.
 
@@
   scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
-  ggtitle("1 weeks ahead") +
+  ggtitle("1 week ahead") +
   theme_light() +
@@
 
-Combine in a 2x2 grid
+Combining maps in a 2x2 grid
 

### 2025-06-18 manuscript_results.Rmd
code review up to line 2000
update

@@
   percentage_improvement = W1_percentage_of_improvement$AUTO_AVG_LB,
-  AUTO_AVG=W1$AUTO_AVG_LB
+  AUTO_AVG_LB=W1$AUTO_AVG_LB
 )
@@
   percentage_improvement = W2_percentage_of_improvement$AUTO_AVG_LB,
-  AUTO_AVG=W2$AUTO_AVG_LB
+  AUTO_AVG_LB=W2$AUTO_AVG_LB
 )
@@
   percentage_improvement = W3_percentage_of_improvement$AUTO_AVG_LB,
-  AUTO_AVG=W3$AUTO_AVG_LB
+  AUTO_AVG_LB=W3$AUTO_AVG_LB
 )
@@
   percentage_improvement = W4_percentage_of_improvement$AUTO_AVG_LB,
-  AUTO_AVG=W4$AUTO_AVG_LB
+  AUTO_AVG_LB=W4$AUTO_AVG_LB
 )
@@
 ES_1WEEK <- ggplot(W1_map) +
-  geom_sf(aes(fill = AUTO_AVG)) +  # Fill based on the best model
+  geom_sf(aes(fill = AUTO_AVG_LB)) +  # Fill based on the best model
   scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
@@
     plot.title = element_text(hjust = 0.5)) +
-  geom_sf_text(data = W1_map, aes(label = round(AUTO_AVG,0)),  # Round to whole numbers
+  geom_sf_text(data = W1_map, aes(label = round(AUTO_AVG_LB,0)),  # Round to whole numbers
                size = 3,
@@
 ES_2WEEK <- ggplot(W2_map) +
-  geom_sf(aes(fill = AUTO_AVG)) +  # Fill based on the best model
+  geom_sf(aes(fill = AUTO_AVG_LB)) +  # Fill based on the best model
   scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
@@
     plot.title = element_text(hjust = 0.5)) +
-  geom_sf_text(data = W2_map, aes(label = round(AUTO_AVG,0)),  # Round to whole numbers
+  geom_sf_text(data = W2_map, aes(label = round(AUTO_AVG_LB,0)),  # Round to whole numbers
                size = 3,
@@
 ES_3WEEK <- ggplot(W3_map) +
-  geom_sf(aes(fill = AUTO_AVG)) +  # Fill based on the best model
+  geom_sf(aes(fill = AUTO_AVG_LB)) +  # Fill based on the best model
   scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", midpoint = 100, limits = c(0, 352)) +  
@@
     plot.title = element_text(hjust = 0.5)) +
-  geom_sf_text(data = W3_map, aes(label = round(AUTO_AVG,0)),  # Round to whole numbers
+  geom_sf_text(data = W3_map, aes(label = round(AUTO_AVG_LB,0)),  # Round to whole numbers
                size = 3,
@@
 ES_4WEEK <- ggplot(W4_map) +
-  geom_sf(aes(fill = AUTO_AVG)) +
+  geom_sf(aes(fill = AUTO_AVG_LB)) +
   scale_fill_gradient2(low = "skyblue", mid = "lightyellow2", high = "red", 
@@
     plot.title = element_text(hjust = 0.5)) +
-  geom_sf_text(data = W4_map, aes(label = round(AUTO_AVG,0)),  
+  geom_sf_text(data = W4_map, aes(label = round(AUTO_AVG_LB,0)),  
                size = 3,
@@
 
-We will run the regression for evaluating if the best model log(WIS) which represent its performance is related to given independent variables.
+Now we run regressions models for evaluating if the best model mean WIS in each state is related to given independent variables.
 
@@
 
-Percentage of AUTO_ADJ mean WIS
-
-AUTO_ADJ WIS x RESIDENT POPULATION 2020
+AUTO_ADJ_LB WIS x RESIDENT POPULATION 2020
 
@@
   # Fitting the regression model
-  model <- lm((WIS_pop_data$AUTO_AVG) ~ (WIS_pop_data$Resident_population_2020))
+  model <- lm((WIS_pop_data$AUTO_AVG_LB) ~ (WIS_pop_data$Resident_population_2020))
   # View the model summary
@@
   # Fit linear model
-  model <- lm(AUTO_AVG ~ Resident_population_2020, data = WIS_pop_data)
+  model <- lm(AUTO_AVG_LB ~ Resident_population_2020, data = WIS_pop_data)
   model_summary <- summary(model)
@@
   # Create plot
-  p <- ggplot(WIS_pop_data, aes(x = Resident_population_2020, y = AUTO_AVG)) +
+  p <- ggplot(WIS_pop_data, aes(x = Resident_population_2020, y = AUTO_AVG_LB)) +
     geom_point(color = "steelblue", size = 2) +
@@
   # Fit the regression model
-  model <- lm((WIS_pop_data$AUTO_AVG) ~ (WIS_pop_data$Population_density_2020))
+  model <- lm((WIS_pop_data$AUTO_AVG_LB) ~ (WIS_pop_data$Population_density_2020))
   # View the model summary
@@
   # Fit the regression model
-  model <- lm((WIS_sovi$AUTO_AVG) ~ (WIS_sovi$weighted_mean))
+  model <- lm((WIS_sovi$AUTO_AVG_LB) ~ (WIS_sovi$weighted_mean))
   # View the model summary
@@
   # Fit the regression model  
-  model <- lm((WIS_bric$AUTO_AVG) ~ (WIS_bric$weighted_mean_z_bric.social))
+  model <- lm((WIS_bric$AUTO_AVG_LB) ~ (WIS_bric$weighted_mean_z_bric.social))
   # View the model summary
@@
   # Fit the regression model
-  model <- lm((WIS_bric$AUTO_AVG) ~ (WIS_bric$weighted_mean_z_bric.economic))
+  model <- lm((WIS_bric$AUTO_AVG_LB) ~ (WIS_bric$weighted_mean_z_bric.economic))
   # View the model summary
@@
   # Fit the regression model
-  model <- lm((WIS_bric$AUTO_AVG) ~ (WIS_bric$weighted_mean_z_bric.infrastructure))
+  model <- lm((WIS_bric$AUTO_AVG_LB) ~ (WIS_bric$weighted_mean_z_bric.infrastructure))
   # View the model summary
@@
   # Fit the regression model
-  model <- lm((WIS_bric$AUTO_AVG) ~ (WIS_bric$weighted_mean_z_bric.institutional))
+  model <- lm((WIS_bric$AUTO_AVG_LB) ~ (WIS_bric$weighted_mean_z_bric.institutional))
   # View the model summary
@@
   # Fit the regression model
-  model <- lm((WIS_bric$AUTO_AVG) ~ (WIS_bric$weighted_mean_z_bric.community))
+  model <- lm((WIS_bric$AUTO_AVG_LB) ~ (WIS_bric$weighted_mean_z_bric.community))
   # View the model summary
@@
   # Fit the regression model
-  model <- lm((WIS_bric$AUTO_AVG) ~ (WIS_bric$weighted_mean_z_bric.environment))
+  model <- lm((WIS_bric$AUTO_AVG_LB) ~ (WIS_bric$weighted_mean_z_bric.environment))
   # View the model summary
@@
   # Fit the regression model
-  model <- lm((WIS_bric$AUTO_AVG) ~ (WIS_bric$weighted_mean_z_bric.total))
+  model <- lm((WIS_bric$AUTO_AVG_LB) ~ (WIS_bric$weighted_mean_z_bric.total))
   # View the model summary
@@
   # Fit the regression model
-  model <- lm((WIS_temp$AUTO_AVG) ~ (WIS_temp$mean))
+  model <- lm((WIS_temp$AUTO_AVG_LB) ~ (WIS_temp$mean))
   # View the model summary
@@
   # Fit the regression model
-  model <- lm((WIS_humidity$AUTO_AVG) ~ (WIS_humidity$mean))
+  model <- lm((WIS_humidity$AUTO_AVG_LB) ~ (WIS_humidity$mean))
   # View the model summary
@@
 
-Let's plot some maps with the with the mean (%) of difference between the AUTO_AVG and the AUTO ARIMA models for the same state.
+Let's plot some maps with the with the mean difference (%) between the AUTO_AVG and the AUTO ARIMA models for the same state.
 
@@
 
-AUTO_AVG WIS x RESIDENT POPULATION 2020
+AUTO_AVG_LB WIS x RESIDENT POPULATION 2020
 
@@
 
-AUTO_AVG WIS x POPULATION DENSITY
+AUTO_AVG_LB WIS x POPULATION DENSITY
 
@@
 
-AUTO_AVG WIS x SOVI
+AUTO_AVG_LB WIS x SOVI
 
@@
 
-AUTO_AVG WIS x BRIC SOCIAL
+AUTO_AVG_LB WIS x BRIC SOCIAL
 
@@
 
-AUTO_AVG WIS x BRIC ECONOMIC
+AUTO_AVG_LB WIS x BRIC ECONOMIC
 
@@
 
-AUTO_AVG WIS x BRIC Infrastructure
+AUTO_AVG_LB WIS x BRIC Infrastructure
 
@@
 
-best regression plot
+Plotting best regression model
 
@@
 
-AUTO_AVG WIS x BRIC institutional
+AUTO_AVG_LB WIS x BRIC institutional
 
@@
 
-AUTO_AVG WIS x BRIC community
+AUTO_AVG_LB WIS x BRIC community
 
@@
 
-AUTO_AVG WIS x BRIC environment
+AUTO_AVG_LB WIS x BRIC environment
 
@@
 
-AUTO_AVG WIS x BRIC total
+AUTO_AVG_LB WIS x BRIC total
 
@@
 
-AUTO_AVG WIS x Temperature ERA5 Data
+AUTO_AVG_LB WIS x Temperature ERA5 Data
 
@@
 
-AUTO_AVG WIS x Specific Humidity ERA5 Data
+AUTO_AVG_LB WIS x Specific Humidity ERA5 Data
 
@@
 ```
-
 SUMMARY OF RESULTS
@@
 ```
+
+Computing and plotting mean WIS improvement by epidemiological weeks.
 
@@
 ```
-hospitalizations
+Total hospitalizations
 
@@
 ```
+Combining model improvement and total hospitalizations
 

### 2025-06-20 manuscript_results.Rmd
code review up to line 2046
code review line 2900

@@
 library("ez")
+library("ggpubr")
 ```
@@
                color = "black",
-               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
+               check_overlap = TRUE,fontface = "bold") +  
   labs(
@@
        y = ""
-  )  # Adding a subtitle
+  )
 
@@
 
-Loading datasets for regression analysis
+Loading datasets for regression models analysis
 
@@
 ```{r}
-library(ggpubr)
-
 # regression results 
@@
 
-Let's plot some maps with the with the mean difference (%) between the AUTO_AVG and the AUTO ARIMA models for the same state.
+Let's plot some maps with the with the mean difference (%) between the AUTO_AVG_LB and the AUTO ARIMA models for the same state.
 
@@
 ES_1WEEK <- ggplot(W1_map) +
-  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
+  geom_sf(aes(fill = percentage_improvement)) +  
   scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
@@
     plot.title = element_text(hjust = 0.5)) +
-  geom_sf_text(data = W1_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
+  geom_sf_text(data = W1_map, aes(label = round(percentage_improvement,0)),
+               size = 3,
+               color = "black",
+               check_overlap = TRUE,fontface = "bold") +  
+  labs(
+    fill = "Mean difference", 
+       x = "",
+       y = ""
+  )
+
+x_limits <- c(-125, -67)  # Set the desired longitude range
+y_limits <- c(25, 50)    # Set the desired latitude range
+
+ES_1WEEK<-ES_1WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
+
+#ggsave(ES_1WEEK,height = 4.5, width = 6.5,  filename="Fig5A.jpg")  
+```
+
+2 weeks ahead
+
+```{r}
+
+ES_2WEEK <- ggplot(W2_map) +
+  geom_sf(aes(fill = percentage_improvement)) + 
+  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
+  ggtitle("2 weeks ahead") +
+  theme_light() +
+  theme(legend.position = "right",
+    plot.title = element_text(hjust = 0.5)) +
+  geom_sf_text(data = W2_map, aes(label = round(percentage_improvement,0)), 
+               size = 3,
+               color = "black",
+               check_overlap = TRUE,fontface = "bold") + 
+  labs(
+    fill = "Mean difference",  
+       x = "",
+       y = ""
+  )  # Adding a subtitle
+
+x_limits <- c(-125, -67)  # Set the desired longitude range
+y_limits <- c(25, 50)    # Set the desired latitude range
+
+ES_2WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
+
+#ggsave(ES_2WEEK, height = 4.5, width = 6.5, filename="Fig5B.jpg")  
+```
+
+3 weeks ahead
+
+```{r}
+
+ES_3WEEK <- ggplot(W3_map) +
+  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
+  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
+  ggtitle("3 weeks ahead") +
+  theme_light() +
+  theme(legend.position = "right",
+    plot.title = element_text(hjust = 0.5)) +
+  geom_sf_text(data = W3_map, aes(label = round(percentage_improvement,0)),  
                size = 3,
@@
 
-ES_1WEEK<-ES_1WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
-
-#ggsave(ES_1WEEK,height = 4.5, width = 6.5,  filename="Fig5A.jpg")  
-```
-2 weeks ahead
-
-```{r}
-
-ES_2WEEK <- ggplot(W2_map) +
+ES_3WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
+
+#ggsave(ES_3WEEK, height = 4.5, width = 6.5, filename="Fig5C.jpg")  
+```
+
+4 weeks ahead
+
+```{r}
+ES_4WEEK <- ggplot(W4_map) +
   geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
   scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
-  ggtitle("2 weeks ahead") +
+  ggtitle("4 weeks ahead") +
   theme_light() +
@@
     plot.title = element_text(hjust = 0.5)) +
-  geom_sf_text(data = W2_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
+  geom_sf_text(data = W4_map, aes(label = round(percentage_improvement,0)), 
                size = 3,
@@
 
-ES_2WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
-
-#ggsave(ES_2WEEK, height = 4.5, width = 6.5, filename="Fig5B.jpg")  
-```
-
-3 weeks ahead
-
-```{r}
-
-ES_3WEEK <- ggplot(W3_map) +
-  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
-  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
-  ggtitle("3 weeks ahead") +
-  theme_light() +
-  theme(legend.position = "right",
-    plot.title = element_text(hjust = 0.5)) +
-  geom_sf_text(data = W3_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
-               size = 3,
-               color = "black",
-               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
-  labs(
-    fill =  "Mean difference",  # Label for the legend
-       x = "",
-       y = ""
-  )  # Adding a subtitle
-
-x_limits <- c(-125, -67)  # Set the desired longitude range
-y_limits <- c(25, 50)    # Set the desired latitude range
-
-ES_3WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
-
-#ggsave(ES_3WEEK, height = 4.5, width = 6.5, filename="Fig5C.jpg")  
-```
-4 weeks ahead
-
-```{r}
-ES_4WEEK <- ggplot(W4_map) +
-  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
-  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
-  ggtitle("4 weeks ahead") +
-  theme_light() +
-  theme(legend.position = "right",
-    plot.title = element_text(hjust = 0.5)) +
-  geom_sf_text(data = W4_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
-               size = 3,
-               color = "black",
-               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
-  labs(
-    fill = "Mean difference",  # Label for the legend
-       x = "",
-       y = ""
-  )  # Adding a subtitle
-
-x_limits <- c(-125, -67)  # Set the desired longitude range
-y_limits <- c(25, 50)    # Set the desired latitude range
-
 ES_4WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
@@
 ```
+
+Combining plot on a 2x2 grid.
 
@@
   )
+
 # Print
@@
 
-We will run the regression for evaluating if the best model log(WIS) which represent its performance is related to given independent variables.
+Now we will run the regression models for evaluating if the best model percentage of improvement in each state is related to given independent variables.
 
@@
 
-Computing and plotting mean WIS improvement by epidemiological weeks.
+Here we will be computing and plotting mean WIS improvement by epidemiological weeks comparing the best model and the basiline model.
 

### 2025-06-20 manuscript_results.Rmd
code review

@@
 
-Here we will be computing and plotting mean WIS improvement by epidemiological weeks comparing the best model and the basiline model.
+Here we will be computing and plotting mean WIS improvement by epidemiological weeks comparing the best model and the baseline model.
 
@@
 weeks <- 1:4
-summed_data_list <- list()
+results_by_epiweek_list <- list()
 
@@
   # Merge the data frames for the current week
-  data__ <- merge(get(paste0("filtered_df_W", week, "_NoLg")),
+  combined_dataframes_ <- merge(get(paste0("filtered_df_W", week, "_NoLg")),
                   get(paste0("filtered_df_W", week)))
@@
   # Sum values by Julian date across all states
-  summed_data <- data__ %>%
+  mean_improvement_ <- combined_dataframes_ %>%
     group_by(Julian_date) %>%
@@
     ) %>%
-    # Calculate the difference ratio
+    # Calculate the difference ratio in percentage
     mutate(Difference =  ((AUTO_AVG_LB / AUTO_AR)-1)*100,
@@
 
-  # Store the summarized data in the list
-  summed_data_list[[week]] <- summed_data
+  # Store the mean improvement by julian date
+  results_by_epiweek_list[[week]] <- mean_improvement_
 }
@@
 # Combine all weeks into a single data frame
-summed_data_all <- bind_rows(summed_data_list)
-
-# Plot the difference using ggplot2
-model_improv<-ggplot(summed_data_all, aes(x = Julian_date, y = (Difference), color = Week)) +
+results_by_epiweek_all <- bind_rows(results_by_epiweek_list)
+
+# Plot the results
+model_improv<-ggplot(results_by_epiweek_all, aes(x = Julian_date, y = (Difference), color = Week)) +
   geom_point(size = 2.5) +
@@
 
+# get total hospitalizations for the 48 states
 total_hospitalizations_by_week <- AUTO_ADJACENT_WEEK1 %>%
-  filter(epiweek >= 40 | epiweek <= 20) %>%  # Apply filter AFTER epiweek exists
+  filter(epiweek >= 40 | epiweek <= 20) %>%  
   group_by(target_end_date) %>%
@@
 
+# plotting results
 hospital_plot <-ggplot(total_hospitalizations_by_week, aes(x = target_end_date, y = mean_cases)) +
@@
 ```{r}
-
-# Combine the plots vertically
+# Combining results in a single plot
 combined_plot <-  model_improv /hospital_plot 
-
-# Display (optional)
 combined_plot

### 2025-06-23 manuscript_results.Rmd
update

@@
 
-# Prepare a list to hold your plots
+# Prepare a list of plots
 bric_plots <- list()
@@
 for (i in 1:4) {
-  # 1) Join the WIS and BRIC data
+  # Join the WIS and BRIC data
   WIS_bric <- inner_join(
@@
 
-  # 2) Fit the linear model
+  # Fit the linear model
   model <- lm(
@@
 
-  # 3) Extract R² and p‐value
+  # Extract R-squared and p-value
   r2 <- round(ms$r.squared, 3)
@@
 
-  # 4) Build the ggplot
+  # Build the ggplot
   p <- ggplot(WIS_bric,
@@
 
-  # 5) Store it
+  # Store it
   bric_plots[[i]] <- p

### 2025-06-23 models_without_logback/ES_TEMPERATURE/TEMPERATURE.Rmd
update name

---
title: "TEMPERATURE ARIMAX MODELS - Influenza hospitalization data"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

This function utilizes ensembles and single automatic ARIMAX models which have mean temperature as exogenous variables. The function fits on a rolling window of previous 104 weeks for the state under analysis and a rolling windows with the same size with 1 week-lag for the exogenous variables to generate forecasts. It return some metrics that evaluate the performance of the models:target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles. The user can choose a  single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE). The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window which is set as 2 years (104 weeks).

```{r}
knitr::opts_chunk$set(echo = TRUE)
```
!!!!!!!!!!!!!!!!!!!!
LOADING THE PACKAGES
!!!!!!!!!!!!!!!!!!!!
```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
```
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
LOADING DATASET AND FUNCTIONS 
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

```{r}
###########################################
#       LOADING AND CLEANING THE DATASET  #        
###########################################

# Loads the model
source("ES_TEMPERATURE_nolog.R", local = TRUE, chdir = TRUE)

# Loads the dataset
my_data = read.csv("treated_influenza_hosp_dataframe_v2.csv")

my_data$target_end_date<-as.Date(my_data$target_end_date) # set the dates as dates

list_of_states <- split(my_data, my_data$state_name)
```

Loading temperature data for 2010_2024 and filtering only the correct dates

```{r}
# ERA5-based temperature dataframe
temperature_data<-read.csv("final_temperature_data_2010_2024.csv")
temperature_data$date<-as.Date(temperature_data$date)

correct_dates <- as.Date(list_of_states$Alabama$target_end_date)

# Filter temperature_data to keep only rows with correct dates
temperature_data <- temperature_data %>%
  filter(date %in% correct_dates)%>% 
  select(-X,-date, -year,-epi_week)

head(temperature_data)
```

AUTO TEMPERATURE WEEK1

```{r}
start_time <- Sys.time()

AUTO_TEMPERATURE_WEEK1_list <- mclapply(list_of_states, ES_TEMPERATURE, auto=TRUE, n_weeks_ahead=1,week_lag=1,temperature_data=temperature_data, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
AUTO_TEMPERATURE_WEEK1 <- bind_rows(AUTO_TEMPERATURE_WEEK1_list, .id = "State")
```

AUTO TEMPERATURE WEEK2

```{r}
start_time <- Sys.time()

AUTO_TEMPERATURE_WEEK2_list <- mclapply(list_of_states, ES_TEMPERATURE, auto=TRUE, n_weeks_ahead=2,week_lag=1,temperature_data=temperature_data, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
AUTO_TEMPERATURE_WEEK2 <- bind_rows(AUTO_TEMPERATURE_WEEK2_list, .id = "State")
```

AUTO TEMPERATURE WEEK3

```{r}
start_time <- Sys.time()

AUTO_TEMPERATURE_WEEK3_list <- mclapply(list_of_states, ES_TEMPERATURE, auto=TRUE, n_weeks_ahead=3,week_lag=1,temperature_data=temperature_data, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
AUTO_TEMPERATURE_WEEK3 <- bind_rows(AUTO_TEMPERATURE_WEEK3_list, .id = "State")
```

AUTO TEMPERATURE WEEK4

```{r}
start_time <- Sys.time()

AUTO_TEMPERATURE_WEEK4_list <- mclapply(list_of_states, ES_TEMPERATURE, auto=TRUE, n_weeks_ahead=4,week_lag=1,temperature_data=temperature_data, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
AUTO_TEMPERATURE_WEEK4 <- bind_rows(AUTO_TEMPERATURE_WEEK4_list, .id = "State")
```

```{r}
save.image("TEMPERATURE_MODELS_influenza_hospitalization_nolog.Rdata")
```

ES27 TEMPERATURE WEEK1

```{r}
start_time <- Sys.time()

ES27_TEMPERATURE_WEEK1_list <- mclapply(list_of_states, ES_TEMPERATURE, ES27=TRUE, n_weeks_ahead=1,week_lag=1,temperature_data=temperature_data, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_TEMPERATURE_WEEK1 <- bind_rows(ES27_TEMPERATURE_WEEK1_list, .id = "State")
```

ES27 TEMPERATURE WEEK2

```{r}
start_time <- Sys.time()

ES27_TEMPERATURE_WEEK2_list <- mclapply(list_of_states, ES_TEMPERATURE, ES27=TRUE, n_weeks_ahead=2,week_lag=1,temperature_data=temperature_data, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_TEMPERATURE_WEEK2 <- bind_rows(ES27_TEMPERATURE_WEEK2_list, .id = "State")
```

ES27 TEMPERATURE WEEK3

```{r}
start_time <- Sys.time()

ES27_TEMPERATURE_WEEK3_list <- mclapply(list_of_states, ES_TEMPERATURE, ES27=TRUE, n_weeks_ahead=3,week_lag=1,temperature_data=temperature_data, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_TEMPERATURE_WEEK3 <- bind_rows(ES27_TEMPERATURE_WEEK3_list, .id = "State")
```

ES27 TEMPERATURE WEEK4

```{r}
start_time <- Sys.time()

ES27_TEMPERATURE_WEEK4_list <- mclapply(list_of_states, ES_TEMPERATURE, ES27=TRUE, n_weeks_ahead=4,week_lag=1,temperature_data=temperature_data, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_TEMPERATURE_WEEK4 <- bind_rows(ES27_TEMPERATURE_WEEK4_list, .id = "State")
```

```{r}
save.image("TEMPERATURE_MODELS_influenza_hospitalization_nolog.Rdata")

```

ES64 TEMPERATURE WEEK1

```{r}
start_time <- Sys.time()

ES64_TEMPERATURE_WEEK1_list <- mclapply(list_of_states, ES_TEMPERATURE, ES64=TRUE, n_weeks_ahead=1,week_lag=1,temperature_data=temperature_data, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_TEMPERATURE_WEEK1 <- bind_rows(ES64_TEMPERATURE_WEEK1_list, .id = "State")
```

ES64 TEMPERATURE WEEK2

```{r}
start_time <- Sys.time()

ES64_TEMPERATURE_WEEK2_list <- mclapply(list_of_states, ES_TEMPERATURE, ES64=TRUE, n_weeks_ahead=2,week_lag=1,temperature_data=temperature_data, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_TEMPERATURE_WEEK2 <- bind_rows(ES64_TEMPERATURE_WEEK2_list, .id = "State")
```

ES64 TEMPERATURE WEEK3

```{r}
start_time <- Sys.time()

ES64_TEMPERATURE_WEEK3_list <- mclapply(list_of_states, ES_TEMPERATURE, ES64=TRUE, n_weeks_ahead=3,week_lag=1,temperature_data=temperature_data, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_TEMPERATURE_WEEK3 <- bind_rows(ES64_TEMPERATURE_WEEK3_list, .id = "State")
```

ES64 TEMPERATURE WEEK4

```{r}
start_time <- Sys.time()

ES64_TEMPERATURE_WEEK4_list <- mclapply(list_of_states, ES_TEMPERATURE, ES64=TRUE, n_weeks_ahead=4,week_lag=1,temperature_data=temperature_data, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_TEMPERATURE_WEEK4 <- bind_rows(ES64_TEMPERATURE_WEEK4_list, .id = "State")
```

```{r}
save.image("TEMPERATURE_MODELS_influenza_hospitalization_nolog.Rdata")
```

### 2025-06-23 models_without_logback/ES_ARIMA/ES_ARIMA_nolog.R

@@ -1,0 +1,387 @@
+#' ES_ARIMA 
+#' This function utilizes ensembles and single automatic ARIMA models.
+#' The function fits a rolling windows of N weeks to generate forecasts.
+#' It return some metrics that evaluate the performance of the models:
+#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
+#' The user defines if it will use an AUTO ARIMA (auto=TRUE), or ensembles of 27 (ES27=TRUE) or 64 models (ES64=TRUE), chooses the number of weeks ahead for each forecast, and the size of the rolling window.
+#' 
+#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. state.
+#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
+#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
+#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
+#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
+#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
+#'
+#' @return A list containing the forecast results and performance metrics.
+
+ES_ARIMA<-function(current_state_tb, auto=FALSE, n_weeks_ahead=1, ES27=TRUE, ES64=FALSE, window=104){
+  set.seed(1)
+  # The model run the ES27 or the ES64.
+  ES27=!ES64 
+  # Empty list that will contain forecasts, predictive quantiles and number of models.
+  results<-listenv()   
+  if(ES27){
+    pdq=c(0,1,2) # Possible ARIMA pdq's.
+    my_order_params<-permutations(3,3,pdq, repeats.allowed = TRUE) # Create 27 permutations of [0,1,2]
+  }
+  if(ES64){
+    pdq=c(0,1,2,3) # Possible ARIMA pdq's.
+    my_order_params<-permutations(4,3,pdq, repeats.allowed = TRUE) # Create 64 permutations of [0,1,2,3]
+  }
+  # Apply the ROLLING_ARIMA function to get results. Window set to 104 weeks (2 years).
+  results[[1]]<- ROLLING_ARIMA(current_state_tb, n_weeks_ahead=n_weeks_ahead, window = window, order_params=my_order_params, auto=auto) 
+  # Resolve from system environment using the future package 
+
[…]

### 2025-06-23 models_without_logback/ES_EPIWEEK/ES_EPIWEEK_nolog.R

@@ -1,0 +1,410 @@
+#' ES_EPIWEEK
+#' This function utilizes ensembles and single automatic ARIMAX models which have mean cases by epidemiological weeks as exogenous variables.
+#' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
+#' It return some metrics that evaluate the performance of the models:
+#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
+#' The user can choose  single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).
+#' The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window.
+#' 
+#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. states.
+#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
+#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
+#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
+#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
+#' @param week_lag An integer specifying the week lag between the exogenous variables and the ILI cases in the state under analysis. Default is \code{1}.
+#' @param list_of_states A list containing Influenza Like Illness tibbles for ALL U.S. states.
+#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
+#'
+#' @return A list containing the forecast results and performance metrics.
+
+
+ES_EPIWEEK<-function(current_state_tb, auto=FALSE, n_weeks_ahead=1, ES27=TRUE, ES64=FALSE, window=104,week_lag=1){
+  set.seed(1)
+  # The model run the ES27 or the ES64.
+  ES27=!ES64 
+  # Empty list that will contain forecasts, predictive quantiles and number of models.
+  results<-listenv()   
+  if(ES27){
+    pdq=c(0,1,2) # Possible ARIMA pdq's.
[…]

### 2025-06-23 NHSH_hospitalizations_seed_nolog_correct/ES_ADJACENT/ES_ADJACENT_48_nolog.R
update

@@ -1,475 +1,0 @@
-#' ES_ADJACENT 
-#' This function utilizes ensembles and single automatic ARIMAX models which have mean cases by adjacent states as exogenous variables.
-#' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
-#' It return some metrics that evaluate the performance of the models:
-#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
-#' The user can choose single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).
-#' The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window.
-#' 
-#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. states.
-#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
-#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
-#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
-#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
-#' @param week_lag An integer specifying the week lag between the exogenous variables and the ILI cases in the state under analysis. Default is \code{1}.
-#' @param list_of_states A list containing Influenza Like Illness tibbles for ALL U.S. states.
-#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
-#'
-#' @return A list containing the forecast results and performance metrics.
-
-ES_ADJACENT<-function(current_state_tb, auto=FALSE, n_weeks_ahead=1, week_lag=1, ES27=TRUE, ES64=FALSE, window=104, list_of_states=list_of_states){
-  
-  set.seed(1)
-  # The model run the ES27 or the ES64.
-  ES27=!ES64 
[…]

### 2025-06-23 NHSH_hospitalizations_seed_nolog_correct/ES_ARIMA/ES_ARIMA_nolog.R

@@ -1,387 +1,0 @@
-#' ES_ARIMA 
-#' This function utilizes ensembles and single automatic ARIMA models.
-#' The function fits a rolling windows of N weeks to generate forecasts.
-#' It return some metrics that evaluate the performance of the models:
-#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
-#' The user defines if it will use an AUTO ARIMA (auto=TRUE), or ensembles of 27 (ES27=TRUE) or 64 models (ES64=TRUE), chooses the number of weeks ahead for each forecast, and the size of the rolling window.
-#' 
-#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. state.
-#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
-#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
-#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
-#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
-#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
-#'
-#' @return A list containing the forecast results and performance metrics.
-
-ES_ARIMA<-function(current_state_tb, auto=FALSE, n_weeks_ahead=1, ES27=TRUE, ES64=FALSE, window=104){
-  set.seed(1)
-  # The model run the ES27 or the ES64.
-  ES27=!ES64 
-  # Empty list that will contain forecasts, predictive quantiles and number of models.
-  results<-listenv()   
-  if(ES27){
-    pdq=c(0,1,2) # Possible ARIMA pdq's.
-    my_order_params<-permutations(3,3,pdq, repeats.allowed = TRUE) # Create 27 permutations of [0,1,2]
-  }
-  if(ES64){
-    pdq=c(0,1,2,3) # Possible ARIMA pdq's.
-    my_order_params<-permutations(4,3,pdq, repeats.allowed = TRUE) # Create 64 permutations of [0,1,2,3]
-  }
-  # Apply the ROLLING_ARIMA function to get results. Window set to 104 weeks (2 years).
-  results[[1]]<- ROLLING_ARIMA(current_state_tb, n_weeks_ahead=n_weeks_ahead, window = window, order_params=my_order_params, auto=auto) 
-  # Resolve from system environment using the future package 
-
[…]

### 2025-06-23 NHSH_hospitalizations_seed_nolog_correct/ES_AVERAGE/ES_AVERAGE_48_nolog.R

@@ -1,477 +1,0 @@
-#' ES_AVERAGE 
-#' This function utilizes ensembles and single automatic ARIMAX models which have mean cases by AVERAGE states as exogenous variables.
-#' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
-#' It return some metrics that evaluate the performance of the models:
-#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
-#' The user can choose single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).
-#' The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window.
-#' 
-#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. states.
-#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
-#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
-#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
-#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
-#' @param week_lag An integer specifying the week lag between the exogenous variables and the ILI cases in the state under analysis. Default is \code{1}.
-#' @param list_of_states A list containing Influenza Like Illness tibbles for ALL U.S. states.
-#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
-#'
-#' @return A list containing the forecast results and performance metrics.
-
-ES_AVERAGE<-function(current_state_tb, auto=FALSE, n_weeks_ahead=1, week_lag=1, ES27=TRUE, ES64=FALSE, window=104, list_of_states=list_of_states){
-  
-  set.seed(1)
-  # The model run the ES27 or the ES64.
-  ES27=!ES64 
-  # Empty list that will contain forecasts, predictive quantiles and number of models.
-  results<-listenv()   
-  if(ES27){
-    pdq=c(0,1,2) # Possible ARIMA pdq's.
[…]

### 2025-06-23 NHSH_hospitalizations_seed_nolog_correct/ES_EPIWEEK/ES_EPIWEEK_nolog.R

@@ -1,410 +1,0 @@
-#' ES_EPIWEEK
-#' This function utilizes ensembles and single automatic ARIMAX models which have mean cases by epidemiological weeks as exogenous variables.
-#' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
-#' It return some metrics that evaluate the performance of the models:
-#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
-#' The user can choose  single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).
-#' The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window.
-#' 
-#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. states.
-#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
-#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
-#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
-#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
-#' @param week_lag An integer specifying the week lag between the exogenous variables and the ILI cases in the state under analysis. Default is \code{1}.
-#' @param list_of_states A list containing Influenza Like Illness tibbles for ALL U.S. states.
-#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
-#'
-#' @return A list containing the forecast results and performance metrics.
-
-
-ES_EPIWEEK<-function(current_state_tb, auto=FALSE, n_weeks_ahead=1, ES27=TRUE, ES64=FALSE, window=104,week_lag=1){
-  set.seed(1)
-  # The model run the ES27 or the ES64.
-  ES27=!ES64 
-  # Empty list that will contain forecasts, predictive quantiles and number of models.
-  results<-listenv()   
-  if(ES27){
-    pdq=c(0,1,2) # Possible ARIMA pdq's.
[…]

### 2025-06-23 NHSH_hospitalizations_seed_nolog_correct/ES_TEMPERATURE/ES_TEMPERATURE_nolog.R

@@ -1,400 +1,0 @@
-#' ES_TEMPERATURE
-#' This function utilizes ensembles and single automatic ARIMAX models which have mean temperature as exogenous variables.
-#' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
-#' It return some metrics that evaluate the performance of the models:
-#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
-#' The user can choose  single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).
-#' The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window.
-#' 
-#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. states.
-#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
-#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
-#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
-#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
-#' @param week_lag An integer specifying the week lag between the exogenous variables and the ILI cases in the state under analysis. Default is \code{1}.
-#' @param temperature_data A data frame containing weekly mean temperature data for ALL U.S. states for the same weeks as the ILI data.
-#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
-#'
-#' @return A list containing the forecast results and performance metrics.
-
-ES_TEMPERATURE<-function(current_state_tb, week_lag=1, temperature_data=NULL, auto=FALSE, n_weeks_ahead=1, ES27=TRUE, ES64=FALSE, window=104){
-  
-  set.seed(1)
-  # The model run the ES27 or the ES64.
-  ES27=!ES64 
-  # Empty list that will contain forecasts, predictive quantiles and number of models.
-  results<-listenv()   
-  if(ES27){
-    pdq=c(0,1,2) # Possible ARIMA pdq's.
[…]

### 2025-06-23 models_with_logback/ES_ADJACENT/ADJACENT.Rmd
update

--- 2025-04-25 NHSH_hospitalizations_seed_nolog_correct/ES_ADJACENT/ADJACENT.Rmd
@@
 # Loads the ADJACENT states models
-source("ES_ADJACENT_48_nolog.R", local = TRUE, chdir = TRUE)
+source("ES_ADJACENT_48.R", local = TRUE, chdir = TRUE)
 
@@
 ```{r}
-save.image("ADJACENT_MODELS_influenza_hospitalization_nolog.Rdata")
+save.image("ADJACENT_MODELS_influenza_hospitalization.Rdata")
 ```

### 2025-06-23 models_with_logback/ES_AVERAGE/AVERAGE.Rmd
update

--- 2025-04-25 NHSH_hospitalizations_seed_nolog_correct/ES_AVERAGE/AVERAGE.Rmd
@@
 ---
-title: "AVERAGE STATES ARIMAX MODEL - Influenza hospitalization data - no log-back transformation"
+title: "AVERAGE STATES ARIMAX MODEL - Influenza hospitalization data"
 author: "Victor Felix"
@@
 # Loads the AVERAGE states models
-source("ES_AVERAGE_48_nolog.R", local = TRUE, chdir = TRUE)
+source("ES_AVERAGE_48.R", local = TRUE, chdir = TRUE)
 
@@
 ```{r}
-save.image("AVERAGE_MODELS_influenza_hospitalization_nolog.Rdata")
+save.image("AVERAGE_MODELS_influenza_hospitalization.Rdata")
 ```

### 2025-06-23 models_with_logback/ES_EPIWEEK/EPIWEEK.Rmd
update

--- 2025-04-25 NHSH_hospitalizations_seed_nolog_correct/ES_EPIWEEK/EPIWEEK.Rmd
@@
 # Loads the ADJACENT states models
-source("ES_EPIWEEK_nolog.R", local = TRUE, chdir = TRUE)
+source("ES_EPIWEEK.R", local = TRUE, chdir = TRUE)
 
@@
 ```{r}
-save.image("EPIWEEK_MODELS_influenza_hospitalization_nolog.Rdata")
+save.image("EPIWEEK_MODELS_influenza_hospitalization.Rdata")
 

### 2025-06-23 models_with_logback/ES_TEMPERATURE/TEMPERATURE.Rmd
update

--- 2025-06-23 models_without_logback/ES_TEMPERATURE/TEMPERATURE.Rmd
@@
 # Loads the model
-source("ES_TEMPERATURE_nolog.R", local = TRUE, chdir = TRUE)
+source("ES_TEMPERATURE.R", local = TRUE, chdir = TRUE)
 
@@
 ```{r}
-save.image("TEMPERATURE_MODELS_influenza_hospitalization_nolog.Rdata")
+save.image("TEMPERATURE_MODELS_influenza_hospitalization.Rdata")
 ```
@@
 ```{r}
-save.image("TEMPERATURE_MODELS_influenza_hospitalization_nolog.Rdata")
+save.image("TEMPERATURE_MODELS_influenza_hospitalization.Rdata")
 

### 2025-06-23 models_with_logback/hospitalization_results.Rmd
update

--- 2025-04-25 NHSH_hospitalizations_seed/hospitalization_results.Rmd
@@
 
-This document summarizes the results of the 12 models that were developed: AUTO ARIMA, AUTO ADJACENT, AUTO EPIWEEK, ES27 TEMPERATURE, ES27 ARIMA, ES27 ADJACENT, ES27 EPIWEEK, ES64 TEMPERATURE, ES64 ARIMA, ES64 ADJACENT, ES64 EPIWEEK, ES64 TEMPERATURE.
+This document summarizes the results of the 12 models with log-back transformations that were developed: AUTO ARIMA, AUTO ADJACENT, AUTO EPIWEEK, ES27 TEMPERATURE, ES27 ARIMA, ES27 ADJACENT, ES27 EPIWEEK, ES64 TEMPERATURE, ES64 ARIMA, ES64 ADJACENT, ES64 EPIWEEK, ES64 TEMPERATURE.
 It has a series of histograms and maps that compare the results of these models based on their WIS.
@@
 # Save the plot with specified size
-ggsave("regression_plot.jpg", width = 10, height = 14, dpi = 300)
+#ggsave("regression_plot.jpg", width = 10, height = 14, dpi = 300)
 ```

### 2025-06-23 models_with_logback/ES_ADJACENT/ES_ADJACENT_48.R

@@ -1,0 +1,475 @@
+#' ES_ADJACENT 
+#' This function use ensembles and single automatic ARIMAX models which have mean cases by adjacent states as exogenous variables.
+#' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
+#' It return some metrics that evaluate the performance of the models:
+#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
+#' The user can choose single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).
+#' The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window.
+#' 
+#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. states.
+#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
+#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
+#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
+#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
+#' @param week_lag An integer specifying the week lag between the exogenous variables and the ILI cases in the state under analysis. Default is \code{1}.
+#' @param list_of_states A list containing Influenza Like Illness tibbles for ALL U.S. states.
+#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
+#'
+#' @return A list containing the forecast results and performance metrics.
+
+ES_ADJACENT<-function(current_state_tb, auto=FALSE, n_weeks_ahead=1, week_lag=1, ES27=TRUE, ES64=FALSE, window=104, list_of_states=list_of_states){
+  
+  set.seed(1)
+  # The model run the ES27 or the ES64.
+  ES27=!ES64 
[…]

### 2025-06-23 models_with_logback/ES_ARIMA/ES_ARIMA.R

@@ -1,0 +1,387 @@
+#' ES_ARIMA 
+#' This function utilizes ensembles and single automatic ARIMA models.
+#' The function fits a rolling windows of N weeks to generate forecasts.
+#' It return some metrics that evaluate the performance of the models:
+#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
+#' The user defines if it will use an AUTO ARIMA (auto=TRUE), or ensembles of 27 (ES27=TRUE) or 64 models (ES64=TRUE), chooses the number of weeks ahead for each forecast, and the size of the rolling window.
+#' 
+#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. state.
+#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
+#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
+#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
+#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
+#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
+#'
+#' @return A list containing the forecast results and performance metrics.
+
+ES_ARIMA<-function(current_state_tb, auto=FALSE, n_weeks_ahead=1, ES27=TRUE, ES64=FALSE, window=104){
+  set.seed(1)
+  # The model run the ES27 or the ES64.
+  ES27=!ES64 
+  # Empty list that will contain forecasts, predictive quantiles and number of models.
+  results<-listenv()   
+  if(ES27){
+    pdq=c(0,1,2) # Possible ARIMA pdq's.
+    my_order_params<-permutations(3,3,pdq, repeats.allowed = TRUE) # Create 27 permutations of [0,1,2]
+  }
+  if(ES64){
+    pdq=c(0,1,2,3) # Possible ARIMA pdq's.
+    my_order_params<-permutations(4,3,pdq, repeats.allowed = TRUE) # Create 64 permutations of [0,1,2,3]
+  }
+  # Apply the ROLLING_ARIMA function to get results. Window set to 104 weeks (2 years).
+  results[[1]]<- ROLLING_ARIMA(current_state_tb, n_weeks_ahead=n_weeks_ahead, window = window, order_params=my_order_params, auto=auto) 
+  # Resolve from system environment using the future package 
+
[…]

### 2025-06-23 models_with_logback/ES_AVERAGE/ES_AVERAGE_48.R

@@ -1,0 +1,477 @@
+#' ES_AVERAGE 
+#' This function utilizes ensembles and single automatic ARIMAX models which have mean cases by AVERAGE states as exogenous variables.
+#' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
+#' It return some metrics that evaluate the performance of the models:
+#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
+#' The user can choose single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).
+#' The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window.
+#' 
+#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. states.
+#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
+#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
+#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
+#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
+#' @param week_lag An integer specifying the week lag between the exogenous variables and the ILI cases in the state under analysis. Default is \code{1}.
+#' @param list_of_states A list containing Influenza Like Illness tibbles for ALL U.S. states.
+#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
+#'
+#' @return A list containing the forecast results and performance metrics.
+
+ES_AVERAGE<-function(current_state_tb, auto=FALSE, n_weeks_ahead=1, week_lag=1, ES27=TRUE, ES64=FALSE, window=104, list_of_states=list_of_states){
+  
+  set.seed(1)
+  # The model run the ES27 or the ES64.
+  ES27=!ES64 
+  # Empty list that will contain forecasts, predictive quantiles and number of models.
+  results<-listenv()   
+  if(ES27){
+    pdq=c(0,1,2) # Possible ARIMA pdq's.
[…]

### 2025-06-23 models_with_logback/ES_EPIWEEK/ES_EPIWEEK.R

@@ -1,0 +1,410 @@
+#' ES_EPIWEEK
+#' This function utilizes ensembles and single automatic ARIMAX models which have mean cases by epidemiological weeks as exogenous variables.
+#' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
+#' It return some metrics that evaluate the performance of the models:
+#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
+#' The user can choose  single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).
+#' The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window.
+#' 
+#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. states.
+#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
+#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
+#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
+#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
+#' @param week_lag An integer specifying the week lag between the exogenous variables and the ILI cases in the state under analysis. Default is \code{1}.
+#' @param list_of_states A list containing Influenza Like Illness tibbles for ALL U.S. states.
+#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
+#'
+#' @return A list containing the forecast results and performance metrics.
+
+
+ES_EPIWEEK<-function(current_state_tb, auto=FALSE, n_weeks_ahead=1, ES27=TRUE, ES64=FALSE, window=104,week_lag=1){
+  set.seed(1)
+  # The model run the ES27 or the ES64.
+  ES27=!ES64 
+  # Empty list that will contain forecasts, predictive quantiles and number of models.
+  results<-listenv()   
+  if(ES27){
+    pdq=c(0,1,2) # Possible ARIMA pdq's.
[…]

### 2025-06-23 manuscript_results.Rmd
delete

@@
 
-load("NHSH_hospitalizations_seed_nolog_correct/ES_ARIMA/ARIMA_MODELS_influenza_hospitalization_nolog.Rdata")
-load("NHSH_hospitalizations_seed_nolog_correct/ES_ADJACENT/ADJACENT_MODELS_influenza_hospitalization_nolog.Rdata")
-load("NHSH_hospitalizations_seed_nolog_correct/ES_EPIWEEK/EPIWEEK_MODELS_influenza_hospitalization_nolog.Rdata")
-load("NHSH_hospitalizations_seed_nolog_correct/ES_TEMPERATURE/TEMPERATURE_MODELS_influenza_hospitalization_nolog.Rdata")
-load("NHSH_hospitalizations_seed_nolog_correct/ES_AVERAGE/AVERAGE_MODELS_influenza_hospitalization_nolog.Rdata")
-
-states <- read_sf("NHSH_hospitalizations_seed_nolog_correct/shapefiles/cb_2018_us_state_500k.shp")
+load("models_without_logback/ES_ARIMA/ARIMA_MODELS_influenza_hospitalization_nolog.Rdata")
+load("models_without_logback/ES_ADJACENT/ADJACENT_MODELS_influenza_hospitalization_nolog.Rdata")
+load("models_without_logback/ES_EPIWEEK/EPIWEEK_MODELS_influenza_hospitalization_nolog.Rdata")
+load("models_without_logback/ES_TEMPERATURE/TEMPERATURE_MODELS_influenza_hospitalization_nolog.Rdata")
+load("models_without_logback/ES_AVERAGE/AVERAGE_MODELS_influenza_hospitalization_nolog.Rdata")
+
+states <- read_sf("models_without_logback/shapefiles/cb_2018_us_state_500k.shp")
 
@@
 ```{r}
-load("NHSH_hospitalizations_seed/ES_ARIMA/ARIMA_MODELS_influenza_hospitalization.Rdata")
-load("NHSH_hospitalizations_seed/ES_ADJACENT/ADJACENT_MODELS_influenza_hospitalization.Rdata")
-load("NHSH_hospitalizations_seed/ES_EPIWEEK/EPIWEEK_MODELS_influenza_hospitalization.Rdata")
-load("NHSH_hospitalizations_seed/ES_TEMPERATURE/TEMPERATURE_MODELS_influenza_hospitalization.Rdata")
-load("NHSH_hospitalizations_seed/ES_AVERAGE/AVERAGE_MODELS_influenza_hospitalization.Rdata")
+load("models_with_logback/ES_ARIMA/ARIMA_MODELS_influenza_hospitalization.Rdata")
+load("models_with_logback/ES_ADJACENT/ADJACENT_MODELS_influenza_hospitalization.Rdata")
+load("models_with_logback/ES_EPIWEEK/EPIWEEK_MODELS_influenza_hospitalization.Rdata")
+load("models_with_logback/ES_TEMPERATURE/TEMPERATURE_MODELS_influenza_hospitalization.Rdata")
+load("models_with_logback/ES_AVERAGE/AVERAGE_MODELS_influenza_hospitalization.Rdata")
 ```

### 2025-06-23 NHSH_hospitalizations_seed/ES_ADJACENT/ES_ADJACENT_48.R

@@ -1,475 +1,0 @@
-#' ES_ADJACENT 
-#' This function use ensembles and single automatic ARIMAX models which have mean cases by adjacent states as exogenous variables.
-#' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
-#' It return some metrics that evaluate the performance of the models:
-#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
-#' The user can choose single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).
-#' The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window.
-#' 
-#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. states.
-#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
-#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
-#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
-#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
-#' @param week_lag An integer specifying the week lag between the exogenous variables and the ILI cases in the state under analysis. Default is \code{1}.
-#' @param list_of_states A list containing Influenza Like Illness tibbles for ALL U.S. states.
-#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
-#'
-#' @return A list containing the forecast results and performance metrics.
-
-ES_ADJACENT<-function(current_state_tb, auto=FALSE, n_weeks_ahead=1, week_lag=1, ES27=TRUE, ES64=FALSE, window=104, list_of_states=list_of_states){
-  
-  set.seed(1)
-  # The model run the ES27 or the ES64.
-  ES27=!ES64 
[…]

### 2025-06-23 NHSH_hospitalizations_seed/ES_ARIMA/ES_ARIMA.R

@@ -1,387 +1,0 @@
-#' ES_ARIMA 
-#' This function utilizes ensembles and single automatic ARIMA models.
-#' The function fits a rolling windows of N weeks to generate forecasts.
-#' It return some metrics that evaluate the performance of the models:
-#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
-#' The user defines if it will use an AUTO ARIMA (auto=TRUE), or ensembles of 27 (ES27=TRUE) or 64 models (ES64=TRUE), chooses the number of weeks ahead for each forecast, and the size of the rolling window.
-#' 
-#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. state.
-#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
-#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
-#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
-#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
-#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
-#'
-#' @return A list containing the forecast results and performance metrics.
-
-ES_ARIMA<-function(current_state_tb, auto=FALSE, n_weeks_ahead=1, ES27=TRUE, ES64=FALSE, window=104){
-  set.seed(1)
-  # The model run the ES27 or the ES64.
-  ES27=!ES64 
-  # Empty list that will contain forecasts, predictive quantiles and number of models.
-  results<-listenv()   
-  if(ES27){
-    pdq=c(0,1,2) # Possible ARIMA pdq's.
-    my_order_params<-permutations(3,3,pdq, repeats.allowed = TRUE) # Create 27 permutations of [0,1,2]
-  }
-  if(ES64){
-    pdq=c(0,1,2,3) # Possible ARIMA pdq's.
-    my_order_params<-permutations(4,3,pdq, repeats.allowed = TRUE) # Create 64 permutations of [0,1,2,3]
-  }
-  # Apply the ROLLING_ARIMA function to get results. Window set to 104 weeks (2 years).
-  results[[1]]<- ROLLING_ARIMA(current_state_tb, n_weeks_ahead=n_weeks_ahead, window = window, order_params=my_order_params, auto=auto) 
-  # Resolve from system environment using the future package 
-
[…]

### 2025-06-23 NHSH_hospitalizations_seed/ES_AVERAGE/ES_AVERAGE_48.R

@@ -1,477 +1,0 @@
-#' ES_AVERAGE 
-#' This function utilizes ensembles and single automatic ARIMAX models which have mean cases by AVERAGE states as exogenous variables.
-#' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
-#' It return some metrics that evaluate the performance of the models:
-#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
-#' The user can choose single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).
-#' The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window.
-#' 
-#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. states.
-#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
-#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
-#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
-#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
-#' @param week_lag An integer specifying the week lag between the exogenous variables and the ILI cases in the state under analysis. Default is \code{1}.
-#' @param list_of_states A list containing Influenza Like Illness tibbles for ALL U.S. states.
-#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
-#'
-#' @return A list containing the forecast results and performance metrics.
-
-ES_AVERAGE<-function(current_state_tb, auto=FALSE, n_weeks_ahead=1, week_lag=1, ES27=TRUE, ES64=FALSE, window=104, list_of_states=list_of_states){
-  
-  set.seed(1)
-  # The model run the ES27 or the ES64.
-  ES27=!ES64 
-  # Empty list that will contain forecasts, predictive quantiles and number of models.
-  results<-listenv()   
-  if(ES27){
-    pdq=c(0,1,2) # Possible ARIMA pdq's.
[…]

### 2025-06-23 NHSH_hospitalizations_seed/ES_EPIWEEK/ES_EPIWEEK.R

@@ -1,410 +1,0 @@
-#' ES_EPIWEEK
-#' This function utilizes ensembles and single automatic ARIMAX models which have mean cases by epidemiological weeks as exogenous variables.
-#' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
-#' It return some metrics that evaluate the performance of the models:
-#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
-#' The user can choose  single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).
-#' The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window.
-#' 
-#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. states.
-#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
-#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
-#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
-#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
-#' @param week_lag An integer specifying the week lag between the exogenous variables and the ILI cases in the state under analysis. Default is \code{1}.
-#' @param list_of_states A list containing Influenza Like Illness tibbles for ALL U.S. states.
-#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
-#'
-#' @return A list containing the forecast results and performance metrics.
-
-
-ES_EPIWEEK<-function(current_state_tb, auto=FALSE, n_weeks_ahead=1, ES27=TRUE, ES64=FALSE, window=104,week_lag=1){
-  set.seed(1)
-  # The model run the ES27 or the ES64.
-  ES27=!ES64 
-  # Empty list that will contain forecasts, predictive quantiles and number of models.
-  results<-listenv()   
-  if(ES27){
-    pdq=c(0,1,2) # Possible ARIMA pdq's.
[…]

### 2025-06-23 NHSH_hospitalizations_seed/ES_TEMPERATURE/ES_TEMPERATURE.R

@@ -1,400 +1,0 @@
-#' ES_TEMPERATURE
-#' This function utilizes ensembles and single automatic ARIMAX models which have mean temperature as exogenous variables.
-#' The function fits rolling windows of N weeks for the state under analysis and rolling windows of the same size with 1 week-lag for the exogenous variables to generate forecasts.
-#' It return some metrics that evaluate the performance of the models:
-#' target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles (%)  
-#' The user can choose  single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).
-#' The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window.
-#' 
-#' @param current_state_tb A list containing Influenza Like Illness tibbles for one or many U.S. states.
-#' @param auto A logical value indicating whether to use AUTO ARIMA. Default is \code{FALSE}.
-#' @param ES27 A logical value indicating whether to use ensembles of 27 models. Default is \code{TRUE}.
-#' @param ES64 A logical value indicating whether to use ensembles of 64 models. Default is \code{FALSE}.
-#' @param n_weeks_ahead An integer specifying the number of weeks ahead for each forecast. Default is \code{1}.
-#' @param week_lag An integer specifying the week lag between the exogenous variables and the ILI cases in the state under analysis. Default is \code{1}.
-#' @param temperature_data A data frame containing weekly mean temperature data for ALL U.S. states for the same weeks as the ILI data.
-#' @param window An integer specifying the size of the rolling window in number of weeks. Default is \code{104}.
-#'
-#' @return A list containing the forecast results and performance metrics.
-
-ES_TEMPERATURE<-function(current_state_tb, week_lag=1, temperature_data=NULL, auto=FALSE, n_weeks_ahead=1, ES27=TRUE, ES64=FALSE, window=104){
-  
-  set.seed(1)
-  # The model run the ES27 or the ES64.
-  ES27=!ES64 
-  # Empty list that will contain forecasts, predictive quantiles and number of models.
-  results<-listenv()   
-  if(ES27){
-    pdq=c(0,1,2) # Possible ARIMA pdq's.
[…]

### 2025-06-23 manuscript_results.Rmd
update

@@
 
-my_data = read.csv("NHSH_hospitalizations_seed/treated_influenza_hosp_dataframe_v2.csv")
+my_data = read.csv("models_with_logback/treated_influenza_hosp_dataframe_v2.csv")
 my_data$target_end_date<-as.Date(my_data$target_end_date) # set the dates as dates
@@
 ```{r}
-pop_data <- read.csv("NHSH_hospitalizations_seed/regression_features/population_data.csv") # resident population and population density
-sovi_data <- read.csv("NHSH_hospitalizations_seed/regression_features/sovi_2010.2014.csv") # SOVI index
-bric_data<-read.csv("NHSH_hospitalizations_seed/regression_features/bric2015.csv") # BRIC index
-humidity_data<-read.csv("NHSH_hospitalizations_seed/regression_features/humidity_climatology_1990_2020.csv") # ERA5 Specific Humidity
-temperature_data<-read.csv("NHSH_hospitalizations_seed/regression_features/temperature_climatology_1990_2020.csv") # ERA5 Specific Temperature
+pop_data <- read.csv("models_with_logback/regression_features/population_data.csv") # resident population and population density
+sovi_data <- read.csv("models_with_logback/regression_features/sovi_2010.2014.csv") # SOVI index
+bric_data<-read.csv("models_with_logback/regression_features/bric2015.csv") # BRIC index
+humidity_data<-read.csv("models_with_logback/regression_features/humidity_climatology_1990_2020.csv") # ERA5 Specific Humidity
+temperature_data<-read.csv("models_with_logback/regression_features/temperature_climatology_1990_2020.csv") # ERA5 Specific Temperature
 

### 2025-06-23 models_with_logback/ES_EPIWEEK/EPIWEEK.Rmd
update

@@
 
-This function utilizes ensembles and single automatic ARIMAX models which have mean cases by epidemiological weeks over the last 2 years as exogenous variables.
+This function utilizes ensembles and single automatic ARIMAX models which have mean hospitalizations by epidemiological weeks over the last 2 years as exogenous variables.
 The function fits on a rolling window of previous 104 weeks for the state under analysis and a rolling windows with the same size with 1 week-lag for the exogenous variables to generate forecasts.

### 2025-06-23 models_without_logback/ES_EPIWEEK/EPIWEEK.Rmd
update

--- 2025-06-23 models_with_logback/ES_EPIWEEK/EPIWEEK.Rmd
@@
 # Loads the ADJACENT states models
-source("ES_EPIWEEK.R", local = TRUE, chdir = TRUE)
+source("ES_EPIWEEK_nolog.R", local = TRUE, chdir = TRUE)
 
@@
 ```{r}
-save.image("EPIWEEK_MODELS_influenza_hospitalization.Rdata")
+save.image("EPIWEEK_MODELS_influenza_hospitalization_nolog.Rdata")
 

### 2025-06-23 manuscript_results.Rmd
uptade

@@
 ```
-
 ------------
@@
 ```
-
 -------------

### 2025-06-23 models_with_logback/ES_ADJACENT/ADJACENT.Rmd
update
comments review

@@
 ---
-title: "ADJACENT STATES ARIMAX MODEL - Influenza hospitalization data"
+title: "ADJACENT STATES ARIMAX MODEL - with log-back transformations"
 author: "Victor Felix"
@@
 
-This function utilizes ensembles and single automatic ARIMAX models which have mean cases by adjacent states as exogenous variables.
-The function fits on a rolling window of previous 104 weeks for the state under analysis and a rolling windows with the same size with 1 week-lag for the exogenous variables to generate forecasts.
-It return some metrics that evaluate the performance of the mod
-els:target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles.
+Here we run ensembles and single automatic ARIMAX models for forecasting weekly hospitalizations in states on the contiguous U.S.
+These model use mean hospitalizations in adjacent states as exogenous variables.
+The models fit on a rolling window of previous 104 weeks for the state under analysis and a rolling windows with the same size with 1 week-lag for the exogenous variables to generate forecasts.
+These models include log-back transformations.
+It return some metrics that evaluate the performance of the models:target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles.
 The user can choose a  single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).

### 2025-06-23 models_with_logback/ES_ARIMA/ES_ARIMA.Rmd
comments review

--- 2025-04-25 NHSH_hospitalizations_seed/ES_ARIMA/ES_ARIMA.Rmd
@@
 ---
-title: "ARIMA ENSEMBLE MODELS - Influenza hospitalization data"
+title: "ARIMA MODELS - with log-back transformations"
 author: "Victor Felix"
@@
 
-This function utilizes ensembles and single automatic ARIMA models.
-The function fits a rolling window of 104 weeks to generate forecasts.
+Here we run ensembles and single automatic ARIMA models for forecasting weekly hospitalizations in states on the contiguous U.S.
+The models fit on a rolling window of previous 104 weeks to generate forecasts.
 It returns some metrics that evaluate the performance of the models: target_end_date, abs_error, cases, forecast, number of models, weighted interval score (WIS) and predictive quantiles (%).

### 2025-06-23 models_without_logback/ES_ADJACENT/ADJACENT.Rmd
update
comments review

--- 2025-04-25 NHSH_hospitalizations_seed_nolog_correct/ES_ADJACENT/ADJACENT.Rmd
@@
 ---
-title: "ADJACENT STATES ARIMAX MODEL - Influenza hospitalization data"
+title: "ADJACENT STATES ARIMAX MODEL - without log-back transformations"
 author: "Victor Felix"
@@
 
-This function utilizes ensembles and single automatic ARIMAX models which have mean cases by adjacent states as exogenous variables.
-The function fits on a rolling window of previous 104 weeks for the state under analysis and a rolling windows with the same size with 1 week-lag for the exogenous variables to generate forecasts.
-It return some metrics that evaluate the performance of the mod
-els:target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles.
+Here we run ensembles and single automatic ARIMAX models for forecasting weekly hospitalizations in states on the contiguous U.S.
+These model use mean hospitalizations in adjacent states as exogenous variables.
+The models fit on a rolling window of previous 104 weeks for the state under analysis and a rolling windows with the same size with 1 week-lag for the exogenous variables to generate forecasts.
+These models do not include log-back transformations.
+It return some metrics that evaluate the performance of the models:target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles.
 The user can choose a  single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).

### 2025-06-23 models_without_logback/ES_ARIMA/ES_ARIMA.Rmd
comments review

--- 2025-06-23 models_with_logback/ES_ARIMA/ES_ARIMA.Rmd
@@
 ---
-title: "ARIMA MODELS - with log-back transformations"
+title: "ARIMA MODELS - without log-back transformations"
 author: "Victor Felix"

### 2025-06-23 models_with_logback/ES_ARIMA/ES_ARIMA.Rmd
update

@@
 
-Here we run ensembles and single automatic ARIMA models for forecasting weekly hospitalizations in states on the contiguous U.S.
+Here we run ensembles and single automatic ARIMA models for forecasting weekly hospitalizations in 48 states on the contiguous U.S.
 The models fit on a rolling window of previous 104 weeks to generate forecasts.
+These models include log-back transformations.
 It returns some metrics that evaluate the performance of the models: target_end_date, abs_error, cases, forecast, number of models, weighted interval score (WIS) and predictive quantiles (%).

### 2025-06-23 models_without_logback/ES_ARIMA/ES_ARIMA.Rmd
update

@@
 
-Here we run ensembles and single automatic ARIMA models for forecasting weekly hospitalizations in states on the contiguous U.S.
+Here we run ensembles and single automatic ARIMA models for forecasting weekly hospitalizations in 48 states on the contiguous U.S.
 The models fit on a rolling window of previous 104 weeks to generate forecasts.
+These models do not include log-back transformations.
 It returns some metrics that evaluate the performance of the models: target_end_date, abs_error, cases, forecast, number of models, weighted interval score (WIS) and predictive quantiles (%).

### 2025-06-23 models_with_logback/ES_AVERAGE/AVERAGE.Rmd
reviewed comments

@@
 ---
-title: "AVERAGE STATES ARIMAX MODEL - Influenza hospitalization data"
+title: "AVERAGE STATES ARIMAX MODEL - with log-back transformations"
 author: "Victor Felix"
@@
 
-This function utilizes ensembles and single automatic ARIMAX models which have mean hospitalization on U.S.
+Here we run ensembles and single automatic ARIMAX models for forecasting weekly hospitalizations in 48 states on the contiguous U.S.
+These model use mean hospitalization across the U.S.
 states as exogenous variables.
 The function fits on a rolling window of previous 104 weeks for the state under analysis and a rolling windows with the same size with 1 week-lag for the exogenous variables to generate forecasts.
-It return some metrics that evaluate the performance of the mod
-els:target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles.
+These models include log-back transformations.
+It return some metrics that evaluate the performance of the models:target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles.
 The user can choose a  single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE).

### 2025-06-23 models_without_logback/ES_AVERAGE/AVERAGE.Rmd
reviewed comments

--- 2025-06-23 models_with_logback/ES_AVERAGE/AVERAGE.Rmd
@@
 ---
-title: "AVERAGE STATES ARIMAX MODEL - with log-back transformations"
+title: "AVERAGE STATES ARIMAX MODEL - without log-back transformations"
 author: "Victor Felix"
@@
 The function fits on a rolling window of previous 104 weeks for the state under analysis and a rolling windows with the same size with 1 week-lag for the exogenous variables to generate forecasts.
-These models include log-back transformations.
+These models do not include log-back transformations.
 It return some metrics that evaluate the performance of the models:target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles.
@@
 # Loads the AVERAGE states models
-source("ES_AVERAGE_48.R", local = TRUE, chdir = TRUE)
+source("ES_AVERAGE_48_nolog.R", local = TRUE, chdir = TRUE)
 
@@
 ```{r}
-save.image("AVERAGE_MODELS_influenza_hospitalization.Rdata")
+save.image("AVERAGE_MODELS_influenza_hospitalization_nolog.Rdata")
 ```

### 2025-06-23 models_with_logback/ES_EPIWEEK/EPIWEEK.Rmd
reviwed comments

--- 2025-06-23 models_with_logback/ES_EPIWEEK/EPIWEEK.Rmd
@@
 ---
-title: "EPIWEEK ARIMAX MODELS - Influenza hospitalization data"
+title: "EPIWEEK ARIMAX MODELS - with log-back transformations"
 author: "Victor Felix"
@@
 
-This function utilizes ensembles and single automatic ARIMAX models which have mean cases by epidemiological weeks over the last 2 years as exogenous variables.
+Here we run ensembles and single automatic ARIMAX models for forecasting weekly hospitalizations in 48 states on the contiguous U.S.
+These model use mean hospitalizations by epidemiological weeks over the last 2 years as exogenous variables.
 The function fits on a rolling window of previous 104 weeks for the state under analysis and a rolling windows with the same size with 1 week-lag for the exogenous variables to generate forecasts.
+These models include log-back transformations.
 It return some metrics that evaluate the performance of the models:target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles.

### 2025-06-23 models_without_logback/ES_EPIWEEK/EPIWEEK.Rmd
reviwed comments

@@
 ---
-title: "EPIWEEK ARIMAX MODELS - Influenza hospitalization data"
+title: "EPIWEEK ARIMAX MODELS - without log-back transformations"
 author: "Victor Felix"
@@
 
-This function utilizes ensembles and single automatic ARIMAX models which have mean hospitalizations by epidemiological weeks over the last 2 years as exogenous variables.
+Here we run ensembles and single automatic ARIMAX models for forecasting weekly hospitalizations in 48 states on the contiguous U.S.
+These model use mean hospitalizations by epidemiological weeks over the last 2 years as exogenous variables.
 The function fits on a rolling window of previous 104 weeks for the state under analysis and a rolling windows with the same size with 1 week-lag for the exogenous variables to generate forecasts.
+These models do not include log-back transformations.
 It return some metrics that evaluate the performance of the models:target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles.

### 2025-06-23 models_with_logback/ES_TEMPERATURE/TEMPERATURE.Rmd
reviwed comments

@@
 ---
-title: "TEMPERATURE ARIMAX MODELS - Influenza hospitalization data"
+title: "TEMPERATURE ARIMAX MODELS - with log-back transformations"
 author: "Victor Felix"
@@
 
-This function utilizes ensembles and single automatic ARIMAX models which have mean temperature as exogenous variables.
+Here we run ensembles and single automatic ARIMAX models for forecasting weekly hospitalizations in 48 states on the contiguous U.S.
+These model use mean temperature as exogenous variables.
 The function fits on a rolling window of previous 104 weeks for the state under analysis and a rolling windows with the same size with 1 week-lag for the exogenous variables to generate forecasts.
+These models include log-back transformations.
 It return some metrics that evaluate the performance of the models:target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles.

### 2025-06-23 models_without_logback/ES_TEMPERATURE/TEMPERATURE.Rmd
reviwed comments

@@
 ---
-title: "TEMPERATURE ARIMAX MODELS - Influenza hospitalization data"
+title: "TEMPERATURE ARIMAX MODELS - without log-back transformations"
 author: "Victor Felix"
@@
 
-This function utilizes ensembles and single automatic ARIMAX models which have mean temperature as exogenous variables.
+Here we run ensembles and single automatic ARIMAX models for forecasting weekly hospitalizations in 48 states on the contiguous U.S.
+These model use mean temperature as exogenous variables.
 The function fits on a rolling window of previous 104 weeks for the state under analysis and a rolling windows with the same size with 1 week-lag for the exogenous variables to generate forecasts.
+These models do not include log-back transformations.
 It return some metrics that evaluate the performance of the models:target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles.

### 2025-06-23 models_with_logback/ES_ADJACENT/ADJACENT.Rmd
update
comments reviewed

@@
 
-Here we run ensembles and single automatic ARIMAX models for forecasting weekly hospitalizations in states on the contiguous U.S.
+Here we run ensembles and single automatic ARIMAX models for forecasting weekly hospitalizations in 48 states on the contiguous U.S.
 These model use mean hospitalizations in adjacent states as exogenous variables.
@@
 AUTO ADJACENT WEEK1 - ALL STATES
-
-IF YOU WANT A FAST RESULT RUN ONLY THE AUTO_example, the others take longer.
 

### 2025-07-22 manuscript_results.Rmd
updated code

--- 2025-06-23 manuscript_results.Rmd
@@
 ```{r}
-plot_forecast <- function(i, ii,
-                          list_of_states,
-                          AUTO_AVERAGE_WEEK1_list,
-                          AUTO_AVERAGE_WEEK2_list,
-                          AUTO_AVERAGE_WEEK3_list,
-                          AUTO_AVERAGE_WEEK4_list) {
+#plot_forecast <- function(i, ii,
+#                          list_of_states,
+#                          AUTO_AVERAGE_WEEK1_list,
+#                          AUTO_AVERAGE_WEEK2_list,
+#                          AUTO_AVERAGE_WEEK3_list,
+#                          AUTO_AVERAGE_WEEK4_list) {
 
   # Define date range for 2-year history
-  end_date <- as.Date(AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i] - 7)
-  start_date <- end_date - (7 * 103)
+#  end_date <- as.Date(AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i] - 7)
+#  start_date <- end_date - (7 * 103)
 
   # Subset observed data
-  subset_data <- subset(list_of_states[[ii]],
-                        target_end_date >= start_date & target_end_date <= end_date)
+#  subset_data <- subset(list_of_states[[ii]],
+#                        target_end_date >= start_date & target_end_date <= end_date)
 
   # Build forecast dataframe
-  forecast_df <- data.frame(
-    Dates = c(
-      AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i],
-      AUTO_AVERAGE_WEEK2_list[[ii]]$target_end_date[i],
-      AUTO_AVERAGE_WEEK3_list[[ii]]$target_end_date[i],
-      AUTO_AVERAGE_WEEK4_list[[ii]]$target_end_date[i]
-    ),
-    Hosp = c(
-      AUTO_AVERAGE_WEEK1_list[[ii]]$cases[i],
-      AUTO_AVERAGE_WEEK2_list[[ii]]$cases[i],
-      AUTO_AVERAGE_WEEK3_list[[ii]]$cases[i],
-      AUTO_AVERAGE_WEEK4_list[[ii]]$cases[i]
-    ),
-    Value = c(
-      AUTO_AVERAGE_WEEK1_list[[ii]]$forecasts[i],
-      AUTO_AVERAGE_WEEK2_list[[ii]]$forecasts[i],
-      AUTO_AVERAGE_WEEK3_list[[ii]]$forecasts[i],
-      AUTO_AVERAGE_WEEK4_list[[ii]]$forecasts[i]
-    ),
-    Lower = c(
-      AUTO_AVERAGE_WEEK1_list[[ii]][["0.025"]][i],
-      AUTO_AVERAGE_WEEK2_list[[ii]][["0.025"]][i],
-      AUTO_AVERAGE_WEEK3_list[[ii]][["0.025"]][i],
-      AUTO_AVERAGE_WEEK4_list[[ii]][["0.025"]][i]
-    ),
-    Upper = c(
-      AUTO_AVERAGE_WEEK1_list[[ii]][["0.975"]][i],
-      AUTO_AVERAGE_WEEK2_list[[ii]][["0.975"]][i],
-      AUTO_AVERAGE_WEEK3_list[[ii]][["0.975"]][i],
-      AUTO_AVERAGE_WEEK4_list[[ii]][["0.975"]][i]
-    )
-  )
+#  forecast_df <- data.frame(
+#    Dates = c(
+#      AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i],
+#      AUTO_AVERAGE_WEEK2_list[[ii]]$target_end_date[i],
+#      AUTO_AVERAGE_WEEK3_list[[ii]]$target_end_date[i],
+#      AUTO_AVERAGE_WEEK4_list[[ii]]$target_end_date[i]
+#    ),
+#    Hosp = c(
+#      AUTO_AVERAGE_WEEK1_list[[ii]]$cases[i],
+#      AUTO_AVERAGE_WEEK2_list[[ii]]$cases[i],
+#      AUTO_AVERAGE_WEEK3_list[[ii]]$cases[i],
+#      AUTO_AVERAGE_WEEK4_list[[ii]]$cases[i]
+#    ),
+#    Value = c(
+#      AUTO_AVERAGE_WEEK1_list[[ii]]$forecasts[i],
+#      AUTO_AVERAGE_WEEK2_list[[ii]]$forecasts[i],
+#      AUTO_AVERAGE_WEEK3_list[[ii]]$forecasts[i],
+#      AUTO_AVERAGE_WEEK4_list[[ii]]$forecasts[i]
+#    ),
+#    Lower = c(
+#      AUTO_AVERAGE_WEEK1_list[[ii]][["0.025"]][i],
+#      AUTO_AVERAGE_WEEK2_list[[ii]][["0.025"]][i],
+#      AUTO_AVERAGE_WEEK3_list[[ii]][["0.025"]][i],
+#      AUTO_AVERAGE_WEEK4_list[[ii]][["0.025"]][i]
+#    ),
+#    Upper = c(
+#      AUTO_AVERAGE_WEEK1_list[[ii]][["0.975"]][i],
+#      AUTO_AVERAGE_WEEK2_list[[ii]][["0.975"]][i],
+#      AUTO_AVERAGE_WEEK3_list[[ii]][["0.975"]][i],
+#      AUTO_AVERAGE_WEEK4_list[[ii]][["0.975"]][i]
+#    )
+#  )
 
   # Build the plot
-  ggplot() +
-    geom_point(data = subset_data,
-               aes(x = target_end_date, y = cases, color = "Fitting data"),
-               size = 2.5) +
-    geom_ribbon(data = forecast_df,
-                aes(x = Dates, ymin = Lower, ymax = Upper),
-                fill = "gray", alpha = 0.4) +
-    geom_point(data = forecast_df,
-               aes(x = Dates, y = Value, color = "Mean forecast"),
-               size = 3) +
-    geom_point(data = forecast_df,
-               aes(x = Dates, y = Value, color = "Mean forecast"),
-               shape = 21, fill = "white", size = 2) +
-    geom_point(data = forecast_df,
-               aes(x = Dates, y = Hosp, color = "Hospitalizations"),
-               shape = 16, size = 2.5) +
-    scale_color_manual(name = "Legend",
-                       values = c("Fitting data" = "red",
-                                  "Mean forecast" = "cyan3",
-                                  "Hospitalizations" = "black")) +
-    scale_x_date(date_breaks = "1 month", date_labels = "%b %y") +
-    labs(title = paste("State:", names(list_of_states)[ii], "- Week Index:", i),
-         subtitle = "Gray ribbon represents 95% confidence interval",
-         x = "", y = "Hospitalizations") +
-    theme_minimal() +
-    theme(plot.title = element_text(size = 18, face = "bold"),
-          plot.subtitle = element_text(size = 14),
-          axis.text.x = element_text(angle = 45, hjust = 1))
-}
-
-for (ii in 1:48) {
-  for (i in 60:70) {
+#  ggplot() +
+#    geom_point(data = subset_data,
+#               aes(x = target_end_date, y = cases, color = "Fitting data"),
+#               size = 2.5) +
+#    geom_ribbon(data = forecast_df,
+#                aes(x = Dates, ymin = Lower, ymax = Upper),
+#                fill = "gray", alpha = 0.4) +
+#    geom_point(data = forecast_df,
+#               aes(x = Dates, y = Value, color = "Mean forecast"),
+#               size = 3) +
+#    geom_point(data = forecast_df,
+#               aes(x = Dates, y = Value, color = "Mean forecast"),
+#               shape = 21, fill = "white", size = 2) +
+#    geom_point(data = forecast_df,
+#               aes(x = Dates, y = Hosp, color = "Hospitalizations"),
+#               shape = 16, size = 2.5) +
+#    scale_color_manual(name = "Legend",
+#                       values = c("Fitting data" = "red",
+#                                  "Mean forecast" = "cyan3",
+#                                  "Hospitalizations" = "black")) +
+#    scale_x_date(date_breaks = "1 month", date_labels = "%b %y") +
+#    labs(title = paste("State:", names(list_of_states)[ii], "- Week Index:", i),
+#         subtitle = "Gray ribbon represents 95% confidence interval",
+#         x = "", y = "Hospitalizations") +
+#    theme_minimal() +
+#    theme(plot.title = element_text(size = 18, face = "bold"),
+#          plot.subtitle = element_text(size = 14),
+#          axis.text.x = element_text(angle = 45, hjust = 1))
+#}
+
+#for (ii in 1:48) {
+#  for (i in 60:70) {
     # Safely check if index exists
-    if (i <= nrow(AUTO_AVERAGE_WEEK1_list[[ii]])) {
-      p <- plot_forecast(i, ii,
-                         list_of_states,
-                         AUTO_AVERAGE_WEEK1_list,
-                         AUTO_AVERAGE_WEEK2_list,
-                         AUTO_AVERAGE_WEEK3_list,
-                         AUTO_AVERAGE_WEEK4_list)
+#    if (i <= nrow(AUTO_AVERAGE_WEEK1_list[[ii]])) {
+#      p <- plot_forecast(i, ii,
+#                         list_of_states,
+#                         AUTO_AVERAGE_WEEK1_list,
+#                         AUTO_AVERAGE_WEEK2_list,
+#                         AUTO_AVERAGE_WEEK3_list,
+#                         AUTO_AVERAGE_WEEK4_list)
       # Show the plot
-      print(p)
-    }
-  }
-}
+#      print(p)
+#    }
+#  }
+#}
 
@@
 
-# Combine your main plot with the custom legend
+# Combine main plot with the custom legend
 figure4 <- combined_plot_clean + 
@@
   patchwork::wrap_elements(legend_only) + 
-  plot_layout(widths = c(4, 0.05, 0.4))  # Adjust widths as needed
+  plot_layout(widths = c(4, 0.05, 0.4))
 
@@
 
-ERASE
-
-```{r}
-# Create a custom color mapping for each variable
-custom_colors <- c(
-  "AUTO_AR_LB" = colors_ar[1], "ES27_AR_LB" = colors_ar[2], "ES64_AR_LB" = colors_ar[3],
-  "AUTO_EPI_LB" = colors_epi[1], "ES27_EPI_LB" = colors_epi[2], "ES64_EPI_LB" = colors_epi[3],
-  "AUTO_TMP_LB" = colors_tmp[1], "ES27_TMP_LB" = colors_tmp[2], "ES64_TMP_LB" = colors_tmp[3],
-  "AUTO_ADJ_LB" = colors_adj[1], "ES27_ADJ_LB" = colors_adj[2], "ES64_ADJ_LB" = colors_adj[3],
-  "AUTO_AVG_LB" = colors_avg[1], "ES27_AVG_LB" = colors_avg[2], "ES64_AVG_LB" = colors_avg[3],
-
-  "AUTO_AR" = colors_ar[1], "ES27_AR" = colors_ar[2], "ES64_AR" = colors_ar[3],
-  "AUTO_EPI" = colors_epi[1], "ES27_EPI" = colors_epi[2], "ES64_EPI" = colors_epi[3],
-  "AUTO_TMP" = colors_tmp[1], "ES27_TMP" = colors_tmp[2], "ES64_TMP" = colors_tmp[3],
-  "AUTO_ADJ" = colors_adj[1], "ES27_ADJ" = colors_adj[2], "ES64_ADJ" = colors_adj[3],
-  "AUTO_AVG" = colors_avg[1], "ES27_AVG" = colors_avg[2], "ES64_AVG" = colors_avg[3]
-)
-
-# --------- WEEK1 MODELS ------------- #
-plot1<- ggplot(W1, aes(x = Best_Result, fill = Best_Result)) +  
-  geom_bar() +  
-  scale_fill_manual(values = custom_colors) +  
-  labs(title = "Best model on the 48 U.S. states (W1)",       
-       x = "", y = "Number of states") +  
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
-  facet_wrap(~ Model_Type, scales = "free_x")
-
-plot1
-
-#ggsave("week1_best_models.jpg", plot1)
-
-# --------- WEEK2 MODELS ------------- #
-plot2<-ggplot(W2, aes(x = Best_Result, fill = Best_Result))  +  
-  geom_bar() +  
-  scale_fill_manual(values = custom_colors) +  
-  labs(title = "Best model on the 48 U.S. states (W2)",       
-       x = "", y = "Number of states") +  
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
-  facet_wrap(~ Model_Type, scales = "free_x")
-
-#ggsave("week2_best_models.jpg", plot2)
-
-plot2
-
-# --------- WEEK3 MODELS ------------- #
-plot3<-ggplot(W3, aes(x = Best_Result, fill = Best_Result))  +  
-  geom_bar() +  
-  scale_fill_manual(values = custom_colors) +  
-  labs(title = "Best model on the 48 U.S. states (W3)",       
-       x = "", y = "Number of states") +  
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
-  facet_wrap(~ Model_Type, scales = "free_x")
-plot3
-
-#ggsave("week3_best_models.jpg", plot3)
-
-# --------- WEEK4 MODELS ------------- #
-plot4<-ggplot(W4, aes(x = Best_Result, fill = Best_Result)) +  
-  geom_bar() +  
-  scale_fill_manual(values = custom_colors) +  
-  labs(title = "Best model on the 48 U.S. states (W4)",       
-       x = "", y = "Number of states") +  
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  
-  facet_wrap(~ Model_Type, scales = "free_x")
-plot4
-
-#ggsave("week4_best_models.jpg", plot4)
-
-```
-
-ERASE
-
-```{r}
-# Add a column to indicate the week ahead
-W1$Week_Ahead <- "W1"
-W2$Week_Ahead <- "W2"
-W3$Week_Ahead <- "W3"
-W4$Week_Ahead <- "W4"
-
-# Combine all weeks into one dataframe
-all_weeks <- bind_rows(W1, W2, W3, W4)
-
-# Create the plot
-combined_plot <- ggplot(all_weeks, aes(x = Best_Result, fill = Best_Result)) +
-  geom_bar() +
-  scale_fill_manual(values = custom_colors) +
-  labs(title = "Best model based on the lowest mean WIS",
-       subtitle = "Results for the 48 states on the contiguous U.S. by target week",
-       x = "", y = "Number of states",
-       fill = "Best model" ) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        plot.title = element_text(size = 16, face = "bold")) +
-  facet_grid(Week_Ahead ~ Model_Type, scales = "free_x")
-
-# Save and display the plot
-#ggsave("Fig4.jpg", combined_plot, width =8, height =5)
-print(combined_plot)
-
-```
-
-Now we use a Wilcox test with Holm p-value adjustment for repeated comparisons to evaluate the differences in model performances.
+Now we use a Wilcoxon test with Holm p-value adjustment for repeated comparisons to evaluate the differences in model performances.
 
@@
 
-# Create a new column "Model_Type" based on model with and withouth log-back transformation
+# Create a new column "Model_Type" based on model with and without log-back transformation
 p_values_and_impr$Model_Type <- ifelse(grepl("_LB$", p_values_and_impr$Model), "LB", "no_LB")
@@
 p_values_and_impr
-```
-ERASE
-
-```{r}
-
-#new_long_data <- summary_impr %>%
-#  mutate(
-#    LogBack = grepl("_LB$", Model),
-#    Family = sub("_.*", "", Model),  # take everything before the first underscore
-#    Method = sub("^.*?_(AR|ADJ|EPI|AVG|TMP)(_LB)?$", "\\1", Model)
-#  )
-
-#new_long_data <- new_long_data %>%
-#  mutate(Model = ifelse(Model == "AUTO_AR", "AUTO_AAR", Model))
-
-#model <- lm(m ~ Model, data = new_long_data)
-#summary(model)
-
-#model
-#new_long_data$Model
-# Get tidy summary of coefficients
-#tidy_model <- broom::tidy(model)
-
-# Filter out intercept for clarity if you want
-#tidy_model_no_intercept <- tidy_model %>% filter(term != "(Intercept)")
-
-# Plot coefficients with 95% confidence intervals
-#tidy_model_no_intercept <- tidy_model_no_intercept %>%
-#  mutate(term = forcats::fct_reorder(term, estimate, .desc = TRUE))
-
-# Plot
-#ggplot(tidy_model_no_intercept, aes(x = term, y = estimate)) +
-#  geom_point() +
-#  geom_errorbar(aes(ymin = estimate - std.error * 1.96,
-#                    ymax = estimate + std.error * 1.96),
-#                width = 0.2) +
-#  coord_flip() +  # flip for better readability
-#  labs(
-#    title = "Effect Size of Model Types on outcome",
-#    subtitle =  "Estimates relative to baseline (intercept)",
-#    y = "Estimate",
-#    x = ""
-#  ) +
-#  theme_minimal()
-
 ```
@@
 ES_1WEEK<-ES_1WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
-
-#ggsave(ES_1WEEK,height = 4.5, width = 6.5,  filename="Fig5A.jpg")  
 ```
@@
 ES_2WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
-
-#ggsave(ES_2WEEK, height = 4.5, width = 6.5, filename="Fig5B.jpg")  
 ```
@@
 
-#ggsave(ES_3WEEK, height = 4.5, width = 6.5, filename="Fig5C.jpg")  
 ```
@@
 
-#ggsave(ES_4WEEK, height = 4.5, width = 6.5,  filename="Fig5D.jpg")  
 ```
@@
 
-#ggsave(model_improv, height = 4.5, width = 9,  filename="Fig10.jpg")  
 ```

### 2025-07-28 manuscript_results.Rmd
update

@@
     summarize(
-      AUTO_AR = mean(AUTO_AR, na.rm = TRUE),  # Sum across all states
-      AUTO_AVG_LB = mean(AUTO_AVG_LB, na.rm = TRUE)           # Sum across all states
+      AUTO_AR = mean(AUTO_AR, na.rm = TRUE),  # Mean across all states
+      AUTO_AVG_LB = mean(AUTO_AVG_LB, na.rm = TRUE) # Mean across all states
     ) %>%

### 2025-07-28 manuscript_results.Rmd
update code

@@
          caption = "m = mean difference, sd = standard deviation",
-       x = "Target Ahead",
+       x = "Target Week",
        y = "") + 
@@
     y = "WIS Difference",
-    x = ""
+    x = "Model components"
   ) +

### 2025-08-05 manuscript_results.Rmd
update

@@
 ```
-Evaluating other forecasts
-
-```{r}
-#plot_forecast <- function(i, ii,
-#                          list_of_states,
-#                          AUTO_AVERAGE_WEEK1_list,
-#                          AUTO_AVERAGE_WEEK2_list,
-#                          AUTO_AVERAGE_WEEK3_list,
-#                          AUTO_AVERAGE_WEEK4_list) {
-
-  # Define date range for 2-year history
-#  end_date <- as.Date(AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i] - 7)
-#  start_date <- end_date - (7 * 103)
-
-  # Subset observed data
-#  subset_data <- subset(list_of_states[[ii]],
-#                        target_end_date >= start_date & target_end_date <= end_date)
-
-  # Build forecast dataframe
-#  forecast_df <- data.frame(
-#    Dates = c(
-#      AUTO_AVERAGE_WEEK1_list[[ii]]$target_end_date[i],
-#      AUTO_AVERAGE_WEEK2_list[[ii]]$target_end_date[i],
-#      AUTO_AVERAGE_WEEK3_list[[ii]]$target_end_date[i],
-#      AUTO_AVERAGE_WEEK4_list[[ii]]$target_end_date[i]
-#    ),
-#    Hosp = c(
-#      AUTO_AVERAGE_WEEK1_list[[ii]]$cases[i],
-#      AUTO_AVERAGE_WEEK2_list[[ii]]$cases[i],
-#      AUTO_AVERAGE_WEEK3_list[[ii]]$cases[i],
-#      AUTO_AVERAGE_WEEK4_list[[ii]]$cases[i]
-#    ),
-#    Value = c(
-#      AUTO_AVERAGE_WEEK1_list[[ii]]$forecasts[i],
-#      AUTO_AVERAGE_WEEK2_list[[ii]]$forecasts[i],
-#      AUTO_AVERAGE_WEEK3_list[[ii]]$forecasts[i],
-#      AUTO_AVERAGE_WEEK4_list[[ii]]$forecasts[i]
-#    ),
-#    Lower = c(
-#      AUTO_AVERAGE_WEEK1_list[[ii]][["0.025"]][i],
-#      AUTO_AVERAGE_WEEK2_list[[ii]][["0.025"]][i],
-#      AUTO_AVERAGE_WEEK3_list[[ii]][["0.025"]][i],
-#      AUTO_AVERAGE_WEEK4_list[[ii]][["0.025"]][i]
-#    ),
-#    Upper = c(
-#      AUTO_AVERAGE_WEEK1_list[[ii]][["0.975"]][i],
-#      AUTO_AVERAGE_WEEK2_list[[ii]][["0.975"]][i],
-#      AUTO_AVERAGE_WEEK3_list[[ii]][["0.975"]][i],
-#      AUTO_AVERAGE_WEEK4_list[[ii]][["0.975"]][i]
-#    )
-#  )
-
-  # Build the plot
-#  ggplot() +
-#    geom_point(data = subset_data,
-#               aes(x = target_end_date, y = cases, color = "Fitting data"),
-#               size = 2.5) +
-#    geom_ribbon(data = forecast_df,
-#                aes(x = Dates, ymin = Lower, ymax = Upper),
-#                fill = "gray", alpha = 0.4) +
-#    geom_point(data = forecast_df,
-#               aes(x = Dates, y = Value, color = "Mean forecast"),
-#               size = 3) +
-#    geom_point(data = forecast_df,
-#               aes(x = Dates, y = Value, color = "Mean forecast"),
-#               shape = 21, fill = "white", size = 2) +
-#    geom_point(data = forecast_df,
-#               aes(x = Dates, y = Hosp, color = "Hospitalizations"),
-#               shape = 16, size = 2.5) +
-#    scale_color_manual(name = "Legend",
-#                       values = c("Fitting data" = "red",
-#                                  "Mean forecast" = "cyan3",
-#                                  "Hospitalizations" = "black")) +
-#    scale_x_date(date_breaks = "1 month", date_labels = "%b %y") +
-#    labs(title = paste("State:", names(list_of_states)[ii], "- Week Index:", i),
-#         subtitle = "Gray ribbon represents 95% confidence interval",
-#         x = "", y = "Hospitalizations") +
-#    theme_minimal() +
-#    theme(plot.title = element_text(size = 18, face = "bold"),
-#          plot.subtitle = element_text(size = 14),
-#          axis.text.x = element_text(angle = 45, hjust = 1))
-#}
-
-#for (ii in 1:48) {
-#  for (i in 60:70) {
-    # Safely check if index exists
-#    if (i <= nrow(AUTO_AVERAGE_WEEK1_list[[ii]])) {
-#      p <- plot_forecast(i, ii,
-#                         list_of_states,
-#                         AUTO_AVERAGE_WEEK1_list,
-#                         AUTO_AVERAGE_WEEK2_list,
-#                         AUTO_AVERAGE_WEEK3_list,
-#                         AUTO_AVERAGE_WEEK4_list)
-      # Show the plot
-#      print(p)
-#    }
-#  }
-#}
-
-```
 
